### 2020-02-06 paper/main.tex
set up .tex

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data,
  April 27, 2020, Heraklion, Crete, Greece}
\acmPrice{15.00}
% \acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Preserving Reciprocal Consistency in Distributed Graph Databases}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Paul Ezhilchelvan}
\email{paul.ezhilchelvan@ncl.ac.uk}
\orcid{0000-0002-6190-5685}
\affiliation{%
  \institution{Newcastle University}
  \city{Newcastle}
  \country{UK}
  \postcode{NE4 5TG}
}

\author{Isi Mitrani}
\email{isi.mitrani@ncl.ac.uk}
\orcid{0000-0002-7797-7755}
\affiliation{%
  \institution{Newcastle University}
  \city{Newcastle}
  \country{UK}
  \postcode{NE4 5TG}
}

\author{Jack Waudby}
\email{j.waudby2@ncl.ac.uk}
\affiliation{%
  \institution{Newcastle University}
  \city{Newcastle}
  \country{UK}
  \postcode{NE4 5TG}
}

\author{Jim Webber}
\email{jim.webber@neo4j.com}
\affiliation{%
  \institution{Neo4j}
  \streetaddress{Union House,182-194 Union Street}
  \city{London}
  \country{UK}
  \postcode{SE1 0LH}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Ezhilchelvan, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A solution for maintaining reciprocal consistency in distributed graph databases is presented.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Data Management~Graph Databases}
\ccsdesc[300]{Data Management~Reciprocal Consistency}

% \ccsdesc[500]{Computer systems organization~Embedded systems}
% \ccsdesc[300]{Computer systems organization~Redundancy}
% \ccsdesc{Computer systems organization~Robotics}
% \ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{reciprocal consistency, graph databases, collision detection}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{sampleteaser}
%   \caption{Seattle Mariners at Spring Training, 2010.}
%   \Description{Enjoying the baseball game from the third-base
%   seats. Ichiro Suzuki preparing to bat.}
%   \label{fig:teaser}
% \end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
In recent years there has been an increased interest in and use of graph technologies. Application areas are wide reaching from healthcare to more traditional use cases such as social networks. Graph systems can be categorised into graph databases and graph processing frameworks. This paper focuses on graph databases, which at a high-level differ from relational databases in that they provide application programmers with a graph data model, as oppose to relations linked through foreign keys. Specifically, graph databases use the labelled property graph data model consisting of vertices and edges, which each can have properties and be group through labels.

As the sizes of graphs grow the limits of single node, replicated graph databases are being reached. This has lead to the development of horizontally scalable distributed graph databases. Moving from a single-node database to one partitioned and distributed across servers introduces several interesting problems regarding data consistency. Graphs place a high demand on the underlying database and maintaining structural integrity across partitions - called \textit{reciprocal consistency} - in a distributed graph database is non-trivial.

Recent work has investigated how violation of reciprocal consistency can lead to semantic corruption at alarmingly rates. In order to prevent this form of inconsistent one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs. The approach taken in this paper is to build on earlier work and develop a protocol that provides weaker consistency guarantees whilst still preserving reciprocal consistency.
\section{Distributed Graph Databases}

In a graph database records represent vertices containing vertex properties and an adjacency list containing edge "pointers" to other vertices along with associated properties. Whilst edges typically have direction they are physically implemented as one would implement an undirected graph with meta-data indicating the direction and some reciprocal information regarding the edge. For example, an edge E joining two vertices A and B results in reciprocal entries for E being written in the adjacency lists for both A and B. If this edge represented the edge between an author and their book, the following reciprocal information would be stored A would store the information \textit{wrote B} and B would store the information \textit{written by A}. This permits bi-directional traversal of the edge, a query reading either A or B would be able to reify the edge correctly. When the adjacency lists for vertices are mutually compatible they are said to be \textit{reciprocally consistent}. The challenge is maintaining reciprocally consistency in a distributed graph database.

A common approach to partitioning a graph is to use a balanced edge-cut. Vertices are evenly assigned to partitions in a manner that minimises the number of edges that span partitions - referred to as \textbf{distributed edges}. That is the number of entries in a vertex's adjacency list that point to vertices on remote servers is minimised. However, regardless of the graph partitioning algorithm used there will exist a non-negligible number of distributed edges. Without sufficient isolation between concurrent updates to a distributed edge, reciprocal consistency can be violated.

To preserve reciprocal consistency distributed edges could be updated through a protocols such as 2-Phase Commit with isolation provided via 2-Phase Locking. However, such an approach is well known to negatively impact scalability and performance. The approach taken here is provide a weaker isolation model whilst still preserving reciprocal consistency.

\section{Preserving Reciprocal Consistency}
Assume a graph database partitioned across servers in a cluster, with each partition being a Raft cluster providing linearizable reads and writes to the data objects it is responsible for. One strawman approach to preserving reciprocal consistency would be for update operations to take an exclusive lock on both records that make up the distributed edge, releasing locks when the update has been completed. Locking is a pessimistic approach and has the implicit assumption that concurrent updates will be common, this is not the case in graph databases. A second approach would be having a read operation read both records to verify the information is consistent. The problem here is that read operations dominant graph workloads and every read of a distributed edge would incur the penalty of traversing the network twice.

The approach developed leverages 1) the frequency of concurrent update operations to the same edge is low 2) a logical write of an edge consists of writes to two physical records 3) read operations only access a single record. We now outline a variant of the \textbf{Collision Detection} mechanism presented in REF and discuss it's combination with Raft. % assume a mechanism to guarantee consistency of operations within partitions
\subsection{Collision Detection Mechanism}

Initially, each write to a distributed edge is provisional. They become permanent only if, and when, the transaction that contains them is allowed to commit.

Provisional writes on a given record can occur only one at a time and each is time-stamped using a combination of the transaction ID and the local host's clock - used to uniquely identify the operation. This is stored in a order-preserving provisional write list. Thus, when a transaction attempts to write a given record, it can identify all other transactions, called predecessors (if any), that have earlier wrote that record provisionally along with the transaction responsible for latest committed value. Currently no higher information regarding specific operation type is assumed. Read operations receive the latest committed version of a record and ignore any provisionally updated values.

\textbf{Collision detection}: An edge has a \textit{tail record} and \textit{head record} - a write operation by a transaction must write to \textit{both} records. The order in which updates occur is not constrained, that is, one update may update the head and then after a network delay the tail, the opposite can also occur with equal chance.

In order to distinguish between which write a transaction they are given labels 1 and 2, which act as 'history' meta-data indicating the partition where a transaction started and completed its write. The meta-data consists of 1) local timestamp, 2) transaction ID and 3) order label.

The following rule is applied: \textbf{Cancellation rule:} If, by the time an operation is performing it's second update, if a previous provisional write labeled 1 has been observed, but the corresponding label 2 has not been observed, then this write is cancelled. In other words, a write is cancelled if it has observed the start of a previous attempt, but not its completion. This prevents writes happening out-of-order. When an update is cancelled, the transaction containing it is aborted and all its provisional records are erased.

\subsection{Alternative Methods}

\section{Evaluation \& Results}

In order to test the throughput and latency of the proposed mechanism a discrete event orientated simulation has been performed. Collision detection has been tested against the simple approach defined in Section 3.

\section{Related Work}

\section{Conclusions}

% \section{Introduction}
% ACM's consolidated article template, introduced in 2017, provides a
% consistent \LaTeX\ style for use across ACM publications, and
% incorporates accessibility and metadata-extraction functionality
% necessary for future Digital Library endeavors. Numerous ACM and
% SIG-specific \LaTeX\ templates have been examined, and their unique
% features incorporated into this single new template.

% If you are new to publishing with ACM, this document is a valuable
% guide to the process of preparing your work for publication. If you
% have published with ACM before, this document provides insight and
% instruction into more recent changes to the article template.

% The ``\verb|acmart|'' document class can be used to prepare articles
% for any ACM publication --- conference or journal, and for any stage
% of publication, from review to final ``camera-ready'' copy, to the
% author's own version, with {\itshape very} few changes to the source.

% % \section{Template Overview}
% % As noted in the introduction, the ``\verb|acmart|'' document class can
% % be used to prepare many different kinds of documentation --- a
% % double-blind initial submission of a full-length technical paper, a
% % two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
% % journal article, a SIGCHI Extended Abstract, and more --- all by
% % selecting the appropriate {\itshape template style} and {\itshape
% %   template parameters}.

% % This document will explain the major features of the document
% % class. For further information, the {\itshape \LaTeX\ User's Guide} is
% % available from
% % \url{https://www.acm.org/publications/proceedings-template}.

% % \subsection{Template Styles}

% % The primary parameter given to the ``\verb|acmart|'' document class is
% % the {\itshape template style} which corresponds to the kind of publication
% % or SIG publishing the work. This parameter is enclosed in square
% % brackets and is a part of the {\verb|documentclass|} command:
% % \begin{verbatim}
% %   \documentclass[STYLE]{acmart}
% % \end{verbatim}

% % Journals use one of three template styles. All but three ACM journals
% % use the {\verb|acmsmall|} template style:
% % \begin{itemize}
% % \item {\verb|acmsmall|}: The default journal template style.
% % \item {\verb|acmlarge|}: Used by JOCCH and TAP.
% % \item {\verb|acmtog|}: Used by TOG.
% % \end{itemize}

% % The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
% % \begin{itemize}
% % \item {\verb|acmconf|}: The default proceedings template style.
% % \item{\verb|sigchi|}: Used for SIGCHI conference articles.
% % \item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
% % \item{\verb|sigplan|}: Used for SIGPLAN conference articles.
% % \end{itemize}

% % \subsection{Template Parameters}

% % In addition to specifying the {\itshape template style} to be used in
% % formatting your work, there are a number of {\itshape template parameters}
% % which modify some part of the applied template style. A complete list
% % of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

% % Frequently-used parameters, or combinations of parameters, include:
% % \begin{itemize}
% % \item {\verb|anonymous,review|}: Suitable for a ``double-blind''
% %   conference submission. Anonymizes the work and includes line
% %   numbers. Use with the \verb|\acmSubmissionID| command to print the
% %   submission's unique ID on each page of the work.
% % \item{\verb|authorversion|}: Produces a version of the work suitable
% %   for posting by the author.
% % \item{\verb|screen|}: Produces colored hyperlinks.
% % \end{itemize}

% % This document uses the following string as the first command in the
% % source file:
% % \begin{verbatim}
% % \documentclass[sigplan,screen]{acmart}
% % \end{verbatim}

% % \section{Modifications}

% % Modifying the template --- including but not limited to: adjusting
% % margins, typeface sizes, line spacing, paragraph and list definitions,
% % and the use of the \verb|\vspace| command to manually adjust the
% % vertical spacing between elements of your work --- is not allowed.

% % {\bfseries Your document will be returned to you for revision if
% %   modifications are discovered.}

% % \section{Typefaces}

% % The ``\verb|acmart|'' document class requires the use of the
% % ``Libertine'' typeface family. Your \TeX\ installation should include
% % this set of packages. Please do not substitute other typefaces. The
% % ``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
% % as they will override the built-in typeface families.

% % \section{Title Information}

% % The title of your work should use capital letters appropriately -
% % \url{https://capitalizemytitle.com/} has useful rules for
% % capitalization. Use the {\verb|title|} command to define the title of
% % your work. If your work has a subtitle, define it with the
% % {\verb|subtitle|} command.  Do not insert line breaks in your title.

% % If your title is lengthy, you must define a short version to be used
% % in the page headers, to prevent overlapping text. The \verb|title|
% % command has a ``short title'' parameter:
% % \begin{verbatim}
% %   \title[short title]{full title}
% % \end{verbatim}

% % \section{Authors and Affiliations}

% % Each author must be defined separately for accurate metadata
% % identification. Multiple authors may share one affiliation. Authors'
% % names should not be abbreviated; use full first names wherever
% % possible. Include authors' e-mail addresses whenever possible.

% % Grouping authors' names or e-mail addresses, or providing an ``e-mail
% % alias,'' as shown below, is not acceptable:
% % \begin{verbatim}
% %   \author{Brooke Aster, David Mehldau}
% %   \email{dave,judy,steve@university.edu}
% %   \email{firstname.lastname@phillips.org}
% % \end{verbatim}

% % The \verb|authornote| and \verb|authornotemark| commands allow a note
% % to apply to multiple authors --- for example, if the first two authors
% % of an article contributed equally to the work.

% % If your author list is lengthy, you must define a shortened version of
% % the list of authors to be used in the page headers, to prevent
% % overlapping text. The following command should be placed just after
% % the last \verb|\author{}| definition:
% % \begin{verbatim}
% %   \renewcommand{\shortauthors}{McCartney, et al.}
% % \end{verbatim}
% % Omitting this command will force the use of a concatenated list of all
% % of the authors' names, which may result in overlapping text in the
% % page headers.

% % The article template's documentation, available at
% % \url{https://www.acm.org/publications/proceedings-template}, has a
% % complete explanation of these commands and tips for their effective
% % use.

% % Note that authors' addresses are mandatory for journal articles.

% % \section{Rights Information}

% % Authors of any work published by ACM will need to complete a rights
% % form. Depending on the kind of work, and the rights management choice
% % made by the author, this may be copyright transfer, permission,
% % license, or an OA (open access) agreement.

% % Regardless of the rights management choice, the author will receive a
% % copy of the completed rights form once it has been submitted. This
% % form contains \LaTeX\ commands that must be copied into the source
% % document. When the document source is compiled, these commands and
% % their parameters add formatted text to several areas of the final
% % document:
% % \begin{itemize}
% % \item the ``ACM Reference Format'' text on the first page.
% % \item the ``rights management'' text on the first page.
% % \item the conference information in the page header(s).
% % \end{itemize}

% % Rights information is unique to the work; if you are preparing several
% % works for an event, make sure to use the correct set of commands with
% % each of the works.

% % The ACM Reference Format text is required for all articles over one
% % page in length, and is optional for one-page articles (abstracts).

% % \section{CCS Concepts and User-Defined Keywords}

% % Two elements of the ``acmart'' document class provide powerful
% % taxonomic tools for you to help readers find your work in an online
% % search.

% % The ACM Computing Classification System ---
% % \url{https://www.acm.org/publications/class-2012} --- is a set of
% % classifiers and concepts that describe the computing
% % discipline. Authors can select entries from this classification
% % system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
% % commands to be included in the \LaTeX\ source.

% % User-defined keywords are a comma-separated list of words and phrases
% % of the authors' choosing, providing a more flexible way of describing
% % the research being presented.

% % CCS concepts and user-defined keywords are required for for all
% % articles over two pages in length, and are optional for one- and
% % two-page articles (or abstracts).

% % \section{Sectioning Commands}

% % Your work should use standard \LaTeX\ sectioning commands:
% % \verb|section|, \verb|subsection|, \verb|subsubsection|, and
% % \verb|paragraph|. They should be numbered; do not remove the numbering
% % from the commands.

% % Simulating a sectioning command by setting the first word or words of
% % a paragraph in boldface or italicized text is {\bfseries not allowed.}

% % \section{Tables}

% % The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
% % package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
% % high-quality tables.

% % Table captions are placed {\itshape above} the table.

% % Because tables cannot be split across pages, the best placement for
% % them is typically the top of the page nearest their initial cite.  To
% % ensure this proper ``floating'' placement of tables, use the
% % environment \textbf{table} to enclose the table's contents and the
% % table caption.  The contents of the table itself must go in the
% % \textbf{tabular} environment, to be aligned properly in rows and
% % columns, with the desired horizontal and vertical rules.  Again,
% % detailed instructions on \textbf{tabular} material are found in the
% % \textit{\LaTeX\ User's Guide}.

% % Immediately following this sentence is the point at which
% % Table~\ref{tab:freq} is included in the input file; compare the
% % placement of the table here with the table in the printed output of
% % this document.

% % \begin{table}
% %   \caption{Frequency of Special Characters}
% %   \label{tab:freq}
% %   \begin{tabular}{ccl}
% %     \toprule
% %     Non-English or Math&Frequency&Comments\\
% %     \midrule
% %     \O & 1 in 1,000& For Swedish names\\
% %     $\pi$ & 1 in 5& Common in math\\
% %     \$ & 4 in 5 & Used in business\\
% %     $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
% %   \bottomrule
% % \end{tabular}
% % \end{table}

% % To set a wider table, which takes up the whole width of the page's
% % live area, use the environment \textbf{table*} to enclose the table's
% % contents and the table caption.  As with a single-column table, this
% % wide table will ``float'' to a location deemed more
% % desirable. Immediately following this sentence is the point at which
% % Table~\ref{tab:commands} is included in the input file; again, it is
% % instructive to compare the placement of the table here with the table
% % in the printed output of this document.

% % \begin{table*}
% %   \caption{Some Typical Commands}
% %   \label{tab:commands}
% %   \begin{tabular}{ccl}
% %     \toprule
% %     Command &A Number & Comments\\
% %     \midrule
% %     \texttt{{\char'134}author} & 100& Author \\
% %     \texttt{{\char'134}table}& 300 & For tables\\
% %     \texttt{{\char'134}table*}& 400& For wider tables\\
% %     \bottomrule
% %   \end{tabular}
% % \end{table*}

% % \section{Math Equations}
% % You may want to display math equations in three distinct styles:
% % inline, numbered or non-numbered display.  Each of the three are
% % discussed in the next sections.

% % \subsection{Inline (In-text) Equations}
% % A formula that appears in the running text is called an inline or
% % in-text formula.  It is produced by the \textbf{math} environment,
% % which can be invoked with the usual
% % \texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
% % the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
% % and structures, from $\alpha$ to $\omega$, available in
% % \LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
% % examples of in-text equations in context. Notice how this equation:
% % \begin{math}
% %   \lim_{n\rightarrow \infty}x=0
% % \end{math},
% % set here in in-line math style, looks slightly different when
% % set in display style.  (See next section).

% % \subsection{Display Equations}
% % A numbered display equation---one set off by vertical space from the
% % text and centered horizontally---is produced by the \textbf{equation}
% % environment. An unnumbered display equation is produced by the
% % \textbf{displaymath} environment.

% % Again, in either environment, you can use any of the symbols and
% % structures available in \LaTeX\@; this section will just give a couple
% % of examples of display equations in context.  First, consider the
% % equation, shown as an inline equation above:
% % \begin{equation}
% %   \lim_{n\rightarrow \infty}x=0
% % \end{equation}
% % Notice how it is formatted somewhat differently in
% % the \textbf{displaymath}
% % environment.  Now, we'll enter an unnumbered equation:
% % \begin{displaymath}
% %   \sum_{i=0}^{\infty} x + 1
% % \end{displaymath}
% % and follow it with another numbered equation:
% % \begin{equation}
% %   \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
% % \end{equation}
% % just to demonstrate \LaTeX's able handling of numbering.

% % \section{Figures}

% % The ``\verb|figure|'' environment should be used for figures. One or
% % more images can be placed within a figure. If your figure contains
% % third-party material, you must clearly identify it as such, as shown
% % in the example below.
% % \begin{figure}[h]
% %   \centering
% %   \includegraphics[width=\linewidth]{sample-franklin}
% %   \caption{1907 Franklin Model D roadster. Photograph by Harris \&
% %     Ewing, Inc. [Public domain], via Wikimedia
% %     Commons. (\url{https://goo.gl/VLCRBB}).}
% %   \Description{The 1907 Franklin Model D roadster.}
% % \end{figure}

% % Your figures should contain a caption which describes the figure to
% % the reader. Figure captions go below the figure. Your figures should
% % {\bfseries also} include a description suitable for screen readers, to
% % assist the visually-challenged to better understand your work.

% % Figure captions are placed {\itshape below} the figure.

% % \subsection{The ``Teaser Figure''}

% % A ``teaser figure'' is an image, or set of images in one figure, that
% % are placed after all author and affiliation information, and before
% % the body of the article, spanning the page. If you wish to have such a
% % figure in your article, place the command immediately before the
% % \verb|\maketitle| command:
% % \begin{verbatim}
% %   \begin{teaserfigure}
% %     \includegraphics[width=\textwidth]{sampleteaser}
% %     \caption{figure caption}
% %     \Description{figure description}
% %   \end{teaserfigure}
% % \end{verbatim}

% % \section{Citations and Bibliographies}

% % The use of \BibTeX\ for the preparation and formatting of one's
% % references is strongly recommended. Authors' names should be complete
% % --- use full first names (``Donald E. Knuth'') not initials
% % (``D. E. Knuth'') --- and the salient identifying features of a
% % reference should be included: title, year, volume, number, pages,
% % article DOI, etc.

% % The bibliography is included in your source document with these two
% % commands, placed just before the \verb|\end{document}| command:
% % \begin{verbatim}
% %   \bibliographystyle{ACM-Reference-Format}
% %   \bibliography{bibfile}
% % \end{verbatim}
% % where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
% % suffix, of the \BibTeX\ file.

% % Citations and references are numbered by default. A small number of
% % ACM publications have citations and references formatted in the
% % ``author year'' style; for these exceptions, please include this
% % command in the {\bfseries preamble} (before
% % ``\verb|\begin{document}|'') of your \LaTeX\ source:
% % \begin{verbatim}
% %   \citestyle{acmauthoryear}
% % \end{verbatim}

% %   Some examples.  A paginated journal article \cite{Abril07}, an
% %   enumerated journal article \cite{Cohen07}, a reference to an entire
% %   issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
% %   monograph/whole book in a series (see 2a in spec. document)
% %   \cite{Harel79}, a divisible-book such as an anthology or compilation
% %   \cite{Editor00} followed by the same example, however we only output
% %   the series if the volume number is given \cite{Editor00a} (so
% %   Editor00a's series should NOT be present since it has no vol. no.),
% %   a chapter in a divisible book \cite{Spector90}, a chapter in a
% %   divisible book in a series \cite{Douglass98}, a multi-volume work as
% %   book \cite{Knuth97}, an article in a proceedings (of a conference,
% %   symposium, workshop for example) (paginated proceedings article)
% %   \cite{Andler79}, a proceedings article with all possible elements
% %   \cite{Smith10}, an example of an enumerated proceedings article
% %   \cite{VanGundy07}, an informally published work \cite{Harel78}, a
% %   doctoral dissertation \cite{Clarkson85}, a master's thesis:
% %   \cite{anisi03}, an online document / world wide web resource
% %   \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1)
% %   \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05} and
% %   (Case 3) a patent \cite{JoeScientist001}, work accepted for
% %   publication \cite{rous08}, 'YYYYb'-test for prolific author
% %   \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
% %   contain 'duplicate' DOI and URLs (some SIAM articles)
% %   \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
% %   multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
% %   couple of citations with DOIs:
% %   \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
% %   citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
% %   \cite{R} and \cite{UMassCitations}.

% % \section{Acknowledgments}

% % Identification of funding sources and other support, and thanks to
% % individuals and groups that assisted in the research and the
% % preparation of the work should be included in an acknowledgment
% % section, which is placed just before the reference section in your
% % document.

% % This section has a special environment:
% % \begin{verbatim}
% %   \begin{acks}
% %   ...
% %   \end{acks}
% % \end{verbatim}
% % so that the information contained therein can be more easily collected
% % during the article metadata extraction phase, and to ensure
% % consistency in the spelling of the section heading.

% % Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

% % \section{Appendices}

% % If your work needs an appendix, add it before the
% % ``\verb|\end{document}|'' command at the conclusion of your source
% % document.

% % Start the appendix with the ``\verb|appendix|'' command:
% % \begin{verbatim}
% %   \appendix
% % \end{verbatim}
% % and note that in the appendix, sections are lettered, not
% % numbered. This document has two appendices, demonstrating the section
% % and subsection identification method.

% % \section{SIGCHI Extended Abstracts}

% % The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
% % not in Word) produces a landscape-orientation formatted article, with
% % a wide left margin. Three environments are available for use with the
% % ``\verb|sigchi-a|'' template style, and produce formatted output in
% % the margin:
% % \begin{itemize}
% % \item {\verb|sidebar|}:  Place formatted text in the margin.
% % \item {\verb|marginfigure|}: Place a figure in the margin.
% % \item {\verb|margintable|}: Place a table in the margin.
% % \end{itemize}

% % %%
% % %% The acknowledgments section is defined using the "acks" environment
% % %% (and NOT an unnumbered section). This ensures the proper
% % %% identification of the section in the article metadata, and the
% % %% consistent spelling of the heading.
% % \begin{acks}
% % To Robert, for the bagels and explaining CMYK and color spaces.
% % \end{acks}

% % %%
% % %% The next two lines define the bibliography style to be used, and
% % %% the bibliography file.
% % \bibliographystyle{ACM-Reference-Format}
% % \bibliography{sample-base}

% % %%
% % %% If your work has an appendix, this is the place to put it.
% % \appendix

% % \section{Research Methods}

% % \subsection{Part One}

% % Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
% % malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
% % sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
% % vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
% % lacinia dolor. Integer ultricies commodo sem nec semper.

% % \subsection{Part Two}

% % Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
% % ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
% % ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
% % eros. Vivamus non purus placerat, scelerisque diam eu, cursus
% % ante. Etiam aliquam tortor auctor efficitur mattis.

% % \section{Online Resources}

% % Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
% % pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
% % enim maximus. Vestibulum gravida massa ut felis suscipit
% % congue. Quisque mattis elit a risus ultrices commodo venenatis eget
% % dui. Etiam sagittis eleifend elementum.

% % Nam interdum magna at lectus dignissim, ac dignissim lorem
% % rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
% % massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2020-02-07 paper/main.tex
refactor paper with uksys abstract

@@
 \setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
+\copyrightyear{2020}
+\acmYear{2020}
 \acmDOI{10.1145/1122445.1122456}
@@
 \section{Introduction}
-In recent years there has been an increased interest in and use of graph technologies.
-Application areas are wide reaching from healthcare to more traditional use cases such as social networks.
-Graph systems can be categorised into graph databases and graph processing frameworks.
-This paper focuses on graph databases, which at a high-level differ from relational databases in that they provide application programmers with a graph data model, as oppose to relations linked through foreign keys.
-Specifically, graph databases use the labelled property graph data model consisting of vertices and edges, which each can have properties and be group through labels.
+In recent years there has been an increased interest in and use of graph technologies[REF].
+Application areas are wide reaching from healthcare to more traditional use cases such as social networks [REF].
+Graph databases model as a \textit{labeled property graph}, vertices represent entities and edges represent the relationship between entities; properties can be stored on both vertices and edges [REF].
 
-As the sizes of graphs grow the limits of single node, replicated graph databases are being reached.
-This has lead to the development of horizontally scalable distributed graph databases.
+In practice graphs can be extremely large [REF], exceeding the storage capacity of a single node database.
+An approach to storing large graphs is to partition graph data over several machines in a cluster [REF].
 Moving from a single-node database to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
@@
 
-Recent work has investigated how violation of reciprocal consistency can lead to semantic corruption at alarmingly rates.
-In order to prevent this form of inconsistent one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs.
+Recent work has investigated how violation of reciprocal consistency can lead to semantic corruption at alarmingly rates [REF].
+In order to prevent this form of inconsistency, one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs [REF].
 The approach taken in this paper is to build on earlier work and develop a protocol that provides weaker consistency guarantees whilst still preserving reciprocal consistency.
+
+\section{Reciprocal Consistency}
+\label{sec:recipr-cons}
+
+In the labelled property graph data model edges have direction, there is a \textit{source} and a \textit{destination} vertex.
+However, for query performance, edge information is stored with \textbf{both} the source and destination vertices.
+This facilitates bi-directional edge traversal.
+Graphs are often implemented as an collection of adjacency lists where a list contains neighbors of a given vertex [REF].
+
+Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
+It is expressed using vertices A and B, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from A (source) to B (destination).
+Adjacency lists of both A and B record information about that edge and this information is mutually reciprocal (or inverse) of each other: A's list will indicate `A \textit{wrote} B' while B's will have `B \textit{written} by A'.
+Thus, a query `list all titles by the author who wrote The Hobbit' can be answered by landing at B and then traversing to A, even though AB is a directed edge at model level abstraction.
+
+The reciprocal nature of information at the ends of an edge not only facilitates bidirectional traversal but also ensures operational correctness.
+Without it, the queries `who wrote The Hobbit?' and `what all did Tolkien write?' could yield in different results.
+So, an edge is said to be \textit{reciprocally consistent}, if the adjacency lists of the vertices it connects have mutually reciprocal information.
+
 \section{Distributed Graph Databases}
-
-In a graph database records represent vertices containing vertex properties and an adjacency list containing edge "pointers" to other vertices along with associated properties.
-Whilst edges typically have direction they are physically implemented as one would implement an undirected graph with meta-data indicating the direction and some reciprocal information regarding the edge.
-For example, an edge E joining two vertices A and B results in reciprocal entries for E being written in the adjacency lists for both A and B.
-If this edge represented the edge between an author and their book, the following reciprocal information would be stored A would store the information \textit{wrote B} and B would store the information \textit{written by A}.
-This permits bi-directional traversal of the edge, a query reading either A or B would be able to reify the edge correctly.
-When the adjacency lists for vertices are mutually compatible they are said to be \textit{reciprocally consistent}.
-The challenge is maintaining reciprocally consistency in a distributed graph database.
 
@@
 Vertices are evenly assigned to partitions in a manner that minimises the number of edges that span partitions - referred to as \textbf{distributed edges}.
-That is the number of entries in a vertex's adjacency list that point to vertices on remote servers is minimised.
-However, regardless of the graph partitioning algorithm used there will exist a non-negligible number of distributed edges.
-Without sufficient isolation between concurrent updates to a distributed edge, reciprocal consistency can be violated.
+Regardless of the partitioning strategy used a non-negligible number of edges are distributed edges.
 
-To preserve reciprocal consistency distributed edges could be updated through a protocols such as 2-Phase Commit with isolation provided via 2-Phase Locking.
-However, such an approach is well known to negatively impact scalability and performance.
-The approach taken here is provide a weaker isolation model whilst still preserving reciprocal consistency.
+When concurrent queries update an edge, maintaining reciprocal consistency gets harder if the edge is distributed, i.e., its vertices are stored in distinct servers of a distributed database system.
+Our earlier work [REF] established that when edges are permitted to become reciprocally inconsistent, independent queries reading inconsistent information from different ends of such edges, do corrupt the database through their subsequent writes, and that a significant degree of corruption can result within durations that are worryingly small compared to database lifetime.
+
+Of course, reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., snapshot isolation) but that incurs a performance cost if users are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed database.
+This paper addresses that
+challenge. It proposes a suite of simple, coordination-free protocols and compares their performance through simulations.
 
 \section{Preserving Reciprocal Consistency}
-Assume a graph database partitioned across servers in a cluster, with each partition being a Raft cluster providing linearizable reads and writes to the data objects it is responsible for.
-One strawman approach to preserving reciprocal consistency would be for update operations to take an exclusive lock on both records that make up the distributed edge, releasing locks when the update has been completed.
-Locking is a pessimistic approach and has the implicit assumption that concurrent updates will be common, this is not the case in graph databases.
-A second approach would be having a read operation read both records to verify the information is consistent.
-The problem here is that read operations dominant graph workloads and every read of a distributed edge would incur the penalty of traversing the network twice.
 
-The approach developed leverages 1) the frequency of concurrent update operations to the same edge is low 2) a logical write of an edge consists of writes to two physical records 3) read operations only access a single record.
-We now outline a variant of the \textbf{Collision Detection} mechanism presented in REF and discuss it's combination with Raft.
-% assume a mechanism to guarantee consistency of operations within partitions
-\subsection{Collision Detection Mechanism}
+Both protocols described below leverage the same fact: at the logical level, updating a distributed edge is a single operation.
+However, it consists of two sub-operations, it must update information at the source and destination vertices an edge connects.
+Note, the order in which the sub-operations take place is not constrained.
+Updating the source then the destination is equally as likely as updating the destination then the source.
 
-Initially, each write to a distributed edge is provisional. They become permanent only if, and when, the transaction that contains them is allowed to commit.
+\subsection{Probabilistic Protocol}
 
-Provisional writes on a given record can occur only one at a time and each is time-stamped using a combination of the transaction ID and the local host's clock - used to uniquely identify the operation.
-This is stored in a order-preserving provisional write list.
-Thus, when a transaction attempts to write a given record, it can identify all other transactions, called predecessors (if any), that have earlier wrote that record provisionally along with the transaction responsible for latest committed value.
-Currently no higher information regarding specific operation type is assumed.
-Read operations receive the latest committed version of a record and ignore any provisionally updated values.
+In the probabilistic approach, an edge update is allowed if an earlier, temporary update is sufficiently early by D as per the local clock time and D is the bound on the server communication delays with probability (1-e).
+(Note that a more pessimistic estimation of D results in a smaller e) In this approach, the smaller the value chosen for D, the more is the extent of contentions permitted, hence the smaller is the number of expected aborts and, on the other hand, the larger is the risk of reciprocal consistency not being ensured and the database being corrupted.
+These are examined through simulations.
 
-\textbf{Collision detection}: An edge has a \textit{tail record} and \textit{head record} - a write operation by a transaction must write to \textit{both} records.
-The order in which updates occur is not constrained, that is, one update may update the head and then after a network delay the tail, the opposite can also occur with equal chance.
+\subsection{Deterministic Protocol}
 
-In order to distinguish between which write a transaction they are given labels 1 and 2, which act as 'history' meta-data indicating the partition where a transaction started and completed its write.
-The meta-data consists of 1) local timestamp, 2) transaction ID and 3) order label.
-
-The following rule is applied: \textbf{Cancellation rule:} If, by the time an operation is performing it's second update, if a previous provisional write labeled 1 has been observed, but the corresponding label 2 has not been observed, then this write is cancelled.
-In other words, a write is cancelled if it has observed the start of a previous attempt, but not its completion.
-This prevents writes happening out-of-order.
-When an update is cancelled, the transaction containing it is aborted and all its provisional records are erased.
-
-\subsection{Alternative Methods}
+In the deterministic approach, queries updating distributed edges are expected to carry with them some state information as they traverse from one end server of a distributed edge to the other.
+Other than this overhead, there is no requirement for estimating bound on traversal delays.
+Our simulations will compare the number of aborts incurred in both approaches.
 
 \section{Evaluation \& Results}
-
-In order to test the throughput and latency of the proposed mechanism a discrete event orientated simulation has been performed.
-Collision detection has been tested against the simple approach defined in Section 3.
 

### 2020-02-07 approximations/plots.R
tidy up time to corruption scripts and general project structure

@@ -1,0 +1,22 @@
+## comparison plot
+comparison <- function(save=F,df){
+  if(save==T){
+    pdf("../images/comparison.pdf")
+  }
+  p=ggplot(data=df, aes(x=lam,y=u,linetype=type,color=type)) +
+    geom_line() +
+    scale_linetype_discrete(name="Type: ",labels=unname(TeX(c('$\\epsilon = 100%$',"$\\epsilon$ = 5%","$\\epsilon$ = 1%")))) +
+    scale_color_discrete(name="Type: ",labels=unname(TeX(c('$\\epsilon = 100%$',"$\\epsilon$ = 5%","$\\epsilon$ = 1%")))) +
+    scale_x_continuous(name=unname(TeX(c("Read-Write Transactions per second (10% TPS)"))),
+                       breaks=seq(100,1500,200)) +
+    scale_y_continuous(name=unname(TeX(c("Time (in hours) for 10% DB Corruption"))),
+                       breaks=seq(0,3000,500)) +
+    theme(legend.text.align = 0)
+  p
+  if(save==T){
+    dev.off()
+  }
+  return(p)
+}
+
+comparison(save=T)

### 2020-02-07 run-no-isolation.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+Rscript approximations/sim-no-isolation.R

### 2020-02-07 run-probabilistic.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+Rscript approximations/sim-probabilistic.R

### 2020-02-07 simulations/deterministic/src/main/java/PRCPSimulation.java

@@ -1,0 +1,116 @@
+import action.*;
+import event.*;
+
+import org.apache.log4j.Logger;
+import state.*;
+import utils.*;
+
+// TODO: unit tests
+// TODO: add functionality to switch between simulating time or completed txns
+// TODO: add arbiter mechanism toggle
+
+public class PRCPSimulation {
+
+    private final static Logger LOGGER = Logger.getLogger(PRCPSimulation.class.getName());
+
+    public static void main(String[] args) {
+
+        int databaseSize = SimulationConfiguration.getInstance().getDatabaseSize();
+
+        // init. database
+        for (int i = 0; i < databaseSize; i++ ) {
+            DistributedEdge e = new DistributedEdge(i);
+            GlobalEdgeListSingleton.getInstance().addEdge(e);
+        }
+
+        // init. first arrival event
+        double initEventTime = SimulationRandom.getInstance().generateNextArrival();
+        GlobalEventListSingleton.getInstance().addEvent(new ArrivalEvent(initEventTime, EventType.ARRIVAL));
+
+        // init. first arbiter event
+        if (SimulationConfiguration.getInstance().getUseArbiter()) {
+            double initArbiter = SimulationRandom.getInstance().generateServiceDelay();
+            GlobalEventListSingleton.getInstance().addEvent(new ArbiterEvent(initArbiter, EventType.ARBITER));
+        }
+        LOGGER.debug("**********************************************************************");
+        LOGGER.debug("**********************************************************************");
+        LOGGER.debug("Initial State");
+        LOGGER.debug("**********************************************************************");
+        LOGGER.debug("Event List: " + GlobalEventListSingleton.getInstance().getEventList());
+        LOGGER.debug("**********************************************************************");
+        LOGGER.debug("**********************************************************************");
+        LOGGER.debug(" ");
+
+        runSimulation();
+
+        LOGGER.info("TPS: " + SystemMetrics.getInstance().getTps());
+        LOGGER.info("Arrivals: " + SystemMetrics.getInstance().getArrivals());
+        LOGGER.info("Completed: " + SystemMetrics.getInstance().getCompleted());
+        LOGGER.info("Total Aborts: " + SystemMetrics.getInstance().getAborts());
+        LOGGER.info("Collisions: " + SystemMetrics.getInstance().getCollisions());
+        LOGGER.info("Arbitration: " + SystemMetrics.getInstance().getArbitration());
[…]

### 2020-02-07 simulations/deterministic/src/main/java/action/AbortCleanUpAction.java

@@ -1,0 +1,31 @@
+package action;
+
+import event.AbortCleanUpEvent;
+import org.apache.log4j.Logger;
+import state.GlobalActiveTransactionListSingleton;
+import state.GlobalEdgeListSingleton;
+import state.Transaction;
+
+/**
+ * This action is invoked by a clear event.
+ */
+public class AbortCleanUpAction {
+
+    private final static Logger LOGGER = Logger.getLogger(AbortCleanUpAction.class.getName());
+
+    public void action(AbortCleanUpEvent abortCleanUpEvent) {
+
+        Transaction transaction = abortCleanUpEvent.getTransaction();
+
+        // clear provisional write
+        GlobalEdgeListSingleton.getInstance().clearProvisionalWrites(transaction.getId());
+
+        // remove from active list + everyone pred lists
+        GlobalActiveTransactionListSingleton.getInstance().removeTransaction(transaction.getId());
+
+
+    }
+
+
+
+}

### 2020-02-07 simulations/deterministic/src/main/java/event/ArrivalEvent.java

@@ -1,0 +1,17 @@
+package event;
+
+public class ArrivalEvent extends AbstractEvent {
+
+    public ArrivalEvent(double eventTime, EventType eventTypeEnum) {
+        super(eventTime, eventTypeEnum);
+    }
+
+    @Override
+    public String toString() {
+        return "{" +
+                "time=" + String.format("%.5f",getEventTime()) +
+                ", " + "type=" + getEventType() +
+                "}";
+    }
+
+}

### 2020-02-07 simulations/deterministic/src/main/java/event/Event.java

@@ -1,0 +1,8 @@
+package event;
+
+public interface Event {
+
+    public double getEventTime();
+    public EventType getEventType();
+
+}

### 2020-02-07 simulations/deterministic/src/main/java/state/DistributedEdge.java

@@ -1,0 +1,70 @@
+package state;
+
+import java.util.ArrayList;
+
+/**
+ * This class represents a distributed edge.
+ * A distributed edge has two distinct side, left and right.
+ * Each side has it's own provisional write list containing transaction IDs.
+ */
+public class DistributedEdge {
+
+    private int edgeId;
+    private ArrayList<ProvisionalWrite> leftProvisionalWrites;
+    private ArrayList<ProvisionalWrite> rightProvisionalWrites;
+
+
+    public DistributedEdge(int edgeId) {
+        this.edgeId = edgeId;
+        leftProvisionalWrites = new ArrayList<>();
+        rightProvisionalWrites = new ArrayList<>();
+    }
+
+    /**
+     * Returns the ID of a given distributed edge.
+     * @return edge ID
+     */
+    public int getEdgeId() {
+        return edgeId;
+    }
+
+    /**
+     * Add a provisional write to a distributed edge
+     * @param leftSide, true corresponds to left and false corresponds to right
+     * @param provisionalWrite, provisional writes by transactions
+     */
+    public void addProvisionalWrite(Boolean leftSide,ProvisionalWrite provisionalWrite) {
+
+        if (leftSide) {
+            leftProvisionalWrites.add(provisionalWrite);
+        } else {
+            rightProvisionalWrites.add(provisionalWrite);
+        }
+
+    }
+
+    public ArrayList<ProvisionalWrite> getLeftProvisionalWrites() {
+        return leftProvisionalWrites;
+    }
+
+    public ArrayList<ProvisionalWrite> getRightProvisionalWrites() {
+        return rightProvisionalWrites;
+    }
+
+    public void removeProvisionalWrites (Integer transactionId) {
+
+        leftProvisionalWrites.removeIf(pw -> (pw.getTransactionId() == transactionId));
+        rightProvisionalWrites.removeIf(pw -> (pw.getTransactionId() == transactionId));
+
+
+    }
+
+    @Override
+    public String toString() {
+        return "{" +
+                "id=" + edgeId +
+                ", leftPW=" + leftProvisionalWrites +
+                ", rightPW=" + rightProvisionalWrites +
+                '}';
+    }
+}

### 2020-02-07 simulations/deterministic/src/main/java/state/GlobalActiveTransactionListSingleton.java

@@ -1,0 +1,37 @@
+package state;
+
+import java.util.ArrayList;
+import java.util.stream.Collectors;
+
+public class GlobalActiveTransactionListSingleton {
+
+    private static final GlobalActiveTransactionListSingleton instance = new GlobalActiveTransactionListSingleton();
+    private ArrayList<Transaction> activeTransactionList;
+
+    private GlobalActiveTransactionListSingleton() {
+        activeTransactionList = new ArrayList<>();
+    }
+
+    public static GlobalActiveTransactionListSingleton getInstance() {
+        return instance;
+    }
+
+    public void addTransaction(Transaction transaction) {
+        activeTransactionList.add(transaction);
+    }
+
+    public void removeTransaction(int transactionId){
+        // remove from list
+        activeTransactionList = (ArrayList<Transaction>) activeTransactionList.stream().filter(transaction -> transaction.getId() != transactionId).collect(Collectors.toList());
+        // remove from all pred lists
+        for (Transaction transaction: activeTransactionList
+             ) {
+            ArrayList<Integer> newPredecessors = (ArrayList<Integer>) transaction.getPredecessors().stream().filter(pred -> pred != transactionId).collect(Collectors.toList());
+            transaction.setPredecessors(newPredecessors);
+        }
+    }
+
+    public ArrayList<Transaction> getActiveTransactionList() {
+        return activeTransactionList;
+    }
+}

### 2020-02-07 simulations/deterministic/src/main/java/utils/SystemMetrics.java

@@ -1,0 +1,123 @@
+package utils;
+
+//import static utils.SimulationConfiguration.getTPS;
+
+public class SystemMetrics {
+
+    private static final SystemMetrics instance = new SystemMetrics();
+
+    private int tps;
+    private int arrivals;       // number of arrivals
+    private int aborts;         // number of total aborts
+    private int collisions;     // aborts by collision detection mechanism
+    private int arbitration;    // aborts by arbitration mechanism
+    private double cumulativeLifetimes;
+    private int commits;
+    private int sentToArbiter;
+
+    private SystemMetrics() {
+        arrivals = 0;
+        aborts = 0;
+        collisions = 0;
+        arbitration = 0;
+        cumulativeLifetimes = 0;
+        commits = 0;
+        sentToArbiter = 0;
+        tps = SimulationConfiguration.getInstance().getTPS();
+    }
+
+    public void reset() {
+        arrivals = 0;
+        aborts = 0;
+        collisions = 0;
+        arbitration = 0;
+        cumulativeLifetimes = 0;
+        commits = 0;
+        sentToArbiter = 0;
+    }
+
+    public int getTps() {
+        return tps;
+    }
+
+    public void setTps(int tps) {
+        this.tps = tps;
+    }
+
+    public static SystemMetrics getInstance() {return instance; }
+
+    public void incrementArrivals() {
+        arrivals = arrivals + 1;
+    }
+
+    public void incrementAborts() {
+        aborts = aborts + 1;
+    }
+
+    public void incrementCommitted() {
+        commits = commits + 1;
+    }
+
+    public void incrementCollisions() {
+        collisions = collisions + 1;
+    }
+
+    public void addLifetime(double txnLifetime) {
+        cumulativeLifetimes = cumulativeLifetimes + txnLifetime;
+    }
+
+    public void incrementArbitration() {
+        arbitration = arbitration + 1;
+    }
+
+    public int getArrivals() {
+        return arrivals;
+    }
+
+    public int getCollisions() {
+        return collisions;
+    }
+
+    public int getAborts() {
+        return aborts;
+    }
+
+    public int getCommits() {
+        return commits;
+    }
+
+    public int getCompleted() {
+        return aborts + commits;
+    }
+
+    public int getSentToArbiter() {
+        return sentToArbiter;
+    }
+
+    public void incrementSentToArbiter() {
+        sentToArbiter = sentToArbiter + 1;
+    }
+
+    public int getArbitration() {
+        return arbitration;
+    }
+
[…]

### 2020-02-07 simulations/probabilistic/src/main/java/event/AbortCleanUpEvent.java

@@ -1,0 +1,27 @@
+package event;
+
+import state.Transaction;
+
+public class AbortCleanUpEvent extends AbstractEvent {
+
+    private Transaction transaction;
+
+    public AbortCleanUpEvent(double eventTime, EventType eventTypeEnum, Transaction transaction) {
+        super(eventTime, eventTypeEnum);
+        this.transaction = transaction;
+    }
+
+    public Transaction getTransaction() {
+        return transaction;
+    }
+
+    @Override
+    public String toString() {
+        return "{" +
+                "transaction=" + transaction.getId() +
+                ", " + "time=" + String.format("%.5f",getEventTime()) +
+                ", " + "type=" + getEventType() +
+                "}";
+    }
+
+}

### 2020-02-07 simulations/probabilistic/src/main/java/event/EventType.java

@@ -1,0 +1,12 @@
+package event;
+
+/**
+ * Defining event types
+ */
+public enum EventType {
+    ARRIVAL,
+    UPDATE_1,
+    UPDATE_2,
+    ARBITER,
+    ABORT_CLEAN_UP
+}

### 2020-02-07 simulations/probabilistic/src/main/java/event/Update1Event.java

@@ -1,0 +1,29 @@
+package event;
+
+import state.Transaction;
+
+/**
+ * An update 1 event contains the transaction performing the update
+ */
+public class Update1Event extends AbstractEvent {
+
+    private Transaction transaction;
+
+    public Update1Event(double eventTime, EventType eventTypeEnum, Transaction transaction) {
+        super(eventTime, eventTypeEnum);
+        this.transaction = transaction;
+    }
+
+    public Transaction getTransaction() {
+        return transaction;
+    }
+
+    @Override
+    public String toString() {
+        return "{" +
+                "transaction=" + transaction.getId() +
+                ", " + "time=" + String.format("%.5f",getEventTime()) +
+                ", " + "type=" + getEventType() +
+                "}";
+    }
+}

### 2020-02-07 simulations/probabilistic/src/main/java/state/ProvisionalWrite.java

@@ -1,0 +1,31 @@
+package state;
+
+/**
+ * This class represents a provisional write made by a transaction on a given distributed edge.
+ */
+public class ProvisionalWrite {
+
+    private int transactionId;
+    private int writeLabel;
+
+    public ProvisionalWrite(int transactionId, int writeLabel) {
+        this.transactionId = transactionId;
+        this.writeLabel = writeLabel;
+    }
+
+    public int getTransactionId() {
+        return transactionId;
+    }
+
+    public int getWriteLabel() {
+        return writeLabel;
+    }
+
+    @Override
+    public String toString() {
+        return "{" +
+                "transactionId=" + transactionId +
+                ", writeLabel=" + writeLabel +
+                '}';
+    }
+}

### 2020-02-07 simulations/probabilistic/src/main/java/state/Transaction.java

@@ -1,0 +1,107 @@
+package state;
+
+import utils.SimulationConfiguration;
+import utils.SimulationRandom;
+
+import java.util.ArrayList;
+
+/**
+ * Represents a transaction in the database
+ */
+public class Transaction {
+
+    private int id;
+    private int updates;
+    private int updatesCompleted;
+    private ArrayList<Integer> predecessors;
+    private ArrayList<ProvisionalWrite> collisionDetection;
+    private double startTime;
+    private double endTime;
+
+    public Transaction(int id) {
+        this.id = id;
+        if (SimulationConfiguration.getInstance().getUseArbiter()) {
+            this.updates = SimulationRandom.getInstance().getTransactionUpdates();
+        } else {
+            this.updates = 1;
+        }
+        this.updatesCompleted = 0;
+        predecessors = new ArrayList<>();
+        collisionDetection = new ArrayList<>();
+    }
+
+    void setPredecessors(ArrayList<Integer> predecessors) {
+        this.predecessors = predecessors;
+    }
+
+    public double getStartTime() {
+        return startTime;
+    }
+
+    public double getEndTime() {
+        return endTime;
+    }
+
+    public void setStartTime(double startTime) {
+        this.startTime = startTime;
+    }
+
+    public void setEndTime(double endTime) {
+        this.endTime = endTime;
+    }
+
+    public double getLifetime() {
+        return endTime - startTime;
+    }
+
+    public int getId() {
+        return id;
+    }
+
+    public int getUpdates() {
+        return updates;
+    }
+
+    public int getUpdatesCompleted() {
+        return updatesCompleted;
+    }
+
+    public ArrayList<Integer> getPredecessors() {
+        return predecessors;
+    }
+
+    public void addPredecessor(int id) {
+        predecessors.add(id);
+    }
+
+    public void addPredecessors(ArrayList<Integer> predecessorList) {
+        predecessors.addAll(predecessorList);
+    }
+
+    public void incrementUpdateCompleted() {
+        updatesCompleted = updatesCompleted + 1;
+    }
+
+    public void addCollisionDetectionInformation(ProvisionalWrite pw) {
+        collisionDetection.add(pw);
+    }
+
+    public void removeCollisionDetectionInformation() {
+        collisionDetection.clear();
+    }
+
+    public ArrayList<ProvisionalWrite> getCollisionDetection() {
+        return collisionDetection;
+    }
+
+    @Override
+    public String toString() {
+        return "Transaction{" +
[…]

### 2020-02-17 paper/main.tex
add algorithm templates
making results

@@
-%%
-%% This is file `sample-sigplan.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigplan')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigplan.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
 \documentclass[sigplan,screen]{acmart}
 
-%%
 %% \BibTeX command to typeset BibTeX logo in the docs
@@
     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
+\usepackage[ruled,vlined]{algorithm2e}
 
@@
 
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
 \begin{abstract}
-  A solution for maintaining reciprocal consistency in distributed graph databases is presented.
+
+  In this paper we formalize the notion of \textit{reciprocal consistency}, a important feature specific to graph databases.
+If reciprocal consistency is not enforced during transaction processing, a scale-free distributed graph database can become semantically corrupt within a short time period relative to database lifetime.
+Reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., snapshot isolation) incurring well established performance costs.
+However, in practice systems are often deployed with weaker consistency, this paper describes two protocols that provide only reciprocal consistency semantics.
+Catering for users that are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed graph database.
+Both protocols presented are compared and contrasted using metrics such as number of aborts per second.
+
 \end{abstract}
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
+% \begin{CCSXML}
+% <ccs2012>
+%  <concept>
+%   <concept_id>10010520.10010553.10010562</concept_id>
+%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+%   <concept_significance>500</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10010520.10010575.10010755</concept_id>
+%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
+%   <concept_significance>300</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10010520.10010553.10010554</concept_id>
+%   <concept_desc>Computer systems organization~Robotics</concept_desc>
+%   <concept_significance>100</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10003033.10003083.10003095</concept_id>
+%   <concept_desc>Networks~Network reliability</concept_desc>
+%   <concept_significance>100</concept_significance>
+%  </concept>
+% </ccs2012>
+% \end{CCSXML}
 
@@
 
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{reciprocal consistency, graph databases, collision detection}
+\keywords{Graph Databases, Reciprocal Consistency}
 
-%% A "teaser" image appears between the author and affiliation
-%% information and the body of the document, and typically spans the
-%% page.
-% \begin{teaserfigure}
-%   \includegraphics[width=\textwidth]{sampleteaser}
-%   \caption{Seattle Mariners at Spring Training, 2010.}
-%   \Description{Enjoying the baseball game from the third-base
-%   seats. Ichiro Suzuki preparing to bat.}
-%   \label{fig:teaser}
-% \end{teaserfigure}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
 \maketitle
@@
 \section{Introduction}
-In recent years there has been an increased interest in and use of graph technologies[REF].
-Application areas are wide reaching from healthcare to more traditional use cases such as social networks [REF].
-Graph databases model as a \textit{labeled property graph}, vertices represent entities and edges represent the relationship between entities; properties can be stored on both vertices and edges [REF].
 
-In practice graphs can be extremely large [REF], exceeding the storage capacity of a single node database.
-An approach to storing large graphs is to partition graph data over several machines in a cluster [REF].
+Recent years have seen a proliferation in the use of graph technologies \cite{Besta2019}.
+Application areas are wide reaching from healthcare, social networks and fraud detection \cite{Eifrem2016}.
+Graph databases model data as a \textit{labeled property graph}, vertices represent entities and edges represent the relationship between entities; properties can be stored on both vertices and edges \cite{Robinson2015}.
+
+In practice graphs can be extremely large \cite{Sahu2017}, exceeding the storage capacity of a single-node database.
+An approach to storing large graphs is to partition graph data over several machines in a cluster.
 Moving from a single-node database to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
@@
 
-Recent work has investigated how violation of reciprocal consistency can lead to semantic corruption at alarmingly rates [REF].
-In order to prevent this form of inconsistency, one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs [REF].
-The approach taken in this paper is to build on earlier work and develop a protocol that provides weaker consistency guarantees whilst still preserving reciprocal consistency.
+In order to preserve reciprocal consistency, one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs.
+To avoid these costs database managements systems (DBMSs) provide several weak isolation models \cite{Berenson1995}.
+In addition, there are a number of systems that eschew transactional guarantees altogether, typically offering eventual consistency \cite{Bailis2013}.
+Recent work \cite{Ezhilchelvan2018} investigated how violation of reciprocal consistency can occur under eventual consistency, leading to semantic corruption at alarmingly rates.
+The contribution of this paper is the description of two protocols that provide only reciprocal consistency, a consistency model between the known extremes that provides suitable performance.
+The performance of each protocol is measured using metrics such as aborts per second.
 
@@
 
-In the labelled property graph data model edges have direction, there is a \textit{source} and a \textit{destination} vertex.
-However, for query performance, edge information is stored with \textbf{both} the source and destination vertices.
-This facilitates bi-directional edge traversal.
-Graphs are often implemented as an collection of adjacency lists where a list contains neighbors of a given vertex [REF].
+In the labeled property graph data model edges have direction, there is a \textit{source} and a \textit{destination} vertex.
+Edge information is stored with \textbf{both} the source and destination vertices.
+This facilitates bi-directional edge traversal and allows for better query performance.
+
+A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB [REF]) is for records to represent vertices containing both data values and an adjacency list containing edge ``pointers'' to other vertices.
+In this representation an edge has ``reciprocal'' entries in the adjacency lists of the vertices the edge connects Figure \ref{adj-list}.
+A query reading either the source or destination vertices should be able to reify the edge correctly, returning consistent results.
+When the adjacency lists for vertices are mutually compatible like this, we say they are \emph{reciprocally consistent}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
+  \caption{Graph representation [REF].}
+  \Description{Graph representation.}
+  \label{adj-list}
+\end{figure}
 
@@
 Adjacency lists of both A and B record information about that edge and this information is mutually reciprocal (or inverse) of each other: A's list will indicate `A \textit{wrote} B' while B's will have `B \textit{written} by A'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered by landing at B and then traversing to A, even though AB is a directed edge at model level abstraction.
-
-The reciprocal nature of information at the ends of an edge not only facilitates bidirectional traversal but also ensures operational correctness.
-Without it, the queries `who wrote The Hobbit?' and `what all did Tolkien write?' could yield in different results.
-So, an edge is said to be \textit{reciprocally consistent}, if the adjacency lists of the vertices it connects have mutually reciprocal information.
+Thus, a query `list all titles by the author who wrote The Hobbit' can be answered by landing at B and then traversing to A; even though the edge \texttt{(A \{name:`Tolkien'\})-[:WROTE]->(B \{title:`The Hobbit'\})} is a directed edge at model level abstraction.
 
@@
 
-A common approach to partitioning a graph is to use a balanced edge-cut.
-Vertices are evenly assigned to partitions in a manner that minimises the number of edges that span partitions - referred to as \textbf{distributed edges}.
-Regardless of the partitioning strategy used a non-negligible number of edges are distributed edges.
+A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
+Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
+The objective of such an approach is to minimize the proportion of edge that span partitions in a manner that balances the distribution of vertices to partitions.
+Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, [FIG].
+The proportion of distributed edges is highly data dependent.
+However, the number of such edges is always non-negligible ranging from 15-30\%.
 
-When concurrent queries update an edge, maintaining reciprocal consistency gets harder if the edge is distributed, i.e., its vertices are stored in distinct servers of a distributed database system.
-Our earlier work [REF] established that when edges are permitted to become reciprocally inconsistent, independent queries reading inconsistent information from different ends of such edges, do corrupt the database through their subsequent writes, and that a significant degree of corruption can result within durations that are worryingly small compared to database lifetime.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{./images/dist-graph}
+  \caption{A partitioned graph assigned to servers in a cluster. Some edges are local and some distributed}
+  \Description{Graph representation.}
+  \label{dist-graph}
+\end{figure}
+
+In a distributed graph database, adjacency lists can contain ``pointers'' to edges on remote servers.
+This makes maintaining distributed edge reciprocal consistency more challenging - especially given a common architecture employed by contemporary distributed graph databases.
+Often they use an existing eventually consistent database (e.g.
+Apache Cassandra [REF]) to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
+Superficially, opting for this design appears to be a good choice: the user has the modeling convenience of graphs with the operational characteristics from the underlying database.
+However, the problem with this design is the (lack of) transactional semantics are inherited from the underlying store.
+For example, Apache Cassandra provides no way of preventing updates from mutual interference in its normal multi-partition use case.
+Earlier work displayed how weak isolation undermines distributed edge reciprocal consistency, causing irreversible corruption that spreads at alarmingly rates \cite{Ezhilchelvan2018}.
 
 Of course, reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., snapshot isolation) but that incurs a performance cost if users are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed database.
-This paper addresses that
-challenge. It proposes a suite of simple, coordination-free protocols and compares their performance through simulations.
+This paper addresses that challenge.
 
@@
 
-Both protocols described below leverage the same fact: at the logical level, updating a distributed edge is a single operation.
-However, it consists of two sub-operations, it must update information at the source and destination vertices an edge connects.
-Note, the order in which the sub-operations take place is not constrained.
-Updating the source then the destination is equally as likely as updating the destination then the source.
+A single logical write to a distributed edge consists of two sub-operations, updating reciprocal information at the source and the destination vertices.
+The order in which sub-operations take place are not constrained.
+For example, when updating an edge between A and B it is equally likely to update A then B as it is to update B then A.
+Without sufficient isolation, concurrent updates can interleave in a manner that violates reciprocal consistency [FIG], such an edge is known as a \emph{half-corrupted} edge.
+Figure [FIG] outlines the interleavings that produce half-corrupted edges.
 
-\subsection{Probabilistic Protocol}
+% transactions consists of reads/writes
 
-In the probabilistic approach, an edge update is allowed if an earlier, temporary update is sufficiently early by D as per the local clock time and D is the bound on the server communication delays with probability (1-e).
-(Note that a more pessimistic estimation of D results in a smaller e) In this approach, the smaller the value chosen for D, the more is the extent of contentions permitted, hence the smaller is the number of expected aborts and, on the other hand, the larger is the risk of reciprocal consistency not being ensured and the database being corrupted.
-These are examined through simulations.
+\subsection{\emph{Delta} Protocol}
 
-\subsection{Deterministic Protocol}
+In order to prevent all interleaving in Figure [FIG] the \emph{Delta} protocol adopts a probabilistic approach.
+It is assumed there is a known $\Delta$, that reflects the bound on server communication delays, with a probability $(1-\epsilon)$ that $\Delta$ is violated.
+Informally, the protocol consists of two-phases.
+In phase 1 all writes by sub-operations are temporary and a sub-operation is allowed to proceed provided there is no other temporary sub-operation preceding it within $\Delta$; measured as per the local clock time.
+If the attempt does not succeed, the transaction aborts, aborting any and every previous tentative write that it may have successfully completed.
+In phase 2, a transaction that succeeds at all its tentative write attempts, decides to commit when it completes its graph traversal; the transaction commits once all its tentative writes are committed.
+This protocol avoids all conflict scenarios in Figure [FIG] provided the bound is not violated.
+If the bound is violate corruption and its spread can still occur, Figure [FIG].
+There exists an interesting relationship between $\Delta$, the proportion of aborts and the spread of corruption.
 
-In the deterministic approach, queries updating distributed edges are expected to carry with them some state information as they traverse from one end server of a distributed edge to the other.
-Other than this overhead, there is no requirement for estimating bound on traversal delays.
-Our simulations will compare the number of aborts incurred in both approaches.
+\begin{algorithm}
+  \SetAlgoLined
+  \eIf{sub-operation=1}{
+    \eIf{$\Delta = \emptyset$}{
+      Perform tentative write operation\;
+      Set $\Delta$\;
+      Issue sub-operation-2\;
+    }{
+      Abort transaction\;
+    }
+  }{
+    \eIf{$\Delta = \emptyset$}{
+      Perform tentative write operation\;
+      Set $\Delta$\;
+      Continue transaction\;
+    }{
+      Abort transaction\;
+    }
+  }
+    \caption{\emph{Delta} Protocol Phase 1}
+\end{algorithm}
 
 \section{Evaluation \& Results}
+
+% Spread Corruption
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{./images/corruption_comparison}
+  \caption{A partitioned graph assigned to servers in a cluster. Some edges are local and some distributed}
+  \Description{Graph representation.}
+  \label{dist-graph}
+\end{figure}
+
+% Aborts/s
 
@@
 % % commands, placed just before the \verb|\end{document}| command:
-% % \begin{verbatim}
-% %   \bibliographystyle{ACM-Reference-Format}
-% %   \bibliography{bibfile}
-% % \end{verbatim}
-% % where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-% % suffix, of the \BibTeX\ file.
+% \begin{verbatim}
+% \bibliographystyle{ACM-Reference-Format}
+% \bibliography{papoc}
+% \end{verbatim}
+% % % where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
+% % % suffix, of t
+%he \BibTeX\ file.
 
@@
 
-% % \section{Appendices}
+% \section{Appendices}
 
-% % If your work needs an appendix, add it before the
-% % ``\verb|\end{document}|'' command at the conclusion of your source
-% % document.
+% If your work needs an appendix, add it before the
+% ``\verb|\end{document}|'' command at the conclusion of your source
+% document.
 
-% % Start the appendix with the ``\verb|appendix|'' command:
-% % \begin{verbatim}
-% %   \appendix
-% % \end{verbatim}
-% % and note that in the appendix, sections are lettered, not
-% % numbered. This document has two appendices, demonstrating the section
-% % and subsection identification method.
+% Start the appendix with the ``\verb|appendix|'' command:
+% \begin{verbatim}
+%   \appendix
+% \end{verbatim}
+% and note that in the appendix, sections are lettered, not
+% numbered. This document has two appendices, demonstrating the section
+% and subsection identification method.
 
-% % \section{SIGCHI Extended Abstracts}
+% %%
+% %% The acknowledgments section is defined using the "acks" environment
+% %% (and NOT an unnumbered section). This ensures the proper
+% %% identification of the section in the article metadata, and the
+% %% consistent spelling of the heading.
+% \begin{acks}
+% To Robert, for the bagels and explaining CMYK and color spaces.
+% \end{acks}
 
-% % The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
-% % not in Word) produces a landscape-orientation formatted article, with
-% % a wide left margin. Three environments are available for use with the
-% % ``\verb|sigchi-a|'' template style, and produce formatted output in
-% % the margin:
-% % \begin{itemize}
-% % \item {\verb|sidebar|}:  Place formatted text in the margin.
-% % \item {\verb|marginfigure|}: Place a figure in the margin.
-% % \item {\verb|margintable|}: Place a table in the margin.
-% % \end{itemize}
-
-% % %%
-% % %% The acknowledgments section is defined using the "acks" environment
-% % %% (and NOT an unnumbered section). This ensures the proper
-% % %% identification of the section in the article metadata, and the
-% % %% consistent spelling of the heading.
-% % \begin{acks}
-% % To Robert, for the bagels and explaining CMYK and color spaces.
-% % \end{acks}
-
-% % %%
-% % %% The next two lines define the bibliography style to be used, and
-% % %% the bibliography file.
-% % \bibliographystyle{ACM-Reference-Format}
-% % \bibliography{sample-base}
+%%
+%% The next two lines define the bibliography style to be used, and
+%% the bibliography file.
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{papoc}
 
@@
 
-% % \section{Research Methods}
-
-% % \subsection{Part One}
-
-% % Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
-% % malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
-% % sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
-% % vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
-% % lacinia dolor. Integer ultricies commodo sem nec semper.
-
-% % \subsection{Part Two}
-
-% % Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
-% % ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
-% % ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
-% % eros. Vivamus non purus placerat, scelerisque diam eu, cursus
-% % ante. Etiam aliquam tortor auctor efficitur mattis.
-
-% % \section{Online Resources}
-
-% % Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
-% % pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
-% % enim maximus. Vestibulum gravida massa ut felis suscipit
-% % congue. Quisque mattis elit a risus ultrices commodo venenatis eget
-% % dui. Etiam sagittis eleifend elementum.
-
-% % Nam interdum magna at lectus dignissim, ac dignissim lorem
-% % rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
-% % massa et mattis lacinia.
-
 \end{document}
 \endinput
-%%
-%% End of file `sample-sigplan.tex'.

### 2020-02-17 approximations/plots.R

@@ -1,22 +1,0 @@
-## comparison plot
-comparison <- function(save=F,df){
-  if(save==T){
-    pdf("../images/comparison.pdf")
-  }
-  p=ggplot(data=df, aes(x=lam,y=u,linetype=type,color=type)) +
-    geom_line() +
-    scale_linetype_discrete(name="Type: ",labels=unname(TeX(c('$\\epsilon = 100%$',"$\\epsilon$ = 5%","$\\epsilon$ = 1%")))) +
-    scale_color_discrete(name="Type: ",labels=unname(TeX(c('$\\epsilon = 100%$',"$\\epsilon$ = 5%","$\\epsilon$ = 1%")))) +
-    scale_x_continuous(name=unname(TeX(c("Read-Write Transactions per second (10% TPS)"))),
-                       breaks=seq(100,1500,200)) +
-    scale_y_continuous(name=unname(TeX(c("Time (in hours) for 10% DB Corruption"))),
-                       breaks=seq(0,3000,500)) +
-    theme(legend.text.align = 0)
-  p
-  if(save==T){
-    dev.off()
-  }
-  return(p)
-}
-
-comparison(save=T)

### 2020-02-17 simulations/corruption/make-plots.R

@@ -1,0 +1,54 @@
+library(ggplot2)
+library(latex2exp)
+# load time to corruption data in 
+no_isolation = read.csv("../../results/corruption/no_isolation.csv")
+type = rep(1,14)
+no_isolation = cbind(no_isolation,type)
+
+e01 = read.csv("../../results/corruption/probabilistic_0_1.csv")
+type = rep(2,14)
+e01 = cbind(e01,type)
+
+e005 = read.csv("../../results/corruption/probabilistic_0_05.csv")
+type = rep(3,14)
+e005 = cbind(e005,type)
+
+e001 = read.csv("../../results/corruption/probabilistic_0_01.csv")
+type = rep(4,14)
+e001 = cbind(e001,type)
+
+
+df = rbind(no_isolation,e005, e01,e001)
+
+df$type = as.factor(df$type)
+
+
+png("../paper/images/corruption_comparison.png")
+
+
+
+p=ggplot(data=df, aes(x=lam,y=U,linetype=type,color=type)) +
+    geom_line() +
+    scale_linetype_discrete(name="Type: ",labels=unname(TeX(c(
+      '$\\epsilon = 100%$',
+      "$\\epsilon$ = 10%",
+      "$\\epsilon$ = 5%",
+      "$\\epsilon$ = 1%")))) +
+    scale_color_discrete(name="Type: ",labels=unname(TeX(c(
+      '$\\epsilon = 100%$',
+      "$\\epsilon$ = 10%",
+      "$\\epsilon$ = 5%",
+      "$\\epsilon$ = 1%")))) +
+    scale_x_continuous(name=unname(TeX(c("Read-Write Transactions per second (10% TPS)"))),
+                       breaks=seq(100,1000,100),limits=c(0,1000)) +
+    scale_y_continuous(name=unname(TeX(c("Time (in hours) for 10% DB Corruption"))),
+                       breaks=seq(0,5000,500)) +
+    theme(legend.text.align = 0)
+
+p
+
+
+dev.off()
+
+
+

### 2020-02-17 simulations/deterministic/src/main/java/DeterministicProtocolSimulation.java

@@ -5,6 +5,8 @@ import org.apache.log4j.Logger;
 import state.*;
 import utils.*;
 
+import java.util.ArrayList;
+
 public class DeterministicProtocolSimulation {
 
     private final static Logger LOGGER = Logger.getLogger(DeterministicProtocolSimulation.class.getName());
@@ -12,8 +14,10 @@ public class DeterministicProtocolSimulation {
     public static void main(String[] args) {
 
         long startTime = System.currentTimeMillis();
-
         int databaseSize = SimulationConfiguration.getInstance().getDatabaseSize();
+        int[] tpsRange = new int[]{ 100,200,300,400,500,600,700,800,900,1000,2000,3000,4000,5000};
+
+        SimulationConfiguration.getInstance().setTPS("200");
 
         // init. database
         for (int i = 0; i < databaseSize; i++ ) {
@@ -24,7 +28,6 @@ public class DeterministicProtocolSimulation {
         // init. first arrival event
         double initEventTime = SimulationRandom.getInstance().generateNextArrival();
         GlobalEventListSingleton.getInstance().addEvent(new ArrivalEvent(initEventTime, EventType.ARRIVAL));
-
 
         LOGGER.debug("**********************************************************************");
         LOGGER.debug("**********************************************************************");

### 2020-02-17 simulations/deterministic/src/main/java/utils/SimulationConfiguration.java

@@ -36,9 +36,14 @@ public class SimulationConfiguration {
      * Get average transaction arrival rate
      * @return double, average transaction arrival rate
      */
-     Integer getTPS() {
+    Integer getTPS() {
         return (Integer.parseInt(simulationConfiguration.getProperty("tps")));
     }
+
+   public void setTPS(String tps) {
+        simulationConfiguration.setProperty("tps",tps);
+    }
+
 
     double getAverageNetworkDelay() {
         return Double.parseDouble(simulationConfiguration.getProperty("av.network.delay"));

### 2020-02-20 paper/main.tex
big commit

@@
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,10pt]{acmart}
 
@@
 \usepackage[ruled,vlined]{algorithm2e}
+\usepackage{subcaption}
+\usepackage{tikz}
+\usepackage{fancyvrb}
+\usepackage{xcolor}
+\usepackage{breqn}
+
+\usetikzlibrary{decorations.pathmorphing}
+\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
+\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
+\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
+\tikzstyle{local} = [thick,->,>=stealth]
+\tikzstyle{dist} = [thick,dashed,->,>=stealth]
+\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
+
+\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
+\definecolor{red}{rgb}{0.7, 0.11, 0.11}
+\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
+\definecolor{green}{rgb}{0.0, 0.42, 0.24}
 
@@
 \acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data,
-  April 27, 2020, Heraklion, Crete, Greece}
+\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
 \acmPrice{15.00}
@@
 
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
 \renewcommand{\shortauthors}{Ezhilchelvan, et al.}
@@
 \begin{abstract}
-
-  In this paper we formalize the notion of \textit{reciprocal consistency}, a important feature specific to graph databases.
-If reciprocal consistency is not enforced during transaction processing, a scale-free distributed graph database can become semantically corrupt within a short time period relative to database lifetime.
-Reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., snapshot isolation) incurring well established performance costs.
-However, in practice systems are often deployed with weaker consistency, this paper describes two protocols that provide only reciprocal consistency semantics.
-Catering for users that are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed graph database.
-Both protocols presented are compared and contrasted using metrics such as number of aborts per second.
-
+  In this paper the notion of \textit{reciprocal consistency}, a important feature specific to graph databases, is formalized.
+If reciprocal consistency is not enforced during transaction processing, a distributed graph database can become operationally corrupt within a short time period relative to database lifetime.
+Reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., Snapshot Isolation) incurring well established performance costs.
+However, in practice systems are often deployed with weaker consistency, this paper describes a protocol that guarantees only reciprocal consistency.
+Catering for application programmers that are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed graph database.
+The protocol is probabilistic ensuring reciprocal consistency provided certain conditions hold.
+The rate of corruption when these conditions no longer hold is investigated.
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:introduction}
 
-Recent years have seen a proliferation in the use of graph technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{labeled property graph}, vertices represent entities and edges represent the relationship between entities; properties can be stored on both vertices and edges \cite{Robinson2015}.
+Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
+Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
+Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationship between entities.
+In addition, properties can be stored on both vertices and edges.
 
-In practice graphs can be extremely large \cite{Sahu2017}, exceeding the storage capacity of a single-node database.
+In practice graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node database management system (DBMS).
 An approach to storing large graphs is to partition graph data over several machines in a cluster.
-Moving from a single-node database to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
+Moving from a single-node DBMS to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
 Graphs place a high demand on the underlying database and maintaining structural integrity across partitions - called \textit{reciprocal consistency} - in a distributed graph database is non-trivial.
@@
 In order to preserve reciprocal consistency, one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs.
-To avoid these costs database managements systems (DBMSs) provide several weak isolation models \cite{Berenson1995}.
-In addition, there are a number of systems that eschew transactional guarantees altogether, typically offering eventual consistency \cite{Bailis2013}.
-Recent work \cite{Ezhilchelvan2018} investigated how violation of reciprocal consistency can occur under eventual consistency, leading to semantic corruption at alarmingly rates.
-The contribution of this paper is the description of two protocols that provide only reciprocal consistency, a consistency model between the known extremes that provides suitable performance.
-The performance of each protocol is measured using metrics such as aborts per second.
+To avoid these costs DBMSs provide application developers with the option of several weaker isolation models \cite{Berenson1995}.
+In addition, there are a number of systems that eschew transactional guarantees altogether, typically offering eventual consistency \cite{Bailis2013}, in order to achieve higher performance.
+
+Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} investigated the use of eventually consistent semantics in a distributed graph database and highlighted how violation of reciprocal consistency can occur under weak isolation.
+Also, due to the \emph{Scale-Free} \cite{ScaleFree} nature of some graphs corruption can propagate through the database at alarmingly rates.
+
+The contribution of this paper is the description of a probabilistic protocol that focuses only on ensuring reciprocal consistency in order to preserve structural integrity of a partitioned graph database in a transaction processing context.
+The protocol's impact on the rate of corruption and its performance is evaluated and discussed.
+
+The key finding is whilst initial corruption of clean records is a rare event, due to the scale-free topology corruption propagates alarmingly quick even for modest transaction arrival rates.
+This motivates the use of deterministic reciprocal consistency protocols or the use of protocols with stronger isolation guarantees, which subsume reciprocal consistency.
 
@@
 
-In the labeled property graph data model edges have direction, there is a \textit{source} and a \textit{destination} vertex.
-Edge information is stored with \textbf{both} the source and destination vertices.
+In the property graph data model edges have direction, there is a \emph{source} and a \emph{destination} vertex.
+In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
 This facilitates bi-directional edge traversal and allows for better query performance.
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB [REF]) is for records to represent vertices containing both data values and an adjacency list containing edge ``pointers'' to other vertices.
-In this representation an edge has ``reciprocal'' entries in the adjacency lists of the vertices the edge connects Figure \ref{adj-list}.
+A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
+In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
 A query reading either the source or destination vertices should be able to reify the edge correctly, returning consistent results.
-When the adjacency lists for vertices are mutually compatible like this, we say they are \emph{reciprocally consistent}.
+When the adjacency lists for vertices are mutually compatible like this, they are said to be \emph{reciprocally consistent}.
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
   \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Graph representation [REF].}
+  \caption{Vertex records containing data values and adjacency lists \cite{janusgraph}.}
   \Description{Graph representation.}
@@
 Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices A and B, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from A (source) to B (destination).
-Adjacency lists of both A and B record information about that edge and this information is mutually reciprocal (or inverse) of each other: A's list will indicate `A \textit{wrote} B' while B's will have `B \textit{written} by A'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered by landing at B and then traversing to A; even though the edge \texttt{(A \{name:`Tolkien'\})-[:WROTE]->(B \{title:`The Hobbit'\})} is a directed edge at model level abstraction.
+It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
+\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
+\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
+\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
+\end{Verbatim}
+
+Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
+Thus, a query `list all titles by the author who wrote The Hobbit' can be answered by landing at \emph{b} and then traversing to \emph{a}; even though the edge is a directed edge at model level abstraction.
 
@@
 Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edge that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, [FIG].
+The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
+Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
 The proportion of distributed edges is highly data dependent.
-However, the number of such edges is always non-negligible ranging from 15-30\%.
+However, the number of such edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
-  \includegraphics[width=\linewidth]{./images/dist-graph}
-  \caption{A partitioned graph assigned to servers in a cluster. Some edges are local and some distributed}
+    \begin{tikzpicture}[node distance=1cm]
+
+    \node (s1) [server,label=$S_1$] {};
+    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
+    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
+
+    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
+    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
+    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
+    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
+    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
+    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
+    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
+    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
+    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
+    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
+    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
+    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
+    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
+    \draw [local] (v2) -- (v1);
+    \draw [local] (v1) -- (v3);
+    \draw [local] (v3) -- (v2);
+    \draw [local] (v2) -- (v5);
+    \draw [local] (v1) -- (v4);
+    \draw [local] (v5) -- (v4);
+    \draw [dist] (v5) -- (v6);
+    \draw [dist] (v4) -- (v7);
+    \draw [dist] (v8) -- (v4);
+    \draw [local] (v8) -- (v9);
+    \draw [local] (v9) -- (v11);
+    \draw [local] (v10) -- (v7);
+    \draw [dist] (v10) -- (v12);
+    \draw [dist] (v12) -- (v13);
+    \draw [local] (v10) -- (v6);
+    \draw [dist] (v10) -- (v11);
+    \draw [local] (v13) -- (v10);
+    \draw [local] (v11) -- (v12);
+    \end{tikzpicture}
+  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
   \Description{Graph representation.}
@@
 
-In a distributed graph database, adjacency lists can contain ``pointers'' to edges on remote servers.
+In a distributed graph database, adjacency lists can contain pointers to edges on remote servers.
 This makes maintaining distributed edge reciprocal consistency more challenging - especially given a common architecture employed by contemporary distributed graph databases.
 Often they use an existing eventually consistent database (e.g.
-Apache Cassandra [REF]) to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
+Apache Cassandra \cite{Cassandra}) to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
 Superficially, opting for this design appears to be a good choice: the user has the modeling convenience of graphs with the operational characteristics from the underlying database.
@@
 For example, Apache Cassandra provides no way of preventing updates from mutual interference in its normal multi-partition use case.
-Earlier work displayed how weak isolation undermines distributed edge reciprocal consistency, causing irreversible corruption that spreads at alarmingly rates \cite{Ezhilchelvan2018}.
+Earlier work displayed how weak isolation undermines distributed edge reciprocal consistency, causing irreversible corruption that spreads at alarmingly rates \cite{Ezhilchelvan2018}, \cite{Webber2019}.
 
-Of course, reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., snapshot isolation) but that incurs a performance cost if users are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed database.
-This paper addresses that challenge.
+\section{Database Corruption}
+\label{sec:db-corruption}
 
-\section{Preserving Reciprocal Consistency}
+\begin{figure*}[ht]
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \draw [thick] (1.3,3) -- (1.7,3);
+      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
+      \draw [thick] (3.3,3) -- (3.7,3);
+      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
+      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
+      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
+    \end{tikzpicture}
+    \caption{}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \draw [thick] (5.3,3) -- (5.7,3);
+      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
+      \draw [thick] (7.3,3) -- (7.7,3);
+      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
+      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
+      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
+    \end{tikzpicture}
+    \caption{}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \draw [thick] (9.3,3) -- (9.7,3);
+      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
+      \draw [thick] (11.3,3) -- (11.7,3);
+      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
+      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
+      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
+    \end{tikzpicture}
+    \caption{}
+  \end{subfigure}%
+  \caption{\emph{Conflict Scenarios.} Possible interleavings between concurrent writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
+In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
+In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
+  \label{conf-scen}
+\end{figure*}
 
-A single logical write to a distributed edge consists of two sub-operations, updating reciprocal information at the source and the destination vertices.
-The order in which sub-operations take place are not constrained.
-For example, when updating an edge between A and B it is equally likely to update A then B as it is to update B then A.
-Without sufficient isolation, concurrent updates can interleave in a manner that violates reciprocal consistency [FIG], such an edge is known as a \emph{half-corrupted} edge.
-Figure [FIG] outlines the interleavings that produce half-corrupted edges.
+We now explain how corruption occurs under weak isolation and propagates.
 
-% transactions consists of reads/writes
+When a given transaction writes a distributed edge it performs two distinct sub-operations, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
 
-\subsection{\emph{Delta} Protocol}
+In a distributed graph database operating with weak isolation, transactions that concurrently modify the same distributed edge can \emph{conflict}.
+That is transaction's sub-operations interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated.
+For such edges there exists a correct and a incorrect entry.
+Figure \ref{conf-scen} outlines the conflict scenarios that produce half-corrupted edges.
+Note, the order in which sub-operations take place are not constrained.
+For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
 
-In order to prevent all interleaving in Figure [FIG] the \emph{Delta} protocol adopts a probabilistic approach.
+If subsequent transactions read the incorrect entry and write further edges, \emph{semantic corruption} has been introduced into the database.
+Further semantic corruption spreads by the same mechanism.
+A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
+
+To illustrate this, consider two transactions $T_x$ and $T_y$. $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
+\textcolor{grey}{// Tx}
+\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
+\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
+\textcolor{blue}{DELETE} w
+
+\textcolor{grey}{// Ty}
+\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
+\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
+\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
+\end{Verbatim}
+Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
+
+\begin{figure}[ht]
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
+
+    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
+
+    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
+        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
+
+    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (2.4,0);
+
+  \end{tikzpicture}
+  \caption{A half-corrupted edge resulting from conflicting transactions.}
+  \label{half-corrupted}
+\end{figure}
+
+\section{\emph{Delta} Protocol}
+
+As stated earlier stronger isolation level such as serializability and snapshot isolation subsumes reciprocal consistency.
+The motivation between the \emph{Delta} protocol was to develop a lightweight protocol permitting weaker isolation, reducing coordination and improving performance.
+It is imagined this protocol could be ``bolted onto'' a distributed graph database employing a NoSQL store with weak isolation for storage.
+
+The \emph{Delta} protocol is designed to mitigate the occurrence of half-corruption and as a result prevent operational corruption.
+It is assumed database servers accept transactions from clients and execute them on their behalf.
+Transactions may consist of a series of read and write operations to any object in the database.
+However, the \emph{Delta} protocol is only invoked when a given transaction writes a distributed edge.
+Therefore, this protocol can be bolted onto existing mechanisms that provide guarantees within partitions but not across database partitions.
+
+In order to prevent all interleavings in Figure \ref{conf-scen} the \emph{Delta} protocol adopts a probabilistic approach.
 It is assumed there is a known $\Delta$, that reflects the bound on server communication delays, with a probability $(1-\epsilon)$ that $\Delta$ is violated.
-Informally, the protocol consists of two-phases.
-In phase 1 all writes by sub-operations are temporary and a sub-operation is allowed to proceed provided there is no other temporary sub-operation preceding it within $\Delta$; measured as per the local clock time.
-If the attempt does not succeed, the transaction aborts, aborting any and every previous tentative write that it may have successfully completed.
-In phase 2, a transaction that succeeds at all its tentative write attempts, decides to commit when it completes its graph traversal; the transaction commits once all its tentative writes are committed.
-This protocol avoids all conflict scenarios in Figure [FIG] provided the bound is not violated.
-If the bound is violate corruption and its spread can still occur, Figure [FIG].
-There exists an interesting relationship between $\Delta$, the proportion of aborts and the spread of corruption.
+All writes by sub-operations are temporary and a sub-operation is allowed to proceed provided there is no other temporary sub-operation preceding it within $\Delta$; measured as per the local clock time.
+Figure \ref{delta-abort} shows an instance of the protocol preventing Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$.
+An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
+A transaction that succeeds at all its tentative write attempts, decides to commit when it completes its all operations; the transaction commits once all its tentative writes are committed.
+This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided the bound  $\Delta$ is not violated.
 
-\begin{algorithm}
-  \SetAlgoLined
-  \eIf{sub-operation=1}{
-    \eIf{$\Delta = \emptyset$}{
-      Perform tentative write operation\;
-      Set $\Delta$\;
-      Issue sub-operation-2\;
-    }{
-      Abort transaction\;
-    }
-  }{
-    \eIf{$\Delta = \emptyset$}{
-      Perform tentative write operation\;
-      Set $\Delta$\;
-      Continue transaction\;
-    }{
-      Abort transaction\;
-    }
-  }
-    \caption{\emph{Delta} Protocol Phase 1}
-\end{algorithm}
+\begin{figure}[ht]
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    % s1
+    \draw [thick] (1.3,3) -- (1.7,3);
+    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
+    % s2
+    \draw [thick] (3.3,3) -- (3.7,3);
+    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
 
-\section{Evaluation \& Results}
+    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
+    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
 
-% Spread Corruption
+    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
+
+  \end{tikzpicture}
+  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
+  \label{delta-abort}
+\end{figure}
+
+However it is possible that the bound $\Delta$ is violated.
+If this happens all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
+Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
+
+\begin{figure}[ht]
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \draw [thick] (1.3,3) -- (1.7,3);
+    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
+    \draw [thick] (3.3,3) -- (3.7,3);
+    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
+        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
+    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
+
+    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.2) node[midway,left] {$> \Delta$};
+    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.5) node[midway,right] {$> \Delta$};
+  \end{tikzpicture}
+  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
+  \label{corruption-again}
+\end{figure}
+
+In short, the protocol attempts to avoid corruption by aborting transactions but cannot do so in all circumstances.
+Two questions naturally arise regards the performance of the protocol i) how does the protocol impact on the time until operational corruption?
+ii) how many transaction are aborted as a result of the protocol?
+
+\section{Modeling}
+\label{sec:modeling}
+
+In order to answer i) the model developed in \cite{Ezhilchelvan2018} and fine-tuned in \cite{Webber2019} was extended to measure the impact of the \emph{Delta} protocol\footnote{An empirical evaluation of existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
+A summary of the model is now provided, before discussing the extensions.
+
+Transactions arrive in a Poisson stream with rate $\lambda$ per second.
+Each transaction contains a random number of read operations, $K$, followed by a single write.
+Edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
+For each type, a fraction $f$ are distributed edges and the remainder are local edges.
+
+At any moment in time an edge can be in one of four states:
+\begin{enumerate}
+\item Local and clean.
+\item Distributed and clean.
+\item Half-corrupted distributed edge arising from interleaved updates.
+\item Semantically corrupted.
+\end{enumerate}
+The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
+
+\begin{figure}[ht]
+  \centering
+    \begin{tikzpicture}[node distance=2cm]
+      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
+      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
+      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
+      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
+      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
+      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
+      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
+      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
+      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
+    \end{tikzpicture}
+    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
+    \label{state-transitions}
+\end{figure}
+
+Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
+Then the probability of conflict $q_i$, of a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
+These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the average time to corruption, $U_{\gamma}$.
+At time $0$, all edges are clean (free from corruption).
+When a certain fraction, $\gamma$, of all edges become corrupted, the database itself is said to be operationally corrupt.
+Note, the model assumes crash-free hardware and bug-free software for simplicity.
+The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
+
+The $\Delta$ protocol influences the rate of corruption reducing the conflict probability, $q_i$.
+To simplify the derivation of the new conflict probability it was assumed.
+This leaves the Figure 6 as the only source of corruption.
+From this the following probability can be formulated:
+$$ q_i = P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$
+The arrival times of $T_x,T_y$ are assumed exponentially distributed , $X \sim \exp (\frac{\lambda P_i}{2N_i})$.
+The transmission times $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
+The complete derivation of $q_i$ is provided in the Appendix.
+
+Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
+The answer depends on several parameters characterizing four systemic aspects:
+\begin{itemize}
+\item \emph{Size and topology of graph database}.
+Size is expressed by the total number of edges $N$, and the fraction $f$ of edges distributed across servers.
+We consider a common edge access patterns or topologies: a Scale-free topology, edges have different access probabilities and those that get accessed more frequently tend to be smaller in number.
+\item \emph{Workload}.
+Measured as transactions per second (TPS).
+Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
+\item \emph{Distributed Write Delays and Choosing $\Delta$}.
+The smaller the delays the less likely the bound $\Delta$ is violated.
+Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
+To choose $\Delta$ one calculates $P(M>\Delta) = 1 - \epsilon$, e.g.
+$\epsilon=0.01$ equates to a $\Delta=2.3s$
+\end{itemize}
+
+To answer ii) an discrete-event simulation was performed.
+This simulation focused on a subset on the total database, the most popular edges to keep the runtime of the simulation to a reasonable time period.
+From the simulation metrics are collected measuring the number of aborts as a result of the \emph{delta} protocol for a range of transaction arrival rates and $\Delta$ values.
+
+\section{Evaluation}
+\label{sec:evaluation}
+
+The model assumes that the distributed graph database  processes many reads-followed-by-write graph queries concurrently, processing both local and distributed edges.
+The simulation is performed on a Scale-Free graph, such as a social network, human brain, or road network, large enough to make processing non-trivial.
+There are approximately 77 million local edges, approximately 33 million distributed edges (in proportion to good graph partitioning algorithms); a graph of this size would have approximately 10 million nodes.
+
+The graph consisted of five edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7, N_5=10^8$ with access probabilities $p_1 =0.5, p_2 =0.26, p_3=0.13, p_4=0.07$ and $p_5 =0.04$.
+The number of read operations per query is distributed geometrically starting at $2$, with an average of $15$.
+In all edge types, a fraction $0.3$ are distributed, the remainder are local.
+The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
+The time taken under operational corruption $U$, is measured in hours.
+$U$ considered for a range of transaction arrival rates, $\lambda = (100, ..., 2000)$.
+
+TODO: discuss results
+
+The second simulation considers the most popular class of edges, $N=10^4$, for 3 values of $\Delta = 1.2, 1.5, 2.3$, measured in seconds, which correspond to a 90,95 and 99 \% chance a given message delivery exceeds $\Delta$.
+
+TODO: discuss results
 
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/corruption_comparison}
-  \caption{A partitioned graph assigned to servers in a cluster. Some edges are local and some distributed}
+  \includegraphics[width=\linewidth]{./images/corruption_comparison2}
+  \caption{Corruption results}
   \Description{Graph representation.}
-  \label{dist-graph}
+  \label{results}
 \end{figure}
 
-% Aborts/s
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{./images/test}
+  \caption{Abort results}
+  \Description{Graph representation.}
+  \label{results2}
+\end{figure}
 
@@
 
+TODO
+
 \section{Conclusions}
 
-% \section{Introduction}
-% ACM's consolidated article template, introduced in 2017, provides a
-% consistent \LaTeX\ style for use across ACM publications, and
-% incorporates accessibility and metadata-extraction functionality
-% necessary for future Digital Library endeavors. Numerous ACM and
-% SIG-specific \LaTeX\ templates have been examined, and their unique
-% features incorporated into this single new template.
+TODO
 
-% If you are new to publishing with ACM, this document is a valuable
-% guide to the process of preparing your work for publication. If you
-% have published with ACM before, this document provides insight and
-% instruction into more recent changes to the article template.
+\section{Acknowledgments}
 
-% The ``\verb|acmart|'' document class can be used to prepare articles
-% for any ACM publication --- conference or journal, and for any stage
-% of publication, from review to final ``camera-ready'' copy, to the
-% author's own version, with {\itshape very} few changes to the source.
+TODO
 
-% % \section{Template Overview}
-% % As noted in the introduction, the ``\verb|acmart|'' document class can
-% % be used to prepare many different kinds of documentation --- a
-% % double-blind initial submission of a full-length technical paper, a
-% % two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
-% % journal article, a SIGCHI Extended Abstract, and more --- all by
-% % selecting the appropriate {\itshape template style} and {\itshape
-% %   template parameters}.
+\appendix
+\section{Appendix}
 
-% % This document will explain the major features of the document
-% % class. For further information, the {\itshape \LaTeX\ User's Guide} is
-% % available from
-% % \url{https://www.acm.org/publications/proceedings-template}.
+Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
 
-% % \subsection{Template Styles}
+Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
+Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
+If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
+Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( M_x >  \Delta + T_y) \cap (M_y > \Delta - T_x) \right]$$.
 
-% % The primary parameter given to the ``\verb|acmart|'' document class is
-% % the {\itshape template style} which corresponds to the kind of publication
-% % or SIG publishing the work. This parameter is enclosed in square
-% % brackets and is a part of the {\verb|documentclass|} command:
-% % \begin{verbatim}
-% %   \documentclass[STYLE]{acmart}
-% % \end{verbatim}
+The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
+Where, $ \rho = \beta = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
+The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
 
-% % Journals use one of three template styles. All but three ACM journals
-% % use the {\verb|acmsmall|} template style:
-% % \begin{itemize}
-% % \item {\verb|acmsmall|}: The default journal template style.
-% % \item {\verb|acmlarge|}: Used by JOCCH and TAP.
-% % \item {\verb|acmtog|}: Used by TOG.
-% % \end{itemize}
+% Therefore,  $$ g(y) = \delta e^{-\delta y} $$
 
-% % The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
-% % \begin{itemize}
-% % \item {\verb|acmconf|}: The default proceedings template style.
-% % \item{\verb|sigchi|}: Used for SIGCHI conference articles.
-% % \item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
-% % \item{\verb|sigplan|}: Used for SIGPLAN conference articles.
-% % \end{itemize}
+Therefore,
+\begin{dmath*}
+  q_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
+  =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
+  =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
+\end{dmath*}
 
-% % \subsection{Template Parameters}
+To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
+\begin{align*}
+  P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
+  1 - e^{- \delta \Delta} & = 1 - \epsilon \\
+  e^{- \delta d} & = \epsilon  \\
+  \Delta & = - \frac{\ln(\epsilon)}{\delta}
+\end{align*}
 
-% % In addition to specifying the {\itshape template style} to be used in
-% % formatting your work, there are a number of {\itshape template parameters}
-% % which modify some part of the applied template style. A complete list
-% % of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}
+Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
+For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
+% \begin{align}
+%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
+% \end{align}
+% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
+% \begin{align}
+%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
+%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
+% \end{align}
 
-% % Frequently-used parameters, or combinations of parameters, include:
-% % \begin{itemize}
-% % \item {\verb|anonymous,review|}: Suitable for a ``double-blind''
-% %   conference submission. Anonymizes the work and includes line
-% %   numbers. Use with the \verb|\acmSubmissionID| command to print the
-% %   submission's unique ID on each page of the work.
-% % \item{\verb|authorversion|}: Produces a version of the work suitable
-% %   for posting by the author.
-% % \item{\verb|screen|}: Produces colored hyperlinks.
-% % \end{itemize}
+%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
+%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
 
-% % This document uses the following string as the first command in the
-% % source file:
-% % \begin{verbatim}
-% % \documentclass[sigplan,screen]{acmart}
-% % \end{verbatim}
+%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
+%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
 
-% % \section{Modifications}
-
-% % Modifying the template --- including but not limited to: adjusting
-% % margins, typeface sizes, line spacing, paragraph and list definitions,
-% % and the use of the \verb|\vspace| command to manually adjust the
-% % vertical spacing between elements of your work --- is not allowed.
-
-% % {\bfseries Your document will be returned to you for revision if
-% %   modifications are discovered.}
-
-% % \section{Typefaces}
-
-% % The ``\verb|acmart|'' document class requires the use of the
-% % ``Libertine'' typeface family. Your \TeX\ installation should include
-% % this set of packages. Please do not substitute other typefaces. The
-% % ``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
-% % as they will override the built-in typeface families.
-
-% % \section{Title Information}
-
-% % The title of your work should use capital letters appropriately -
-% % \url{https://capitalizemytitle.com/} has useful rules for
-% % capitalization. Use the {\verb|title|} command to define the title of
-% % your work. If your work has a subtitle, define it with the
-% % {\verb|subtitle|} command.  Do not insert line breaks in your title.
-
-% % If your title is lengthy, you must define a short version to be used
-% % in the page headers, to prevent overlapping text. The \verb|title|
-% % command has a ``short title'' parameter:
-% % \begin{verbatim}
-% %   \title[short title]{full title}
-% % \end{verbatim}
-
-% % \section{Authors and Affiliations}
-
-% % Each author must be defined separately for accurate metadata
-% % identification. Multiple authors may share one affiliation. Authors'
-% % names should not be abbreviated; use full first names wherever
-% % possible. Include authors' e-mail addresses whenever possible.
-
-% % Grouping authors' names or e-mail addresses, or providing an ``e-mail
-% % alias,'' as shown below, is not acceptable:
-% % \begin{verbatim}
-% %   \author{Brooke Aster, David Mehldau}
-% %   \email{dave,judy,steve@university.edu}
-% %   \email{firstname.lastname@phillips.org}
-% % \end{verbatim}
-
-% % The \verb|authornote| and \verb|authornotemark| commands allow a note
-% % to apply to multiple authors --- for example, if the first two authors
-% % of an article contributed equally to the work.
-
-% % If your author list is lengthy, you must define a shortened version of
-% % the list of authors to be used in the page headers, to prevent
-% % overlapping text. The following command should be placed just after
-% % the last \verb|\author{}| definition:
-% % \begin{verbatim}
-% %   \renewcommand{\shortauthors}{McCartney, et al.}
-% % \end{verbatim}
-% % Omitting this command will force the use of a concatenated list of all
-% % of the authors' names, which may result in overlapping text in the
-% % page headers.
-
-% % The article template's documentation, available at
-% % \url{https://www.acm.org/publications/proceedings-template}, has a
-% % complete explanation of these commands and tips for their effective
-% % use.
-
-% % Note that authors' addresses are mandatory for journal articles.
-
-% % \section{Rights Information}
-
-% % Authors of any work published by ACM will need to complete a rights
-% % form. Depending on the kind of work, and the rights management choice
-% % made by the author, this may be copyright transfer, permission,
-% % license, or an OA (open access) agreement.
-
-% % Regardless of the rights management choice, the author will receive a
-% % copy of the completed rights form once it has been submitted. This
-% % form contains \LaTeX\ commands that must be copied into the source
-% % document. When the document source is compiled, these commands and
-% % their parameters add formatted text to several areas of the final
-% % document:
-% % \begin{itemize}
-% % \item the ``ACM Reference Format'' text on the first page.
-% % \item the ``rights management'' text on the first page.
-% % \item the conference information in the page header(s).
-% % \end{itemize}
-
-% % Rights information is unique to the work; if you are preparing several
-% % works for an event, make sure to use the correct set of commands with
-% % each of the works.
-
-% % The ACM Reference Format text is required for all articles over one
-% % page in length, and is optional for one-page articles (abstracts).
-
-% % \section{CCS Concepts and User-Defined Keywords}
-
-% % Two elements of the ``acmart'' document class provide powerful
-% % taxonomic tools for you to help readers find your work in an online
-% % search.
-
-% % The ACM Computing Classification System ---
-% % \url{https://www.acm.org/publications/class-2012} --- is a set of
-% % classifiers and concepts that describe the computing
-% % discipline. Authors can select entries from this classification
-% % system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
-% % commands to be included in the \LaTeX\ source.
-
-% % User-defined keywords are a comma-separated list of words and phrases
-% % of the authors' choosing, providing a more flexible way of describing
-% % the research being presented.
-
-% % CCS concepts and user-defined keywords are required for for all
-% % articles over two pages in length, and are optional for one- and
-% % two-page articles (or abstracts).
-
-% % \section{Sectioning Commands}
-
-% % Your work should use standard \LaTeX\ sectioning commands:
-% % \verb|section|, \verb|subsection|, \verb|subsubsection|, and
-% % \verb|paragraph|. They should be numbered; do not remove the numbering
-% % from the commands.
-
-% % Simulating a sectioning command by setting the first word or words of
-% % a paragraph in boldface or italicized text is {\bfseries not allowed.}
-
-% % \section{Tables}
-
-% % The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
-% % package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
-% % high-quality tables.
-
-% % Table captions are placed {\itshape above} the table.
-
-% % Because tables cannot be split across pages, the best placement for
-% % them is typically the top of the page nearest their initial cite.  To
-% % ensure this proper ``floating'' placement of tables, use the
-% % environment \textbf{table} to enclose the table's contents and the
-% % table caption.  The contents of the table itself must go in the
-% % \textbf{tabular} environment, to be aligned properly in rows and
-% % columns, with the desired horizontal and vertical rules.  Again,
-% % detailed instructions on \textbf{tabular} material are found in the
-% % \textit{\LaTeX\ User's Guide}.
-
-% % Immediately following this sentence is the point at which
-% % Table~\ref{tab:freq} is included in the input file; compare the
-% % placement of the table here with the table in the printed output of
-% % this document.
-
-% % \begin{table}
-% %   \caption{Frequency of Special Characters}
-% %   \label{tab:freq}
-% %   \begin{tabular}{ccl}
-% %     \toprule
-% %     Non-English or Math&Frequency&Comments\\
-% %     \midrule
-% %     \O & 1 in 1,000& For Swedish names\\
-% %     $\pi$ & 1 in 5& Common in math\\
-% %     \$ & 4 in 5 & Used in business\\
-% %     $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-% %   \bottomrule
-% % \end{tabular}
-% % \end{table}
-
-% % To set a wider table, which takes up the whole width of the page's
-% % live area, use the environment \textbf{table*} to enclose the table's
-% % contents and the table caption.  As with a single-column table, this
-% % wide table will ``float'' to a location deemed more
-% % desirable. Immediately following this sentence is the point at which
-% % Table~\ref{tab:commands} is included in the input file; again, it is
-% % instructive to compare the placement of the table here with the table
-% % in the printed output of this document.
-
-% % \begin{table*}
-% %   \caption{Some Typical Commands}
-% %   \label{tab:commands}
-% %   \begin{tabular}{ccl}
-% %     \toprule
-% %     Command &A Number & Comments\\
-% %     \midrule
-% %     \texttt{{\char'134}author} & 100& Author \\
-% %     \texttt{{\char'134}table}& 300 & For tables\\
-% %     \texttt{{\char'134}table*}& 400& For wider tables\\
-% %     \bottomrule
-% %   \end{tabular}
-% % \end{table*}
-
-% % \section{Math Equations}
-% % You may want to display math equations in three distinct styles:
-% % inline, numbered or non-numbered display.  Each of the three are
-% % discussed in the next sections.
-
-% % \subsection{Inline (In-text) Equations}
-% % A formula that appears in the running text is called an inline or
-% % in-text formula.  It is produced by the \textbf{math} environment,
-% % which can be invoked with the usual
-% % \texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
-% % the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
-% % and structures, from $\alpha$ to $\omega$, available in
-% % \LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
-% % examples of in-text equations in context. Notice how this equation:
-% % \begin{math}
-% %   \lim_{n\rightarrow \infty}x=0
-% % \end{math},
-% % set here in in-line math style, looks slightly different when
-% % set in display style.  (See next section).
-
-% % \subsection{Display Equations}
-% % A numbered display equation---one set off by vertical space from the
-% % text and centered horizontally---is produced by the \textbf{equation}
-% % environment. An unnumbered display equation is produced by the
-% % \textbf{displaymath} environment.
-
-% % Again, in either environment, you can use any of the symbols and
-% % structures available in \LaTeX\@; this section will just give a couple
-% % of examples of display equations in context.  First, consider the
-% % equation, shown as an inline equation above:
-% % \begin{equation}
-% %   \lim_{n\rightarrow \infty}x=0
-% % \end{equation}
-% % Notice how it is formatted somewhat differently in
-% % the \textbf{displaymath}
-% % environment.  Now, we'll enter an unnumbered equation:
-% % \begin{displaymath}
-% %   \sum_{i=0}^{\infty} x + 1
-% % \end{displaymath}
-% % and follow it with another numbered equation:
-% % \begin{equation}
-% %   \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
-% % \end{equation}
-% % just to demonstrate \LaTeX's able handling of numbering.
-
-% % \section{Figures}
-
-% % The ``\verb|figure|'' environment should be used for figures. One or
-% % more images can be placed within a figure. If your figure contains
-% % third-party material, you must clearly identify it as such, as shown
-% % in the example below.
-% % \begin{figure}[h]
-% %   \centering
-% %   \includegraphics[width=\linewidth]{sample-franklin}
-% %   \caption{1907 Franklin Model D roadster. Photograph by Harris \&
-% %     Ewing, Inc. [Public domain], via Wikimedia
-% %     Commons. (\url{https://goo.gl/VLCRBB}).}
-% %   \Description{The 1907 Franklin Model D roadster.}
-% % \end{figure}
-
-% % Your figures should contain a caption which describes the figure to
-% % the reader. Figure captions go below the figure. Your figures should
-% % {\bfseries also} include a description suitable for screen readers, to
-% % assist the visually-challenged to better understand your work.
-
-% % Figure captions are placed {\itshape below} the figure.
-
-% % \subsection{The ``Teaser Figure''}
-
-% % A ``teaser figure'' is an image, or set of images in one figure, that
-% % are placed after all author and affiliation information, and before
-% % the body of the article, spanning the page. If you wish to have such a
-% % figure in your article, place the command immediately before the
-% % \verb|\maketitle| command:
-% % \begin{verbatim}
-% %   \begin{teaserfigure}
-% %     \includegraphics[width=\textwidth]{sampleteaser}
-% %     \caption{figure caption}
-% %     \Description{figure description}
-% %   \end{teaserfigure}
-% % \end{verbatim}
-
-% % \section{Citations and Bibliographies}
-
-% % The use of \BibTeX\ for the preparation and formatting of one's
-% % references is strongly recommended. Authors' names should be complete
-% % --- use full first names (``Donald E. Knuth'') not initials
-% % (``D. E. Knuth'') --- and the salient identifying features of a
-% % reference should be included: title, year, volume, number, pages,
-% % article DOI, etc.
-
-% % The bibliography is included in your source document with these two
-% % commands, placed just before the \verb|\end{document}| command:
-% \begin{verbatim}
-% \bibliographystyle{ACM-Reference-Format}
-% \bibliography{papoc}
-% \end{verbatim}
-% % % where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-% % % suffix, of t
-%he \BibTeX\ file.
-
-% % Citations and references are numbered by default. A small number of
-% % ACM publications have citations and references formatted in the
-% % ``author year'' style; for these exceptions, please include this
-% % command in the {\bfseries preamble} (before
-% % ``\verb|\begin{document}|'') of your \LaTeX\ source:
-% % \begin{verbatim}
-% %   \citestyle{acmauthoryear}
-% % \end{verbatim}
-
-% %   Some examples.  A paginated journal article \cite{Abril07}, an
-% %   enumerated journal article \cite{Cohen07}, a reference to an entire
-% %   issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
-% %   monograph/whole book in a series (see 2a in spec. document)
-% %   \cite{Harel79}, a divisible-book such as an anthology or compilation
-% %   \cite{Editor00} followed by the same example, however we only output
-% %   the series if the volume number is given \cite{Editor00a} (so
-% %   Editor00a's series should NOT be present since it has no vol. no.),
-% %   a chapter in a divisible book \cite{Spector90}, a chapter in a
-% %   divisible book in a series \cite{Douglass98}, a multi-volume work as
-% %   book \cite{Knuth97}, an article in a proceedings (of a conference,
-% %   symposium, workshop for example) (paginated proceedings article)
-% %   \cite{Andler79}, a proceedings article with all possible elements
-% %   \cite{Smith10}, an example of an enumerated proceedings article
-% %   \cite{VanGundy07}, an informally published work \cite{Harel78}, a
-% %   doctoral dissertation \cite{Clarkson85}, a master's thesis:
-% %   \cite{anisi03}, an online document / world wide web resource
-% %   \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1)
-% %   \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05} and
-% %   (Case 3) a patent \cite{JoeScientist001}, work accepted for
-% %   publication \cite{rous08}, 'YYYYb'-test for prolific author
-% %   \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
-% %   contain 'duplicate' DOI and URLs (some SIAM articles)
-% %   \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
-% %   multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
-% %   couple of citations with DOIs:
-% %   \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
-% %   citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
-% %   \cite{R} and \cite{UMassCitations}.
-
-% % \section{Acknowledgments}
-
-% % Identification of funding sources and other support, and thanks to
-% % individuals and groups that assisted in the research and the
-% % preparation of the work should be included in an acknowledgment
-% % section, which is placed just before the reference section in your
-% % document.
-
-% % This section has a special environment:
-% % \begin{verbatim}
-% %   \begin{acks}
-% %   ...
-% %   \end{acks}
-% % \end{verbatim}
-% % so that the information contained therein can be more easily collected
-% % during the article metadata extraction phase, and to ensure
-% % consistency in the spelling of the section heading.
-
-% % Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.
-
-% \section{Appendices}
-
-% If your work needs an appendix, add it before the
-% ``\verb|\end{document}|'' command at the conclusion of your source
-% document.
-
-% Start the appendix with the ``\verb|appendix|'' command:
-% \begin{verbatim}
-%   \appendix
-% \end{verbatim}
-% and note that in the appendix, sections are lettered, not
-% numbered. This document has two appendices, demonstrating the section
-% and subsection identification method.
-
-% %%
-% %% The acknowledgments section is defined using the "acks" environment
-% %% (and NOT an unnumbered section). This ensures the proper
-% %% identification of the section in the article metadata, and the
-% %% consistent spelling of the heading.
-% \begin{acks}
-% To Robert, for the bagels and explaining CMYK and color spaces.
-% \end{acks}
-
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
 \bibliographystyle{ACM-Reference-Format}
@@
 
-% % %%
-% % %% If your work has an appendix, this is the place to put it.
-% % \appendix
-
 \end{document}

### 2020-02-20 delta.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+Rscript simulations/corruption/delta.R $1

### 2020-02-20 run-no-isolation.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-Rscript approximations/sim-no-isolation.R
+Rscript simulations/corruption/sim-no-isolation.R

### 2020-02-20 run-probabilistic.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-Rscript approximations/sim-probabilistic.R $1
+Rscript simulations/corruption/sim-probabilistic.R $1

### 2020-02-20 simulations/corruption/calculate_delta.R

@@ -1,0 +1,5 @@
+# calculating d using eps and delta
+calc_d <- function(eps,del){
+  d=-log(eps)/del
+  return(d)
+}

### 2020-02-20 simulations/corruption/delta.R

@@ -1,0 +1,8 @@
+## command line args
+args = commandArgs(trailingOnly=TRUE)
+eps = as.numeric(args[1])
+
+source(file= "./simulations/corruption/calculate_delta.R")
+
+calc_d(eps=eps,del=200)
+

### 2020-02-20 simulations/corruption/func_no_isolation.R

@@ -1,19 +1,30 @@
-no_isolation <- function(N,p,f,lam,r,gam){
-  t=0
-  n=matrix(rep(0,15),ncol=3,nrow=5)
-  tr=matrix(rep(0,25),ncol=5,nrow=5)
-  Nt=sum(N)
-  n[,1]=(1-f)*N
-  n[,2]=f*N
-  q=1-exp(-0.5*lam*0.005*(p/N))
+# N - vector of edges
+# p - access probs
+# f - fraction dist
+# lam - arrival rate
+# r - geom
+# gam - corrupted prop
+
+no_isolation <- function(N,p,f,lam,r,gam,del){
+
+  t=0 # init. time
+  n=matrix(0,ncol=3,nrow=length(N)) # db state
+  tr=matrix(0,ncol=5,nrow=length(N)) # transition matrix
+  Nt=sum(N) # total number of edges
+  n[,1]=(1-f)*N # init. local
+  n[,2]=f*N # init. dist
+
+  #q=1-exp(-0.5*lam*0.005*(p/N))
+  q=(lam*del / (2*N + lam*del)) # conflict prob
+
   while(sum(n)>Nt*(1-gam)){
-    a=rep(0,5)
-    for(i in 1:5){
+    a=rep(0,length(N))
+    for(i in 1:length(N)){
       a[i]=(p[i]/N[i])*(n[i,1]+n[i,2]+0.5*n[i,3])
     }
-    a=sum(a)
-    b=a^2*r/(1-a+a*r)
-    for(i in 1:5){
+    a=sum(a) # read prob
+    b=a^2*r/(1-a+a*r) # all reads can prob
+    for(i in 1:length(N)){
       tr[i,1]=((lam*p[i]*n[i,2])/N[i])*q[i]*b*b
       tr[i,2]=((lam*p[i]*n[i,3])/N[i])*(1-q[i])*b
       tr[i,3]=((lam*p[i]*n[i,3])/N[i])*(1-b)
@@ -22,7 +33,7 @@ no_isolation <- function(N,p,f,lam,r,gam){
     }
     tt=sum(tr)
     t = t + 1/tt
-    i = sample(c(1, 2, 3, 4, 5), size = 1, replace = TRUE, prob = rowSums(tr)/tt)
+    i = sample(seq(1,length(N)), size = 1, replace = TRUE, prob = rowSums(tr)/tt)
[…]

### 2020-02-20 simulations/corruption/func_probabilistic.R

@@ -1,12 +1,11 @@
-probabilistic <- function(N,p,f,lam,r,gam,eps){
+probabilistic <- function(N,p,f,lam,r,gam,eps,del){
   t=0
   n=matrix(rep(0,15),ncol=3,nrow=5)
   tr=matrix(rep(0,25),ncol=5,nrow=5)
   Nt=sum(N)
   n[,1]=(1-f)*N
   n[,2]=f*N
-  eps=eps
-  del=200
+  eps=eps   
   d=-log(eps)/del
   pn=(lam*p)/(N*2)
   q=rep(0,5)
@@ -45,6 +44,6 @@ probabilistic <- function(N,p,f,lam,r,gam,eps){
       n[i,1] = n[i,1] - 1
     }
   }
-  t = t/3600
+  t = t/(3600*24)
   return(t)
 }

### 2020-02-20 simulations/corruption/make-plots.R

@@ -1,5 +1,7 @@
 library(ggplot2)
 library(latex2exp)
+library(ggthemes)
+
 # load time to corruption data in 
 no_isolation = read.csv("../../results/corruption/no_isolation.csv")
 type = rep(1,14)
@@ -23,27 +25,17 @@ df = rbind(no_isolation,e005, e01,e001)
 df$type = as.factor(df$type)
 
 
-png("../paper/images/corruption_comparison.png")
+png("../../paper/images/corruption_comparison.png")
 
 
 
 p=ggplot(data=df, aes(x=lam,y=U,linetype=type,color=type)) +
     geom_line() +
-    scale_linetype_discrete(name="Type: ",labels=unname(TeX(c(
-      '$\\epsilon = 100%$',
-      "$\\epsilon$ = 10%",
-      "$\\epsilon$ = 5%",
-      "$\\epsilon$ = 1%")))) +
-    scale_color_discrete(name="Type: ",labels=unname(TeX(c(
-      '$\\epsilon = 100%$',
-      "$\\epsilon$ = 10%",
-      "$\\epsilon$ = 5%",
-      "$\\epsilon$ = 1%")))) +
-    scale_x_continuous(name=unname(TeX(c("Read-Write Transactions per second (10% TPS)"))),
-                       breaks=seq(100,1000,100),limits=c(0,1000)) +
-    scale_y_continuous(name=unname(TeX(c("Time (in hours) for 10% DB Corruption"))),
-                       breaks=seq(0,5000,500)) +
-    theme(legend.text.align = 0)
+    scale_linetype_discrete(name="Delta: ",labels=unname(TeX(c("2.3s","1.5s","1.2s","0s")))) +
+    scale_color_discrete(name="Delta: ",labels=unname(TeX(c("2.3s","1.5s","1.2s","0s")))) +
+    scale_x_continuous(name="TPS",breaks=seq(0,2000,200),limits=c(0,2000)) +
+    scale_y_continuous(name="U",breaks=seq(0,5000,500)) + 
+    theme_few()
 
 p
 

### 2020-02-20 simulations/corruption/sim-no-isolation.R

@@ -1,35 +1,53 @@
-## clear environment
-remove(list=ls())
+args = commandArgs(trailingOnly=TRUE)
+lam = as.numeric(args[1])
 
 ## load function
-source(file = "./approximations/func_no_isolation.R")
+## source(file = "./simulations/corruption/func_no_isolation.R")
+source(file = "./func_no_isolation.R")
 
-## arrival rate 
-lam = c(100,200,300,400,500,600,700,800,900,1000,2000,3000,4000,5000)
+## arrival rate
+## lam = c(100,200,300,400,500,600,700,800,900,1000,2000,3000,4000,5000)
 
 ## time until corruption
 U = rep(0,length(lam))
 
+## parameters
+N=c(10^4,10^5,10^6,10^7,10^8,10^9,10^10)
+p=c(0.5,0.25,0.13,0.06,0.03,0.02,0.01)
+f=0.3
+r=0.07
+gam=0.1
+del=200
+
 print("lam  time taken  U")
-write("lam U",file="./log/no_isolation.log",append=TRUE)
+#write("lam U",file="./log/no_isolation.log",append=TRUE)
 
 for (i in 1:length(lam)) {
   start_time = Sys.time()
   U[i] = no_isolation(
-    N=c(10^4,10^5,10^6,10^7,10^8),
-    p=c(0.5,0.26,0.13,0.07,0.04),
-    f=0.3,
-    r=0.07,
+    N=N,
+    p=p,
+    f=f,
+    r=r,
     lam=lam[i],
-    gam=0.1)
+    gam=gam,
+    del=del)
   end_time = Sys.time()
   duration = end_time - start_time
   print(paste0(lam[i]," ",duration," ",U[i]))
-  write(paste0(lam[i]," ",duration),file="./log/no_isolation.log",append=TRUE)
+  #write(paste0(lam[i]," ",duration),file="./log/no_isolation.log",append=TRUE)
 }
 
 results = cbind(lam,U)
-write.csv(results,'./results/no_isolation.csv', row.names=FALSE)
+## write.csv(results,'./results/no_isolation.csv', row.names=FALSE)
+filename=paste0("./no_isolation_",lam,".csv")
+write.csv(results,'./no_isolation.csv', row.names=FALSE)
 
-
-  
+# ArXiv paper parameters
+# N=c(10^4,10^5,10^6,10^7,10^8,10^9,10^10)
[…]

### 2020-02-20 simulations/corruption/sim-probabilistic.R

@@ -2,13 +2,14 @@
 remove(list=ls())
 
 ## load function
-source(file = "./approximations/func_probabilistic.R")
+source(file = "./simulations/corruption/func_probabilistic.R")
+source(file= "./simulations/corruption/calculate_delta.R")
 
 ## command line args
 args = commandArgs(trailingOnly=TRUE)
 eps = as.numeric(args[1])
 
-## arrival rate 
+## arrival rate
 lam = c(100,200,300,400,500,600,700,800,900,1000,2000,3000,4000,5000)
 
 ## time until corruption
@@ -16,15 +17,6 @@ U = rep(0,length(lam))
 
 write(paste0(eps),file="./log/probabilistic.log",append=TRUE)
 write("lam U",file="./log/probabilistic.log",append=TRUE)
-
-# calculating d using eps and delta
-calc_d <- function(eps,del){
-     d=-log(eps)/del
-     return(d)
-}
-
-# calc_d(eps=0.01,del=200)
-# calc_d(eps=0.05,del=200)
 
 print(paste0("eps = ",args[1]))
 print(paste0("delta = ",calc_d(eps=eps,del=200)))
@@ -39,7 +31,8 @@ for (i in 1:length(lam)) {
         r=0.07,
         lam=lam[i],
         gam=0.1,
-        eps=eps)
+        eps=eps,
+        del=200)
     end_time = Sys.time()
     duration = end_time - start_time
     print(paste0(lam[i]," ",duration," ",U[i]))
@@ -48,17 +41,3 @@ for (i in 1:length(lam)) {
 
 results = cbind(lam,U)
 write.csv(results,paste0('./results/corruption/probabilistic_',gsub("\\.","\\_",eps),'.csv'), row.names=FALSE)
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2020-02-20 simulations/corruption/stepsFP.m

@@ -1,0 +1,60 @@
+%function U = stepsFP(T,N,p,lam,r,h,f,gam)
+T = 7;
+N=[10^4,10^5,10^6,10^7,10^8,10^9,10^10];
+p=[0.5,0.26,0.13,0.05,0.03,0.02,0.01];
+lam=1000;
+r=0.07;
+h=3600;
+f=0.3;
+gam=0.1;
+
+%Find corruption time U and vectors (ni0,ni1,ni2);
+%use Fixed-Point iterations at each step to find beta
+%T=no. of types; N=vector of edge numbers; p=distribution of accesses;
+%lam=arr rate; r=geometric distr. of reads r(1-r)^(i-2);  
+%h= time increment; f=fract. distr. edges; gam=fract. corrupted
+n0 = zeros(T,3);                   %Last ni0, ni1, ni2
+n1 = zeros(T,3);                   %Next ni0, ni1, ni2
+nn = sum(N);                       %total no. of edges
+q = 1 - exp(-0.5*lam*0.005*p./N)'; %conflict probability vector
+%q = zeros(7,1);
+z = (lam*h*p./N)';                 %column of common factors
+n0(:,[1,2]) = [(1-f)*N',f*N'];     %initial nij
+U = 0;
+b = 1;                             %initial estimate of beta
+n1(:,1) = n0(:,1)-(1-b)*z.*n0(:,1);                            %next ni0
+n1(:,2) = n0(:,2)+z.*(b*(1-q).*n0(:,3)-(b^2*q+1-b).*n0(:,2));  %next ni1
+n1(:,3) = n0(:,3)+z.*(b^2*q.*n0(:,2)-(1-q*b).*n0(:,3));        %next ni2
+n = (n0 + n1)/2;                        %averages
+while nn-sum(sum(n1))<gam*nn
+    a = (p./N)*(n(:,1)+n(:,2)+0.5*n(:,3));  %alpha
+%    bA = a^2*r/(1-a*(1-r));                 %ave. beta
+    bA = a*r/(1-a*(1-r));                   %ave. beta
+    while abs(b-bA) > 0.000001
+        b = bA;
+        x = max(1-(1-b)*z,0);                     %non-negative multiplier
+        n1(:,1) = n0(:,1).*x;                        %next ni0
+        x = max(1-(b^2*q+1-b).*z,0);              %non-negative multiplier
+        n1(:,2) = n0(:,2).*x + b*(1-q).*z.*n0(:,3);  %next ni1
[…]

### 2020-02-20 simulations/probabilistic/src/main/java/ProbabilisticProtocolSimulation.java

@@ -5,9 +5,8 @@ import org.apache.log4j.Logger;
 import state.*;
 import utils.*;
 
-// TODO: strip out arbiter
-// TODO: strip out unused config
-// TODO: add mechanism
+
+
 
 // txn arrives, updates 4 edges
 // for each edge
@@ -24,62 +23,84 @@ public class ProbabilisticProtocolSimulation {
 
     public static void main(String[] args) {
 
-        long startTime = System.currentTimeMillis();
+        int[] tpsRange = {100000};
+
+        String[] headers = {"tps","arrivals","collisions","commits","duration"};
+
+        WriteOutResults.writeOut(headers);
+        for (int j = 0; j<tpsRange.length; j++) {
+
+            //clear database
+            //clear event list
+
+            long startTime = System.currentTimeMillis();
+
+            int databaseSize = SimulationConfiguration.getInstance().getDatabaseSize();
+
+            SimulationConfiguration.getInstance().setTPS(tpsRange[j]);
+
+            // initialise database
+            for (int i = 0; i < databaseSize; i++) {
+                DistributedEdge e = new DistributedEdge(i);
+                GlobalEdgeListSingleton.getInstance().addEdge(e);
+            }
+
+            // init. first arrival event
+            double initEventTime = SimulationRandom.getInstance().generateNextArrival();
+            GlobalEventListSingleton.getInstance().addEvent(new ArrivalEvent(initEventTime, EventType.ARRIVAL));
+
+            LOGGER.debug("**********************************************************************");
+            LOGGER.debug("**********************************************************************");
+            LOGGER.debug("Initial State");
+            LOGGER.debug("**********************************************************************");
+            LOGGER.debug("Event List: " + GlobalEventListSingleton.getInstance().getEventList());
+            LOGGER.debug("**********************************************************************");
+            LOGGER.debug("**********************************************************************");
+            LOGGER.debug(" ");
+
+            runSimulation();
+
+            System.out.println(" ");
+            LOGGER.info("TPS: " + SimulationConfiguration.getInstance().getTPS());
+            LOGGER.info("Arrivals: " + SystemMetrics.getInstance().getArrivals());
+            LOGGER.info("Completed: " + SystemMetrics.getInstance().getCompleted());
+            LOGGER.info("Collisions: " + SystemMetrics.getInstance().getAborts());
+            LOGGER.info("Total Commits: " + SystemMetrics.getInstance().getCommits());
[…]

### 2020-02-20 simulations/probabilistic/src/main/java/action/ArrivalAction.java

@@ -25,7 +25,6 @@ public class ArrivalAction {
 
         // create transaction
         Transaction transaction = new Transaction(GlobalEventListSingleton.getInstance().getTransactionsArrived());
-        transaction.setStartTime(arrivalEvent.getEventTime());
         LOGGER.debug("Arrival: " + transaction);
         GlobalEventListSingleton.getInstance().incrementTransactionsArrived();
 

### 2020-02-20 simulations/probabilistic/src/main/java/action/Update1Action.java

@@ -8,8 +8,6 @@ import utils.SimulationConfiguration;
 import utils.SimulationRandom;
 import utils.SystemMetrics;
 
-import java.util.ArrayList;
-
 /**
  * The first update of a distributed edge.
  * Choose source or destination
@@ -48,9 +46,7 @@ public class Update1Action {
                 GlobalEventListSingleton.getInstance().addEvent(update2Event);
             } else {
                 LOGGER.debug("Transaction Aborted");
-                update1Event.getTransaction().setEndTime(GlobalClockSingleton.getInstance().getGlobalClock());
-                SystemMetrics.getInstance().incrementCollisions();
-                SystemMetrics.getInstance().addLifetime(update1Event.getTransaction().getLifetime());
+                SystemMetrics.getInstance().incrementAborts();
                 // remove from active list
                 GlobalActiveTransactionListSingleton.getInstance().removeTransaction(update1Event.getTransaction().getId());
             }
@@ -68,9 +64,7 @@ public class Update1Action {
                 GlobalEventListSingleton.getInstance().addEvent(update2Event);
             } else {
                 LOGGER.debug("Transaction Aborted");
-                update1Event.getTransaction().setEndTime(GlobalClockSingleton.getInstance().getGlobalClock());
-                SystemMetrics.getInstance().incrementCollisions();
-                SystemMetrics.getInstance().addLifetime(update1Event.getTransaction().getLifetime());
+                SystemMetrics.getInstance().incrementAborts();
                 // remove from active list
                 GlobalActiveTransactionListSingleton.getInstance().removeTransaction(update1Event.getTransaction().getId());
             }

### 2020-02-20 simulations/probabilistic/src/main/java/action/Update2Action.java

@@ -1,8 +1,5 @@
 package action;
 
-import event.AbortCleanUpEvent;
-import event.EventType;
-import event.Update1Event;
 import event.Update2Event;
 import org.apache.log4j.Logger;
 import state.*;
@@ -36,17 +33,13 @@ public class Update2Action {
                 distributedEdge.setSourceLock(update2Event.getEventTime() + delta);
                 if (transaction.getUpdatesCompleted() == transaction.getUpdates()) { // finished updating
                     LOGGER.debug("Transaction Successful");
-                    transaction.setEndTime(update2Event.getEventTime()); // set end time
                     SystemMetrics.getInstance().incrementCommitted();
-                    SystemMetrics.getInstance().addLifetime(transaction.getLifetime());
                     GlobalActiveTransactionListSingleton.getInstance().removeTransaction(transaction.getId());
                 }
             } else {
 
                 LOGGER.debug("Transaction Aborted");
-                update2Event.getTransaction().setEndTime(GlobalClockSingleton.getInstance().getGlobalClock());
-                SystemMetrics.getInstance().incrementCollisions();
-                SystemMetrics.getInstance().addLifetime(update2Event.getTransaction().getLifetime());
+                SystemMetrics.getInstance().incrementAborts();
                 // remove from active list
                 GlobalActiveTransactionListSingleton.getInstance().removeTransaction(update2Event.getTransaction().getId());
 
@@ -60,18 +53,14 @@ public class Update2Action {
                 distributedEdge.setDestinationLock(update2Event.getEventTime() + delta);
                 if (transaction.getUpdatesCompleted() == transaction.getUpdates()) { // finished updating
                     LOGGER.debug("Transaction Successful");
-                    transaction.setEndTime(update2Event.getEventTime()); // set end time
                     SystemMetrics.getInstance().incrementCommitted();
-                    SystemMetrics.getInstance().addLifetime(transaction.getLifetime());
                     GlobalActiveTransactionListSingleton.getInstance().removeTransaction(transaction.getId());
 
                 }
             } else {
 
                 LOGGER.debug("Transaction Aborted");
-                update2Event.getTransaction().setEndTime(GlobalClockSingleton.getInstance().getGlobalClock());
-                SystemMetrics.getInstance().incrementCollisions();
-                SystemMetrics.getInstance().addLifetime(update2Event.getTransaction().getLifetime());
+                SystemMetrics.getInstance().incrementAborts();
                 // remove from active list
[…]

### 2020-02-20 simulations/probabilistic/src/main/java/event/AbortCleanUpEvent.java

@@ -1,27 +1,0 @@
-package event;
-
-import state.Transaction;
-
-public class AbortCleanUpEvent extends AbstractEvent {
-
-    private Transaction transaction;
-
-    public AbortCleanUpEvent(double eventTime, EventType eventTypeEnum, Transaction transaction) {
-        super(eventTime, eventTypeEnum);
-        this.transaction = transaction;
-    }
-
-    public Transaction getTransaction() {
-        return transaction;
-    }
-
-    @Override
-    public String toString() {
-        return "{" +
-                "transaction=" + transaction.getId() +
-                ", " + "time=" + String.format("%.5f",getEventTime()) +
-                ", " + "type=" + getEventType() +
-                "}";
-    }
-
-}

### 2020-02-20 simulations/probabilistic/src/main/java/event/EventType.java

@@ -7,6 +7,4 @@ public enum EventType {
     ARRIVAL,
     UPDATE_1,
     UPDATE_2,
-    ARBITER,
-    ABORT_CLEAN_UP
 }

### 2020-02-20 simulations/probabilistic/src/main/java/state/GlobalActiveTransactionListSingleton.java

@@ -3,6 +3,9 @@ package state;
 import java.util.ArrayList;
 import java.util.stream.Collectors;
 
+/**
+ * This class maintains a list of all transaction's active during the simulation
+ */
 public class GlobalActiveTransactionListSingleton {
 
     private static final GlobalActiveTransactionListSingleton instance = new GlobalActiveTransactionListSingleton();
@@ -26,7 +29,4 @@ public class GlobalActiveTransactionListSingleton {
 
     }
 
-    public ArrayList<Transaction> getActiveTransactionList() {
-        return activeTransactionList;
-    }
 }

### 2020-02-20 simulations/probabilistic/src/main/java/state/GlobalClockSingleton.java

@@ -1,5 +1,8 @@
 package state;
 
+/**
+ * This class represents the simulation's clock.
+ */
 public class GlobalClockSingleton {
 
     private static final GlobalClockSingleton instance = new GlobalClockSingleton();
@@ -19,5 +22,9 @@ public class GlobalClockSingleton {
 
     public double getGlobalClock() {
         return globalClock;
+    }
+
+    public void resetClock() {
+        globalClock = 0;
     }
 }

### 2020-02-20 simulations/probabilistic/src/main/java/state/GlobalEdgeListSingleton.java

@@ -55,6 +55,8 @@ public class GlobalEdgeListSingleton {
         return edgeList;
     }
 
-
+    public void clearDatabase() {
+        edgeList.clear();
+    }
 
 }

### 2020-02-20 simulations/probabilistic/src/main/java/state/GlobalEventListSingleton.java

@@ -5,6 +5,9 @@ import event.AbstractEvent;
 import java.util.ArrayList;
 import java.util.Collections;
 
+/**
+ * This class represents the simulation's event list
+ */
 public class GlobalEventListSingleton {
 
     private int transactionsArrived = 0;
@@ -23,8 +26,8 @@ public class GlobalEventListSingleton {
         eventList.add(event);
     }
 
-    public void removeEvents() {
-
+    public void clearEvents() {
+        eventList.clear();
     }
 
     public AbstractEvent getNextEvent() {

### 2020-02-20 simulations/probabilistic/src/main/java/state/Transaction.java

@@ -1,9 +1,6 @@
 package state;
 
-import utils.SimulationConfiguration;
 import utils.SimulationRandom;
-
-import java.util.ArrayList;
 
 /**
  * Represents a transaction in the database
@@ -13,38 +10,12 @@ public class Transaction {
     private int id;
     private int updates;
     private int updatesCompleted;
-    private double startTime;
-    private double endTime;
 
     public Transaction(int id) {
         this.id = id;
-        if (SimulationConfiguration.getInstance().getUseArbiter()) {
-            this.updates = SimulationRandom.getInstance().getTransactionUpdates();
-        } else {
-            this.updates = 1;
-        }
+//        this.updates = SimulationRandom.getInstance().getTransactionUpdates();
+        this.updates = 1;
         this.updatesCompleted = 0;
-    }
-
-
-    public double getStartTime() {
-        return startTime;
-    }
-
-    public double getEndTime() {
-        return endTime;
-    }
-
-    public void setStartTime(double startTime) {
-        this.startTime = startTime;
-    }
-
-    public void setEndTime(double endTime) {
-        this.endTime = endTime;
-    }
-
-    public double getLifetime() {
-        return endTime - startTime;
     }
 
     public int getId() {

### 2020-02-20 simulations/probabilistic/src/main/java/utils/SimulationConfiguration.java

@@ -31,8 +31,12 @@ public class SimulationConfiguration {
         }
     }
 
-     Integer getTPS() {
+    public Integer getTPS() {
         return (Integer.parseInt(simulationConfiguration.getProperty("tps")));
+    }
+
+    public void setTPS(long TPS) {
+        simulationConfiguration.setProperty("tps", String.valueOf(TPS));
     }
 
     double getAverageNetworkDelay() {
@@ -41,10 +45,6 @@ public class SimulationConfiguration {
 
     public int getDatabaseSize() {
         return Integer.parseInt(simulationConfiguration.getProperty("database.size"));
-    }
-
-    double getAverageArbiterServiceTime() {
-        return Double.parseDouble(simulationConfiguration.getProperty("arbiter.service"));
     }
 
     public int getTxnLimit(){
@@ -73,10 +73,6 @@ public class SimulationConfiguration {
 
     public double getDelta(){
         return Double.parseDouble(simulationConfiguration.getProperty("delta"));
-    }
-
-    public boolean getUseArbiter() {
-        return Boolean.parseBoolean(simulationConfiguration.getProperty("use.arbiter"));
     }
 
     public boolean saveResults() {

### 2020-02-20 simulations/probabilistic/src/main/java/utils/SimulationRandom.java

@@ -13,10 +13,6 @@ public class SimulationRandom {
     private Random side;
     private Random txn;
 
-    private int length; // geometric sequence length
-    private double startValue; // geometric sequence start value
-    private double commonRatio; // geometric sequence common ratio
-    private double[] geometricSequence;
     private double[] cumulativeSum;
 
     private static final SimulationRandom instance = new SimulationRandom();
@@ -27,7 +23,7 @@ public class SimulationRandom {
 
         // init random number generators
         networkDelayDistribution = new ExponentialDistribution(averageNetworkDelay);
-        double arrivalRate = (double) 1/SystemMetrics.getInstance().getTps();
+        double arrivalRate = (double) 1/SimulationConfiguration.getInstance().getTPS();
         arrivalDistribution = new ExponentialDistribution(arrivalRate);
         edgeId = new Random();
         side = new Random();
@@ -44,17 +40,20 @@ public class SimulationRandom {
         }
 
         // init geometric sequence
-        length = SimulationConfiguration.getInstance().getLength();
-        startValue = SimulationConfiguration.getInstance().getStartValue();
-        commonRatio = SimulationConfiguration.getInstance().getRatio();
-        geometricSequence = new double[length];
+        // geometric sequence length
+        int length = SimulationConfiguration.getInstance().getLength();
+        // geometric sequence start value
+        double startValue = SimulationConfiguration.getInstance().getStartValue();
+        // geometric sequence common ratio
+        double commonRatio = SimulationConfiguration.getInstance().getRatio();
+        double[] geometricSequence = new double[length];
         geometricSequence[0] = startValue;
-        for (int i = 1; i<length; i++) {
+        for (int i = 1; i< length; i++) {
             geometricSequence[i] = geometricSequence[i-1] * commonRatio;
         }
         cumulativeSum = new double[length];
         cumulativeSum[0] = geometricSequence[0];
-        for (int i = 1; i < length;i++) {
+        for (int i = 1; i < length; i++) {
 
             cumulativeSum[i] = cumulativeSum[i-1] + geometricSequence[i];
         }

### 2020-02-20 simulations/probabilistic/src/main/java/utils/SystemMetrics.java

@@ -4,30 +4,22 @@ public class SystemMetrics {
 
     private static final SystemMetrics instance = new SystemMetrics();
 
-    private int tps;
     private int arrivals;       // number of arrivals
-    private int collisions;     // aborts by delta mechanism
-    private double cumulativeLifetimes;
+    private int aborts;     // aborts by delta mechanism
     private int commits;
-    private long duration;
+    private long simulationDuration;
 
     private SystemMetrics() {
         arrivals = 0;
-        collisions = 0;
-        cumulativeLifetimes = 0;
+        aborts = 0;
         commits = 0;
-        tps = SimulationConfiguration.getInstance().getTPS();
+        simulationDuration = 0;
     }
 
     public void reset() {
         arrivals = 0;
-        collisions = 0;
-        cumulativeLifetimes = 0;
+        aborts = 0;
         commits = 0;
-    }
-
-    public int getTps() {
-        return tps;
     }
 
     public static SystemMetrics getInstance() {return instance; }
@@ -40,20 +32,16 @@ public class SystemMetrics {
         commits = commits + 1;
     }
 
-    public void incrementCollisions() {
-        collisions = collisions + 1;
-    }
-
-    public void addLifetime(double txnLifetime) {
-        cumulativeLifetimes = cumulativeLifetimes + txnLifetime;
+    public void incrementAborts() {
+        aborts = aborts + 1;
     }
 
     public int getArrivals() {
         return arrivals;
     }
 
-    public int getCollisions() {
-        return collisions;
+    public int getAborts() {
+        return aborts;
     }
 
     public int getCommits() {
@@ -61,23 +49,22 @@ public class SystemMetrics {
     }
 
     public int getCompleted() {
-        return collisions + commits;
+        return aborts + commits;
     }
 
-    public void setDuration(long duration) {
-        this.duration = duration;
+    public long getSimulationDuration() {
+        return simulationDuration;
     }
 
-    public double getAverageResponseTime() {
-        return cumulativeLifetimes / (collisions + commits);
+    public void setSimulationDuration(long simulationDuration) {
+        this.simulationDuration = simulationDuration;
     }
 
     @Override
     public String toString() {
-        return tps +
-                "," + arrivals +
-                "," + collisions +
+        return arrivals +
+                "," + aborts +
                 "," + commits +
-                "," + duration;
+                "," + simulationDuration;
     }
 }

### 2020-02-20 simulations/probabilistic/src/main/java/utils/WriteOutResults.java

@@ -9,28 +9,41 @@ import java.io.IOException;
  */
 public class WriteOutResults {
 
-    public static void writeOutResults() {
+//    public static void fileHeader() {
+//
+//        String[] headers = {"tps","arrivals","collisions","commits","duration"};
+//
+//        StringBuilder headerStringBuilder = new StringBuilder();
+//        for(String header : headers){
+//            headerStringBuilder.append(header).append(",");
+//        }
+//
+//        String headerString = headerStringBuilder.toString();
+//        if( headerString.length() > 0 ) // remove trailing comma
+//            headerString = headerString.substring(0, headerString.length() - 1);
+//
+//        writeOut(headerString);
+//    }
 
-        String[] headers = {"tps","arrivals","collisions","commits","duration"};
+    public static void writeOut(String[] string) {
 
         StringBuilder headerStringBuilder = new StringBuilder();
-        for(String header : headers){
+        for(String header : string){
             headerStringBuilder.append(header).append(",");
         }
+
         String headerString = headerStringBuilder.toString();
         if( headerString.length() > 0 ) // remove trailing comma
             headerString = headerString.substring(0, headerString.length() - 1);
 
-        String results = SystemMetrics.getInstance().toString();
 
         BufferedWriter outputStream = null;
         FileWriter fileWriter;
+        double delta = SimulationConfiguration.getInstance().getDelta();
         try {
-            fileWriter = new FileWriter("./results/probabilistic_1000.csv");
+            fileWriter = new FileWriter("./results/probabilistic/probabilistic_" + delta + ".csv",true);
             outputStream = new BufferedWriter(fileWriter);
             outputStream.append(headerString);
-            outputStream.append("\n");
-            outputStream.append(results);
             outputStream.append("\n");
         } catch (IOException e) {
             e.printStackTrace();

### 2020-02-23 paper/main.tex
init

@@
 
-%% \BibTeX command to typeset BibTeX logo in the docs
 \AtBeginDocument{%
@@
 \begin{abstract}
-  In this paper the notion of \textit{reciprocal consistency}, a important feature specific to graph databases, is formalized.
-If reciprocal consistency is not enforced during transaction processing, a distributed graph database can become operationally corrupt within a short time period relative to database lifetime.
+  In this paper the notion of \textit{reciprocal consistency}, a important constraint specific to graph databases, is formalized.
+If reciprocal consistency is not ensured during transaction processing, a distributed graph database can become operationally corrupt within a short time period relative to database lifetime.
 Reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., Snapshot Isolation) incurring well established performance costs.
-However, in practice systems are often deployed with weaker consistency, this paper describes a protocol that guarantees only reciprocal consistency.
-Catering for application programmers that are interested only in ensuring reciprocal consistency to preserve the integrity of a distributed graph database.
-The protocol is probabilistic ensuring reciprocal consistency provided certain conditions hold.
-The rate of corruption when these conditions no longer hold is investigated.
+However, in practice distributed graph databases are often built atop BASE databases with few or no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
+A lightweight protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
+Protocol performance is evaluated through simulations.
 \end{abstract}
@@
 Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationship between entities.
+Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
 In addition, properties can be stored on both vertices and edges.
 
-In practice graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node database management system (DBMS).
+In practice graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node database.
 An approach to storing large graphs is to partition graph data over several machines in a cluster.
-Moving from a single-node DBMS to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
-Graphs place a high demand on the underlying database and maintaining structural integrity across partitions - called \textit{reciprocal consistency} - in a distributed graph database is non-trivial.
+Moving from a single-node database to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
+Graphs place a high demand on the underlying database and maintaining structural integrity, specifically across partitions, is non-trivial.
 
-In order to preserve reciprocal consistency, one could implement a heavyweight coordination protocol across partitions providing Serializability or Snapshot Isolation, which have associated performance costs.
-To avoid these costs DBMSs provide application developers with the option of several weaker isolation models \cite{Berenson1995}.
+Edges are subject to a constraint called \emph{reciprocal consistency}.
+In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
+This allows for bi-directional traversal and improved query performance.
+Preserving reciprocal consistency when these reciprocal pointers reside on different partitions requires some degree of coordination between partitions in the presence of concurrent modifications.
+
+In order to preserve reciprocal consistency, one could implement a heavyweight coordination protocol across partitions providing Serializability \cite{Bernstein1987}, which has associated performance costs.
+To avoid these costs databases provide application developers with the option of several weaker isolation models \cite{Berenson1995}.
 In addition, there are a number of systems that eschew transactional guarantees altogether, typically offering eventual consistency \cite{Bailis2013}, in order to achieve higher performance.
+Such systems are referred as  BASE (basically available, soft state, eventually consistent) databases \cite{Pritchett2008}.
 
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} investigated the use of eventually consistent semantics in a distributed graph database and highlighted how violation of reciprocal consistency can occur under weak isolation.
-Also, due to the \emph{Scale-Free} \cite{ScaleFree} nature of some graphs corruption can propagate through the database at alarmingly rates.
+A common design pattern for building a distributed graph database is to use a BASE database for storage, adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
+Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency can easily be violated in such a design introducing corruption into the database.
+Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
 
-The contribution of this paper is the description of a probabilistic protocol that focuses only on ensuring reciprocal consistency in order to preserve structural integrity of a partitioned graph database in a transaction processing context.
+The contribution of this paper is the description of a protocol that ensures reciprocal consistency, preserving structural integrity in a distributed graph database assuming a transaction processing context.
+The protocol maintains these guarantee given certain condition hold.
+If these conditions are violated corruption can still occur.
 The protocol's impact on the rate of corruption and its performance is evaluated and discussed.
 
-The key finding is whilst initial corruption of clean records is a rare event, due to the scale-free topology corruption propagates alarmingly quick even for modest transaction arrival rates.
-This motivates the use of deterministic reciprocal consistency protocols or the use of protocols with stronger isolation guarantees, which subsume reciprocal consistency.
+The key finding is whilst the protocol reduces the chance of initial corruption of clean database records, due to the scale-free topology corruption propagates alarmingly quick even for modest transaction arrival rates.
+This motivates the use of deterministic reciprocal consistency protocols or the use of protocols with stronger isolation guarantees, which subsume reciprocal consistency (\cite{Bailis2014}, \cite{Berenson1995}, \cite{Bernstein1987}).
 
@@
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
+A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
 In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
 A query reading either the source or destination vertices should be able to reify the edge correctly, returning consistent results.
-When the adjacency lists for vertices are mutually compatible like this, they are said to be \emph{reciprocally consistent}.
+When the adjacency lists for vertices are mutually compatible like this, they are said to be \emph{reciprocally consistent}, a form of referential integrity.
 
@@
   \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Vertex records containing data values and adjacency lists \cite{janusgraph}.}
-  \Description{Graph representation.}
+  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
   \label{adj-list}
@@
 It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
+
+Using openCypher \cite{openCypher} this can be represented by:
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
@@
 
-In a distributed graph database, adjacency lists can contain pointers to edges on remote servers.
-This makes maintaining distributed edge reciprocal consistency more challenging - especially given a common architecture employed by contemporary distributed graph databases.
-Often they use an existing eventually consistent database (e.g.
+Adjacency lists can now contain edge pointers to vertices on remote servers.
+Maintaining reciprocal consistency for such distributed edges is challenging - especially given a common architecture employed by contemporary distributed graph databases.
+Often they use an existing BASE database (e.g.
 Apache Cassandra \cite{Cassandra}) to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
@@
 For example, Apache Cassandra provides no way of preventing updates from mutual interference in its normal multi-partition use case.
-Earlier work displayed how weak isolation undermines distributed edge reciprocal consistency, causing irreversible corruption that spreads at alarmingly rates \cite{Ezhilchelvan2018}, \cite{Webber2019}.
+Therefore, whilst maintaining reciprocal consistency for local edges is straightforward, this is not true for distributed edges.
+It is possible that concurrent updates can interleave in a manner that violates reciprocal consistency.
 
-\section{Database Corruption}
+\section{Corruption in BASE  Distributed Graph Databases}
 \label{sec:db-corruption}
@@
   \end{subfigure}%
-  \caption{\emph{Conflict Scenarios.} Possible interleavings between concurrent writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
+  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
 In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
@@
 
-We now explain how corruption occurs under weak isolation and propagates.
+Earlier work investigated how weak isolation in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates \cite{Ezhilchelvan2018}, \cite{Webber2019}.
 
-When a given transaction writes a distributed edge it performs two distinct sub-operations, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
+When a given transaction writes a distributed edge it must perform two distinct sub-operations, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
+Concurrent transaction's sub-operations interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated Figure \ref{conf-scen}.
+For such edges there exists a correct and a incorrect entry.
+Note, the order in which sub-operations take place is not constrained.
+For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
+A graph with half-corrupted edges has suffered \emph{structural corruption}.
 
-In a distributed graph database operating with weak isolation, transactions that concurrently modify the same distributed edge can \emph{conflict}.
-That is transaction's sub-operations interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated.
-For such edges there exists a correct and a incorrect entry.
-Figure \ref{conf-scen} outlines the conflict scenarios that produce half-corrupted edges.
-Note, the order in which sub-operations take place are not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
+This is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
+Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
 
-If subsequent transactions read the incorrect entry and write further edges, \emph{semantic corruption} has been introduced into the database.
+In a BASE distributed graph database there is no mechanism for repairing half-corrupt distributed edges.
+Structural corruption once occurred tends to persist\footnote{It is possible that subsequent modification could overwrite the corrupted information.}.
+
+Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
 Further semantic corruption spreads by the same mechanism.
@@
 
-To illustrate this, consider two transactions $T_x$ and $T_y$. $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
+To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$. $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
 
+TODO: Add an example of a subsequent R/W transaction that introduces semantic corruption
+
 \section{\emph{Delta} Protocol}
 
-As stated earlier stronger isolation level such as serializability and snapshot isolation subsumes reciprocal consistency.
-The motivation between the \emph{Delta} protocol was to develop a lightweight protocol permitting weaker isolation, reducing coordination and improving performance.
-It is imagined this protocol could be ``bolted onto'' a distributed graph database employing a NoSQL store with weak isolation for storage.
+As stated earlier stronger isolation levels such as Serializability and Snapshot Isolation ensure reciprocal consistency.
+The motivation behind the \emph{Delta} protocol was to develop a lightweight protocol that maintains structural integrity of a distributed graph database at a cheaper cost then protocols with stronger guarantees.
+One could imagine such a protocol be ``bolted onto'' a BASE distributed graph database.
 
-The \emph{Delta} protocol is designed to mitigate the occurrence of half-corruption and as a result prevent operational corruption.
-It is assumed database servers accept transactions from clients and execute them on their behalf.
-Transactions may consist of a series of read and write operations to any object in the database.
-However, the \emph{Delta} protocol is only invoked when a given transaction writes a distributed edge.
-Therefore, this protocol can be bolted onto existing mechanisms that provide guarantees within partitions but not across database partitions.
+The \emph{Delta} protocol is designed to prevent the occurrence of dirty writes to maintain structural integrity of the graph.
+A client-server model is assumed, with transactions comprised of a graph traversal (set of read operations) followed by a series of writes.
+There are two distinct phases, in the first phase transactions attempt perform reads and write provisionally.
+If successful an atomic commitment protocol is invoked, else the transaction is aborted and tentative writes are removed.
 
-In order to prevent all interleavings in Figure \ref{conf-scen} the \emph{Delta} protocol adopts a probabilistic approach.
-It is assumed there is a known $\Delta$, that reflects the bound on server communication delays, with a probability $(1-\epsilon)$ that $\Delta$ is violated.
-All writes by sub-operations are temporary and a sub-operation is allowed to proceed provided there is no other temporary sub-operation preceding it within $\Delta$; measured as per the local clock time.
+In order to prevent all interleavings in Figure \ref{conf-scen} it is assumed there is a known $\Delta$, that reflects the time taken to perform both sub-operations in the first phase, with a probability $(1-\epsilon)$ that $\Delta$ is exceeded.
+
+A distributed edge write sub-operation is allowed to proceed provided there is no other temporary write sub-operation preceding it within $\Delta$; measured as per the local clock time.
 Figure \ref{delta-abort} shows an instance of the protocol preventing Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$.
@@
 A transaction that succeeds at all its tentative write attempts, decides to commit when it completes its all operations; the transaction commits once all its tentative writes are committed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided the bound  $\Delta$ is not violated.
+This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is exceeded.
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
 
-However it is possible that the bound $\Delta$ is violated.
+However it is possible that $\Delta$ is exceeded.
 If this happens all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
@@
 Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of conflict $q_i$, of a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
+Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
 These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the average time to corruption, $U_{\gamma}$.
@@
 
-The $\Delta$ protocol influences the rate of corruption reducing the conflict probability, $q_i$.
+The \emph{Delta} protocol influences the rate of corruption reducing the probability a transaction corrupts an edge, $q^{new}_i$.
 To simplify the derivation of the new conflict probability it was assumed.
@@
 From this the following probability can be formulated:
-$$ q_i = P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$
+$$ q^{new}_i = P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$
 The arrival times of $T_x,T_y$ are assumed exponentially distributed , $X \sim \exp (\frac{\lambda P_i}{2N_i})$.
 The transmission times $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-The complete derivation of $q_i$ is provided in the Appendix.
+The complete derivation of $q_i$ is provided in Appendix A.
 
@@
 Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-To choose $\Delta$ one calculates $P(M>\Delta) = 1 - \epsilon$, e.g.
-$\epsilon=0.01$ equates to a $\Delta=2.3s$
+To choose $\Delta$ one calculates $P(M>\Delta) = 1 - \epsilon$.
 \end{itemize}
 
-To answer ii) an discrete-event simulation was performed.
-This simulation focused on a subset on the total database, the most popular edges to keep the runtime of the simulation to a reasonable time period.
-From the simulation metrics are collected measuring the number of aborts as a result of the \emph{delta} protocol for a range of transaction arrival rates and $\Delta$ values.
+TODO: Method for working out number of aborts ii)
 
@@
 The time taken under operational corruption $U$, is measured in hours.
-$U$ considered for a range of transaction arrival rates, $\lambda = (100, ..., 2000)$.
-
-TODO: discuss results
-
-The second simulation considers the most popular class of edges, $N=10^4$, for 3 values of $\Delta = 1.2, 1.5, 2.3$, measured in seconds, which correspond to a 90,95 and 99 \% chance a given message delivery exceeds $\Delta$.
-
-TODO: discuss results
+$U$ considered for a range of transaction arrival rates, $\lambda = (100, ..., 1000)$.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/corruption_comparison2}
+  \includegraphics[width=\linewidth]{./images/results}
   \caption{Corruption results}
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/test}
-  \caption{Abort results}
-  \Description{Graph representation.}
-  \label{results2}
-\end{figure}
+TODO: add aborts/s
+TODO: discuss results
 

### 2020-02-23 run-delta-protocol-sim.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+Rscript simulations/corruption/delta-protocol-sim.R $1 $2 $3

### 2020-02-23 run-no-isolation.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-Rscript simulations/corruption/sim-no-isolation.R
+Rscript simulations/corruption/sim-no-isolation.R $1 $2

### 2020-02-23 run-probabilistic.sh

@@ -1,3 +1,0 @@
-#!/bin/bash
-
-Rscript simulations/corruption/sim-probabilistic.R $1

### 2020-02-23 run-range-delta-protocol-sim.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+for i in 100 250 500 750 1000 2000 3000 4000 5000;
+do docker run -v `pwd`/:/usr/local/src/scripts/ -d delta-protocol ./run-delta-protocol-sim.sh $i 0.1 $1;
+done

### 2020-02-23 sim-deterministic.sh

@@ -1,3 +1,0 @@
-#!/bin/bash
-
-java -cp "simulations/deterministic/target/uber-deterministic-protocol-1.0-SNAPSHOT.jar" DeterministicProtocolSimulation

### 2020-02-23 sim-probabilistic.sh

@@ -1,3 +1,0 @@
-#!/bin/bash
-
-java -cp "simulations/probabilistic/target/uber-probabilistic-1.0-SNAPSHOT.jar" ProbabilisticProtocolSimulation

### 2020-02-23 simulations/corruption/delta-protocol-sim.R

@@ -1,11 +1,14 @@
+## command line args
 args = commandArgs(trailingOnly=TRUE)
 lam = as.numeric(args[1])
 gam = as.numeric(args[2])
+eps = as.numeric(args[3])
 
 setwd("./simulations/corruption")
 
 ## load function
-source(file = "./func_no_isolation.R")
+source(file = "./func_probabilistic.R")
+source(file= "./calculate_delta.R")
 
 ## time until corruption
 U = 0
@@ -18,19 +21,21 @@ r=0.07
 gam=gam
 del=200
 
-
 start_time = Sys.time()
-U = no_isolation(
+U = probabilistic(
     N=N,
     p=p,
     f=f,
     r=r,
     lam=lam,
     gam=gam,
-    del=del)
-  end_time = Sys.time()
-  duration = (end_time - start_time)/1000/60
+    eps=eps,
+    del=200)
+
+end_time = Sys.time()
+duration = (end_time - start_time)/1000/60
+
 
 results = cbind(lam,U,duration)
-filename=paste0("../../results/corruption/no_isolation_",lam,".csv")
-write.csv(results,filename, row.names=FALSE)
+filename = paste0('../../results/probabilistic_',lam,'_',gsub("\\.","\\_",eps),'.csv')
+write.csv(results,filename,row.names=FALSE)

### 2020-02-23 simulations/corruption/make-plots.R

@@ -31,8 +31,8 @@ png("../../paper/images/corruption_comparison.png")
 
 p=ggplot(data=df, aes(x=lam,y=U,linetype=type,color=type)) +
     geom_line() +
-    scale_linetype_discrete(name="Delta: ",labels=unname(TeX(c("2.3s","1.5s","1.2s","0s")))) +
-    scale_color_discrete(name="Delta: ",labels=unname(TeX(c("2.3s","1.5s","1.2s","0s")))) +
+    scale_linetype_discrete(name="Eps: ",labels=unname(TeX(c("NA","0.1","0.05","0.01")))) +
+    scale_color_discrete(name="Eps: ",labels=unname(TeX(c("NA","0.1","0.05","0.01")))) +
     scale_x_continuous(name="TPS",breaks=seq(0,2000,200),limits=c(0,2000)) +
     scale_y_continuous(name="U",breaks=seq(0,5000,500)) + 
     theme_few()

### 2020-02-23 simulations/corruption/sim-probabilistic.R

@@ -1,43 +1,0 @@
-## clear environment
-remove(list=ls())
-
-## load function
-source(file = "./simulations/corruption/func_probabilistic.R")
-source(file= "./simulations/corruption/calculate_delta.R")
-
-## command line args
-args = commandArgs(trailingOnly=TRUE)
-eps = as.numeric(args[1])
-
-## arrival rate
-lam = c(100,200,300,400,500,600,700,800,900,1000,2000,3000,4000,5000)
-
-## time until corruption
-U = rep(0,length(lam))
-
-write(paste0(eps),file="./log/probabilistic.log",append=TRUE)
-write("lam U",file="./log/probabilistic.log",append=TRUE)
-
-print(paste0("eps = ",args[1]))
-print(paste0("delta = ",calc_d(eps=eps,del=200)))
-print("lam  time-taken  U")
-
-for (i in 1:length(lam)) {
-    start_time = Sys.time()
-    U[i] = probabilistic(
-        N=c(10^4,10^5,10^6,10^7,10^8),
-        p=c(0.5,0.26,0.13,0.07,0.04),
-        f=0.3,
-        r=0.07,
-        lam=lam[i],
-        gam=0.1,
-        eps=eps,
-        del=200)
-    end_time = Sys.time()
-    duration = end_time - start_time
-    print(paste0(lam[i]," ",duration," ",U[i]))
-    write(paste0(lam[i]," ",duration),file="./log/probabilistic.log",append=TRUE)
-}
-
-results = cbind(lam,U)
-write.csv(results,paste0('./results/corruption/probabilistic_',gsub("\\.","\\_",eps),'.csv'), row.names=FALSE)

### 2020-02-23 simulations/corruption/stepsFP.m

@@ -1,60 +1,0 @@
-%function U = stepsFP(T,N,p,lam,r,h,f,gam)
-T = 7;
-N=[10^4,10^5,10^6,10^7,10^8,10^9,10^10];
-p=[0.5,0.26,0.13,0.05,0.03,0.02,0.01];
-lam=1000;
-r=0.07;
-h=3600;
-f=0.3;
-gam=0.1;
-
-%Find corruption time U and vectors (ni0,ni1,ni2);
-%use Fixed-Point iterations at each step to find beta
-%T=no. of types; N=vector of edge numbers; p=distribution of accesses;
-%lam=arr rate; r=geometric distr. of reads r(1-r)^(i-2);  
-%h= time increment; f=fract. distr. edges; gam=fract. corrupted
-n0 = zeros(T,3);                   %Last ni0, ni1, ni2
-n1 = zeros(T,3);                   %Next ni0, ni1, ni2
-nn = sum(N);                       %total no. of edges
-q = 1 - exp(-0.5*lam*0.005*p./N)'; %conflict probability vector
-%q = zeros(7,1);
-z = (lam*h*p./N)';                 %column of common factors
-n0(:,[1,2]) = [(1-f)*N',f*N'];     %initial nij
-U = 0;
-b = 1;                             %initial estimate of beta
-n1(:,1) = n0(:,1)-(1-b)*z.*n0(:,1);                            %next ni0
-n1(:,2) = n0(:,2)+z.*(b*(1-q).*n0(:,3)-(b^2*q+1-b).*n0(:,2));  %next ni1
-n1(:,3) = n0(:,3)+z.*(b^2*q.*n0(:,2)-(1-q*b).*n0(:,3));        %next ni2
-n = (n0 + n1)/2;                        %averages
-while nn-sum(sum(n1))<gam*nn
-    a = (p./N)*(n(:,1)+n(:,2)+0.5*n(:,3));  %alpha
-%    bA = a^2*r/(1-a*(1-r));                 %ave. beta
-    bA = a*r/(1-a*(1-r));                   %ave. beta
-    while abs(b-bA) > 0.000001
-        b = bA;
-        x = max(1-(1-b)*z,0);                     %non-negative multiplier
-        n1(:,1) = n0(:,1).*x;                        %next ni0
-        x = max(1-(b^2*q+1-b).*z,0);              %non-negative multiplier
-        n1(:,2) = n0(:,2).*x + b*(1-q).*z.*n0(:,3);  %next ni1
[…]

### 2020-02-25 paper/main.tex
dockerise delta

@@
 The \emph{Delta} protocol is designed to prevent the occurrence of dirty writes to maintain structural integrity of the graph.
-A client-server model is assumed, with transactions comprised of a graph traversal (set of read operations) followed by a series of writes.
+A client-server model is assumed, with transactions comprised of a graph traversal (set of read operations) followed by a series of writes - for simplicity a single write is assumed.
 There are two distinct phases, in the first phase transactions attempt perform reads and write provisionally.
@@
 The transmission times $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-The complete derivation of $q_i$ is provided in Appendix A.
+The complete derivation of $q^{new}_i$ is provided in Appendix A.
 
@@
 The time taken under operational corruption $U$, is measured in hours.
-$U$ considered for a range of transaction arrival rates, $\lambda = (100, ..., 1000)$.
+$U$ considered for a range of transaction arrival rates, $\lambda = (100, ..., 2000)$.
 
@@
 
-TODO: add aborts/s
+TODO: add aborts/s \\
 TODO: discuss results

### 2020-02-25 delta.sh

@@ -1,3 +1,0 @@
-#!/bin/bash
-
-Rscript simulations/corruption/delta.R $1

### 2020-02-25 run-delta-protocol-sim.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-Rscript simulations/corruption/delta-protocol-sim.R $1 $2 $3
+Rscript simulations/corruption/delta-sim.R $1 $2 $3

### 2020-02-25 run-no-isolation.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-Rscript simulations/corruption/sim-no-isolation.R $1 $2
+Rscript simulations/corruption/no-isolation-sim.R $1 $2

### 2020-02-25 run-range-delta-protocol-sim.sh

@@ -1,5 +1,5 @@
 #!/bin/bash
 
-for i in 100 250 500 750 1000 2000 3000 4000 5000;
-do docker run -v `pwd`/:/usr/local/src/scripts/ -d delta-protocol ./run-delta-protocol-sim.sh $i 0.1 $1;
+for i in 1000 2000 3000 4000 5000 6000 7000 8000 9000 10000;
+do docker run -v `pwd`/:/usr/local/src/scripts/ -d delta-protocol ./run-range-delta-protocol-sim.sh $i 0.1 $1;
 done

### 2020-02-25 setup.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+sudo apt-get update
+sudo apt-get install     apt-transport-https     ca-certificates     curl     gnupg-agent     software-properties-common
+curl -fsSL https://download.docker.com/linux/ubuntu/gpg | sudo apt-key add -
+sudo apt-key fingerprint 0EBFCD88
+sudo add-apt-repository    "deb [arch=bionic] https://download.docker.com/linux/ubuntu \
+$(lsb_release -cs) \
+stable"
+sudo add-apt-repository "deb [arch=amd64] https://download.docker.com/linux/ubuntu bionic stable"
+apt-cache madison docker-ce
+sudo apt-get install docker-ce docker-ce-cli containerd.io
+sudo docker run hello-world
+sudo usermod -aG docker ${USER}

### 2020-02-25 simulations/corruption/calculate_delta.R

@@ -1,5 +1,0 @@
-# calculating d using eps and delta
-calc_d <- function(eps,del){
-  d=-log(eps)/del
-  return(d)
-}

### 2020-02-25 simulations/corruption/delta-sim.R

@@ -2,26 +2,26 @@
 args = commandArgs(trailingOnly=TRUE)
 lam = as.numeric(args[1])
 gam = as.numeric(args[2])
-eps = as.numeric(args[3])
+D = as.numeric(args[3])
 
 setwd("./simulations/corruption")
 
 ## load function
-source(file = "./func_probabilistic.R")
-source(file= "./calculate_delta.R")
+source(file = "./func_delta.R")
 
 ## time until corruption
 U = 0
 
 ## parameters
-N=c(10^4,10^5,10^6,10^7,10^8)
-p=c(0.5,0.26,0.13,0.07,0.04)
+N=c(10^4,10^5,10^6,10^7,10^8,10^9,10^10)
+p=c(0.5,0.25,0.13,0.06,0.03,0.02,0.01)
 f=0.3
 r=0.07
 gam=gam
 del=200
 
 start_time = Sys.time()
+
 U = probabilistic(
     N=N,
     p=p,
@@ -29,13 +29,13 @@ U = probabilistic(
     r=r,
     lam=lam,
     gam=gam,
-    eps=eps,
-    del=200)
+    D=D,
+    del=del)
 
 end_time = Sys.time()
-duration = (end_time - start_time)/1000/60
+duration = (end_time - start_time)/60
 
 
 results = cbind(lam,U,duration)
-filename = paste0('../../results/probabilistic_',lam,'_',gsub("\\.","\\_",eps),'.csv')
+filename = paste0('../../results/delta_',lam,'_',gsub("\\.","\\_",D),'.csv')
 write.csv(results,filename,row.names=FALSE)

### 2020-02-25 simulations/corruption/delta.R

@@ -1,8 +1,0 @@
-## command line args
-args = commandArgs(trailingOnly=TRUE)
-eps = as.numeric(args[1])
-
-source(file= "./simulations/corruption/calculate_delta.R")
-
-calc_d(eps=eps,del=200)
-

### 2020-02-25 simulations/corruption/func_delta.R

@@ -1,25 +1,23 @@
-probabilistic <- function(N,p,f,lam,r,gam,eps,del){
+probabilistic <- function(N,p,f,lam,r,gam,D,del){
   t=0
-  n=matrix(rep(0,15),ncol=3,nrow=5)
-  tr=matrix(rep(0,25),ncol=5,nrow=5)
+  n=matrix(0,ncol=3,nrow=length(N))
+  tr=matrix(0,ncol=5,nrow=length(N))
   Nt=sum(N)
   n[,1]=(1-f)*N
   n[,2]=f*N
-  eps=eps   
-  d=-log(eps)/del
   pn=(lam*p)/(N*2)
   q=rep(0,5)
-  for(i in 1:5){
-    q[i]=eps*eps*(1-(del/(pn[i]+del))*exp(-pn[i]*d))
+  for(i in 1:length(N)){
+    q[i]=exp(-2*D*del) - (del/(pn[i] + del)) * exp(-D*(pn[i]+2*del))
   }
   while(sum(n)>Nt*(1-gam)){
-    a=rep(0,5)
-    for(i in 1:5){
+    a=rep(0,length(N))
+    for(i in 1:length(N)){
       a[i]=(p[i]/N[i])*(n[i,1]+n[i,2]+0.5*n[i,3])
     }
     a=sum(a)
     b=a^2*r/(1-a+a*r)
-    for(i in 1:5){
+    for(i in 1:length(N)){
       tr[i,1]=((lam*p[i]*n[i,2])/N[i])*q[i]*b*b
       tr[i,2]=((lam*p[i]*n[i,3])/N[i])*(1-q[i])*b
       tr[i,3]=((lam*p[i]*n[i,3])/N[i])*(1-b)
@@ -28,7 +26,7 @@ probabilistic <- function(N,p,f,lam,r,gam,eps,del){
     }
     tt=sum(tr)
     t = t + 1/tt
-    i = sample(c(1, 2, 3, 4, 5), size = 1, replace = TRUE, prob = rowSums(tr)/tt)
+    i = sample(seq(1,length(N)), size = 1, replace = TRUE, prob = rowSums(tr)/tt)
     j = sample(c(1, 2, 3, 4, 5), size = 1, replace = TRUE, prob = tr[i,]/sum(tr[i,]))
     if(j==1){
       n[i,2] = n[i,2] - 1

### 2020-02-25 simulations/corruption/func_no_isolation.R

@@ -14,7 +14,6 @@ no_isolation <- function(N,p,f,lam,r,gam,del){
   n[,1]=(1-f)*N # init. local
   n[,2]=f*N # init. dist
 
-  #q=1-exp(-0.5*lam*0.005*(p/N))
   q=(lam*del / (2*N + lam*del)) # conflict prob
 
   while(sum(n)>Nt*(1-gam)){

### 2020-02-25 simulations/corruption/no-isolation-sim.R

@@ -11,8 +11,10 @@ source(file = "./func_no_isolation.R")
 U = 0
 
 ## parameters
-N=c(10^4,10^5,10^6,10^7,10^8)
-p=c(0.5,0.26,0.13,0.07,0.04)
+
+N=c(10^4,10^5,10^6,10^7,10^8,10^9)
+#p=c(0.5,0.26,0.13,0.07,0.04)
+p=c(0.5,0.25,0.12,0.07,0.04,0.02)
 f=0.3
 r=0.07
 gam=gam

### 2020-02-25 simulations/deterministic/src/main/java/DeterministicProtocolSimulation.java

@@ -1,110 +1,0 @@
-import action.*;
-import event.*;
-
-import org.apache.log4j.Logger;
-import state.*;
-import utils.*;
-
-import java.util.ArrayList;
-
-public class DeterministicProtocolSimulation {
-
-    private final static Logger LOGGER = Logger.getLogger(DeterministicProtocolSimulation.class.getName());
-
-    public static void main(String[] args) {
-
-        long startTime = System.currentTimeMillis();
-        int databaseSize = SimulationConfiguration.getInstance().getDatabaseSize();
-        int[] tpsRange = new int[]{ 100,200,300,400,500,600,700,800,900,1000,2000,3000,4000,5000};
-
-        SimulationConfiguration.getInstance().setTPS("200");
-
-        // init. database
-        for (int i = 0; i < databaseSize; i++ ) {
-            DistributedEdge e = new DistributedEdge(i);
-            GlobalEdgeListSingleton.getInstance().addEdge(e);
-        }
-
-        // init. first arrival event
-        double initEventTime = SimulationRandom.getInstance().generateNextArrival();
-        GlobalEventListSingleton.getInstance().addEvent(new ArrivalEvent(initEventTime, EventType.ARRIVAL));
-
-        LOGGER.debug("**********************************************************************");
-        LOGGER.debug("**********************************************************************");
-        LOGGER.debug("Initial State");
-        LOGGER.debug("**********************************************************************");
-        LOGGER.debug("Event List: " + GlobalEventListSingleton.getInstance().getEventList());
-        LOGGER.debug("**********************************************************************");
-        LOGGER.debug("**********************************************************************");
-        LOGGER.debug(" ");
-
-        runSimulation();
-
-        LOGGER.info("TPS: " + SystemMetrics.getInstance().getTps());
-        LOGGER.info("Arrivals: " + SystemMetrics.getInstance().getArrivals());
-        LOGGER.info("Completed: " + SystemMetrics.getInstance().getCompleted());
-        LOGGER.info("Collisions/Aborts: " + SystemMetrics.getInstance().getCollisions());
-        LOGGER.info("Commits: " + SystemMetrics.getInstance().getCommits());
-        LOGGER.info("Av Resp. Time: " +  SystemMetrics.getInstance().getAverageResponseTime());
-
-        long endTime = System.currentTimeMillis();
-
-        long duration = (endTime - startTime)/1000/60;
-
-        SystemMetrics.getInstance().setRuntime(duration);
-
-        if (SimulationConfiguration.getInstance().saveResults()) {
[…]

### 2020-02-25 simulations/deterministic/src/main/java/action/AbortCleanUpAction.java

@@ -1,31 +1,0 @@
-package action;
-
-import event.AbortCleanUpEvent;
-import org.apache.log4j.Logger;
-import state.GlobalActiveTransactionListSingleton;
-import state.GlobalEdgeListSingleton;
-import state.Transaction;
-
-/**
- * This action is invoked by a clear event.
- */
-public class AbortCleanUpAction {
-
-    private final static Logger LOGGER = Logger.getLogger(AbortCleanUpAction.class.getName());
-
-    public void action(AbortCleanUpEvent abortCleanUpEvent) {
-
-        Transaction transaction = abortCleanUpEvent.getTransaction();
-
-        // clear provisional write
-        GlobalEdgeListSingleton.getInstance().clearProvisionalWrites(transaction.getId());
-
-        // remove from active list + everyone pred lists
-        GlobalActiveTransactionListSingleton.getInstance().removeTransaction(transaction.getId());
-
-
-    }
-
-
-
-}

### 2020-02-25 simulations/deterministic/src/main/java/action/ArrivalAction.java

@@ -1,45 +1,0 @@
-package action;
-
-import event.ArrivalEvent;
-import event.EventType;
-import event.Update1Event;
-import org.apache.log4j.Logger;
-import state.GlobalActiveTransactionListSingleton;
-import state.GlobalClockSingleton;
-import state.Transaction;
-import utils.SimulationRandom;
-import state.GlobalEventListSingleton;
-
-
-public class ArrivalAction {
-
-    private final static Logger LOGGER = Logger.getLogger(ArrivalAction.class.getName());
-
-    public void action(ArrivalEvent arrivalEvent) {
-
-        // creating next event
-        double nextArrivalEventTime = SimulationRandom.getInstance().generateNextArrival(); // next arrival time
-        ArrivalEvent nextArrivalEvent =
-                new ArrivalEvent(nextArrivalEventTime,EventType.ARRIVAL); // create event
-        GlobalEventListSingleton.getInstance().addEvent(nextArrivalEvent); // add to event list
-
-        // create transaction
-        Transaction transaction = new Transaction(GlobalEventListSingleton.getInstance().getTransactionsArrived());
-        transaction.setStartTime(arrivalEvent.getEventTime());
-        LOGGER.debug("Arrival: " + transaction);
-        GlobalEventListSingleton.getInstance().incrementTransactionsArrived();
-
-        // add to list
-        GlobalActiveTransactionListSingleton.getInstance().addTransaction(transaction);
-
-//        Update1Event update1Event = new Update1Event(SimulationRandom.getInstance().generateNetworkDelay(), EventType.UPDATE_1, transaction);
-        Update1Event update1Event =
-                new Update1Event(
-                        GlobalClockSingleton.getInstance().getGlobalClock(),
-                        EventType.UPDATE_1,
-                        transaction
-                );
-        GlobalEventListSingleton.getInstance().addEvent(update1Event); // add to event list
-
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/action/Update1Action.java

@@ -1,73 +1,0 @@
-package action;
-import event.EventType;
-import event.Update1Event;
-import event.Update2Event;
-import org.apache.log4j.Logger;
-import state.ProvisionalWrite;
-import utils.SimulationRandom;
-import state.GlobalEventListSingleton;
-import state.GlobalEdgeListSingleton;
-
-import java.util.ArrayList;
-
-/**
- * The first update of a distributed edge.
- * Choose an edge and side to update.
- * Collect collision meta data.
- * Add provisional write.
- * Create update 2 event.
- */
-public class Update1Action {
-
-    private final static Logger LOGGER = Logger.getLogger(Update1Action.class.getName());
-
-    public void action(Update1Event update1Event) {
-
-        LOGGER.debug("Transaction ID: " + update1Event.getTransaction().getId());
-
-        int edgeId = SimulationRandom.getInstance().getEdgeId(); // select a distributed edge at random
-        LOGGER.debug("Edge ID: " + edgeId);
-
-        int side = SimulationRandom.getInstance().getSide(); // choose side of distributed edge to update
-        LOGGER.debug("Edge Side: " + side); // sides indicated by 0 (left) or 1 (right)
-
-        // collect provisional write info from edge for collision detection at second update
-        ArrayList<ProvisionalWrite> pws;
-        if (side == 0) { // get information from left side
-            pws = GlobalEdgeListSingleton.getInstance().getEdge(edgeId).getLeftProvisionalWrites();
-        } else { // get information from right  side
-            pws = GlobalEdgeListSingleton.getInstance().getEdge(edgeId).getRightProvisionalWrites();
-        }
-        if (!pws.isEmpty()) { // if there is some information
-            for (ProvisionalWrite pw : pws // add to transaction's collision information
-            ) {
-                update1Event.getTransaction().addCollisionDetectionInformation(pw);
-            }
-        }
-        LOGGER.debug("Collision Detection: " + update1Event.getTransaction().getCollisionDetection());
-
-
-        LOGGER.debug("Add provisional write to edge " + edgeId);
-        ProvisionalWrite provisionalWrite =
-                new ProvisionalWrite(update1Event.getTransaction().getId(),1); // provisional write with label 1
-
-        if (side == 0) { // add provisional to the side transaction is updating
-            GlobalEdgeListSingleton.getInstance().getEdge(edgeId).addProvisionalWrite(true,provisionalWrite);
-        } else {
[…]

### 2020-02-25 simulations/deterministic/src/main/java/action/Update2Action.java

@@ -1,96 +1,0 @@
-package action;
-
-import event.AbortCleanUpEvent;
-import event.EventType;
-import event.Update1Event;
-import event.Update2Event;
-import org.apache.log4j.Logger;
-import state.*;
-import utils.*;
-
-public class Update2Action  {
-
-    private final static Logger LOGGER = Logger.getLogger(Update2Action.class.getName());
-
-    public void action(Update2Event update2Event) {
-
-        Transaction transaction = update2Event.getTransaction(); // get transaction
-        LOGGER.debug("Transaction: "  + transaction.getId());
-        LOGGER.debug("Collision Detection: "  + transaction.getCollisionDetection());
-
-        DistributedEdge distributedEdge = // get edge to be updated
-                GlobalEdgeListSingleton.getInstance().getEdge(update2Event.getEdgeId());
-        LOGGER.debug("Distributed Edge: "  + distributedEdge);
-
-        int side = update2Event.getSide(); // get side to be updated
-        LOGGER.debug("Side: "  + side);
-
-
-        boolean outcome =  // perform collision detection test
-                CollisionDetection.collisionDetection(distributedEdge,transaction,side);
-
-        if (outcome) { // passes test
-            LOGGER.debug("Collision Detection: PASS" );
-
-            transaction.removeCollisionDetectionInformation(); // clear collision meta data
-            LOGGER.debug("Clear Collection Detection: "  + transaction.getCollisionDetection());
-
-            transaction.addPredecessors(GlobalEdgeListSingleton.getInstance().collectPredecessors(distributedEdge.getEdgeId(),
-                    side)); // collect predecessor
-            LOGGER.debug("Collect Predecessors: "  + transaction.getPredecessors());
-
-            transaction.incrementUpdateCompleted(); // increment number complete
-            LOGGER.debug("Updates Completed: " + transaction.getUpdatesCompleted());
-
-            // put tentative write in the other side
-            ProvisionalWrite provisionalWrite = new ProvisionalWrite(update2Event.getTransaction().getId(),2);
-            if (side == 0) {
-                GlobalEdgeListSingleton.getInstance().getEdge(update2Event.getEdgeId()).
-                        addProvisionalWrite(true,provisionalWrite);
-            } else {
-                GlobalEdgeListSingleton.getInstance().getEdge(update2Event.getEdgeId()).
-                        addProvisionalWrite(false,provisionalWrite);
-            }
-
-            if (transaction.getUpdatesCompleted() == transaction.getUpdates()) { // finished updating
-
[…]

### 2020-02-25 simulations/deterministic/src/main/java/event/AbstractEvent.java

@@ -1,35 +1,0 @@
-package event;
-
-public abstract class AbstractEvent implements Event, Comparable<AbstractEvent> {
-
-    private double eventTime;           // event time
-    private EventType eventTypeEnum;        // event type
-
-    AbstractEvent(double eventTime, EventType eventTypeEnum) {
-        this.eventTime = eventTime;
-        this.eventTypeEnum = eventTypeEnum;
-    }
-
-    @Override
-    public double getEventTime() {
-        return eventTime;
-    }
-
-    @Override
-    public EventType getEventType() {
-        return eventTypeEnum;
-    }
-
-    @Override
-    public int compareTo(AbstractEvent event) {
-        int result;
-        if((this.eventTime - event.getEventTime()) < 0) {
-            result = -1;
-        } else if ((this.eventTime - event.getEventTime()) > 0) {
-            result = 1;
-        } else {
-            result = 0;
-        }
-        return result;
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/event/ArrivalEvent.java

@@ -1,17 +1,0 @@
-package event;
-
-public class ArrivalEvent extends AbstractEvent {
-
-    public ArrivalEvent(double eventTime, EventType eventTypeEnum) {
-        super(eventTime, eventTypeEnum);
-    }
-
-    @Override
-    public String toString() {
-        return "{" +
-                "time=" + String.format("%.5f",getEventTime()) +
-                ", " + "type=" + getEventType() +
-                "}";
-    }
-
-}

### 2020-02-25 simulations/deterministic/src/main/java/event/Event.java

@@ -1,8 +1,0 @@
-package event;
-
-public interface Event {
-
-    public double getEventTime();
-    public EventType getEventType();
-
-}

### 2020-02-25 simulations/deterministic/src/main/java/event/EventType.java

@@ -1,12 +1,0 @@
-package event;
-
-/**
- * Defining event types
- */
-public enum EventType {
-    ARRIVAL,
-    UPDATE_1,
-    UPDATE_2,
-    ARBITER,
-    ABORT_CLEAN_UP
-}

### 2020-02-25 simulations/deterministic/src/main/java/event/Update1Event.java

@@ -1,29 +1,0 @@
-package event;
-
-import state.Transaction;
-
-/**
- * An update 1 event contains the transaction performing the update
- */
-public class Update1Event extends AbstractEvent {
-
-    private Transaction transaction;
-
-    public Update1Event(double eventTime, EventType eventTypeEnum, Transaction transaction) {
-        super(eventTime, eventTypeEnum);
-        this.transaction = transaction;
-    }
-
-    public Transaction getTransaction() {
-        return transaction;
-    }
-
-    @Override
-    public String toString() {
-        return "{" +
-                "transaction=" + transaction.getId() +
-                ", " + "time=" + String.format("%.5f",getEventTime()) +
-                ", " + "type=" + getEventType() +
-                "}";
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/event/Update2Event.java

@@ -1,40 +1,0 @@
-package event;
-
-import state.Transaction;
-
-
-public class Update2Event extends AbstractEvent {
-
-    private Transaction transaction;
-    private int edgeId;
-    private int side;
-
-    public Update2Event(double eventTime, EventType eventTypeEnum, Transaction transaction, int edgeId, int side) {
-        super(eventTime, eventTypeEnum);
-        this.transaction = transaction;
-        this.edgeId = edgeId;
-        this.side = side;
-    }
-
-    public Transaction getTransaction() {
-        return transaction;
-    }
-
-    public int getEdgeId() {
-        return edgeId;
-    }
-
-    public int getSide() {
-        return side;
-    }
-
-    @Override
-    public String toString() {
-        return "{" +
-                "transaction=" + transaction.getId() +
-                ", edgeId=" + edgeId +
-                ", " + "time=" + String.format("%.5f",getEventTime()) +
-                ", " + "type=" + getEventType() +
-                "}";
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/state/DistributedEdge.java

@@ -1,70 +1,0 @@
-package state;
-
-import java.util.ArrayList;
-
-/**
- * This class represents a distributed edge.
- * A distributed edge has two distinct side, left and right.
- * Each side has it's own provisional write list containing transaction IDs.
- */
-public class DistributedEdge {
-
-    private int edgeId;
-    private ArrayList<ProvisionalWrite> leftProvisionalWrites;
-    private ArrayList<ProvisionalWrite> rightProvisionalWrites;
-
-
-    public DistributedEdge(int edgeId) {
-        this.edgeId = edgeId;
-        leftProvisionalWrites = new ArrayList<>();
-        rightProvisionalWrites = new ArrayList<>();
-    }
-
-    /**
-     * Returns the ID of a given distributed edge.
-     * @return edge ID
-     */
-    public int getEdgeId() {
-        return edgeId;
-    }
-
-    /**
-     * Add a provisional write to a distributed edge
-     * @param leftSide, true corresponds to left and false corresponds to right
-     * @param provisionalWrite, provisional writes by transactions
-     */
-    public void addProvisionalWrite(Boolean leftSide,ProvisionalWrite provisionalWrite) {
-
-        if (leftSide) {
-            leftProvisionalWrites.add(provisionalWrite);
-        } else {
-            rightProvisionalWrites.add(provisionalWrite);
-        }
-
-    }
-
-    public ArrayList<ProvisionalWrite> getLeftProvisionalWrites() {
-        return leftProvisionalWrites;
-    }
-
-    public ArrayList<ProvisionalWrite> getRightProvisionalWrites() {
-        return rightProvisionalWrites;
-    }
-
-    public void removeProvisionalWrites (Integer transactionId) {
-
-        leftProvisionalWrites.removeIf(pw -> (pw.getTransactionId() == transactionId));
-        rightProvisionalWrites.removeIf(pw -> (pw.getTransactionId() == transactionId));
-
-
-    }
-
-    @Override
-    public String toString() {
-        return "{" +
-                "id=" + edgeId +
-                ", leftPW=" + leftProvisionalWrites +
-                ", rightPW=" + rightProvisionalWrites +
-                '}';
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/state/GlobalActiveTransactionListSingleton.java

@@ -1,37 +1,0 @@
-package state;
-
-import java.util.ArrayList;
-import java.util.stream.Collectors;
-
-public class GlobalActiveTransactionListSingleton {
-
-    private static final GlobalActiveTransactionListSingleton instance = new GlobalActiveTransactionListSingleton();
-    private ArrayList<Transaction> activeTransactionList;
-
-    private GlobalActiveTransactionListSingleton() {
-        activeTransactionList = new ArrayList<>();
-    }
-
-    public static GlobalActiveTransactionListSingleton getInstance() {
-        return instance;
-    }
-
-    public void addTransaction(Transaction transaction) {
-        activeTransactionList.add(transaction);
-    }
-
-    public void removeTransaction(int transactionId){
-        // remove from list
-        activeTransactionList = (ArrayList<Transaction>) activeTransactionList.stream().filter(transaction -> transaction.getId() != transactionId).collect(Collectors.toList());
-        // remove from all pred lists
-        for (Transaction transaction: activeTransactionList
-             ) {
-            ArrayList<Integer> newPredecessors = (ArrayList<Integer>) transaction.getPredecessors().stream().filter(pred -> pred != transactionId).collect(Collectors.toList());
-            transaction.setPredecessors(newPredecessors);
-        }
-    }
-
-    public ArrayList<Transaction> getActiveTransactionList() {
-        return activeTransactionList;
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/state/GlobalClockSingleton.java

@@ -1,23 +1,0 @@
-package state;
-
-public class GlobalClockSingleton {
-
-    private static final GlobalClockSingleton instance = new GlobalClockSingleton();
-    private double globalClock;
-
-    private GlobalClockSingleton() {
-        globalClock = 0;
-    }
-
-    public static GlobalClockSingleton getInstance() {
-        return instance;
-    }
-
-    public void setGlobalClock(double eventTime) {
-        globalClock = eventTime;
-    }
-
-    public double getGlobalClock() {
-        return globalClock;
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/state/GlobalEdgeListSingleton.java

@@ -1,96 +1,0 @@
-package state;
-
-import java.util.ArrayList;
-import java.util.stream.Collectors;
-
-
-/**
- * This class represents all of the distributed edges in the database.
- */
-public class GlobalEdgeListSingleton {
-
-    private static final GlobalEdgeListSingleton instance = new GlobalEdgeListSingleton();
-
-    private ArrayList<DistributedEdge> edgeList;
-
-    private GlobalEdgeListSingleton() {
-
-        this.edgeList = new ArrayList<>();
-
-    }
-
-    /**
-     * Global point of access to edge list
-     * @return GlobalEdgeListSingleton, edge list
-     */
-    public static GlobalEdgeListSingleton getInstance() {return instance; }
-
-    /**
-     * Add edge to edge list
-     * @param edge distributed edge
-     */
-    public void addEdge(DistributedEdge edge) {
-        edgeList.add(edge);
-    }
-
-    /**
-     * Get distributed edge by id
-     * @param edgeId distributed edge id
-     * @return distributed edge
-     */
-    public DistributedEdge getEdge(int edgeId) {
-        ArrayList<DistributedEdge> result =
-                (ArrayList<DistributedEdge>) edgeList.
-                        stream().
-                        filter(e -> e.getEdgeId() == edgeId).
-                        collect(Collectors.toList());
-        return result.get(0);
-    }
-
-    /**
-     * Get all distributed edges in the database
-     * @return array list of distributed edges
-     */
-    public ArrayList<DistributedEdge> getEdgeList() {
-        return edgeList;
-    }
-
-    /**
-     * Clears the provisional writes at all edges for a given transaction id
-     * @param transactionId transaction whose provisional writes are cleared
-     */
-    public void clearProvisionalWrites(int transactionId) {
-
-        for (DistributedEdge distributedEdge: edgeList
-             ) {
-            distributedEdge.removeProvisionalWrites(transactionId);
-        }
-
-    }
-
-    /**
-     * Collects predecessors (transaction ids) for a given side of a distributed edge
-     * @param edgeId distributed edge id
-     * @param side side of the distributed edge
-     * @return
-     */
-    public ArrayList<Integer> collectPredecessors(int edgeId, int side) {
-        ArrayList<Integer> result = new ArrayList<>();
-        ArrayList<ProvisionalWrite> provisionalWrites;
-        if (side == 0) {
-            provisionalWrites =
-                    getEdge(edgeId).getLeftProvisionalWrites();
[…]

### 2020-02-25 simulations/deterministic/src/main/java/state/GlobalEventListSingleton.java

@@ -1,47 +1,0 @@
-package state;
-
-import event.AbstractEvent;
-
-import java.util.ArrayList;
-import java.util.Collections;
-
-public class GlobalEventListSingleton {
-
-    private int transactionsArrived = 0;
-
-    private static final GlobalEventListSingleton instance = new GlobalEventListSingleton();
-    private ArrayList<AbstractEvent> eventList;
-    private GlobalEventListSingleton() {
-        eventList = new ArrayList<>();
-    }
-
-    public static GlobalEventListSingleton getInstance() {
-        return instance;
-    }
-
-    public void addEvent(AbstractEvent event) {
-        eventList.add(event);
-    }
-
-    public void removeEvents() {
-
-    }
-
-    public AbstractEvent getNextEvent() {
-        Collections.sort(eventList);
-        return eventList.remove(0);
-    }
-
-    public ArrayList<AbstractEvent> getEventList() {
-        Collections.sort(eventList);
-        return eventList;
-    }
-
-    public void incrementTransactionsArrived() {
-        transactionsArrived = transactionsArrived + 1;
-    }
-
-    public int getTransactionsArrived() {
-        return transactionsArrived;
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/state/Transaction.java

@@ -1,103 +1,0 @@
-package state;
-
-import utils.SimulationConfiguration;
-import utils.SimulationRandom;
-
-import java.util.ArrayList;
-
-/**
- * Represents a transaction in the database
- */
-public class Transaction {
-
-    private int id;
-    private int updates;
-    private int updatesCompleted;
-    private ArrayList<Integer> predecessors;
-    private ArrayList<ProvisionalWrite> collisionDetection;
-    private double startTime;
-    private double endTime;
-
-    public Transaction(int id) {
-        this.id = id;
-        this.updates = SimulationRandom.getInstance().getTransactionUpdates();
-        this.updatesCompleted = 0;
-        predecessors = new ArrayList<>();
-        collisionDetection = new ArrayList<>();
-    }
-
-    void setPredecessors(ArrayList<Integer> predecessors) {
-        this.predecessors = predecessors;
-    }
-
-    public double getStartTime() {
-        return startTime;
-    }
-
-    public double getEndTime() {
-        return endTime;
-    }
-
-    public void setStartTime(double startTime) {
-        this.startTime = startTime;
-    }
-
-    public void setEndTime(double endTime) {
-        this.endTime = endTime;
-    }
-
-    public double getLifetime() {
-        return endTime - startTime;
-    }
-
-    public int getId() {
-        return id;
-    }
-
-    public int getUpdates() {
-        return updates;
-    }
-
-    public int getUpdatesCompleted() {
-        return updatesCompleted;
-    }
-
-    public ArrayList<Integer> getPredecessors() {
-        return predecessors;
-    }
-
-    public void addPredecessor(int id) {
-        predecessors.add(id);
-    }
-
-    public void addPredecessors(ArrayList<Integer> predecessorList) {
-        predecessors.addAll(predecessorList);
-    }
-
-    public void incrementUpdateCompleted() {
-        updatesCompleted = updatesCompleted + 1;
-    }
-
-    public void addCollisionDetectionInformation(ProvisionalWrite pw) {
-        collisionDetection.add(pw);
-    }
-
-    public void removeCollisionDetectionInformation() {
-        collisionDetection.clear();
-    }
-
-    public ArrayList<ProvisionalWrite> getCollisionDetection() {
-        return collisionDetection;
-    }
-
-    @Override
-    public String toString() {
-        return "Transaction{" +
-                "id=" + id +
-                ", updates=" + updates +
-                ", completed=" + updatesCompleted +
[…]

### 2020-02-25 simulations/deterministic/src/main/java/utils/CollisionDetection.java

@@ -1,100 +1,0 @@
-package utils;
-
-import org.apache.log4j.Logger;
-import state.DistributedEdge;
-import state.ProvisionalWrite;
-import state.Transaction;
-
-import java.util.ArrayList;
-import java.util.stream.Collectors;
-
-/**
- * Collision detection mechanism performed during update 2.
- * Note, side refers to the side the edge is updating, e.g. 1 implies the edge is updating the right side second.
- */
-public class CollisionDetection {
-
-    private final static Logger LOGGER = Logger.getLogger(CollisionDetection.class.getName());
-
-
-    public static boolean collisionDetection(DistributedEdge distributedEdge, Transaction transaction, int side){
-
-        // split edge information into 1 and 2
-        ArrayList<ProvisionalWrite> edgeCheck1;
-        ArrayList<ProvisionalWrite> edgeCheck2;
-
-        // if left side, need 1 and 2s from the left to check against
-        if (side == 0) {
-            edgeCheck1 = (ArrayList<ProvisionalWrite>) distributedEdge.getLeftProvisionalWrites().
-                    stream().filter(pw -> pw.getWriteLabel() == 1).collect(Collectors.toList());
-            edgeCheck2 = (ArrayList<ProvisionalWrite>) distributedEdge.getLeftProvisionalWrites().
-                    stream().filter(pw -> pw.getWriteLabel() == 2).collect(Collectors.toList());
-        } else {        // if right side, need 1 and 2s from the right to check against
-
-            edgeCheck1 = (ArrayList<ProvisionalWrite>) distributedEdge.getRightProvisionalWrites().
-                    stream().filter(pw -> pw.getWriteLabel() == 1).collect(Collectors.toList());
-            edgeCheck2 = (ArrayList<ProvisionalWrite>) distributedEdge.getRightProvisionalWrites().
-                    stream().filter(pw -> pw.getWriteLabel() == 2).collect(Collectors.toList());
-        }
-
-        // split collision detection information into 1s and 2s
-        ArrayList<ProvisionalWrite> transactionCheck1 = (ArrayList<ProvisionalWrite>) transaction.getCollisionDetection().
-                stream().filter(pw -> pw.getWriteLabel() == 1).collect(Collectors.toList());
-        ArrayList<ProvisionalWrite> transactionCheck2 = (ArrayList<ProvisionalWrite>) transaction.getCollisionDetection().
[…]

### 2020-02-25 simulations/deterministic/src/main/java/utils/SimulationConfiguration.java

@@ -1,84 +1,0 @@
-package utils;
-
-import org.apache.commons.configuration.Configuration;
-
-import java.io.InputStream;
-import java.util.Properties;
-
-/**
- * Reads in configuration .properties file
- */
-public class SimulationConfiguration {
-
-    private static final SimulationConfiguration instance = new SimulationConfiguration();
-
-    private SimulationConfiguration(){
-
-    }
-
-    public static SimulationConfiguration getInstance() {
-        return instance;
-    }
-
-    private static Properties simulationConfiguration = new Properties();
-
-    static {
-        try {
-            InputStream in = Configuration.class.getResourceAsStream("/simulation.properties");
-            simulationConfiguration.load(in);
-        } catch (Exception e) {
-            e.printStackTrace();
-        }
-    }
-
-
-    /**
-     * Get average transaction arrival rate
-     * @return double, average transaction arrival rate
-     */
-    Integer getTPS() {
-        return (Integer.parseInt(simulationConfiguration.getProperty("tps")));
-    }
-
-   public void setTPS(String tps) {
-        simulationConfiguration.setProperty("tps",tps);
-    }
-
-
-    double getAverageNetworkDelay() {
-        return Double.parseDouble(simulationConfiguration.getProperty("av.network.delay"));
-    }
-
-    public int getDatabaseSize() {
-        return Integer.parseInt(simulationConfiguration.getProperty("database.size"));
-    }
-
-    public int getTxnLimit(){
-        return Integer.parseInt(simulationConfiguration.getProperty("txn.limit"));
-    }
-
-    long getSeedValue() {
-        return Long.parseLong(simulationConfiguration.getProperty("seed.value"));
-    }
-
-    boolean isSeedSet() {
-        return Boolean.parseBoolean(simulationConfiguration.getProperty("fix.seed"));
-    }
-
-    public boolean saveResults() {
-        return Boolean.parseBoolean(simulationConfiguration.getProperty("save.results"));
-    }
-
-    int getLength() {
-     return Integer.parseInt(simulationConfiguration.getProperty("length"));
-    }
-
-    double getStartValue() {
-        return Double.parseDouble(simulationConfiguration.getProperty("start.value"));
-    }
-
-    double getRatio(){
-        return Double.parseDouble(simulationConfiguration.getProperty("common.ratio"));
-    }
-
-}

### 2020-02-25 simulations/deterministic/src/main/java/utils/SimulationRandom.java

@@ -1,103 +1,0 @@
-package utils;
-
-import org.apache.commons.math3.distribution.ExponentialDistribution;
-import state.GlobalClockSingleton;
-
-import java.util.Random;
-
-public class SimulationRandom {
-
-    private ExponentialDistribution networkDelayDistribution;
-    private ExponentialDistribution arrivalDistribution;
-    private Random edgeId;
-    private Random side;
-    private Random txn;
-
-    private int length; // geometric sequence length
-    private double startValue; // geometric sequence start value
-    private double commonRatio; // geometric sequence common ratio
-    private double[] geometricSequence;
-    private double[] cumulativeSum;
-
-    private static final SimulationRandom instance = new SimulationRandom();
-
-    public SimulationRandom() {
-
-        double averageNetworkDelay = SimulationConfiguration.getInstance().getAverageNetworkDelay();
-
-        // init random number generators
-        networkDelayDistribution = new ExponentialDistribution(averageNetworkDelay);
-        double arrivalRate = (double) 1/SystemMetrics.getInstance().getTps();
-        arrivalDistribution = new ExponentialDistribution(arrivalRate);
-        edgeId = new Random();
-        side = new Random();
-        txn = new Random();
-
-        // set seeds
-        if (SimulationConfiguration.getInstance().isSeedSet()) {
-            long seedValue = SimulationConfiguration.getInstance().getSeedValue();
-            networkDelayDistribution.reseedRandomGenerator(seedValue);
-            arrivalDistribution.reseedRandomGenerator(seedValue);
-            edgeId.setSeed(seedValue);
-            side.setSeed(seedValue);
-            txn.setSeed(seedValue);
-        }
-
-        // init geometric sequence
-        length = SimulationConfiguration.getInstance().getLength();
-        startValue = SimulationConfiguration.getInstance().getStartValue();
-        commonRatio = SimulationConfiguration.getInstance().getRatio();
-        geometricSequence = new double[length];
-        geometricSequence[0] = startValue;
-        for (int i = 1; i<length; i++) {
-            geometricSequence[i] = geometricSequence[i-1] * commonRatio;
-        }
-        cumulativeSum = new double[length];
-        cumulativeSum[0] = geometricSequence[0];
-        for (int i = 1; i < length;i++) {
-
-            cumulativeSum[i] = cumulativeSum[i-1] + geometricSequence[i];
-        }
-    }
-
-    public static SimulationRandom getInstance() {
[…]

### 2020-02-25 simulations/deterministic/src/main/java/utils/SystemMetrics.java

@@ -1,86 +1,0 @@
-package utils;
-
-//import static utils.SimulationConfiguration.getTPS;
-
-public class SystemMetrics {
-
-    private static final SystemMetrics instance = new SystemMetrics();
-
-    private int tps;
-    private int arrivals;       // number of arrivals
-    private int collisions;     // aborts by collision detection mechanism
-    private double cumulativeLifetimes;
-    private int commits;
-    private long runtime;
-
-    private SystemMetrics() {
-        arrivals = 0;
-        collisions = 0;
-        cumulativeLifetimes = 0;
-        commits = 0;
-        tps = SimulationConfiguration.getInstance().getTPS();
-        runtime = 0;
-    }
-
-    public void reset() {
-        arrivals = 0;
-        collisions = 0;
-        cumulativeLifetimes = 0;
-        commits = 0;
-    }
-
-    public int getTps() {
-        return tps;
-    }
-
-    public static SystemMetrics getInstance() {return instance; }
-
-    public void incrementArrivals() {
-        arrivals = arrivals + 1;
-    }
-
-    public void incrementCommitted() {
-        commits = commits + 1;
-    }
-
-    public void incrementCollisions() {
-        collisions = collisions + 1;
-    }
-
-    public void addLifetime(double txnLifetime) {
-        cumulativeLifetimes = cumulativeLifetimes + txnLifetime;
-    }
-
-    public int getArrivals() {
-        return arrivals;
-    }
-
-    public int getCollisions() {
-        return collisions;
-    }
-
-    public int getCommits() {
-        return commits;
-    }
-
-    public int getCompleted() {
-        return collisions + commits;
-    }
-
-    public double getAverageResponseTime() {
-        return cumulativeLifetimes / (collisions + commits);
-    }
-
-    public void setRuntime(long runtime) {
-        this.runtime = runtime;
-    }
-
-    @Override
-    public String toString() {
-        return tps +
-                "," + arrivals +
-                "," + collisions +
-                "," + commits +
-                "," + runtime;
-    }
-}

### 2020-02-25 simulations/deterministic/src/main/java/utils/WriteOutResults.java

@@ -1,50 +1,0 @@
-package utils;
-
-import java.io.BufferedWriter;
-import java.io.FileWriter;
-import java.io.IOException;
-
-/**
- * This class write the system metrics to a CSV file called results.csv.
- */
-public class WriteOutResults {
-
-    public static void writeOutResults() {
-
-        String[] headers = {"tps","arrivals","collisions","commits","time"};
-
-        StringBuilder headerStringBuilder = new StringBuilder();
-        for(String header : headers){
-            headerStringBuilder.append(header).append(",");
-        }
-        String headerString = headerStringBuilder.toString();
-        if( headerString.length() > 0 ) // remove trailing comma
-            headerString = headerString.substring(0, headerString.length() - 1);
-
-        String results = SystemMetrics.getInstance().toString();
-
-        BufferedWriter outputStream = null;
-        FileWriter fileWriter;
-        try {
-            fileWriter = new FileWriter("./results/deterministic_1000.csv");
-            outputStream = new BufferedWriter(fileWriter);
-            outputStream.append(headerString);
-            outputStream.append("\n");
-            outputStream.append(results);
-            outputStream.append("\n");
-        } catch (IOException e) {
-            e.printStackTrace();
-        } finally {
-            if (outputStream != null) {
-                try {
-                    outputStream.flush();
-                    outputStream.close();
-
-                } catch (IOException e) {
-                    e.printStackTrace();
-                }
-            }
-        }
-    }
-
-}

### 2020-02-25 simulations/probabilistic/src/main/java/ProbabilisticProtocolSimulation.java

@@ -23,7 +23,7 @@ public class ProbabilisticProtocolSimulation {
 
     public static void main(String[] args) {
 
-        int[] tpsRange = {100000};
+        int[] tpsRange = {1000};
 
         String[] headers = {"tps","arrivals","collisions","commits","duration"};
 
@@ -96,7 +96,7 @@ public class ProbabilisticProtocolSimulation {
 
     private static void runSimulation() {
 
-        while (SystemMetrics.getInstance().getCompleted() < 1000000) {
+        while (SystemMetrics.getInstance().getCompleted() < 100000) {
 
             if (SystemMetrics.getInstance().getArrivals() % 10 == 0){
                 System.out.print("Completed: " + SystemMetrics.getInstance().getCompleted() +

### 2020-02-25 paper/main.tex
dockerise aborts

@@
 
-The contribution of this paper is the description of a protocol that ensures reciprocal consistency, preserving structural integrity in a distributed graph database assuming a transaction processing context.
-The protocol maintains these guarantee given certain condition hold.
+The contribution of this paper is the description of a protocol that ensures reciprocal consistency, preserving structural integrity in a distributed graph database in a transaction processing setting.
+The protocol maintains this guarantee provided certain condition hold.
 If these conditions are violated corruption can still occur.
-The protocol's impact on the rate of corruption and its performance is evaluated and discussed.
+The protocol's impact on the rate of database corruption and its performance is evaluated and discussed.
 
@@
 
-Earlier work investigated how weak isolation in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates \cite{Ezhilchelvan2018}, \cite{Webber2019}.
+Earlier work investigated how weak isolation in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
 
 When a given transaction writes a distributed edge it must perform two distinct sub-operations, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Concurrent transaction's sub-operations interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated Figure \ref{conf-scen}.
+Concurrent transaction's sub-operations interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
 For such edges there exists a correct and a incorrect entry.
@@
 In a BASE distributed graph database there is no mechanism for repairing half-corrupt distributed edges.
-Structural corruption once occurred tends to persist\footnote{It is possible that subsequent modification could overwrite the corrupted information.}.
+Structural corruption once occurred tends to persist\footnote{It is possible that subsequent modifications could correctly overwrite the corrupted information.}.
 
@@
 
-TODO: Add an example of a subsequent R/W transaction that introduces semantic corruption
+Assume, the correct ordering of transactions is $T_y$ then $T_x$.
+One could now imagine a read-write transaction, $T_z$ that introduces new information based on the incorrect information at vertex \emph{a}.
 
@@
 
-The \emph{Delta} protocol is designed to prevent the occurrence of dirty writes to maintain structural integrity of the graph.
+The \emph{Delta} protocol is designed to prevent the occurrence of dirty writes across partitions in order to maintain structural integrity of the graph.
 A client-server model is assumed, with transactions comprised of a graph traversal (set of read operations) followed by a series of writes - for simplicity a single write is assumed.
 There are two distinct phases, in the first phase transactions attempt perform reads and write provisionally.
-If successful an atomic commitment protocol is invoked, else the transaction is aborted and tentative writes are removed.
+In the second phase, if successful an atomic commitment protocol is invoked, else the transaction is aborted and tentative writes are removed.
 
-In order to prevent all interleavings in Figure \ref{conf-scen} it is assumed there is a known $\Delta$, that reflects the time taken to perform both sub-operations in the first phase, with a probability $(1-\epsilon)$ that $\Delta$ is exceeded.
+In order to prevent all interleavings in Figure \ref{conf-scen} it is assumed there is a known $\Delta$, that reflects the time taken to perform both sub-operations at the ends of a distributed edge in the first phase.
 
 A distributed edge write sub-operation is allowed to proceed provided there is no other temporary write sub-operation preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$.
+Figure \ref{delta-abort} shows an instance of the protocol preventing Figure \ref{conf-scen}(c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$.
 An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
 A transaction that succeeds at all its tentative write attempts, decides to commit when it completes its all operations; the transaction commits once all its tentative writes are committed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is exceeded.
+This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
 
@@
 
-However it is possible that $\Delta$ is exceeded.
-If this happens all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
+If $\Delta$ is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
 Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
@@
 
-In short, the protocol attempts to avoid corruption by aborting transactions but cannot do so in all circumstances.
+In summary, the protocol attempts to avoid corruption by aborting transactions but cannot do so in all circumstances.
 Two questions naturally arise regards the performance of the protocol i) how does the protocol impact on the time until operational corruption?
 ii) how many transaction are aborted as a result of the protocol?
+
+% more details
 
@@
 
+Providing only reciprocal consistency is a weak guarantee yet vital to the integrity of a distributed graph database.
+
+RAMP is the closest thing.
+
+IPA/CRDT defining which edge wins in an application is not trivial
+
 TODO

### 2020-02-25 simulations/probabilistic/run.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+docker build -f Dockerfile -t abort-sim .
+
+for i in 1 5
+do
+for j in 1000 2000 3000 4000 5000 6000 7000 8000 9000 10000
+do docker run -v `pwd`/:/usr/local/src/scripts -d abort-sim java -cp "./target/uber-probabilistic-1.0-SNAPSHOT.jar" ProbabilisticProtocolSimulation $i $j
+done
+done

### 2020-02-25 simulations/probabilistic/src/main/java/ProbabilisticProtocolSimulation.java

@@ -6,97 +6,84 @@ import state.*;
 import utils.*;
 
 
-
-
-// txn arrives, updates 4 edges
-// for each edge
-// get source or destination
-// if d < write time then do write and set timer
-// else d aborts and returns
-
-// do not unset timer when aborted
-
-
 public class ProbabilisticProtocolSimulation {
 
     private final static Logger LOGGER = Logger.getLogger(ProbabilisticProtocolSimulation.class.getName());
 
     public static void main(String[] args) {
 
-        int[] tpsRange = {1000};
+        int delta = Integer.parseInt(args[0]);
+        SimulationConfiguration.getInstance().setDelta(delta);
 
-        String[] headers = {"tps","arrivals","collisions","commits","duration"};
+        int tps = Integer.parseInt(args[1]);
+        SimulationConfiguration.getInstance().setTPS(tps);
+
+        String[] headers = {"tps", "arrivals", "collisions", "commits", "duration"};
 
         WriteOutResults.writeOut(headers);
-        for (int j = 0; j<tpsRange.length; j++) {
 
-            //clear database
-            //clear event list
+        long startTime = System.currentTimeMillis();
+        int databaseSize = SimulationConfiguration.getInstance().getDatabaseSize();
 
-            long startTime = System.currentTimeMillis();
-
-            int databaseSize = SimulationConfiguration.getInstance().getDatabaseSize();
-
-            SimulationConfiguration.getInstance().setTPS(tpsRange[j]);
-
-            // initialise database
-            for (int i = 0; i < databaseSize; i++) {
-                DistributedEdge e = new DistributedEdge(i);
-                GlobalEdgeListSingleton.getInstance().addEdge(e);
-            }
-
-            // init. first arrival event
-            double initEventTime = SimulationRandom.getInstance().generateNextArrival();
-            GlobalEventListSingleton.getInstance().addEvent(new ArrivalEvent(initEventTime, EventType.ARRIVAL));
-
-            LOGGER.debug("**********************************************************************");
-            LOGGER.debug("**********************************************************************");
-            LOGGER.debug("Initial State");
-            LOGGER.debug("**********************************************************************");
-            LOGGER.debug("Event List: " + GlobalEventListSingleton.getInstance().getEventList());
-            LOGGER.debug("**********************************************************************");
-            LOGGER.debug("**********************************************************************");
[…]

### 2020-02-25 simulations/probabilistic/src/main/java/utils/SimulationConfiguration.java

@@ -47,10 +47,6 @@ public class SimulationConfiguration {
         return Integer.parseInt(simulationConfiguration.getProperty("database.size"));
     }
 
-    public int getTxnLimit(){
-        return Integer.parseInt(simulationConfiguration.getProperty("txn.limit"));
-    }
-
     long getSeedValue() {
         return Long.parseLong(simulationConfiguration.getProperty("seed.value"));
     }
@@ -71,9 +67,22 @@ public class SimulationConfiguration {
         return Double.parseDouble(simulationConfiguration.getProperty("common.ratio"));
     }
 
-    public double getDelta(){
+    double getDeltaInMS(){
         return Double.parseDouble(simulationConfiguration.getProperty("delta"));
     }
+
+    public double getDelta(){
+        return Double.parseDouble(simulationConfiguration.getProperty("delta"))/1000;
+    }
+
+    public void setDelta(double delta){
+        simulationConfiguration.setProperty("delta",String.valueOf(delta));
+    }
+
+    public int getTxnLimit(){
+        return Integer.parseInt(simulationConfiguration.getProperty("txn.limit"));
+    }
+
 
     public boolean saveResults() {
         return Boolean.parseBoolean(simulationConfiguration.getProperty("save.results"));

### 2020-02-25 simulations/probabilistic/src/main/java/utils/WriteOutResults.java

@@ -9,22 +9,6 @@ import java.io.IOException;
  */
 public class WriteOutResults {
 
-//    public static void fileHeader() {
-//
-//        String[] headers = {"tps","arrivals","collisions","commits","duration"};
-//
-//        StringBuilder headerStringBuilder = new StringBuilder();
-//        for(String header : headers){
-//            headerStringBuilder.append(header).append(",");
-//        }
-//
-//        String headerString = headerStringBuilder.toString();
-//        if( headerString.length() > 0 ) // remove trailing comma
-//            headerString = headerString.substring(0, headerString.length() - 1);
-//
-//        writeOut(headerString);
-//    }
-
     public static void writeOut(String[] string) {
 
         StringBuilder headerStringBuilder = new StringBuilder();
@@ -39,9 +23,10 @@ public class WriteOutResults {
 
         BufferedWriter outputStream = null;
         FileWriter fileWriter;
-        double delta = SimulationConfiguration.getInstance().getDelta();
+        int delta = (int) SimulationConfiguration.getInstance().getDeltaInMS();
+        int tps = SimulationConfiguration.getInstance().getTPS();
         try {
-            fileWriter = new FileWriter("./results/probabilistic/probabilistic_" + delta + ".csv",true);
+            fileWriter = new FileWriter("./delta_" + delta + "_" + tps + ".csv",true);
             outputStream = new BufferedWriter(fileWriter);
             outputStream.append(headerString);
             outputStream.append("\n");

### 2020-02-27 paper/main.tex
second draft

@@
 
-\usepackage[ruled,vlined]{algorithm2e}
 \usepackage{subcaption}
@@
 %% used to denote shared contribution to the research.
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
+\author{Jack Waudby}
+\email{j.waudby2@ncl.ac.uk}
 \affiliation{%
@@
 
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
+\author{Paul Ezhilchelvan}
+\email{paul.ezhilchelvan@ncl.ac.uk}
+\orcid{0000-0002-6190-5685}
 \affiliation{%
@@
 
+\author{Isi Mitrani}
+\email{isi.mitrani@ncl.ac.uk}
+\orcid{0000-0002-7797-7755}
+\affiliation{%
+  \institution{Newcastle University}
+  \city{Newcastle}
+  \country{UK}
+  \postcode{NE4 5TG}
+}
+
 \renewcommand{\shortauthors}{Ezhilchelvan, et al.}
@@
 \begin{abstract}
-  In this paper the notion of \textit{reciprocal consistency}, a important constraint specific to graph databases, is formalized.
+% In this paper the notion of \textit{reciprocal consistency}, a important constraint specific to graph databases, is formalized.
 If reciprocal consistency is not ensured during transaction processing, a distributed graph database can become operationally corrupt within a short time period relative to database lifetime.
 Reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., Snapshot Isolation) incurring well established performance costs.
+
+In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
+If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
+Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
 However, in practice distributed graph databases are often built atop BASE databases with few or no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
@@
 
-In practice graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node database.
-An approach to storing large graphs is to partition graph data over several machines in a cluster.
-Moving from a single-node database to one partitioned and distributed across servers introduces several interesting problems regarding data consistency.
-Graphs place a high demand on the underlying database and maintaining structural integrity, specifically across partitions, is non-trivial.
-
-Edges are subject to a constraint called \emph{reciprocal consistency}.
 In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
 This allows for bi-directional traversal and improved query performance.
-Preserving reciprocal consistency when these reciprocal pointers reside on different partitions requires some degree of coordination between partitions in the presence of concurrent modifications.
+An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
 
-In order to preserve reciprocal consistency, one could implement a heavyweight coordination protocol across partitions providing Serializability \cite{Bernstein1987}, which has associated performance costs.
-To avoid these costs databases provide application developers with the option of several weaker isolation models \cite{Berenson1995}.
-In addition, there are a number of systems that eschew transactional guarantees altogether, typically offering eventual consistency \cite{Bailis2013}, in order to achieve higher performance.
-Such systems are referred as  BASE (basically available, soft state, eventually consistent) databases \cite{Pritchett2008}.
+In practice graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node database and motivating the need for distributed graph databases.
+A common design pattern is to partition graph data over several machines in a cluster and use a BASE database \cite{Pritchett2008} for storage, adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
+A caveat with this approach is BASE databases in practice eschew transactional guarantees in order to achieve higher performance.
 
-A common design pattern for building a distributed graph database is to use a BASE database for storage, adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
 Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency can easily be violated in such a design introducing corruption into the database.
@@
 
-The contribution of this paper is the description of a protocol that ensures reciprocal consistency, preserving structural integrity in a distributed graph database in a transaction processing setting.
+Preserving reciprocal consistency requires some degree of coordination between partitions in the presence of concurrent modifications.
+This paper proposes and analyses a light-weight protocol that ensures reciprocal consistency in operational contexts where concurrency control mechanisms are done away with for sake of performance.
+
+%The contribution of this paper is the description of a protocol that ensures reciprocal consistency, preserving structural integrity in a distributed graph database in a transaction processing setting.
 The protocol maintains this guarantee provided certain condition hold.
@@
 
-The key finding is whilst the protocol reduces the chance of initial corruption of clean database records, due to the scale-free topology corruption propagates alarmingly quick even for modest transaction arrival rates.
+%The key finding is whilst the protocol reduces the chance of initial corruption of clean database records, due to the scale-free topology corruption propagates alarmingly quick even for modest transaction arrival rates.
 This motivates the use of deterministic reciprocal consistency protocols or the use of protocols with stronger isolation guarantees, which subsume reciprocal consistency (\cite{Bailis2014}, \cite{Berenson1995}, \cite{Bernstein1987}).
@@
 
-In the property graph data model edges have direction, there is a \emph{source} and a \emph{destination} vertex.
+In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
 In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
@@
 In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertices should be able to reify the edge correctly, returning consistent results.
-When the adjacency lists for vertices are mutually compatible like this, they are said to be \emph{reciprocally consistent}, a form of referential integrity.
+A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
+When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
 
@@
 Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered by landing at \emph{b} and then traversing to \emph{a}; even though the edge is a directed edge at model level abstraction.
+Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
 
@@
 Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is highly data dependent.
-However, the number of such edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
+The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
 
@@
 Maintaining reciprocal consistency for such distributed edges is challenging - especially given a common architecture employed by contemporary distributed graph databases.
-Often they use an existing BASE database (e.g.
-Apache Cassandra \cite{Cassandra}) to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
+Often they use an existing BASE database to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
 Superficially, opting for this design appears to be a good choice: the user has the modeling convenience of graphs with the operational characteristics from the underlying database.
 However, the problem with this design is the (lack of) transactional semantics are inherited from the underlying store.
-For example, Apache Cassandra provides no way of preventing updates from mutual interference in its normal multi-partition use case.
-Therefore, whilst maintaining reciprocal consistency for local edges is straightforward, this is not true for distributed edges.
-It is possible that concurrent updates can interleave in a manner that violates reciprocal consistency.
+Using the primitives provided by BASE databases maintaining reciprocal consistency for local edges is straightforward, this is not true for distributed edges.
+The lack of concurrency control across partitions means it is possible that concurrent updates can interleave in a manner that violates reciprocal consistency.
 
@@
 
-Earlier work investigated how weak isolation in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+Earlier work investigated how weak isolation across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
 
-When a given transaction writes a distributed edge it must perform two distinct sub-operations, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Concurrent transaction's sub-operations interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
+When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
+Concurrent transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
 For such edges there exists a correct and a incorrect entry.
-Note, the order in which sub-operations take place is not constrained.
+Note, the order in which the two writes take place is not constrained.
 For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
 A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-This is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-In a BASE distributed graph database there is no mechanism for repairing half-corrupt distributed edges.
-Structural corruption once occurred tends to persist\footnote{It is possible that subsequent modifications could correctly overwrite the corrupted information.}.
 
@@
 
-Assume, the correct ordering of transactions is $T_y$ then $T_x$.
-One could now imagine a read-write transaction, $T_z$ that introduces new information based on the incorrect information at vertex \emph{a}.
+Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
+The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an unknown author vertex.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
+\textcolor{grey}{// Tz}
+\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
+\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
+\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
+\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
+\end{Verbatim}
+%One could now imagine a read-write transaction, $T_z$ that introduces new information based on the incorrect information at vertex \emph{a}.
 
@@
 
-As stated earlier stronger isolation levels such as Serializability and Snapshot Isolation ensure reciprocal consistency.
-The motivation behind the \emph{Delta} protocol was to develop a lightweight protocol that maintains structural integrity of a distributed graph database at a cheaper cost then protocols with stronger guarantees.
+A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
+Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
+This is can be implemented by transactions taking long duration write locks\cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
+To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used \cite{Harding2017}.
+This approach is problematic in a distributed graph database as a subset of edges are traversed and modified with a high frequency e.g.
+critical sections of motorway in a road network, leading to high contention.
+Taking long duration locks on these edges significantly limits concurrency and throughput.
+
+The motivation behind the \emph{Delta} protocol was to develop a lightweight protocol that prevents half-corruption at a cheaper cost.
 One could imagine such a protocol be ``bolted onto'' a BASE distributed graph database.
+Note, this protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are left for future work, along with issues surrounding replication and atomicity.
 
-The \emph{Delta} protocol is designed to prevent the occurrence of dirty writes across partitions in order to maintain structural integrity of the graph.
-A client-server model is assumed, with transactions comprised of a graph traversal (set of read operations) followed by a series of writes - for simplicity a single write is assumed.
-There are two distinct phases, in the first phase transactions attempt perform reads and write provisionally.
-In the second phase, if successful an atomic commitment protocol is invoked, else the transaction is aborted and tentative writes are removed.
-
-In order to prevent all interleavings in Figure \ref{conf-scen} it is assumed there is a known $\Delta$, that reflects the time taken to perform both sub-operations at the ends of a distributed edge in the first phase.
-
-A distributed edge write sub-operation is allowed to proceed provided there is no other temporary write sub-operation preceding it within $\Delta$; measured as per the local clock time.
+We leverage the fact that a transaction that writes at one end of a distributed edge, must then write the other.
+We then assumed that the network delay between two servers can be predicted to be some $\Delta$.
+All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
+From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
 Figure \ref{delta-abort} shows an instance of the protocol preventing Figure \ref{conf-scen}(c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$.
 An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-A transaction that succeeds at all its tentative write attempts, decides to commit when it completes its all operations; the transaction commits once all its tentative writes are committed.
 This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
+It is helpful to think of this protocol as a locking protocol with a timer, in the extreme when $\Delta$ is set to infinity the protocol is equivalent to long-duration write locks with \texttt{NO_WAIT} deadlock detection.
 
@@
 In summary, the protocol attempts to avoid corruption by aborting transactions but cannot do so in all circumstances.
-Two questions naturally arise regards the performance of the protocol i) how does the protocol impact on the time until operational corruption?
-ii) how many transaction are aborted as a result of the protocol?
-
-% more details
+Two questions naturally arise regards the performance of the protocol for different values of  $\Delta$, i) how does the protocol impact on the time until operational corruption?
+ii) how many transactions are aborted as a result of the protocol?
 
@@
 The \emph{Delta} protocol influences the rate of corruption reducing the probability a transaction corrupts an edge, $q^{new}_i$.
-To simplify the derivation of the new conflict probability it was assumed.
+To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order, Figure 3(c).
 This leaves the Figure 6 as the only source of corruption.
@@
 $$ q^{new}_i = P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$
-The arrival times of $T_x,T_y$ are assumed exponentially distributed , $X \sim \exp (\frac{\lambda P_i}{2N_i})$.
+The arrival times of $T_x,T_y$ are assumed exponentially distributed , $X \sim \exp (\rho)$, where $\rho = \frac{\lambda P_i}{2N_i}$, the probability of accessing a given end edge of type $i$.
 The transmission times $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
@@
 Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-To choose $\Delta$ one calculates $P(M>\Delta) = 1 - \epsilon$.
 \end{itemize}
 
-TODO: Method for working out number of aborts ii)
+In order to answer (ii), calculating the number of aborts per second for a given $\Delta$ an discrete event-based simulation was constructed.
+The simulation, focuses on the subset of edges with the highest access probability.
 
@@
 The simulation is performed on a Scale-Free graph, such as a social network, human brain, or road network, large enough to make processing non-trivial.
-There are approximately 77 million local edges, approximately 33 million distributed edges (in proportion to good graph partitioning algorithms); a graph of this size would have approximately 10 million nodes.
+There are approximately 7.7 billion local edges, approximately 3.3 million distributed edges (in proportion to good graph partitioning algorithms); a graph of this size would have approximately 1 billion nodes.
 
-The graph consisted of five edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7, N_5=10^8$ with access probabilities $p_1 =0.5, p_2 =0.26, p_3=0.13, p_4=0.07$ and $p_5 =0.04$.
-The number of read operations per query is distributed geometrically starting at $2$, with an average of $15$.
+The graph consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^10$ with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$.
+The number of read operations per query is distributed geometrically starting at $2$, with an average of $15$, before a write.
 In all edge types, a fraction $0.3$ are distributed, the remainder are local.
+The network delay between servers is exponential distributed with a rate of $200ms$.
 The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken under operational corruption $U$, is measured in hours.
-$U$ considered for a range of transaction arrival rates, $\lambda = (100, ..., 2000)$.
+The time taken until operational corruption $U$, is measured in hours.
+$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects the range $\lambda = (10000, ..., 100000)$.
+We considered 3 values of $\Delta = 1, 5, 10ms$.
 
@@
   \includegraphics[width=\linewidth]{./images/results}
-  \caption{Corruption results}
+  \caption{Time until operational corruption under delta protocol}
   \Description{Graph representation.}
-  \label{results}
+  \label{time-to-corruption-results}
 \end{figure}
 
-TODO: add aborts/s \\
-TODO: discuss results
+The results for measuring the impact of time until operational corruption are given in Figure \ref{time-to-corruption-results}.
+%The $\Delta$ has a significant on increasing the time to operational corruption for modest transaction arrival rates ($<1000$).
 
-\section{Related Work}
+The abort rates for $\Delta = 1, 5, 10ms$ is given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
+The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
 
-Providing only reciprocal consistency is a weak guarantee yet vital to the integrity of a distributed graph database.
-
-RAMP is the closest thing.
-
-IPA/CRDT defining which edge wins in an application is not trivial
-
-TODO
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{./images/aborts}
+  \caption{Aborts per seconds under protocol}
+  \Description{Graph representation.}
+  \label{aborts-results}
+\end{figure}
 
@@
 
-TODO
+In this paper a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database has been presented.
+The $\Delta$ protocol leverages the fact writes to edges always consists of two sequential writes.
+The result is guarantees much weaker than even Read Uncommitted isolation, the weakest ANSI isolation level.
+However, a mechanism to provide solely reciprocal consistency is believed to be valuable in practice.
+Concurrency control has been a long researched area, however to the best of our knowledge no other work has presented protocols bespoke to maintaining distributed graph databases.
 
+TODO: More details regards protocol performance.
 \section{Acknowledgments}
@@
 If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( M_x >  \Delta + T_y) \cap (M_y > \Delta - T_x) \right]$$.
+Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
 
 The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho = \beta = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
+Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
 The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
@@
 \begin{dmath*}
-  q_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
+  q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
@@
 
-To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-\begin{align*}
-  P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-  1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-  e^{- \delta d} & = \epsilon  \\
-  \Delta & = - \frac{\ln(\epsilon)}{\delta}
-\end{align*}
+% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
+% \begin{align*}
+%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
+%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
+%   e^{- \delta d} & = \epsilon  \\
+%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
+% \end{align*}
 
-Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
+% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
 For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$

### 2020-02-27 make-plots.R

@@ -1,0 +1,83 @@
+library(ggplot2)
+library(latex2exp)
+library(ggthemes)
+
+# aborts/s 
+
+ms1 = read.csv("./results/aborts_delta_1.csv")
+ms1$collisions = (ms1$collisions/10)
+delta = rep(1,10)
+ms1 = cbind(ms1,delta)
+ms1 = rbind(c(100,100,0,0,0,1),ms1)
+
+ms5 = read.csv("./results/aborts_delta_5.csv")
+ms5$collisions = (ms5$collisions/10)
+delta = rep(5,10)
+ms5 = cbind(ms5,delta)
+ms5= rbind(c(100,100,0,0,0,5),ms5)
+
+
+ms10 = read.csv("./results/aborts_delta_10.csv")
+ms10$collisions = (ms10$collisions/10)
+delta = rep(10,10)
+ms10 = cbind(ms10,delta)
+ms10= rbind(c(100,100,0.01,0,0,10),ms10)
+
+
+aborts = rbind(ms1,ms5,ms10)
+aborts$delta = as.factor(aborts$delta)
+
+a=ggplot(data=aborts, aes(x=tps,y=collisions,linetype=delta,color=delta)) +
+    geom_line() +
+     scale_linetype_discrete(name="Delta (ms): ",labels=unname(TeX(c("1","5","10")))) +
+     scale_color_discrete(name="Delta (ms): ",labels=unname(TeX(c("1","5","10")))) +
+     scale_x_continuous(name="TPS",breaks=seq(0,10000,1000)) +
+     scale_y_continuous(name="Aborts/s",breaks=seq(0,150,25)) + 
+    theme_few()
+a
+
+
+
+# load time to corruption data in 
+no_isolation = read.csv("./results/no_isolation.csv")
+no_isolation = rbind(c(100,4991.678),no_isolation)
+type = rep(4,11)
+no_isolation = cbind(no_isolation,type)
+
+d1 = read.csv("./results/delta_1.csv")
+type = rep(1,11)
+d1 = cbind(d1,type)
+
+d5 = read.csv("./results/delta_5.csv")
+type = rep(2,11)
+d5 = cbind(d5,type)
+
+d10 = read.csv("./results/delta_10.csv")
+type = rep(3,11)
+d10 = cbind(d10,type)
+
+
+df = rbind(no_isolation,d1,d5,d10)
+
+df$type = as.factor(df$type)
+
+
[…]

### 2020-02-27 simulations/corruption/func_delta.R

@@ -45,3 +45,53 @@ probabilistic <- function(N,p,f,lam,r,gam,D,del){
   t = t/(3600*24)
   return(t)
 }
+
+prcp <- function(N,p,f,lam,r,gam,eps,del){
+  t=0
+  n=matrix(0,ncol=3,nrow=length(N))
+  tr=matrix(0,ncol=5,nrow=length(N))
+  Nt=sum(N)
+  n[,1]=(1-f)*N
+  n[,2]=f*N
+  d=-log(eps)/del
+  pn=(lam*p)/(N*2)
+  q=rep(0,length(N))
+  for(i in 1:length(N)){
+    q[i]=eps*eps*(1-(del/(pn[i]+del))*exp(-pn[i]*d))
+  }
+  while(sum(n)>Nt*(1-gam)){
+    a=rep(0,length(N))
+    for(i in 1:length(N)){
+      a[i]=(p[i]/N[i])*(n[i,1]+n[i,2]+0.5*n[i,3])
+    }
+    a=sum(a)
+    b=a^2*r/(1-a+a*r)
+    for(i in 1:length(N)){
+      tr[i,1]=((lam*p[i]*n[i,2])/N[i])*q[i]*b*b
+      tr[i,2]=((lam*p[i]*n[i,3])/N[i])*(1-q[i])*b
+      tr[i,3]=((lam*p[i]*n[i,3])/N[i])*(1-b)
+      tr[i,4]=((lam*p[i]*n[i,2])/N[i])*(1-b)
+      tr[i,5]=((lam*p[i]*n[i,1])/N[i])*(1-b)
+    }
+    tt=sum(tr)
+    t = t + 1/tt
+    i = sample(seq(1,length(N)), size = 1, replace = TRUE, prob = rowSums(tr)/tt)
+    j = sample(c(1, 2, 3, 4, 5), size = 1, replace = TRUE, prob = tr[i,]/sum(tr[i,]))
+    if(j==1){
+      n[i,2] = n[i,2] - 1
+      n[i,3] = n[i,3] + 1
+    } else if(j==2) {
+      n[i,2] = n[i,2] + 1
+      n[i,3] = n[i,3] - 1
+    } else if(j==3) {
+      n[i,3] = n[i,3] - 1
+    } else if(j==4) {
+      n[i,2] = n[i,2] - 1
+    } else {
+      n[i,1] = n[i,1] - 1
+    }
+  }
+  t = t/(3600*24)
+  return(t)
+}
+

### 2020-02-27 simulations/corruption/make-plots.R

@@ -1,46 +1,0 @@
-library(ggplot2)
-library(latex2exp)
-library(ggthemes)
-
-# load time to corruption data in 
-no_isolation = read.csv("../../results/corruption/no_isolation.csv")
-type = rep(1,14)
-no_isolation = cbind(no_isolation,type)
-
-e01 = read.csv("../../results/corruption/probabilistic_0_1.csv")
-type = rep(2,14)
-e01 = cbind(e01,type)
-
-e005 = read.csv("../../results/corruption/probabilistic_0_05.csv")
-type = rep(3,14)
-e005 = cbind(e005,type)
-
-e001 = read.csv("../../results/corruption/probabilistic_0_01.csv")
-type = rep(4,14)
-e001 = cbind(e001,type)
-
-
-df = rbind(no_isolation,e005, e01,e001)
-
-df$type = as.factor(df$type)
-
-
-png("../../paper/images/corruption_comparison.png")
-
-
-
-p=ggplot(data=df, aes(x=lam,y=U,linetype=type,color=type)) +
-    geom_line() +
-    scale_linetype_discrete(name="Eps: ",labels=unname(TeX(c("NA","0.1","0.05","0.01")))) +
-    scale_color_discrete(name="Eps: ",labels=unname(TeX(c("NA","0.1","0.05","0.01")))) +
-    scale_x_continuous(name="TPS",breaks=seq(0,2000,200),limits=c(0,2000)) +
-    scale_y_continuous(name="U",breaks=seq(0,5000,500)) + 
-    theme_few()
-
-p
-
-
-dev.off()
-
-
-

### 2020-03-20 paper/main.tex
final draft

@@
 
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
+%
+% Submission ID.
+% Use this when submitting an article to a sponsored event. You'll
+% receive a unique submission ID from the organizers
+% of the event, and this ID should be used as the parameter to this command.
+%\acmSubmissionID{123-A56-BU3}
 
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
 \title{Preserving Reciprocal Consistency in Distributed Graph Databases}
 
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
 \author{Jack Waudby}
@@
 
-\renewcommand{\shortauthors}{Ezhilchelvan, et al.}
+\renewcommand{\shortauthors}{Waudby, et al.}
 
 \begin{abstract}
-% In this paper the notion of \textit{reciprocal consistency}, a important constraint specific to graph databases, is formalized.
-If reciprocal consistency is not ensured during transaction processing, a distributed graph database can become operationally corrupt within a short time period relative to database lifetime.
-Reciprocal consistency can be maintained as a part of enforcing any of several known strong isolation guarantees (e.g., Snapshot Isolation) incurring well established performance costs.
 
@@
 Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with few or no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
+However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
+A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
 Protocol performance is evaluated through simulations.
@@
 \ccsdesc[300]{Data Management~Reciprocal Consistency}
+\ccsdesc[300]{Data Management~Concurrency Control}
 
-% \ccsdesc[500]{Computer systems organization~Embedded systems}
-% \ccsdesc[300]{Computer systems organization~Redundancy}
-% \ccsdesc{Computer systems organization~Robotics}
-% \ccsdesc[100]{Networks~Network reliability}
-
-\keywords{Graph Databases, Reciprocal Consistency}
+\keywords{Graph Databases, Reciprocal Consistency, BASE}
 
@@
 In addition, properties can be stored on both vertices and edges.
-
 In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance.
+This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
 An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
 
-In practice graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node database and motivating the need for distributed graph databases.
-A common design pattern is to partition graph data over several machines in a cluster and use a BASE database \cite{Pritchett2008} for storage, adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases in practice eschew transactional guarantees in order to achieve higher performance.
+In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
+A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
+Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
+A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
 
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency can easily be violated in such a design introducing corruption into the database.
+Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
 Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-
-Preserving reciprocal consistency requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a light-weight protocol that ensures reciprocal consistency in operational contexts where concurrency control mechanisms are done away with for sake of performance.
-
-%The contribution of this paper is the description of a protocol that ensures reciprocal consistency, preserving structural integrity in a distributed graph database in a transaction processing setting.
-The protocol maintains this guarantee provided certain condition hold.
-If these conditions are violated corruption can still occur.
-The protocol's impact on the rate of database corruption and its performance is evaluated and discussed.
-
-%The key finding is whilst the protocol reduces the chance of initial corruption of clean database records, due to the scale-free topology corruption propagates alarmingly quick even for modest transaction arrival rates.
-This motivates the use of deterministic reciprocal consistency protocols or the use of protocols with stronger isolation guarantees, which subsume reciprocal consistency (\cite{Bailis2014}, \cite{Berenson1995}, \cite{Bernstein1987}).
+Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
+This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
 
@@
 In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This facilitates bi-directional edge traversal and allows for better query performance.
+This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
 
@@
 Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for such distributed edges is challenging - especially given a common architecture employed by contemporary distributed graph databases.
-Often they use an existing BASE database to store data, which has been adapted with a programmatic API or query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
-Superficially, opting for this design appears to be a good choice: the user has the modeling convenience of graphs with the operational characteristics from the underlying database.
-However, the problem with this design is the (lack of) transactional semantics are inherited from the underlying store.
-Using the primitives provided by BASE databases maintaining reciprocal consistency for local edges is straightforward, this is not true for distributed edges.
-The lack of concurrency control across partitions means it is possible that concurrent updates can interleave in a manner that violates reciprocal consistency.
+Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
+Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
+Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
 
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
+However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
+BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
+This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
+Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+This process is explained in Section \ref{sec:db-corruption}.
 
@@
 
-Earlier work investigated how weak isolation across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+\section{Corruption in BASE  Distributed Graph Databases}
+\label{sec:db-corruption}
 
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal information at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Concurrent transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
+When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
+Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
 For such edges there exists a correct and a incorrect entry.
@@
 A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$. $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
+To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
+$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
@@
 Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
+Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
+The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
+Further corruption spreads by the same mechanism.
 
-\begin{figure}[H]
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
+\textcolor{grey}{// Tz}
+\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
+\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
+\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
+\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
+\end{Verbatim}
+
+\begin{figure}[ht!]
   \centering
@@
 
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (2.4,0);
+    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
 
   \end{tikzpicture}
-  \caption{A half-corrupted edge resulting from conflicting transactions.}
+  \caption{A half-corrupted edge.}
   \label{half-corrupted}
 \end{figure}
-
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an unknown author vertex.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-%One could now imagine a read-write transaction, $T_z$ that introduces new information based on the incorrect information at vertex \emph{a}.
 
@@
 Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-This is can be implemented by transactions taking long duration write locks\cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used \cite{Harding2017}.
-This approach is problematic in a distributed graph database as a subset of edges are traversed and modified with a high frequency e.g.
+
+A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
+To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
+
+This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
 critical sections of motorway in a road network, leading to high contention.
-Taking long duration locks on these edges significantly limits concurrency and throughput.
+In addition, graph transactions tend to be longer-lived than transactions in relational databases.
+Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
+With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
+We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
 
-The motivation behind the \emph{Delta} protocol was to develop a lightweight protocol that prevents half-corruption at a cheaper cost.
-One could imagine such a protocol be ``bolted onto'' a BASE distributed graph database.
-Note, this protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are left for future work, along with issues surrounding replication and atomicity.
-
-We leverage the fact that a transaction that writes at one end of a distributed edge, must then write the other.
-We then assumed that the network delay between two servers can be predicted to be some $\Delta$.
+The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
+It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
 All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
 From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing Figure \ref{conf-scen}(c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$.
+Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
 An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
 This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-It is helpful to think of this protocol as a locking protocol with a timer, in the extreme when $\Delta$ is set to infinity the protocol is equivalent to long-duration write locks with \texttt{NO_WAIT} deadlock detection.
 
+This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
+The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
+However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
+
+% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
+
+%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
+%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
+
+%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
 \begin{figure}[H]
@@
 
-If $\Delta$ is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
+Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
 Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
+The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
 
@@
 
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.5) node[midway,right] {$> \Delta$};
+    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
+    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
   \end{tikzpicture}
@@
 
-In summary, the protocol attempts to avoid corruption by aborting transactions but cannot do so in all circumstances.
-Two questions naturally arise regards the performance of the protocol for different values of  $\Delta$, i) how does the protocol impact on the time until operational corruption?
-ii) how many transactions are aborted as a result of the protocol?
+In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
+Two questions naturally arise regards the performance of the protocol.
+For different values of $\Delta$:
+\begin{itemize}
+\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
+\item How many transactions are aborted as a result of the protocol?
+\end{itemize}
 
@@
 
-In order to answer i) the model developed in \cite{Ezhilchelvan2018} and fine-tuned in \cite{Webber2019} was extended to measure the impact of the \emph{Delta} protocol\footnote{An empirical evaluation of existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
+To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
 A summary of the model is now provided, before discussing the extensions.
 
-Transactions arrive in a Poisson stream with rate $\lambda$ per second.
-Each transaction contains a random number of read operations, $K$, followed by a single write.
-Edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
+The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
+To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
+To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
 For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-
+The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
 At any moment in time an edge can be in one of four states:
@@
 \item Distributed and clean.
-\item Half-corrupted distributed edge arising from interleaved updates.
+\item Half-corrupted distributed edge arising from a dirty write.
 \item Semantically corrupted.
@@
 Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the average time to corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption).
-When a certain fraction, $\gamma$, of all edges become corrupted, the database itself is said to be operationally corrupt.
-Note, the model assumes crash-free hardware and bug-free software for simplicity.
+These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
+At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
 The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
 
-The \emph{Delta} protocol influences the rate of corruption reducing the probability a transaction corrupts an edge, $q^{new}_i$.
-To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order, Figure 3(c).
-This leaves the Figure 6 as the only source of corruption.
-From this the following probability can be formulated:
-$$ q^{new}_i = P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$
-The arrival times of $T_x,T_y$ are assumed exponentially distributed , $X \sim \exp (\rho)$, where $\rho = \frac{\lambda P_i}{2N_i}$, the probability of accessing a given end edge of type $i$.
-The transmission times $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-The complete derivation of $q^{new}_i$ is provided in Appendix A.
+% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
+the interleaving in Figure 3(c) cannot occur.}.
 
+The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
 Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-The answer depends on several parameters characterizing four systemic aspects:
+Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
+The answers to these questions depends on several parameters characterizing the following systemic aspects:
 \begin{itemize}
-\item \emph{Size and topology of graph database}.
-Size is expressed by the total number of edges $N$, and the fraction $f$ of edges distributed across servers.
-We consider a common edge access patterns or topologies: a Scale-free topology, edges have different access probabilities and those that get accessed more frequently tend to be smaller in number.
+\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
 \item \emph{Workload}.
@@
 
-In order to answer (ii), calculating the number of aborts per second for a given $\Delta$ an discrete event-based simulation was constructed.
-The simulation, focuses on the subset of edges with the highest access probability.
-
 \section{Evaluation}
@@
 
-The model assumes that the distributed graph database  processes many reads-followed-by-write graph queries concurrently, processing both local and distributed edges.
-The simulation is performed on a Scale-Free graph, such as a social network, human brain, or road network, large enough to make processing non-trivial.
-There are approximately 7.7 billion local edges, approximately 3.3 million distributed edges (in proportion to good graph partitioning algorithms); a graph of this size would have approximately 1 billion nodes.
-
-The graph consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^10$ with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$.
-The number of read operations per query is distributed geometrically starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local.
+The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
+The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
+In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
 The network delay between servers is exponential distributed with a rate of $200ms$.
 The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in hours.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects the range $\lambda = (10000, ..., 100000)$.
-We considered 3 values of $\Delta = 1, 5, 10ms$.
+The time taken until operational corruption $U$, is measured in days.
+$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
+The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
+
+The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
+Under no isolation, $U$ ranges between 500-50 days.
+For $50ms$ this increases to 74-1 years.
+For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/results}
-  \caption{Time until operational corruption under delta protocol}
+  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
+  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
   \Description{Graph representation.}
@@
 
-The results for measuring the impact of time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-%The $\Delta$ has a significant on increasing the time to operational corruption for modest transaction arrival rates ($<1000$).
-
-The abort rates for $\Delta = 1, 5, 10ms$ is given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
+The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
 The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
+For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/aborts}
-  \caption{Aborts per seconds under protocol}
+  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
+  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
   \Description{Graph representation.}
@@
 
-In this paper a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database has been presented.
-The $\Delta$ protocol leverages the fact writes to edges always consists of two sequential writes.
-The result is guarantees much weaker than even Read Uncommitted isolation, the weakest ANSI isolation level.
-However, a mechanism to provide solely reciprocal consistency is believed to be valuable in practice.
-Concurrency control has been a long researched area, however to the best of our knowledge no other work has presented protocols bespoke to maintaining distributed graph databases.
+Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
+We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
+The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
+The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
+However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
+Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
+For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
+Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
 
-TODO: More details regards protocol performance.
-\section{Acknowledgments}
+% \appendix
+% \section{Appendix}
 
-TODO
+% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
 
-\appendix
-\section{Appendix}
-
-Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
+% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
 Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
@@
 
-The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
+% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
 Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
@@
 
-% Therefore,  $$ g(y) = \delta e^{-\delta y} $$
+% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
 
-Therefore,
-\begin{dmath*}
-  q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-  =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-  =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-\end{dmath*}
+% Therefore,
+% \begin{dmath*}
+%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
+%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
+%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
+% \end{dmath*}
 

### 2020-03-20 make-plots.R

@@ -1,83 +1,0 @@
-library(ggplot2)
-library(latex2exp)
-library(ggthemes)
-
-# aborts/s 
-
-ms1 = read.csv("./results/aborts_delta_1.csv")
-ms1$collisions = (ms1$collisions/10)
-delta = rep(1,10)
-ms1 = cbind(ms1,delta)
-ms1 = rbind(c(100,100,0,0,0,1),ms1)
-
-ms5 = read.csv("./results/aborts_delta_5.csv")
-ms5$collisions = (ms5$collisions/10)
-delta = rep(5,10)
-ms5 = cbind(ms5,delta)
-ms5= rbind(c(100,100,0,0,0,5),ms5)
-
-
-ms10 = read.csv("./results/aborts_delta_10.csv")
-ms10$collisions = (ms10$collisions/10)
-delta = rep(10,10)
-ms10 = cbind(ms10,delta)
-ms10= rbind(c(100,100,0.01,0,0,10),ms10)
-
-
-aborts = rbind(ms1,ms5,ms10)
-aborts$delta = as.factor(aborts$delta)
-
-a=ggplot(data=aborts, aes(x=tps,y=collisions,linetype=delta,color=delta)) +
-    geom_line() +
-     scale_linetype_discrete(name="Delta (ms): ",labels=unname(TeX(c("1","5","10")))) +
-     scale_color_discrete(name="Delta (ms): ",labels=unname(TeX(c("1","5","10")))) +
-     scale_x_continuous(name="TPS",breaks=seq(0,10000,1000)) +
-     scale_y_continuous(name="Aborts/s",breaks=seq(0,150,25)) + 
-    theme_few()
-a
-
-
-
-# load time to corruption data in 
-no_isolation = read.csv("./results/no_isolation.csv")
-no_isolation = rbind(c(100,4991.678),no_isolation)
-type = rep(4,11)
-no_isolation = cbind(no_isolation,type)
-
-d1 = read.csv("./results/delta_1.csv")
-type = rep(1,11)
-d1 = cbind(d1,type)
-
-d5 = read.csv("./results/delta_5.csv")
-type = rep(2,11)
-d5 = cbind(d5,type)
-
-d10 = read.csv("./results/delta_10.csv")
-type = rep(3,11)
-d10 = cbind(d10,type)
-
-
-df = rbind(no_isolation,d1,d5,d10)
-
-df$type = as.factor(df$type)
-
-
[…]

### 2020-03-20 scripts/make-plots.R

@@ -1,0 +1,87 @@
+library(ggplot2)
+library(latex2exp)
+library(ggthemes)
+
+# aborts/s 
+ms50 = read.csv("./results/aborts_50.csv")
+ms50$collisions = (ms50$collisions/10)
+delta = rep(50,11)
+ms50 = cbind(ms50,delta)
+
+
+ms75 = read.csv("./results/aborts_75.csv")
+ms75$collisions = (ms75$collisions/10)
+delta = rep(75,11)
+ms75 = cbind(ms75,delta)
+
+
+ms100 = read.csv("./results/aborts_100.csv")
+ms100$collisions = (ms100$collisions/10)
+delta = rep(100,11)
+ms100 = cbind(ms100,delta)
+
+max(ms100$collisions)
+
+aborts = rbind(ms50,ms75,ms100)
+aborts$delta = as.factor(aborts$delta)
+
+a=ggplot(data=aborts, aes(x=tps,y=collisions,linetype=delta,color=delta)) +
+    geom_line() +
+     scale_linetype_discrete(name="Delta (ms): ",labels=unname(TeX(c("50","75","100")))) +
+     scale_color_discrete(name="Delta (ms): ",labels=unname(TeX(c("50","75","100")))) +
+     scale_x_continuous(name="TPS",breaks=seq(0,10000,1000)) +
+     scale_y_continuous(name="Aborts/s",breaks=seq(0,1000,100)) + 
+    theme_few()
+a
+
+
+
+# load time to corruption data in 
+no_isolation = read.csv("./results/no_isolation.csv")
+#no_isolation = rbind(c(100,no_isolation$U[1]*1.99),no_isolation)
+type = rep(1,10)
+no_isolation = cbind(no_isolation,type)
+
+ 
+d1 = read.csv("./results/delta_50.csv")
+d1 = d1[-1,]
+type = rep(2,5)
+d1 = cbind(d1,type)
+
+d5 = read.csv("./results/delta_75.csv")
+d5 = d5[-1,]
+type = rep(3,5)
+d5 = cbind(d5,type)
+
+d10 = read.csv("./results/delta_100.csv")
+d10 = d10[-1,]
+type = rep(4,5)
+d10 = cbind(d10,type)
+
+
+
+df = rbind(no_isolation,d1[,c(1,2,4)],d5[,c(1,2,4)],d10[,c(1,2,4)])
+
+df$U = log(df$U)
+
+df$type = as.factor(df$type)
+
+
[…]

### 2020-04-05 paper/main.tex
fix up

@@
 
-\usepackage{subcaption}
+\usepackage{xspace}
+\usepackage[labelformat=simple]{subcaption}
+\renewcommand\thesubfigure{(\alph{subfigure})}
 \usepackage{tikz}
@@
 \usepackage{breqn}
+
+\newcommand{\tDelta}{\textsf{Delta}\xspace}
 
@@
 This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
+An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (details in Section \ref{sec:recipr-cons}).
 
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
+In practice, graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
 A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
+Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database.
+Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
 
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
+When, for example, a BASE database \cite{Pritchett2008} is adapted with a graph processing layer, then violations of reciprocal consistency will occur if that adaptation provides no concurrency control for operations that span partitions in order to offer higher performance.
+This paper proposes a simple concurrency control protocol, called \tDelta protocol, that does not impede performance adversely.
+That is because the protocol is exclusively designed for one purpose only: reciprocal consistency in distributed edges.
+Its design leverages the fact that the updating of end pointers of a given distributed edge immediately follow each other and the small interval between them is the window for possible conflicts between concurrent updates.
+The protocol ensures that any two updates on any given end pointer of any given edge are at least $\Delta$ time apart by aborting any write attempt that is within .
+It ensures reciprocal consistency when $\Delta$ is chosen to be larger than the length of the conflict window.
+
+The paper is organized as follows: Section \ref{sec:recipr-cons} presents reciprocal consistency, Sections \ref{sec:distr-graph-datab}-\ref{sec:db-corruption} discuss the architecture of distributed graph databases and describes how corruption occurs.
+Section \ref{sec:tdelta-protocol} outlines the \tDelta protocol.
+Sections \ref{sec:modeling}-\ref{sec:evaluation} presents the model used to evaluate protocol performance.
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
+  \caption{Database records representing vertices, containing properties and a sequence of edges that comprise the adjacency list of a vertex \cite{janusgraph}.}
+  \label{adj-list}
+\end{figure}
 
@@
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
+A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
 Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
@@
 
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
+Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
 Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
+When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
+A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.
+
+\begin{figure}[ht!]
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
+
+    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
+
+    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
+        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
+
+    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (4.2,0);
+  \end{tikzpicture}
+  \caption{A reciprocally consistent edge.}
+  \label{reciprocally-consistent-edge}
+\end{figure}
 
 \section{Distributed Graph Databases}
+\label{sec:distr-graph-datab}
 
@@
 This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
 This process is explained in Section \ref{sec:db-corruption}.
@@
     \caption{}
+    \label{fig:tx-ty}
   \end{subfigure}%
@@
     \caption{}
+\label{fig:ty-tx}
   \end{subfigure}%
@@
     \caption{}
+    \label{fig:same-side}
   \end{subfigure}%
   \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
+In \subref{fig:tx-ty} $T_y$ begins writing to the distributed edge before $T_x$, in \subref{fig:ty-tx} the converse is true, else they are equivalent.
+In \subref{fig:same-side} both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
   \label{conf-scen}
@@
 
-\section{Corruption in BASE  Distributed Graph Databases}
+\begin{figure}[H]
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
+
+    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
+
+    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
+        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
+
+    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
+  \end{tikzpicture}
+  \caption{A half-corrupted edge.}
+  \label{half-corrupted}
+\end{figure}
+
+\section{Corruption in the Absence of Concurrency Control}
 \label{sec:db-corruption}
@@
 When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
+Without concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated.
+Possible interleavings of concurrent writes to a distributed edge are given in Figure \ref{conf-scen}.
+A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
+Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
+
+For such half-corrupted edges there exists a correct and a incorrect entry.
 Note, the order in which the two writes take place is not constrained.
@@
 \textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
+\textcolor{blue}{AND} u.\textcolor{red}{name} = 'unknown'
 \textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
@@
 
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
+\section{\tDelta Protocol}
+\label{sec:tdelta-protocol}
 
@@
 
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
+Our \tDelta protocol employs principles behind all these well-tested strategies but has two crucial differences:
+\begin{itemize}
+\item No locks are used.
+\item A write operation need not await until the preceding write commits but can proceed if at least $\Delta$ duration (measured in local clock) has elapsed.
+\end{itemize}
+
+These differences lead to several advantages in the context of graph databases.
+Firstly, a subset of edges are traversed and modified with a high frequency e.g.
 critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
+Secondly, graph transactions tend to be longer-lived than transactions in relational databases.
+Waiting for earlier writes on edges by long running transactions to commit, these frequently accessed edges significantly limit concurrency and reduce throughput.
+With these concerns in mind we developed the \tDelta protocol, which aimed at preventing edges becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
 
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
+\subsection{Protocol Description}
+\label{sec:protocol-description}
+
+The \tDelta Protocol has five rules:
+\begin{enumerate}
+\item A transaction's update on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
+\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative update was done at least $\Delta$ time before (where time is measured as per local clock).
+\item If a transaction successfully performs all its tentative updates, then it is permitted to commit; otherwise, it must abort.
+\item A transaction commits, when all its tentative updates are made permanent, e.g., using an atomic commitment protocol.
+\item Tentative updates of an aborting transactions are ignored.
+An ignored tentative update can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
+\end{enumerate}
+
+\subsection{Correctness Reasoning}
+\label{sec:corr-reas}
+
+Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction updating one end of a distributed edge at one server and updating the other end of the same edge at another server.
+
+Let $\Delta$ be chosen such that $\Delta > \delta$.
+Now consider the interleaving in Figure \ref{fig:tx-ty} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
+Note $t_y < t_x$, i.e.
+$(t_x - t_y) > 0$.
+
+Say, $T_y$ reaches $S_i$ at time $t_y + d$, where $d$ is the actual time elapsed between completing a tentative write at one end and starting at another end, let us assume that $d \leq \delta$.
+When $T_y$ arrives at $S_i$ it will find a tentative write done at time $t_x$.
+In this case, $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge.
+A similar arguments can be made for the scenario in with Figure \ref{fig:ty-tx}.
+For the interleaving in Figure \ref{fig:same-side}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$.
+
+Assume that $d > \Delta$, i.e., the estimate $\delta$ does not hold at this moment.
+In Figure \ref{fig:tx-ty} interfering updates are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$ otherwise, an edge becomes half-corrupted.
+This holds for the interleaving in \ref{fig:ty-tx}.
+Thus, in the extreme case $t_x = t_y$, reciprocal consistency is not guaranteed if $d > \Delta$ when the former exceeds its upper bound estimate $\delta$.
+For Figure \ref{fig:same-side}, if $d$ for $T_x$ is larger than $\Delta$, reciprocal consistency can be violated.
+
+This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
+The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
+
+In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur.
+Two questions naturally arise regards the performance of the protocol.
+For different values of $\Delta$:
+\begin{itemize}
+\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
+\item How many transactions are aborted as a result of the protocol?
+\end{itemize}
+
+%The two paragraphs (The protocol leverages... edge is lost) can be commented out.
+
+%Figures 5 and 6 can be removed.
+
+% The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
 It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
@@
 
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
+% \begin{figure}[H]
+%   \centering
+%   \begin{tikzpicture}[node distance=2cm]
+%     % s1
+%     \draw [thick] (1.3,3) -- (1.7,3);
+%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
+%     % s2
+%     \draw [thick] (3.3,3) -- (3.7,3);
+%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
 
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
+%     \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
+%     \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
 
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
+%     \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
 
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
+%   \end{tikzpicture}
+%   \caption{An example of the \tDelta protocol preserving reciprocal consistency.}
+%   \label{delta-abort}
+% \end{figure}
 
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
+% Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
+Figure \ref{corruption-again} displays an example when the \tDelta protocol fails to maintain reciprocal consistency.
 The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
 
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
+% \begin{figure}[H]
+%   \centering
+%   \begin{tikzpicture}[node distance=2cm]
+%     \draw [thick] (1.3,3) -- (1.7,3);
+%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
+%     \draw [thick] (3.3,3) -- (3.7,3);
+%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
+%         \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
+%     \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
 
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
+%     \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
+%     \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
+%   \end{tikzpicture}
+%   \caption{An example of the \tDelta protocol failing to preserve reciprocal consistency.}
+%   \label{corruption-again}
+% \end{figure}
 
@@
 \end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
+% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
 
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
+% \begin{figure}[ht]
+%   \centering
+%     \begin{tikzpicture}[node distance=2cm]
+%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
+%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
+%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
+%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
+%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
+%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
+%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
+%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
+%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
+%     \end{tikzpicture}
+%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
+%     \label{state-transitions}
+% \end{figure}
 
@@
 
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
+The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
+Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol?
 Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
@@
 We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
+The \tDelta protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
 The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
@@
 
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
+% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.
 

### 2020-04-06 paper/conclusion.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
 \section{Conclusions}
@@
 Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/corruption.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
+\section{Corruption in the Absence of Concurrency Control}
 \label{sec:db-corruption}
@@
 $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
+\input{cypher/delete-edge.tex}
 Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
@@
 
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput
+\input{cypher/spread.tex}

### 2020-04-06 paper/cypher/delete-edge.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
@@
 \end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/delta.tex
refactor dir

--- 2020-04-05 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{xspace}
-\usepackage[labelformat=simple]{subcaption}
-\renewcommand\thesubfigure{(\alph{subfigure})}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\newcommand{\tDelta}{\textsf{Delta}\xspace}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (details in Section \ref{sec:recipr-cons}).
-
-In practice, graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-
-When, for example, a BASE database \cite{Pritchett2008} is adapted with a graph processing layer, then violations of reciprocal consistency will occur if that adaptation provides no concurrency control for operations that span partitions in order to offer higher performance.
-This paper proposes a simple concurrency control protocol, called \tDelta protocol, that does not impede performance adversely.
-That is because the protocol is exclusively designed for one purpose only: reciprocal consistency in distributed edges.
-Its design leverages the fact that the updating of end pointers of a given distributed edge immediately follow each other and the small interval between them is the window for possible conflicts between concurrent updates.
-The protocol ensures that any two updates on any given end pointer of any given edge are at least $\Delta$ time apart by aborting any write attempt that is within .
-It ensures reciprocal consistency when $\Delta$ is chosen to be larger than the length of the conflict window.
-
-The paper is organized as follows: Section \ref{sec:recipr-cons} presents reciprocal consistency, Sections \ref{sec:distr-graph-datab}-\ref{sec:db-corruption} discuss the architecture of distributed graph databases and describes how corruption occurs.
-Section \ref{sec:tdelta-protocol} outlines the \tDelta protocol.
-Sections \ref{sec:modeling}-\ref{sec:evaluation} presents the model used to evaluate protocol performance.
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing properties and a sequence of edges that comprise the adjacency list of a vertex \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (4.2,0);
-  \end{tikzpicture}
-  \caption{A reciprocally consistent edge.}
-  \label{reciprocally-consistent-edge}
-\end{figure}
-
-\section{Distributed Graph Databases}
-\label{sec:distr-graph-datab}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:tx-ty}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-\label{fig:ty-tx}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:same-side}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In \subref{fig:tx-ty} $T_y$ begins writing to the distributed edge before $T_x$, in \subref{fig:ty-tx} the converse is true, else they are equivalent.
-In \subref{fig:same-side} both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{Corruption in the Absence of Concurrency Control}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated.
-Possible interleavings of concurrent writes to a distributed edge are given in Figure \ref{conf-scen}.
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-For such half-corrupted edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} u.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
 \section{\tDelta Protocol}
@@
 % \end{figure}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-% \begin{figure}[ht]
-%   \centering
-%     \begin{tikzpicture}[node distance=2cm]
-%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-%     \end{tikzpicture}
-%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-%     \label{state-transitions}
-% \end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \tDelta protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/dist-graphdb.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
 \section{Distributed Graph Databases}
+\label{sec:distr-graph-datab}
 
@@
 
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
+\input{figures/dist-graph}
 
@@
 This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
 This process is explained in Section \ref{sec:db-corruption}.
 
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
+\input{figures/conflict-scenarios}
 
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput
+\input{figures/half-corrupted-edge}

### 2020-04-06 paper/evaluation.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
 \section{Evaluation}
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
+  \includegraphics[width=\linewidth]{./figures/delta_50_75_100}
   \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
+  \includegraphics[width=\linewidth]{./figures/aborts_50_75_100}
   \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
@@
 \end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/figures/dist-graph.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
 \begin{figure}[ht]
@@
 \end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/figures/half-corrupted-edge.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
     \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
   \end{tikzpicture}
@@
 \end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/figures/reciprocally-consistent-edge.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{subcaption}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
-
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
-
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
 \begin{figure}[ht!]
@@
 
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
+    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (4.2,0);
   \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
+  \caption{A reciprocally consistent edge.}
+  \label{reciprocally-consistent-edge}
 \end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \emph{Delta} protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/introduction.tex
refactor dir

--- 2020-04-05 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{xspace}
-\usepackage[labelformat=simple]{subcaption}
-\renewcommand\thesubfigure{(\alph{subfigure})}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\newcommand{\tDelta}{\textsf{Delta}\xspace}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
 \section{Introduction}
@@
   \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
+  \includegraphics[width=\linewidth]{./figures/janusgraph-adj-list}
   \caption{Database records representing vertices, containing properties and a sequence of edges that comprise the adjacency list of a vertex \cite{janusgraph}.}
@@
 \end{figure}
-
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
-
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
-
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (4.2,0);
-  \end{tikzpicture}
-  \caption{A reciprocally consistent edge.}
-  \label{reciprocally-consistent-edge}
-\end{figure}
-
-\section{Distributed Graph Databases}
-\label{sec:distr-graph-datab}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:tx-ty}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-\label{fig:ty-tx}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:same-side}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In \subref{fig:tx-ty} $T_y$ begins writing to the distributed edge before $T_x$, in \subref{fig:ty-tx} the converse is true, else they are equivalent.
-In \subref{fig:same-side} both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{Corruption in the Absence of Concurrency Control}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated.
-Possible interleavings of concurrent writes to a distributed edge are given in Figure \ref{conf-scen}.
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-For such half-corrupted edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} u.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\section{\tDelta Protocol}
-\label{sec:tdelta-protocol}
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-Our \tDelta protocol employs principles behind all these well-tested strategies but has two crucial differences:
-\begin{itemize}
-\item No locks are used.
-\item A write operation need not await until the preceding write commits but can proceed if at least $\Delta$ duration (measured in local clock) has elapsed.
-\end{itemize}
-
-These differences lead to several advantages in the context of graph databases.
-Firstly, a subset of edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-Secondly, graph transactions tend to be longer-lived than transactions in relational databases.
-Waiting for earlier writes on edges by long running transactions to commit, these frequently accessed edges significantly limit concurrency and reduce throughput.
-With these concerns in mind we developed the \tDelta protocol, which aimed at preventing edges becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-\subsection{Protocol Description}
-\label{sec:protocol-description}
-
-The \tDelta Protocol has five rules:
-\begin{enumerate}
-\item A transaction's update on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
-\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative update was done at least $\Delta$ time before (where time is measured as per local clock).
-\item If a transaction successfully performs all its tentative updates, then it is permitted to commit; otherwise, it must abort.
-\item A transaction commits, when all its tentative updates are made permanent, e.g., using an atomic commitment protocol.
-\item Tentative updates of an aborting transactions are ignored.
-An ignored tentative update can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
-\end{enumerate}
-
-\subsection{Correctness Reasoning}
-\label{sec:corr-reas}
-
-Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction updating one end of a distributed edge at one server and updating the other end of the same edge at another server.
-
-Let $\Delta$ be chosen such that $\Delta > \delta$.
-Now consider the interleaving in Figure \ref{fig:tx-ty} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
-Note $t_y < t_x$, i.e.
-$(t_x - t_y) > 0$.
-
-Say, $T_y$ reaches $S_i$ at time $t_y + d$, where $d$ is the actual time elapsed between completing a tentative write at one end and starting at another end, let us assume that $d \leq \delta$.
-When $T_y$ arrives at $S_i$ it will find a tentative write done at time $t_x$.
-In this case, $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge.
-A similar arguments can be made for the scenario in with Figure \ref{fig:ty-tx}.
-For the interleaving in Figure \ref{fig:same-side}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$.
-
-Assume that $d > \Delta$, i.e., the estimate $\delta$ does not hold at this moment.
-In Figure \ref{fig:tx-ty} interfering updates are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$ otherwise, an edge becomes half-corrupted.
-This holds for the interleaving in \ref{fig:ty-tx}.
-Thus, in the extreme case $t_x = t_y$, reciprocal consistency is not guaranteed if $d > \Delta$ when the former exceeds its upper bound estimate $\delta$.
-For Figure \ref{fig:same-side}, if $d$ for $T_x$ is larger than $\Delta$, reciprocal consistency can be violated.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-
-In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-%The two paragraphs (The protocol leverages... edge is lost) can be commented out.
-
-%Figures 5 and 6 can be removed.
-
-% The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-% \begin{figure}[H]
-%   \centering
-%   \begin{tikzpicture}[node distance=2cm]
-%     % s1
-%     \draw [thick] (1.3,3) -- (1.7,3);
-%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-%     % s2
-%     \draw [thick] (3.3,3) -- (3.7,3);
-%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-%     \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-%     \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-%     \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-%   \end{tikzpicture}
-%   \caption{An example of the \tDelta protocol preserving reciprocal consistency.}
-%   \label{delta-abort}
-% \end{figure}
-
-% Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \tDelta protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-% \begin{figure}[H]
-%   \centering
-%   \begin{tikzpicture}[node distance=2cm]
-%     \draw [thick] (1.3,3) -- (1.7,3);
-%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-%     \draw [thick] (3.3,3) -- (3.7,3);
-%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-%         \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-%     \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-%     \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-%     \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-%   \end{tikzpicture}
-%   \caption{An example of the \tDelta protocol failing to preserve reciprocal consistency.}
-%   \label{corruption-again}
-% \end{figure}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-% \begin{figure}[ht]
-%   \centering
-%     \begin{tikzpicture}[node distance=2cm]
-%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-%     \end{tikzpicture}
-%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-%     \label{state-transitions}
-% \end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \tDelta protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/main.tex
refactor dir

--- 2020-03-20 paper/main.tex
@@
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
 \setcopyright{acmcopyright}
@@
 
-\section{Introduction}
-\label{sec:introduction}
+\input{introduction}
 
-Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}.
-Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}.
-Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities.
-In addition, properties can be stored on both vertices and edges.
-In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects.
-This allows for bi-directional traversal and improved query performance \cite{Robinson2015}.
-An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).
+\input{reciprocal-consistency}
 
-In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
-A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}).
-A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.
+\input{dist-graphdb}
 
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database.
-Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
-Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications.
-This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.
+\input{corruption}
 
-\section{Reciprocal Consistency}
-\label{sec:recipr-cons}
+\input{delta}
 
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
+\input{modeling}
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects.
-A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
+\input{evaluation}
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}
-
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person}), (b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{CREATE} (a)-[w:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-
-\section{Distributed Graph Databases}
-
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
-Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=1cm]
-
-    \node (s1) [server,label=$S_1$] {};
-    \node (s2) [server, right of=s1,xshift=1.5cm,yshift=1cm,label=$S_2$] {};
-    \node (s3) [server, right of=s1,xshift=4cm,label=$S_3$] {};
-
-    \node (v1) [vertex,xshift=-0.5cm,yshift=-0.8cm] {};
-    \node (v2) [vertex, above of=v1,xshift=-0.1cm,yshift=0.7cm] {};
-    \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-    \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.25cm] {};
-    \node (v5) [vertex, above right of=v3,xshift=-0.1cm,yshift=-0.2cm] {};
-    \node (v6) [vertex, right of=v5,xshift=0.5cm,yshift=0.8cm] {};
-    \node (v7) [vertex, below of=v6, xshift=-0.15cm,yshift=0cm] {};
-    \node (v8) [vertex, right of=v4, yshift=-0.1cm,xshift=4cm] {};
-    \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-    \node (v10) [vertex, above of=v9,xshift=-1.7cm,yshift=-0.33cm] {};
-    \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-    \node (v12) [vertex, above of=v11,yshift=-0.3cm,xshift=-1cm] {};
-    \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-    \draw [local] (v2) -- (v1);
-    \draw [local] (v1) -- (v3);
-    \draw [local] (v3) -- (v2);
-    \draw [local] (v2) -- (v5);
-    \draw [local] (v1) -- (v4);
-    \draw [local] (v5) -- (v4);
-    \draw [dist] (v5) -- (v6);
-    \draw [dist] (v4) -- (v7);
-    \draw [dist] (v8) -- (v4);
-    \draw [local] (v8) -- (v9);
-    \draw [local] (v9) -- (v11);
-    \draw [local] (v10) -- (v7);
-    \draw [dist] (v10) -- (v12);
-    \draw [dist] (v12) -- (v13);
-    \draw [local] (v10) -- (v6);
-    \draw [dist] (v10) -- (v11);
-    \draw [local] (v13) -- (v10);
-    \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  \caption{A graph partitioned across $k=3$ servers in a cluster. Dashed lines represent distributed edges and solid lines represent local edges.}
-  \Description{Graph representation.}
-  \label{dist-graph}
-\end{figure}
-
-Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
-
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (1.3,3) -- (1.7,3);
-      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (3.3,3) -- (3.7,3);
-      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (5.3,3) -- (5.7,3);
-      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (7.3,3) -- (7.7,3);
-      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
-      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \draw [thick] (9.3,3) -- (9.7,3);
-      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
-      \draw [thick] (11.3,3) -- (11.7,3);
-      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
-      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
-      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
-    \end{tikzpicture}
-    \caption{}
-  \end{subfigure}%
-  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$.
-In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent.
-In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
-  \label{conf-scen}
-\end{figure*}
-
-\section{Corruption in BASE  Distributed Graph Databases}
-\label{sec:db-corruption}
-
-When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively.
-Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}.
-For such edges there exists a correct and a incorrect entry.
-Note, the order in which the two writes take place is not constrained.
-For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}.
-A graph with half-corrupted edges has suffered \emph{structural corruption}.
-
-Now, if subsequent transactions read the incorrect entry of a half-corrupted edge and write further edges, \emph{semantic corruption} has been introduced into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
-To illustrate the process of corruption, consider two transactions $T_x$ and $T_y$.
-$T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tx}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{DELETE} w
-
-\textcolor{grey}{// Ty}
-\textcolor{blue}{MATCH} (a:\textcolor{green}{Person})-[w:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{WHERE} a.\textcolor{red}{name} = 'Tolkien' \textcolor{blue}{AND} b.\textcolor{red}{title} = 'The Hobbit'
-\textcolor{blue}{SET} w.\textcolor{red}{year} = 1937
-\end{Verbatim}
-Any interleaving in Figure \ref{conf-scen} will result in the half-corrupted edge displayed in Figure \ref{half-corrupted}.
-Assume, the correct ordering of transactions is $T_y$ then $T_x$ and the edge between \emph{a} and \emph{b} should still exist.
-The following transaction, $T_z$, introduces semantic corruption if it starts at \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
-Further corruption spreads by the same mechanism.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
-\textcolor{grey}{// Tz}
-\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
-\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
-\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
-\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
-\end{Verbatim}
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
-
-    \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
-        \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
-
-  \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}
-
-\section{\emph{Delta} Protocol}
-
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}.
-Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.
-
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
-
-This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-In addition, graph transactions tend to be longer-lived than transactions in relational databases.
-Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput.
-With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption.
-We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
-
-The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
-However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.
-
-% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:
-
-%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
-%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$
-
-%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    % s1
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    % s2
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
-  \label{delta-abort}
-\end{figure}
-
-Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-\begin{figure}[H]
-  \centering
-  \begin{tikzpicture}[node distance=2cm]
-    \draw [thick] (1.3,3) -- (1.7,3);
-    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-    \draw [thick] (3.3,3) -- (3.7,3);
-    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-  \end{tikzpicture}
-  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
-  \label{corruption-again}
-\end{figure}
-
-In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
-\begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
-\end{itemize}
-
-\section{Modeling}
-\label{sec:modeling}
-
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
-
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
-\begin{enumerate}
-\item Local and clean.
-\item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
-\end{enumerate}
-The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
-
-\begin{figure}[ht]
-  \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-      \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-      \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-      \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-      \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-      \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-      \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-      \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-    \end{tikzpicture}
-    \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-    \label{state-transitions}
-\end{figure}
-
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
-
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
-
-The \emph{Delta} protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \emph{Delta} protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
-\begin{itemize}
-\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
-\item \emph{Workload}.
-Measured as transactions per second (TPS).
-Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
-\item \emph{Distributed Write Delays and Choosing $\Delta$}.
-The smaller the delays the less likely the bound $\Delta$ is violated.
-Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
-\end{itemize}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
-The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{time-to-corruption-results}
-\end{figure}
-
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{./images/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
-  \label{aborts-results}
-\end{figure}
-
-\section{Conclusions}
-
-Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at a developing protocol specific for distributed graph databases.
-We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.
-The \emph{Delta} protocol leverages the fact writes to distributed edges always consists of two sequential writes to entries in the adjacency lists of vertices the edge connects.
-The protocol provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).
-However, such a mechanism is believed to be valuable in practice, given the popularity of BASE distributed graph databases and the rate at which corruption can spread if left unchecked.
-Simulations indicate the protocol rules out corruption resulting from half-corrupted distributed edges in realistic database lifetime, with the abort rate being in a reasonable range.
-For future work, we intend on implementing the protocol to assess the validity of the simulations and measure performance against a BASE distributed graph database operating without any concurrency control.
-Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.
+\input{conclusion}
 

### 2020-04-06 paper/modeling.tex
refactor dir

\section{Modeling}
\label{sec:modeling}

To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}. A summary of the model is now provided, before discussing the extensions.

The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second. To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write. To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$. For each type, a fraction $f$ are distributed edges and the remainder are local edges. The network delay between servers was assumed to be exponentially distributed with rate $\rho$. At any moment in time an edge can be in one of four states:
\begin{enumerate}
\item Local and clean.
\item Distributed and clean.
\item Half-corrupted distributed edge arising from a dirty write.
\item Semantically corrupted.
\end{enumerate}
% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.

% \begin{figure}[ht]
%   \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
%     \end{tikzpicture}
%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
%     \label{state-transitions}
% \end{figure}

Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers. Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived. These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$. At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt. The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.

% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e. the interleaving in Figure 3(c) cannot occur.}.

The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$. Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol? Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed. The answers to these questions depends on several parameters characterizing the following systemic aspects:
\begin{itemize}
\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
\item \emph{Workload}. Measured as transactions per second (TPS). Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
\item \emph{Distributed Write Delays and Choosing $\Delta$}. The smaller the delays the less likely the bound $\Delta$ is violated. Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
\end{itemize}

### 2020-04-06 paper/reciprocal-consistency.tex
refactor dir

\section{Reciprocal Consistency}
\label{sec:recipr-cons}

In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices. In the storage layer, edge information is stored with \textbf{both} the source and destination vertices. This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.

A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}. Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit. It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).

Using openCypher \cite{openCypher} this can be represented by:
\input{cypher/create-edge}

Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}. Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction. When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity. A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.

\input{figures/reciprocally-consistent-edge}

### 2020-04-06 paper/abstract.tex
more refactoring

--- 2020-04-06 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{xspace}
-\usepackage[labelformat=simple]{subcaption}
-\renewcommand\thesubfigure{(\alph{subfigure})}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
-\newcommand{\tDelta}{\textsf{Delta}\xspace}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
 \begin{abstract}
-
 In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
@@
 \end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\input{introduction}
-
-\input{reciprocal-consistency}
-
-\input{dist-graphdb}
-
-\input{corruption}
-
-\input{delta}
-
-\input{modeling}
-
-\input{evaluation}
-
-\input{conclusion}
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/commands.tex
more refactoring

--- 2020-04-06 paper/main.tex
@@
-\documentclass[sigplan,10pt]{acmart}
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\usepackage{xspace}
-\usepackage[labelformat=simple]{subcaption}
+% commands
 \renewcommand\thesubfigure{(\alph{subfigure})}
-\usepackage{tikz}
-\usepackage{fancyvrb}
-\usepackage{xcolor}
-\usepackage{breqn}
-
 \newcommand{\tDelta}{\textsf{Delta}\xspace}
 
+% tikz
 \usetikzlibrary{decorations.pathmorphing}
@@
 
+% colours
 \definecolor{grey}{rgb}{0.52, 0.52, 0.51}
@@
 \definecolor{green}{rgb}{0.0, 0.42, 0.24}
-
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
-\acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-\begin{document}
-
-\title{Preserving Reciprocal Consistency in Distributed Graph Databases}
-
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\renewcommand{\shortauthors}{Waudby, et al.}
-
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
-
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
-
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
-
-\maketitle
-
-\input{introduction}
-
-\input{reciprocal-consistency}
-
-\input{dist-graphdb}
-
-\input{corruption}
-
-\input{delta}
-
-\input{modeling}
-
-\input{evaluation}
-
-\input{conclusion}
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{papoc}
-
-\end{document}
-\endinput

### 2020-04-06 paper/main.tex
more refactoring

@@
 \usepackage[labelformat=simple]{subcaption}
-\renewcommand\thesubfigure{(\alph{subfigure})}
 \usepackage{tikz}
@@
 \usepackage{breqn}
-
-\newcommand{\tDelta}{\textsf{Delta}\xspace}
-
-\usetikzlibrary{decorations.pathmorphing}
-\tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{node}=[circle,draw,radius=0.3, align=center]
-
-\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
-\definecolor{red}{rgb}{0.7, 0.11, 0.11}
-\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
-\definecolor{green}{rgb}{0.0, 0.42, 0.24}
 
@@
 
-%
 % Submission ID.
@@
 
-\author{Jack Waudby}
-\email{j.waudby2@ncl.ac.uk}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Paul Ezhilchelvan}
-\email{paul.ezhilchelvan@ncl.ac.uk}
-\orcid{0000-0002-6190-5685}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
-
-\author{Jim Webber}
-\email{jim.webber@neo4j.com}
-\affiliation{%
-  \institution{Neo4j}
-  \streetaddress{Union House,182-194 Union Street}
-  \city{London}
-  \country{UK}
-  \postcode{SE1 0LH}}
-
-\author{Isi Mitrani}
-\email{isi.mitrani@ncl.ac.uk}
-\orcid{0000-0002-7797-7755}
-\affiliation{%
-  \institution{Newcastle University}
-  \city{Newcastle}
-  \country{UK}
-  \postcode{NE4 5TG}
-}
+\input{authors}
 
@@
 
-\begin{abstract}
-
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
-\end{abstract}
+\input{abstract}
 

### 2020-04-06 paper/dist-graphdb.tex
some paul changes

@@
 The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
+Figure \ref{dist-graph} depicts a graph database partitioned across 3 servers, $S1$, $S2$ and $S3$.
+Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges} (shown with dashed lines in Figure \ref{dist-graph}).
 The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.

### 2020-04-06 paper/reciprocal-consistency.tex
some paul changes

@@
 In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
+This, as we explain next, allows an edge to be traversed in both directions and thus facilitates improved query performance.
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
+A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is given in Figure \ref{adj-list}.
+A vertex is represented by a record that contains one or more vertex properties and a sequence of edge pointers (each shown simply as edge in Figure \ref{adj-list}) pointing to other vertices.
+This sequence of edge pointers is referred to as an adjacency list.
+
 Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
@@
 
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
+Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
 Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.

### 2020-04-06 paper/figures/half-corrupted-edge.tex
rename tikz node

@@
   \begin{tikzpicture}[node distance=2cm]
-    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
+    \node (v1) [big_vertex,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
 
-    \node (v2) [vertex2,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
+    \node (v2) [big_vertex,xshift=5cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
 

### 2020-04-06 scripts/make-plots.R
fix image stain

@@ -34,6 +34,7 @@ a=ggplot(data=aborts, aes(x=tps,y=collisions,linetype=delta,color=delta)) +
     theme_few()
 a
 
+ggsave(a, filename = "aborts.pdf",  bg = "transparent")
 
 
 # load time to corruption data in 
@@ -79,6 +80,7 @@ p=ggplot(data=df, aes(x=lam,y=U,linetype=type,color=type)) +
     scale_y_continuous(name="log(U)",breaks=seq(0,40,5)) + 
     theme_few()
 p
+ggsave(p, filename = "./paper/figures/delta.pdf",  bg = "transparent")
 
 
 dev.off()

### 2020-04-07 paper/commands.tex
rename tikz node
section 2

@@
     \tikzstyle{vertex} = [circle, minimum width=10pt,draw,inner sep=0pt]
-\tikzstyle{vertex2} = [circle, minimum width=40pt,draw,inner sep=0pt]
+    \tikzstyle{big_vertex} = [circle, minimum width=40pt,draw,inner sep=0pt]
     \tikzstyle{server} = [rectangle, rounded corners,minimum width=2cm,minimum height=2.5cm,draw=black]
@@
     \definecolor{green}{rgb}{0.0, 0.42, 0.24}
+
+            \def\record(#1,#2,#3,#4,#5,#6){
+              \node (v) [draw,black,minimum height=0.7cm,minimum width=1cm,text width=0.2\linewidth,xshift=-6cm,align=center,yshift=#1cm] {#2};
+
+            \node (b) [draw,black,minimum height=0.7cm,minimum width=11.4cm,text width=0.2\linewidth,xshift=1.4cm,yshift=#1cm] {};
+
+            \draw[thick,->] (v) -- (b);
+
+            \node (i1)  [draw,black,minimum height=0.5cm,minimum width=2cm,xshift=-2.8cm,text width=0.3\linewidth,yshift=#1cm,align=center] {#3};
+
+            \node (i2)  [draw,black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.9cm,right of=i1,yshift=0cm,text width=0.3\linewidth,align=center] {#4};
+
+            \node (i3)  [draw,black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.75cm,right of=i2,yshift=0cm,align=center] {#5};
+
+            \node (i4)  [draw,black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.6cm,right of=i3,yshift=0cm,align=center] {#6};
+
+        }
+
+\def\spaceRecord(#1){
+\node (v) [black,minimum height=0.7cm,minimum width=1cm,text width=0.2\linewidth,xshift=-6cm,align=center,yshift=#1cm] {$\vdots$};
+\node (b) [black,minimum height=0.7cm,minimum width=11.4cm,text width=0.2\linewidth,xshift=1.4cm,yshift=#1cm] {};
+\node (i1)  [black,minimum height=0.5cm,minimum width=2cm,xshift=-2.8cm,text width=0.3\linewidth,yshift=#1cm,align=center] {$\vdots$};
+\node (i2)  [black,minimum height=0.5cm,minimum width=2.5cm,xshift=0.9cm,right of=i1,yshift=0cm,text width=0.3\linewidth,align=center] {$\vdots$};
+\node (i3)  [minimum height=0.5cm,minimum width=2.5cm,xshift=0.75cm,right of=i2,yshift=0cm,align=center] {$\vdots$};
+\node (i4)  [minimum height=0.5cm,minimum width=2.5cm,xshift=0.6cm,right of=i3,yshift=0cm,align=center] {$\vdots$};
+}

### 2020-04-07 paper/corruption.tex
section 2

@@
 \label{sec:db-corruption}
+
+This section must be re-structured for ease of understanding.
+First talk about transactions interleaving - narrative using fig 4.
+Explain: Interleavings lead to half-corruption.
+Exemplify half corruption using the first half of the code example.
+Relate to dirty write..
+Explain how half-corruption then leads to  sematic corruption and operational corruption.
+Exemplify it using the second half of the coded example - unknown author.
+Finally relate half corruption to Read Uncommitted.
+
+\input{figures/hc}
 

### 2020-04-07 paper/appendix.tex
camera-ready

\section{Derivation of Conflict Probability}
\label{sec:deriv-confl-prob}

Under the \tDelta protocol each interleaving in Figure \ref{fig:conf} can be given a probability of occurring. Figures \ref{fig:a} and \ref{fig:b} are equivalent and can be summarized by the same probability. Letting $t_x = 0$, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + D) \cap (T_y > \Delta-D) \right]$$ The arrival times of $T_x,T_y$ are assumed exponentially distributed, $T \sim \exp (\rho)$.
Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
The probability of $d$ exceeding $\Delta$ is exponentially distributed with mean $\mu$, $D\sim \exp(\mu)$.
Therefore,
\begin{dmath*}
   q^{new}_{i} =  {P \left[ ( T_1 >  \Delta + D) \cap (T_2 > \Delta-D)  \right]} \\
   =  \int_{0}^{\Delta}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} d} e^{-\mu (\Delta+d)} e^{-\mu (\Delta-d)} dd + \int_{\Delta}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} d} e^{-\mu (\Delta+d)} dd  \\
   =  e^{-2 d \mu} - \left( \frac{\mu}{\frac{\lambda P_i}{2 N_i} + \mu} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\mu)d} \\
 \end{dmath*}
To simply the derivation of the conflict probability we assume FIFO which rules out the interleaving in Figure \ref{fig:c}.
 % To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
 % \begin{align*}
 %   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
 %   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
 %   e^{- \delta d} & = \epsilon  \\
 %   \Delta & = - \frac{\ln(\epsilon)}{\delta}
 % \end{align*}

%  Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$. For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
% \begin{align}
%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
% \end{align}
% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
% \begin{align}
%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
% \end{align}

% Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
% Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$

%  % =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
%  %        =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\

### 2020-04-07 paper/sec1-introduction.tex
camera-ready

--- 2020-04-06 paper/introduction.tex
@@
 Its design leverages the fact that the updating of end pointers of a given distributed edge immediately follow each other and the small interval between them is the window for possible conflicts between concurrent updates.
-The protocol ensures that any two updates on any given end pointer of any given edge are at least $\Delta$ time apart by aborting any write attempt that is within .
-It ensures reciprocal consistency when $\Delta$ is chosen to be larger than the length of the conflict window.
 
@@
 Sections \ref{sec:modeling}-\ref{sec:evaluation} presents the model used to evaluate protocol performance.
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\linewidth]{./figures/janusgraph-adj-list}
-  \caption{Database records representing vertices, containing properties and a sequence of edges that comprise the adjacency list of a vertex \cite{janusgraph}.}
-  \label{adj-list}
-\end{figure}

### 2020-04-07 paper/sec2-reciprocal-consistency.tex
camera-ready

--- 2020-04-06 paper/reciprocal-consistency.tex
@@
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is given in Figure \ref{adj-list}.
-A vertex is represented by a record that contains one or more vertex properties and a sequence of edge pointers (each shown simply as edge in Figure \ref{adj-list}) pointing to other vertices.
-This sequence of edge pointers is referred to as an adjacency list.
-
 Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
 It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
+The logical representation of the edge is given in Figure \ref{rc-edge}.
 
@@
 
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
+A common internal representation of this graph (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is given in Figure \ref{db-rep}.
+A vertex is represented by a record that contains one or more vertex properties and a sequence of edge pointers pointing to other vertices.
+This sequence of edge pointers is referred to as an adjacency list.
+
+From Figure \ref{db-rep} it can be seen adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
 Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
@@
 
-\input{figures/reciprocally-consistent-edge}
+\input{figures/rc}

### 2020-04-07 paper/sec4-corruption.tex
camera-ready

\section{Corruption in the Absence of Concurrency Control}
\label{sec:db-corruption}

Consider the edge \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} spans servers $S_i$ and $S_j$.
When a transaction writes a distributed edge there are two important facts to consider:
\begin{enumerate}
\item Two write operations are performed, writing reciprocal entries in the adjacency lists of \emph{a} and \emph{b}.
\item Write order is unconstrained, a transaction is equally likely to write \emph{a} then \emph{b} as it is to write \emph{b} then \emph{a}.
\end{enumerate}

Without concurrency control, transactions $T_x$ and $T_y$ that write the distributed edge can interleave in the following ways:
\begin{enumerate}[(i)]
\item $T_x$ begins writing the distributed edge before $T_y$ but at opposite ends, transactions cross in the network and at $S_i$, $T_x \rightarrow T_y$ and at $S_y$, $T_y \rightarrow T_x$, Figure \ref{fig:a}.
\item $T_y$ begins writing the distributed edge before $T_x$ but at opposite ends, transactions cross in the network and at $S_i$, $T_x \rightarrow T_y$ and at $S_j$, $T_y \rightarrow T_x$, Figure \ref{fig:b}.
\item $T_x$ and $T_y$ perform their first writes at $S_i$, $T_x \rightarrow T_y$, but overlap in the network and arrive out-of-order at $S_j$, $T_y \rightarrow T_x$, Figure \ref{fig:c}.
\end{enumerate}

Each interleaving in Figure \ref{fig:conf} results in a \emph{half-corrupted} distributed edge - reciprocal consistency has been violated.
A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000})
To illustrate the process of half-corruption, consider $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
\input{cypher/delete-edge.tex}
\input{figures/hc}
Each interleaving results in the distributed edge in Figure \ref{hc-edge}, reciprocal consistency is clearly violated.
We say a graph with half-corrupted edges has suffered \emph{structural corruption}.

Assume we know the correct state of the distributed edge to be \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} with \texttt{\textcolor{red}{year} = 1937}, then there exists correct and incorrect entries for edge, visible in Figure \ref{hc-db-rep}.
When a transaction reads an edge it does not check both entries that constitute the edge are reciprocal consistent.
Therefore, a subsequent transactions can read the incorrect entry of a half-corrupted edge and write further edges, introducing \emph{semantic corruption} into the database.
Further semantic corruption spreads by the same mechanism.
A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.

For example, $T_z$, introduces semantic corruption if it reads \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.

\input{cypher/spread.tex}

If the database provides the ANSI isolation level \textbf{Read Uncommitted}, the database consistently orders writes from concurrent transactions, which would prevent all interleavings in Figure \ref{fig:conf} and the spread of semantic corruption

### 2020-04-07 paper/sec5-delta.tex
camera-ready

--- 2020-04-06 paper/delta.tex
@@
 \begin{enumerate}
-\item A transaction's update on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
-\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative update was done at least $\Delta$ time before (where time is measured as per local clock).
-\item If a transaction successfully performs all its tentative updates, then it is permitted to commit; otherwise, it must abort.
-\item A transaction commits, when all its tentative updates are made permanent, e.g., using an atomic commitment protocol.
-\item Tentative updates of an aborting transactions are ignored.
-An ignored tentative update can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
+\item A transaction's write on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
+\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative write was done at least $\Delta$ time before (where time is measured as per local clock).
+\item If a transaction successfully performs all its tentative writes, then it is permitted to commit; otherwise, it must abort.
+\item A transaction commits, when all its tentative writes are made permanent, e.g., using an atomic commitment protocol.
+\item Tentative writes of an aborting transactions are ignored.
+An ignored tentative write can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
 \end{enumerate}
@@
 
-Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction updating one end of a distributed edge at one server and updating the other end of the same edge at another server.
+Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction writing one end of a distributed edge at one server and writing the other end of the same edge at another server.
 
 Let $\Delta$ be chosen such that $\Delta > \delta$.
-Now consider the interleaving in Figure \ref{fig:tx-ty} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
-Note $t_y < t_x$, i.e.
-$(t_x - t_y) > 0$.
+Now consider the interleaving in Figure \ref{fig:b} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
+Note $t_y < t_x$, i.e. $(t_x - t_y) > 0$.
 
@@
 In this case, $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge.
-A similar arguments can be made for the scenario in with Figure \ref{fig:ty-tx}.
+A similar arguments can be made for the scenario in with Figure \ref{fig:b}, where $t_y > t_x$ and $T_x$ will abort because it will find out, on reaching server $S_j$ for the final tentative write, that $t_x +d - t_y = d - (t_y-t_x) < d <= \delta < \Delta$.
 For the interleaving in Figure \ref{fig:same-side}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$.
 
-Assume that $d > \Delta$, i.e., the estimate $\delta$ does not hold at this moment.
-In Figure \ref{fig:tx-ty} interfering updates are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$ otherwise, an edge becomes half-corrupted.
-This holds for the interleaving in \ref{fig:ty-tx}.
-Thus, in the extreme case $t_x = t_y$, reciprocal consistency is not guaranteed if $d > \Delta$ when the former exceeds its upper bound estimate $\delta$.
-For Figure \ref{fig:same-side}, if $d$ for $T_x$ is larger than $\Delta$, reciprocal consistency can be violated.
+Assume that $d > \delta$, i.e., the estimate $\delta$ does not hold at this moment.
+In Figure \ref{fig:b} interfering writes are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d < \Delta$ otherwise, reciprocal inconsistency occurs.
+In Figure \ref{fig:a} interfering writes are avoided only if $t_x + d - t_y = d - (t_y-t_x) < d < \Delta$ otherwise, reciprocal inconsistency occurs.
+Thus, in the extreme case $t_x = t_y$ and reciprocal consistency is not guaranteed if $\Delta < d$ when the latter exceeds its upper bound estimate $\delta$.
+For Figure \ref{fig:c}, interfering writes are avoided only if $(t_y + d) - (t_x + d) = (t_y- t_x) < d < \Delta$ other reciprocal inconsistency occurs.
 
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
+% This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
 The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
 
-In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur.
+% In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur.
 Two questions naturally arise regards the performance of the protocol.
 For different values of $\Delta$:
+% \begin{itemize}
+% \item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
+% \item How many transactions are aborted as a result of the protocol?
+% \end{itemize}
+
+In summary, the \tDelta protocol eliminates interleaving of transactions during edge writes, so long as $\Delta$ remains larger than the interval $d$ that elapses between a transaction completing its write at one end of a distributed edge and starting at the other end.
+Since the exact value of $d$ taken by a transaction cannot be precisely known in advance, its bound $\delta$ is estimated with the best effort and we choose $\Delta > \delta$.
+
+The larger the value of $\Delta$ used, the more likely is that $\Delta > d$ holds and half-corruption and thereby operational corruption are averted; also, the more likely is that one or more of concurrent transactions will find their tentative write within $\Delta$ time of each other and choose to abort.
+Our performance evaluation of the protocol will therefore involve measuring the folowing two metrics for various values of $\Delta$:
 \begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
+\item Time taken for 10\% of a large database to be corrupted due to the introduction of half-corruption, and
+\item Number of transactions being aborted per second.
 \end{itemize}
+e let $d$ be an exponentially distributed random variable with mean $\mu$, $D\sim \exp(\mu)$. Therefore, the probability of $d$ exceeding $\Delta$ is given by $P(D > \Delta) = e^{(-\mu \Delta)}$.
 

### 2020-04-07 paper/sec6-modeling.tex
camera-ready

--- 2020-04-06 paper/modeling.tex
@@
 For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
+The delay $d$ between a transaction completing a tentative write at one end and starting at another end is exponential distributed with a mean of $\mu$.
 At any moment in time an edge can be in one of four states:
@@
 
-The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
+The \tDelta protocol impacts the rate of corruption by reducing the probability a given transaction half-corrupts a distributed edge, $q^{new}_i$, see Appendix \ref{sec:deriv-confl-prob} for the derivation.
 Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol?

### 2020-04-07 paper/sec7-evaluation.tex
camera-ready

--- 2020-04-06 paper/evaluation.tex
@@
 In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
-The network delay between servers is exponential distributed with a rate of $200ms$.
+The delay $d$ between a transaction completing a tentative write at one end and starting at another end is exponential distributed with a mean of $5ms$.
 The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
@@
 The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
+For each $\Delta$ the probability that $d$ exceeds $\Delta$ is $P(D > \Delta) = 4.5 \time 10^{-5}, 3.1 \times 10^{-7}, 2.1 \times 10^{-9}$ respectively.
 
 The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 500-50 days.
-For $50ms$ this increases to 74-1 years.
+Under no isolation, $U$ ranges between 50-500 days.
+For $50ms$ this increases to 1-75 years.
 For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
@@
   \centering
-  \includegraphics[width=\linewidth]{./figures/delta_50_75_100}
-  \caption{Time, $\log(U)$, until operational corruption ($\gamma = 0.1$) for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
+  \includegraphics[width=\linewidth]{./figures/delta}
+  \caption{Time until operational corruption.}
   \label{time-to-corruption-results}
@@
 
-The abort rates for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
+The number of aborts per second for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
 The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
@@
   \centering
-  \includegraphics[width=\linewidth]{./figures/aborts_50_75_100}
-  \caption{Aborts per seconds for $\Delta = 50, 75, 100ms$.}
-  \Description{Graph representation.}
+  \includegraphics[width=\linewidth]{./figures/aborts}
+  \caption{Number of aborts per second.}
   \label{aborts-results}

### 2020-04-10 extras/proof.tex
camera-ready

\let\oldref\ref
\renewcommand{\ref}[1]{(\oldref{#1})}

\subsection{Correctness Proof}
\label{sec:correctness-proof}

Lets us first define formally the \emph{interference-free} writing of a distributed edge.
Consider an edge \emph{e} connecting a vertex at server $S_i$ to another vertex at $S_j$.

Let $T_x$ (tentatively) write the ends of \emph{e} without aborting.
Let $Pred_{x}^{i}$ and $Pred_{x}^{j}$ denote the set of transactions that $T_x$ observes to have already carried out (tentatively) writes of \emph{e} at $S_i$ and $S_j$ respectively.
Let \emph{aborted} denote the set of all transactions that have aborted until any given moment

\paragraph{Definition}
$T_x$ updating \emph{e} is interference-free if and only if  $Pred_{x}^{i}-aborted =  Pred_{x}^{j}-aborted$.
Intuitively, the set of transactions that precede $T_x$ in writing \emph{e} and do not abort, must be identical at both ends of \emph{e}.
It is easy to see that when all updates of \emph{e} are interference-free as per this definition, they are identically ordered at both ends of \emph{e}.
(A simple proof by induction is possible and omitted here).

Referring to Figure \ref{fig:conf}, $Pred_{x}^{i} = \{ \phi \} \neq \{ T_y\} =  Pred_{x}^{j}$ for all three cases that depict the three possible interleavings of transactions.

\paragraph{Lemma}
If a transaction $T_x$ follows \tDelta protocol in writing a distributed edge \emph{e}, then its write of \emph{e} is guaranteed to be interference-free provided (i) the bound estimate $\delta$ holds and (ii) $\Delta > \delta$.

\paragraph{Proof by Contradiction}
Let us hypothesize that there exists a $T_y : T_y \in (Pred_{x}^{i}-aborted) \land T_y \in (Pred_{x}^{j}-aborted)$.

Let $t_{x}^{i}$ and $t_{x}^{j}$ be the (global) time instances when $T_x$ writes \emph{e} at $S_i$ and $S_j$ respectively.
Let $t_{y}^{i}$ and $t_{y}^{j}$ be the corresponding instances for $T_y$.
By the hypothesis, $t_{x}^{i} - t_{y}^{i} \geq \Delta$.
Since $\delta$ holds,
\begin{align}
  & t_{x}^{i} - \delta <  t_{x}^{j} <  t_{x}^{i} + \delta  \label{eq:1} \\
  & t_{y}^{i} - \delta <  t_{y}^{j} <  t_{y}^{i} + \delta \label{eq:2}
\end{align}
Negating inequality \ref{eq:2} and re-arranging,
\begin{equation}
  - t_{y}^{i} - \delta < - t_{y}^{j} < - t_{y}^{i} + \delta \label{eq:3}
\end{equation}
Combining \ref{eq:1} and \ref{eq:3}:
\begin{equation}
  t_{x}^{i} - t_{y}^{i} - 2 \delta < t_{x}^{j} - t_{y}^{j} < t_{x}^{i} - t_{y}^{i} + 2 \delta\label{eq:4}
\end{equation}
Let $k = t_{x}^{i} - t_{y}^{i} + 2 \delta$ and $k>0$ because $t_{x}^{i} - t_{y}^{i} \geq \Delta$ and $2\delta>0$. So,
\begin{equation}
  t_{x}^{i} - t_{y}^{i} - 2 \delta \geq \Delta - 2 \delta
\end{equation}
Rewriting \ref{eq:4}:
\begin{equation}
  \Delta  -2 \delta < t_{x}^{j} - t_{y}^{j} < k \label{eq:5}
\end{equation}
$(\Delta - 2 \delta) \geq 0 $ only if $\Delta$ is chosen to be $\geq 2\delta$; otherwise $(\Delta - 2 \delta) < 0 $.
So there are two cases in analyzing \ref{eq:5}:
\paragraph{Case 1}
When $\Delta > 2 \delta$, \ref{eq:5} reduces to $0 < t_{x}^{j} - t_{y}^{j} < k $ and implies that $T_x$ must observe $T_y$ as its predecessor at $S_j$ -- contradicting the hypothesis that $T_y \not\in Pred_{x}^{j} and T_y \not\in aborted$.
\paragraph{Case 2}
When  $\Delta < 2 \delta$, \ref{eq:5} becomes $0 < t_{y}^{j} - t_{x}^{j} < 2 \delta - \Delta$.
Since $\Delta > \delta, (2\delta - \Delta) < \delta < \Delta$.
This means that $T_y$ observes $T_x$ as its predecessor at $S_j$ but within $\Delta$ and must therefore abort. But the hypothesis that $T_y \in Pred_{x}^{i}-aborted$ assumes $T_y \not\in aborted$. So, the supposed hypothesis cannot be correct and \tDelta protocol is proven correct.

### 2020-04-10 paper/abstract.tex
camera-ready

@@
 \begin{abstract}
-In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases.
-If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime.
-Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs.
-However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption.
-A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database.
-Protocol performance is evaluated through simulations.
+  Our earlier work identifies \emph{reciprocal consistency} as an important property that must be preserved in distributed graph databases.
+  It also demonstrates that a failure to do so seriously undermines the integrity of the database itself in the long term.
+  Reciprocal consistency can be maintained as a part of enforcing any known isolation guarantee and such an enforcement is also known to lead to reduction in performance.
+Therefore, in practice, distributed graph databases are often built atop BASE databases with no isolation guarantees, benefiting from good performance but leaving them susceptible to corruption due to violations of reciprocal consistency.
+  This paper designs and presents a lightweight, locking-free protocol and then evaluates the protocol's abilities to preserve reciprocal consistency and also offer good throughput.
+Our evaluations establish that  the protocol can offer both integrity guarantees and sound performance when the value of its parameter is chosen appropriately.
 \end{abstract}

### 2020-04-10 paper/appendix.tex
camera-ready

@@
 
-Under the \tDelta protocol each interleaving in Figure \ref{fig:conf} can be given a probability of occurring.
-Figures \ref{fig:a} and \ref{fig:b} are equivalent and can be summarized by the same probability.
-Letting $t_x = 0$, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + D) \cap (T_y > \Delta-D) \right]$$ The arrival times of $T_x,T_y$ are assumed exponentially distributed, $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The probability of $d$ exceeding $\Delta$ is exponentially distributed with mean $\mu$, $D\sim \exp(\mu)$.
-Therefore,
+Edges are divided into $T$ types, $i= 1,2, ..., T$. The number of edges of type $i$ is represented by $n_i$. 
+The probability an edge of type $i$ is accessed by a given operation is represented by $p_i$.  (Exact values taken by $n_i$ and $p_i$ are given in Section \ref{sec:evaluation}.)
+
+The arrival time, $a$, of a transaction, $T$, is assumed exponentially distributed with rate $\rho$. 
+Where $ \rho  = \frac{\lambda p_i}{2n_i}$ is the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
+
+The time interval, $d$, that elapses between transaction $T$ completing its tentative write at one end of a distributed edge and starting at another end, is assumed exponentially distributed with rate $\mu$.
+
+Consider the interleaving in Fig \ref{fig:a} and assume $T_x$ arrives at $S_i$ at time $0$.
+Then, $T_x$ arrives at $S_j$ after $d_x$.
+Assume $T_y$ arrives at $S_j$ at some time $a$.
+Then, $T_y$ arrives at $S_i$ at time $a  + d_y$.
+Half-corruption occurs under the following conditions:
+\begin{enumerate}[(i)]
+    \item At $S_j$, $d_x > a + \Delta$
+    \item At $S_i$, $a  + d_y > \Delta$
+\end{enumerate}
+The conflict probability, $q^{new}_{i}$, for edge type $i$ is given by,
+
 \begin{dmath*}
-   q^{new}_{i} =  {P \left[ ( T_1 >  \Delta + D) \cap (T_2 > \Delta-D)  \right]} \\
-   =  \int_{0}^{\Delta}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} d} e^{-\mu (\Delta+d)} e^{-\mu (\Delta-d)} dd + \int_{\Delta}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} d} e^{-\mu (\Delta+d)} dd  \\
-   =  e^{-2 d \mu} - \left( \frac{\mu}{\frac{\lambda P_i}{2 N_i} + \mu} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\mu)d} \\
+   q^{new}_{i} =  {P \left[ (d_x > a + \Delta) \cap (d_y > \Delta - a)  \right]} \\
+   =  \int_{0}^{\Delta}  \frac{\lambda p_i}{2 n_i} e^{-\frac{\lambda p_i}{2 n_i} a} e^{-\mu (\Delta+a)} e^{-\mu (\Delta-a)} da + \int_{\Delta}^{\infty} \frac{\lambda p_i}{2 n_i} e^{-\frac{\lambda p_i}{2 n_i} a} e^{-\mu (\Delta+a)} da \\
+   =  e^{-2 a \mu} - \left( \frac{\mu}{\frac{\lambda p_i}{2 n_i} + \mu} \right) e^{-(\frac{\lambda p_i}{2 n_i}+ 2\mu)a} \\
  \end{dmath*}
-To simply the derivation of the conflict probability we assume FIFO which rules out the interleaving in Figure \ref{fig:c}.
+%The interleaving in Fig \ref{fig:a} is equivalent to Fig \ref{fig:b} where $T_y$ arrives at $S_j$ at time $0$ and $S_i$ at $a$ while $T_x$ starts in $S_i$ at $a$ and reaches $S_j$ at $a+d_x$.
+
+From the perspective of computing $q^{new}_{i}$, Fig \ref{fig:b} is equivalent to Fig \ref{fig:a} and the above expression holds.  
+Fig \ref{fig:c} is ignored as it can be avoided by having all first tentative updates on a given edge at a given server (e.g., at $S_i$) be numbered sequentially and having the second tentative attempts  processed (at $S_j$) as per this sequence number.
+
+%To simply the derivation of the conflict probability we assume FIFO which rules out the interleaving in Figure \ref{fig:c}.
+
  % To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.

### 2020-04-10 paper/figures/hc-edge.tex
section 2
camera-ready

--- 2020-04-06 paper/figures/half-corrupted-edge.tex
@@
-\begin{figure}[H]
-  \centering
 \begin{tikzpicture}[node distance=2cm]
@@
 
-    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
+  \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year}:1937}}} (2.4,0);
 \end{tikzpicture}
-  \caption{A half-corrupted edge.}
-  \label{half-corrupted}
-\end{figure}

### 2020-04-10 paper/main.tex
more refactoring++
camera-ready

@@
 
+\usepackage{enumerate}
 \usepackage{xspace}
@@
 
-\setcopyright{acmcopyright}
+\input{commands}
+
 \copyrightyear{2020}
 \acmYear{2020}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
-\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
+\setcopyright{acmcopyright}
+\acmConference[PaPoC '20]{7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Greece}
+\acmBooktitle{7th Workshop on Principles and Practice of Consistency for Distributed Data (PaPoC '20), April 27, 2020, Heraklion, Greece}
 \acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll
-% receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
+\acmDOI{10.1145/3380787.3393675}
+\acmISBN{978-1-4503-7524-5/20/04}
 
@@
 
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10002951.10002952.10003190.10003193</concept_id>
+       <concept_desc>Information systems~Database transaction processing</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10002951.10002952.10002953.10010146</concept_id>
+       <concept_desc>Information systems~Graph-based database models</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
 
-\ccsdesc[500]{Data Management~Graph Databases}
-\ccsdesc[300]{Data Management~Reciprocal Consistency}
-\ccsdesc[300]{Data Management~Concurrency Control}
+\ccsdesc[500]{Information systems~Database transaction processing}
+\ccsdesc[500]{Information systems~Graph-based database models}
 
-\keywords{Graph Databases, Reciprocal Consistency, BASE}
+\keywords{Graph Databases, Concurrency Control, Reciprocal Consistency}
 
@@
 
-\input{introduction}
+\input{sec1-introduction}
 
-\input{reciprocal-consistency}
+\input{sec2-reciprocal-consistency}
 
-\input{dist-graphdb}
+\input{sec3-dist-graphdb}
 
-\input{corruption}
+\input{sec4-corruption}
 
-\input{delta}
+\input{sec5-delta}
 
-\input{modeling}
+\input{sec6-modeling}
 
-\input{evaluation}
+\input{sec7-evaluation}
 
-\input{conclusion}
-
-% \appendix
-% \section{Appendix}
-
-% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.
-
-% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$.
-Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$.
-If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted.
-Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.
-
-% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$.
-Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
-The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.
-
-% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$
-
-% Therefore,
-% \begin{dmath*}
-%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
-%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
-%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
-% \end{dmath*}
-
-% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
-% \begin{align*}
-%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
-%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
-%   e^{- \delta d} & = \epsilon  \\
-%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
-% \end{align*}
-
-% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$.
-For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
-% \begin{align}
-%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
-% \end{align}
-% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
-% \begin{align}
-%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
-%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
-% \end{align}
-
-%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
-%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$
-
-%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
-%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\
+\input{sec8-conclusion}
 
@@
 
+\appendix
+
+\input{appendix}
+
 \end{document}

### 2020-04-10 paper/sec1-introduction.tex
camera-ready

@@
 A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
-Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database.
+Recent work \cite{Ezhilchelvan2018} and \cite{Ezhilchelvan2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database.
 Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.
@@
 When, for example, a BASE database \cite{Pritchett2008} is adapted with a graph processing layer, then violations of reciprocal consistency will occur if that adaptation provides no concurrency control for operations that span partitions in order to offer higher performance.
-This paper proposes a simple concurrency control protocol, called \tDelta protocol, that does not impede performance adversely.
+This paper proposes a simple concurrency control protocol, called the \tDelta protocol, that does not impede performance adversely.
 That is because the protocol is exclusively designed for one purpose only: reciprocal consistency in distributed edges.
-Its design leverages the fact that the updating of end pointers of a given distributed edge immediately follow each other and the small interval between them is the window for possible conflicts between concurrent updates.
+Its design leverages the fact that the updating of end pointers of a distributed edge \emph{must} immediately follow each other and the small interval between them is the sole \emph{raison d'\^etre} for reciprocal concurrency violations.
 
-The paper is organized as follows: Section \ref{sec:recipr-cons} presents reciprocal consistency, Sections \ref{sec:distr-graph-datab}-\ref{sec:db-corruption} discuss the architecture of distributed graph databases and describes how corruption occurs.
-Section \ref{sec:tdelta-protocol} outlines the \tDelta protocol.
-Sections \ref{sec:modeling}-\ref{sec:evaluation} presents the model used to evaluate protocol performance.
+The paper is organized as follows: Section \ref{sec:recipr-cons} describes how edges in a graph database are stored and the notion of reciprocal consistency.
+This notion is examined in the context of distributed graph databases in Section \ref{sec:distr-graph-datab} and the root causes of consistency violations are explained in the following section.
+Section \ref{sec:tdelta-protocol} presents the \tDelta protocol together with correctness reasoning.
+Sections  \ref{sec:modeling} and \ref{sec:evaluation} are devoted to evaluating the protocol using two metrics that measuring the protocol's ability to avoid inconsistencies (ensuring \emph{safety}) and unnecessary aborts (enhancing \emph{throughput}).
+Section  \ref{sec:modeling} explains the strategies used for evaluating these metrics and the latter presents the results for various values chosen for the protocol parameter ($\Delta$).
+Section \ref{sec:Conclude} concludes the paper.

### 2020-04-10 paper/sec2-reciprocal-consistency.tex
camera-ready

--- 2020-04-06 paper/reciprocal-consistency.tex
@@
 
-In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
-In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
-This allows an edge to be traversed in both directions at the same cost, facilitating improved query performance.
+In the property graph data model, edges have direction and each edge runs from a \emph{source} vertex to a \emph{destination} vertex.
+In the storage layer, however, edge directionality does not exist;  \textbf{both} the source and the destination vertices store information about each other.
+This allows edge traversal to be bidirectional and speeds up query performance.
 
-A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}.
-Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
-It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
-
-Using openCypher \cite{openCypher} this can be represented by:
+Consider, for example, the statement: ``Tolkien \textit{wrote} The Hobbit''.
+It is expressed using vertex  \emph{a} for Tolkien and vertex \emph{b} for The Hobbit, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).
+Corresponding openCypher \cite{openCypher} code is given below and Fig \ref{rc-edge} shows the model level view.% of edge $ab$.
 \input{cypher/create-edge}
 
-Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}.
-Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
-When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
-A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.
+Fig \ref{db-rep} depicts the internal representation of a graph arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}.
+A vertex, such as \emph{a}, is represented by a record that contains one or more properties of that vertex, followed by a sequence of \emph{edge pointers} pointing to all those vertices to which this vertex is related either as a source or a destination.
+The sequence of edge pointers is also called the \emph{adjacency list}.
 
-\input{figures/reciprocally-consistent-edge}
+It can be seen in Fig \ref{db-rep} that \emph{a}'s adjacency list has an edge pointer entry that stores `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s list has a corresponding entry storing the reciprocal (or inverse) information `\emph{b} \emph{written} by \emph{a}'.
+When the adjacency list entries for a given edge refer to each other in a complementary manner like this, that edge is said to be \emph{reciprocally consistent}.
+
+Consider a query: `list all titles by the author who wrote The Hobbit'.
+This query needs to start from \emph{b} which represents the only entity specified explicitly in it.
+Thanks to the reciprocal information in \emph{b}, it can reach \emph{a} from \emph{b}, even though edge \emph{ab} is  ``directed'' from \emph{a} to \emph{b}, and then compile the necessary list from \emph{a}.
+Note that reciprocal consistency is assumed to be prevail when a query reads only the source or destination vertex of an edge.
+
+\input{figures/rc}

### 2020-04-10 paper/sec3-dist-graphdb.tex
camera-ready

--- 2020-04-06 paper/dist-graphdb.tex
@@
 
-A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
+A distributed graph database employs a shared-nothing architecture, partitioning a graph among loosely cooperating servers.
 Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
-The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
-Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges}, Figure \ref{dist-graph}.
-The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.
+The objective of such an approach is to minimize the proportion of edges that span partitions and also to balance the distribution of vertices to partitions.
+Fig \ref{dist-graph} depicts a graph database partitioned across 3 servers, $S_i, i = 1, 2$ and $3$.\footnote{Typically, each partition would be replicated for fault tolerance and availability, but these issues are beyond the scope of this paper.}
+
+Intra-partition and inter-partition edges are respectively referred to as \textit{local edges}  and  \textit{distributed edges} (shown using dashed lines in Fig \ref{dist-graph}).
+The proportion of distributed edges is not negligible and can range from 25-75\% \cite{Huang2016}.
 
@@
 Adjacency lists can now contain edge pointers to vertices on remote servers.
-Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}).
-Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}.
-Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.
+Guaranteeing reciprocal consistency for distributed edges is challenging, especially when an existing BASE database is used for storage and is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
+Opting for this design appears to be a good choice: it offers the application programmer the modeling convenience of graphs together with the operational characteristics of the underlying BASE database.
 
-However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database.
-BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
-This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}.
-Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
-This process is explained in Section \ref{sec:db-corruption}.
+However, a major problem with this design option is the lack of transactional semantics from the underlying BASE databases.
+The latter seldom provide guarantees for multi-operation, multi-object transactions that span partitions.
+Without concurrency control across partitions, concurrent updates of distributed edges can interleave and violate reciprocal consistency.\footnote{The concurrency control primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}
+Our earlier work showed that undermining reciprocal consistency in distributed edges can lead to irreversible corruption of the distributed database itself and the corruption rate can be worryingly high (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+Origins of reciprocal inconsistency in distributed edges are explained in Section \ref{sec:db-corruption}.
 
 \input{figures/conflict-scenarios}
-
-\input{figures/half-corrupted-edge}

### 2020-04-10 paper/sec4-corruption.tex
camera-ready

@@
-\section{Corruption in the Absence of Concurrency Control}
+\section{Reciprocally Inconsistent Distributed Edges}
 \label{sec:db-corruption}
 
-Consider the edge \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} spans servers $S_i$ and $S_j$.
-When a transaction writes a distributed edge there are two important facts to consider:
+Suppose that the edge \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} is a distributed edge, with vertices \emph{a} and \emph{b} in servers $S_i$ and $S_j , j\neq i$,  respectively.
+When a transaction writes this edge $ab$,
 \begin{enumerate}
-\item Two write operations are performed, writing reciprocal entries in the adjacency lists of \emph{a} and \emph{b}.
-\item Write order is unconstrained, a transaction is equally likely to write \emph{a} then \emph{b} as it is to write \emph{b} then \emph{a}.
+\item two writes are performed: reciprocal entries in the adjacency lists of both \emph{a} and \emph{b} are updated, and
+\item write order is unconstrained: a transaction is equally likely to write \emph{a} then \emph{b} as it is to write \emph{b} then \emph{a}.
 \end{enumerate}
 
-Without concurrency control, transactions $T_x$ and $T_y$ that write the distributed edge can interleave in the following ways:
-\begin{enumerate}[(i)]
-\item $T_x$ begins writing the distributed edge before $T_y$ but at opposite ends, transactions cross in the network and at $S_i$, $T_x \rightarrow T_y$ and at $S_y$, $T_y \rightarrow T_x$, Figure \ref{fig:a}.
-\item $T_y$ begins writing the distributed edge before $T_x$ but at opposite ends, transactions cross in the network and at $S_i$, $T_x \rightarrow T_y$ and at $S_j$, $T_y \rightarrow T_x$, Figure \ref{fig:b}.
-\item $T_x$ and $T_y$ perform their first writes at $S_i$, $T_x \rightarrow T_y$, but overlap in the network and arrive out-of-order at $S_j$, $T_y \rightarrow T_x$, Figure \ref{fig:c}.
+Concurrent transactions $T_x$ and $T_y$ can interleave in the following three ways and each one is depicted in Fig \ref{fig:conf}:
+\begin{enumerate}[(a)]
+\item $T_x$ starts before $T_y$; it writes $a$ at $S_i$ first and then proceeds to $S_j$ across the network; $T_y$ operates the other way round, beginning with $S_j$ and proceeding to $S_i$ (see Fig \ref{fig:a}).
+The net effect is:  $T_x \rightarrow T_y$ at $S_i$ and at  $T_y \rightarrow T_x$ at $S_j$, where  $T \rightarrow T'$ at $S$ denotes that $T$ \emph{precedes} $T'$ at server $S$.
+\item Same as the previous case, except that $T_y$ starts earlier than $T_x$ (see Fig \ref{fig:b}).
+\item Same net effect as in previous two cases, except that both $T_x$ and $T_y$ start their first writes at $S_i$, $T_x \rightarrow T_y$, but $T_y$ overtakes $T_x$ in reaching $S_j$ where $T_y \rightarrow T_x$ (see Fig \ref{fig:c}).
 \end{enumerate}
 
-Each interleaving in Figure \ref{fig:conf} results in a \emph{half-corrupted} distributed edge - reciprocal consistency has been violated.
-A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000})
-To illustrate the process of half-corruption, consider $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
+We could envisage three more corresponding cases (a') - (c') where the roles of $T_x$ and $T_y$ in cases (a) - (c) are simply interchanged; e.g., in case (a'), $T_y$ starts before $T_x$, writes $a$ at $S_i$ first and then proceeds to $S_j$ across the network; $T_x$ operates the other way round, beginning with $S_j$ and proceeding to $S_i$.
+Thus, there are only 6 ways concurrent $T_x$ and $T_y$ can interleave.
+The arguments we make based on cases (a) - (c) of Fig \ref{fig:conf} equally apply, by symmetry, to cases (a') - (c') and so, for brevity, we will not consider the latter.
+
+At the end of each case in Fig \ref{fig:conf}, the last update on $a$ is by $T_y$ and that on $b$ is by $T_x$.
+Unless updates of $T_x$ and $T_y$ are commutative, $ab$ cannot be reciprocally consistent.
+As an example, suppose that $T_x$ deletes the \emph{wrote} edge while $T_y$ concurrently appends a property \emph{year}:
 \input{cypher/delete-edge.tex}
 \input{figures/hc}
-Each interleaving results in the distributed edge in Figure \ref{hc-edge}, reciprocal consistency is clearly violated.
-We say a graph with half-corrupted edges has suffered \emph{structural corruption}.
+The interleaving patterns depicted in Fig \ref{fig:conf} leave $ab$ reciprocally inconsistent, as shown in Fig \ref{hc}.
 
-Assume we know the correct state of the distributed edge to be \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} with \texttt{\textcolor{red}{year} = 1937}, then there exists correct and incorrect entries for edge, visible in Figure \ref{hc-db-rep}.
-When a transaction reads an edge it does not check both entries that constitute the edge are reciprocal consistent.
-Therefore, a subsequent transactions can read the incorrect entry of a half-corrupted edge and write further edges, introducing \emph{semantic corruption} into the database.
-Further semantic corruption spreads by the same mechanism.
-A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.
+When $T_x$ and $T_y$ do not interleave,  either $T_y \rightarrow T_x$ or $T_x \rightarrow T_y$ holds at both servers $S_i$ and $S_j$; in that case, the last update on \emph{both} $a$ and $b$ would be by either $T_x$ or $T_y$, respectively.
+That is, when transactions update at both ends of $ab$  in \emph{some} arbitrarily chosen but identical order, $ab$ is left  reciprocally consistent after each transaction's update.
 
-For example, $T_z$, introduces semantic corruption if it reads \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.
+When interleaving updates leave $ab$ reciprocally inconsistent, $ab$ can be said to have become \emph{half-corrupted} because if $T_y \rightarrow T_x$ is the chosen order between $T_y$ and $T_x$, then the edge pointer in $b$ of $ab$ is in error; otherwise, the edge pointer in $a$ is erroneous.
+Thus, a reciprocally inconsistent edge $ab$ certainly has a corrupt half but the question of which half is corrupt has been left open.
 
-\input{cypher/spread.tex}
+Suppose that a future transaction $T_w$ \emph{first} reads the edge pointer of reciprocally in consistent $ab$, say, at vertex $a$.
+(Note that when $T_w$ reads any edge, it does not check for reciprocally consistency).
+At that moment, $T_w$ (implicitly) chooses the order $T_x \rightarrow T_y$ and thereby invalidates the other order $T_y \rightarrow T_x$ that prevails at vertex $b$.
+Thus, from that moment onward, the $b$ end of edge $ab$ becomes the corrupt end.
 
-If the database provides the ANSI isolation level \textbf{Read Uncommitted}, the database consistently orders writes from concurrent transactions, which would prevent all interleavings in Figure \ref{fig:conf} and the spread of semantic corruption
+If no transaction ever reads the edge pointer at vertex $b$, then the order $T_x \rightarrow T_y$ effectively prevails and the half-corruption of $ab$ remains invisible to the rest of the database.
+However, if $T_z$ is to subsequently read the edge pointer at $b$ and write another edge based on what it read, i.e., The Hobbit has unknown author, then it is introducing updates not consistent with what $T_w$ read earlier; it thus introduces \emph{semantic corruption} into the database.
+Further writes based on reading semantically corrupt data also spread corruption.
+
+A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupt state.
+As stated earlier, we had shown that 10\%  of a distributed graph database can become semantically corrupt well within the system lifetime itself (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
+
+Two more relevant remarks on past works:
+a half-corrupted edge is due to a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}) in the context of distributed graph databases.
+If the database provides the ANSI isolation level \emph{Read Uncommitted}, it will identically order the writes of concurrent transactions and this would prevent all interleaving patterns shown in Fig \ref{fig:conf} and thus avert half-corruption altogether.

### 2020-04-10 paper/sec5-delta.tex
camera-ready

--- 2020-04-06 paper/delta.tex
@@
 
-A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted.
-To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
+A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only after the acquiring transaction has committed or aborted.
+To prevent deadlock, a policy such as \texttt{NO_WAIT} deadlock avoidance is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.
 
@@
 
-These differences lead to several advantages in the context of graph databases.
-Firstly, a subset of edges are traversed and modified with a high frequency e.g.
-critical sections of motorway in a road network, leading to high contention.
-Secondly, graph transactions tend to be longer-lived than transactions in relational databases.
-Waiting for earlier writes on edges by long running transactions to commit, these frequently accessed edges significantly limit concurrency and reduce throughput.
-With these concerns in mind we developed the \tDelta protocol, which aimed at preventing edges becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}
+These differences lead to several advantages in the context of graph databases which are characterized by the following two aspects.
+First, a subset of edges are traversed and modified with a very high frequency e.g. critical sections of motorway in a road network, leading to high contention.
+Secondly, graph transactions tend to be longer-lived than those in other databases.
+So, having to wait for earlier writes to terminate while transactions are long lived, will severely limit the scope for concurrent processing and reduce throughput in a highly contentious environment.
+
+With these concerns in mind we developed the \tDelta protocol, which aims at preventing edges from becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is a concurrency control mechanism only for distributed edge updates; it does not concern itself with updates on  vertices.}
 
@@
 \begin{enumerate}
-\item A transaction's update on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
-\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative update was done at least $\Delta$ time before (where time is measured as per local clock).
-\item If a transaction successfully performs all its tentative updates, then it is permitted to commit; otherwise, it must abort.
-\item A transaction commits, when all its tentative updates are made permanent, e.g., using an atomic commitment protocol.
-\item Tentative updates of an aborting transactions are ignored.
-An ignored tentative update can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
+\item A transaction's write on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
+\item A tentative write is possible only if the end pointer is either in a committed state or the immediately preceding tentative write was done at least $\Delta$ time before (where time is measured as per local clock).
+\item If a transaction performs all its tentative writes, then it is permitted to commit; otherwise, it must abort.
+\item A transaction commits, when all its tentative writes are made permanent, e.g., by using an atomic commit protocol.
+\item Tentative writes of an aborting transactions are ignored.
+An ignored tentative write can make a new transaction abort unnecessarily for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
 \end{enumerate}
@@
 
-Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction updating one end of a distributed edge at one server and updating the other end of the same edge at another server.
+Let us define $\delta$ as the bound \emph{estimate} on the interval that may elapse between a transaction completing its update at one end of a distributed edge at one server and starting its update at the other end of the same edge at another server.
 
 Let $\Delta$ be chosen such that $\Delta > \delta$.
-Now consider the interleaving in Figure \ref{fig:tx-ty} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
-Note $t_y < t_x$, i.e.
-$(t_x - t_y) > 0$.
+Now consider the interleaving in Fig \ref{fig:b} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
+Since $T_x$ starts after $T_y$ in Fig \ref{fig:b}, $t_x > t_y$, i.e. $(t_x - t_y) > 0$.
 
-Say, $T_y$ reaches $S_i$ at time $t_y + d$, where $d$ is the actual time elapsed between completing a tentative write at one end and starting at another end, let us assume that $d \leq \delta$.
-When $T_y$ arrives at $S_i$ it will find a tentative write done at time $t_x$.
-In this case, $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge.
-A similar arguments can be made for the scenario in with Figure \ref{fig:ty-tx}.
-For the interleaving in Figure \ref{fig:same-side}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$.
+Say, $T_y$ reaches $S_i$ at time $t_y + d_y$, where $d_y$ is the \emph{actual} time elapsed between $T_y$ completing a tentative write at one end and starting at another end, let us assume that $d_y \leq \delta$.
+When $T_y$ arrives at $S_i$ it will find a tentative write already done at time $t_x$.
+In this case, $t_y + d_y - t_x = d_y - (t_x-t_y) < d_y \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge.
+Similar arguments can be made for the scenario in Fig \ref{fig:a}, if $d_x \leq \delta$ where $d_x$ is the \emph{actual} time elapsed between $T_x$ completing a tentative write at one end and starting at another end.
+Note that $t_y > t_x$ and $T_x$ will abort because it will find out, on reaching server $S_j$ for its next tentative write, that $t_x +d_x - t_y = d_x - (t_y-t_x) < d_x \leq \delta < \Delta$.
+Finally, for the interleaving in Fig \ref{fig:c}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$, if $d_x \leq \delta$ holds.
 
-Assume that $d > \Delta$, i.e., the estimate $\delta$ does not hold at this moment.
-In Figure \ref{fig:tx-ty} interfering updates are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$ otherwise, an edge becomes half-corrupted.
-This holds for the interleaving in \ref{fig:ty-tx}.
-Thus, in the extreme case $t_x = t_y$, reciprocal consistency is not guaranteed if $d > \Delta$ when the former exceeds its upper bound estimate $\delta$.
-For Figure \ref{fig:same-side}, if $d$ for $T_x$ is larger than $\Delta$, reciprocal consistency can be violated.
+Assume that $d_x > \delta$ or $d_y > \delta$ is possible, i.e., the estimate $\delta$ can fail to hold.
+In Fig \ref{fig:b}, interleaving  writes are avoided only if $t_y + d_y - t_x = d_y - (t_x-t_y) < d_y < \Delta$; otherwise, reciprocal inconsistency can occur.
+Similarly, interleaving  writes of Fig \ref{fig:a} are avoided only if $t_x + d_x - t_y = d_x - (t_y-t_x) < d_x < \Delta$; otherwise, reciprocal inconsistency can occur.
+Thus, in the extreme case $t_x = t_y$ and reciprocal consistency is not guaranteed if $\Delta \leq $ max $\{d_x , d_y \}$.
 
-This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency.
-The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.
+For Fig \ref{fig:c}, interleaving  writes are avoided only if $(t_x + d_x) - (t_y + d_y)  < \Delta$; given that $(t_y - t_x) < \Delta$, when $d_x$ exceeds $ \delta$, the interleaving writes of Fig \ref{fig:c} are avoided only if $d_x \leq 2 \Delta$; otherwise, reciprocal inconsistency can occur.
 
-In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur.
-Two questions naturally arise regards the performance of the protocol.
-For different values of $\Delta$:
+In summary, the \tDelta protocol eliminates interleaving of transactions during edge writes, so long as $\Delta$ remains larger than the interval $d$ that elapses between a transaction completing its write at one end of a distributed edge and starting at the other end.
+Since the exact value of $d$ taken by a transaction cannot be known in advance, its bound $\delta$ is estimated with the best effort and $\Delta$ is chosen to be $\Delta > \delta$.
+
+The larger the value of $\Delta$ used, the more likely is that $\Delta > d$ holds and half-corruption and thereby operational corruption are averted; also, on the downside, the more likely is that non-interleaving transactions will find their tentative writes within $\Delta$ time of each other and the later ones choose to abort \emph{unnecessarily}.
+In the extreme case, choosing a very large $\Delta$ ($\Delta \approx \infty $) totally eliminates any risk of reciprocal inconsistency but does not allow any tentative write until the preceding one is made permanent.
+It is equivalent to enforcing \texttt{NO_WAIT} policy, wherein a requesting transaction that finds the requested record being locked, must abort.
+
+Our performance evaluation will therefore  measure the following two metrics for various values of $\Delta$:
 \begin{itemize}
-\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
-\item How many transactions are aborted as a result of the protocol?
+\item Time taken for 10\% of a large database to be corrupt,
+\item Number of transactions aborted per second.
 \end{itemize}
-
-%The two paragraphs (The protocol leverages... edge is lost) can be commented out.
-
-%Figures 5 and 6 can be removed.
-
-% The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end.
-It is then assumed that the network delay between two servers can be predicted to be some $\Delta$.
-All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed.
-From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time.
-Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$.
-An aborted transaction, aborts any and every previous tentative write that it may have successfully completed.
-This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.
-
-% \begin{figure}[H]
-%   \centering
-%   \begin{tikzpicture}[node distance=2cm]
-%     % s1
-%     \draw [thick] (1.3,3) -- (1.7,3);
-%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-%     % s2
-%     \draw [thick] (3.3,3) -- (3.7,3);
-%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-
-%     \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
-%     \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);
-
-%     \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};
-
-%   \end{tikzpicture}
-%   \caption{An example of the \tDelta protocol preserving reciprocal consistency.}
-%   \label{delta-abort}
-% \end{figure}
-
-% Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption.
-Figure \ref{corruption-again} displays an example when the \tDelta protocol fails to maintain reciprocal consistency.
-The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.
-
-% \begin{figure}[H]
-%   \centering
-%   \begin{tikzpicture}[node distance=2cm]
-%     \draw [thick] (1.3,3) -- (1.7,3);
-%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
-%     \draw [thick] (3.3,3) -- (3.7,3);
-%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
-%         \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
-%     \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};
-
-%     \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
-%     \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
-%   \end{tikzpicture}
-%   \caption{An example of the \tDelta protocol failing to preserve reciprocal consistency.}
-%   \label{corruption-again}
-% \end{figure}
+If $d$ is exponentially distributed with mean $1/ \mu$, the probability of $d$ exceeding $\Delta$ is $ e^{(-\mu \Delta)}$.
+The values of $\Delta$ chosen will explore a range of probabilities of $\Delta$ being exceeded.

### 2020-04-10 paper/sec6-modeling.tex
camera-ready

--- 2020-04-06 paper/modeling.tex
@@
-\section{Modeling}
+\section{Performance Evaluation Strategies}
 \label{sec:modeling}
 
-To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}.
-A summary of the model is now provided, before discussing the extensions.
+To assess the time to operational corruption, we reused the model that we developed in \cite{Ezhilchelvan2018} and adapted it for the \tDelta protocol.
+For completeness, the model is briefly explained first before we discuss the adaptation.
+We note that an empirical evaluation of the time to operational corruption using a real system would be impractical due to the sheer length of time and the cost it would take to run the experiment.
+(For certain values of $\Delta$, the model predicts anywhere between 1-75 years for operational corruption! Details are to follow.)
 
-The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second.
-To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write.
-To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$.
-For each type, a fraction $f$ are distributed edges and the remainder are local edges.
-The network delay between servers was assumed to be exponentially distributed with rate $\rho$.
-At any moment in time an edge can be in one of four states:
+The model of \cite{Ezhilchelvan2018} assumes that transactions arrive in a Poisson stream with rate $\lambda$ transactions per second (TPS, for short).
+Each transaction performs a random number ($K$) of read operations and then updates a single edge.
+
+To model a scale-free graph, edges in the distributed database are divided into $T$ types, $i = 1, 2, \ldots, T$.
+Type-1 edges represent the most popular edges to be accessed by transactions and type-$T$ edges are the least popular. Popular edges are smaller in number compared to the less popular ones.
+$N$ denotes the total number of edges in the database.
+For all edge types, a fraction $f$ are distributed.
+
+At time $0$, all edges are assumed to be clean (free from corruption); thereafter, an edge can be in one of four states:
 \begin{enumerate}
@@
 \item Distributed and clean.
-\item Half-corrupted distributed edge arising from a dirty write.
-\item Semantically corrupted.
+\item Distributed and half-corrupted.
+\item Local or distributed and semantically corrupted.
 \end{enumerate}
-% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.
 
-% \begin{figure}[ht]
-%   \centering
-%     \begin{tikzpicture}[node distance=2cm]
-%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
-%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
-%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
-%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
-%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
-%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
-%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
-%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
-%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
-%     \end{tikzpicture}
-%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
-%     \label{state-transitions}
-% \end{figure}
+The delay $d$ is the time interval that elapses between a transaction completing its tentative write at one end of a distributed edge and starting at another end.
+It constitutes the window for concurrent transactions to interleave as portrayed in Fig \ref{fig:conf}.
+It is assumed to be exponentially distributed with mean $ 1/ \mu$ (i.e., at rate $\mu$).
 
-Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers.
-Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived.
-These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$.
-At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt.
-The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
+The model parameters now enable us to compute the probability $q_i$ that a transaction updating a clean type-$i$ distributed edge, leaves it half-corrupted, in the absence of any concurrency control mechanism.
+(Our earlier work in \cite{Ezhilchelvan2018} assumes no concurrency control).
+These probabilities, $q_i, 1 \leq i \leq T$, are then used to compute transition rates $a_{j,k}$ between state $j$ to $k$, $1 \leq j,k \leq 4$.
+(Since local and distributed edges are fixed, $a_{1,2} = a_{2,1} = 0$.)
 
-% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e.
-the interleaving in Figure 3(c) cannot occur.}.
+The rates $a_{j,k}$ are in turn used to simulate the spread of corruption within the database to estimate the first passage time $U_{\gamma}$  for a fraction ${\gamma}$ of $N$ edges to enter the state 4.
+(When $\gamma N$ edges become semantically corrupt, the database becomes operationally corrupt.)  The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.
 
-The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$.
-Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol?
-Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed.
-The answers to these questions depends on several parameters characterizing the following systemic aspects:
+\paragraph{\tDelta Protocol Adaptation}
+Our concurrency protocol reduces the probabilities $q_i, 1 \leq i \leq T$ and we compute $q^{new}_i$ as shown in Appendix \ref{sec:deriv-confl-prob}.
+The new probabilities $q^{new}_i$ are used in the model and simulations of \cite{Ezhilchelvan2018} to estimate $U_{\gamma}$.
+
+\paragraph{Number of aborts per second.} To evaluate this metric for various values of $\Delta$, a second simulation that focuses specifically on the subset of most frequently accessed distributed edges was performed.
+
+Note that both metrics that we set out to evaluate will be influenced by several parameters that characterize the database and other aspects:
 \begin{itemize}

### 2020-04-10 paper/sec7-evaluation.tex
camera-ready

@@
 
-The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
-The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
-In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
+The following parameter choices are inline in with industry experiences.
+The graph analyzed consisted of seven edge types, $n_1=10^4, n_2=10^5, n_3=10^6, n_4=10^7,  n_5=10^8, n_6=10^9, n_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
+The number of read operations before a write per query is geometrically distributed starting at $2$, with an average of $15$.
+In all edge types, $f=0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
+
 The delay $d$ between a transaction completing a tentative write at one end and starting at another end is exponential distributed with a mean of $5ms$.
-The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
-The time taken until operational corruption $U$, is measured in days.
-$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
+The database is initially clean and considered to be operationally corrupted when $10$\% ($\gamma = 0.1$) of all edges are semantically corrupted.
+The time taken until operational corruption, $U$, is measured in days.
+We consider a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload $(10000, ..., 100000)$.
 The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
-For each $\Delta$ the probability that $d$ exceeds $\Delta$ is $P(D > \Delta) = 4.5 \time 10^{-5}, 3.1 \times 10^{-7}, 2.1 \times 10^{-9}$ respectively.
+For each $\Delta$ the probability that $d$ exceeds $\Delta$ is $P(d > \Delta) = 4.5 \time 10^{-5}, 3.1 \times 10^{-7}, 2.1 \times 10^{-9}$ respectively.
 
-The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
-Under no isolation, $U$ ranges between 50-500 days.
-For $50ms$ this increases to 1-75 years.
-For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.
+The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Fig \ref{time-to-corruption-results} (where the $y$-axis in log scale).
+With no concurrency control, $U$ ranges between 50-500 days.
+For $\Delta = 50ms$,  $U$ increases to 1-75 years. For $\Delta = 75, 100ms$ the time to corruption is significantly large as shown Fig \ref{time-to-corruption-results}.
 
-\begin{figure}[h]
+To evaluate the number $\alpha$ of aborts occurred per second, simulations were run for 10 seconds for each arrival rate, $\lambda = (1000, ..., 10000)$.
+Fig \ref{aborts-results} reports the fraction $\frac{\alpha }{\lambda}$ for various values of $\lambda$.
+This fraction is also the probability that an incoming transaction is aborted due to the $\tDelta$ protocol.
+For $\Delta = 50$ ms, the abort probability is between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
+
+\begin{figure}[htbp]
   \centering
@@
 
-The number of aborts per second for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$.
-The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$.
-For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.
-
-\begin{figure}[h]
+\begin{figure}[htbp]
   \centering
   \includegraphics[width=\linewidth]{./figures/aborts}
-  \caption{Number of aborts per second.}
+  \caption{Fraction of aborts.}
   \label{aborts-results}

### 2020-04-10 paper/sec8-conclusion.tex
camera-ready

\section{Conclusions}\label{sec:Conclude}

Database concurrency control has been a long researched area, however to the best of our knowledge this is first attempt at developing a protocol specific for distributed graph databases.
We presented a lightweight protocol for providing reciprocal consistency and mitigating the problem of high contention in a distributed graph database.

The \tDelta protocol leverages the fact that writes to distributed edges always consist of two sequential writes to entries in the adjacency lists of vertices the edge connects.
Since it is concerned only with edges (and not vertices) in a graph, it provides guarantees weaker than Read Uncommitted isolation (the weakest ANSI isolation level).

The protocol that is presented and performance-evaluated here, we believe,  valuable in practice given the popularity of BASE distributed graph databases and the rate at which semantic corruption can spread if reciprocal consistency is left unchecked.
Simulations indicate that when $\Delta$ values are chosen to be reasonably large, the protocol rules out corruption resulting from half-corrupted distributed edges while keeping the abort rate considerably small.

For future work, we intend on implementing the protocol to assess the validity of the simulations and to enable comparisons with traditional lock-based approaches.
Moreover, we plan on investigating the suitability of higher isolation levels in a distributed graph database, \textbf{Read Atomic} isolation \cite{Bailis2014} seems particularly well suited.

### 2020-04-10 scripts/make-plots.R

@@ -4,45 +4,38 @@ library(ggthemes)
 
 # aborts/s 
 ms50 = read.csv("./results/aborts_50.csv")
-ms50$collisions = (ms50$collisions/10)
+ms50$collisions = (ms50$collisions/ms50$arrivals)
 delta = rep(50,11)
 ms50 = cbind(ms50,delta)
 
-
 ms75 = read.csv("./results/aborts_75.csv")
-ms75$collisions = (ms75$collisions/10)
+ms75$collisions = (ms75$collisions/ms75$arrivals)
 delta = rep(75,11)
 ms75 = cbind(ms75,delta)
 
-
 ms100 = read.csv("./results/aborts_100.csv")
-ms100$collisions = (ms100$collisions/10)
+ms100$collisions = (ms100$collisions/ms100$arrivals)
 delta = rep(100,11)
 ms100 = cbind(ms100,delta)
-
-max(ms100$collisions)
 
 aborts = rbind(ms50,ms75,ms100)
 aborts$delta = as.factor(aborts$delta)
 
 a=ggplot(data=aborts, aes(x=tps,y=collisions,linetype=delta,color=delta)) +
     geom_line() +
-     scale_linetype_discrete(name="Delta (ms): ",labels=unname(TeX(c("50","75","100")))) +
-     scale_color_discrete(name="Delta (ms): ",labels=unname(TeX(c("50","75","100")))) +
-     scale_x_continuous(name="TPS",breaks=seq(0,10000,1000)) +
-     scale_y_continuous(name="Aborts/s",breaks=seq(0,1000,100)) + 
+     scale_linetype_discrete(name=TeX('$\\Delta$ (ms):'),labels=unname(TeX(c("50","75","100")))) +
+     scale_color_discrete(name=TeX('$\\Delta$ (ms):'),labels=unname(TeX(c("50","75","100")))) +
+     scale_x_continuous(name=TeX('$\\lambda$ (TPS)'),breaks=seq(0,10000,1000)) +
+     scale_y_continuous(name="Abort probability",breaks=seq(0,0.1,0.01)) + 
     theme_few()
 a
 
-ggsave(a, filename = "aborts.pdf",  bg = "transparent")
+ggsave(a, filename = "./paper/figures/aborts.pdf",  bg = "transparent",width = 6.5, height = 4.5)
 
-
-# load time to corruption data in 
+# corruption
[…]

### 2020-04-16 slides/figures/graph.tex
copy epew slides as base

\begin{figure}[h!]
  \centering
  \resizebox{0.6\textwidth}{!}{
    \begin{tikzpicture}[node distance=2cm]
      \node (v1) [vertex] {};
      \node (v2) [vertex, above of=v1,xshift=0.3cm] {};
      \node (v3) [vertex, above right of=v1,yshift=-0.3cm] {};
      \node (v4) [vertex, right of=v3,yshift=-0.3cm] {};
      \node (v5) [vertex, above left of=v4,xshift=0.7cm] {};
      \node (v6) [vertex, right of=v5,xshift=-0.8cm,yshift=1.2cm] {};
      \node (v7) [vertex, below right of=v6, yshift=-0.8cm] {};
      \node (v8) [vertex, below of=v7, yshift=0.3cm,xshift=0.3cm] {};
      \node (v9) [vertex, right of=v7,xshift=-0.4cm,yshift=0cm] {};
      \node (v10) [vertex, above left of=v9,xshift=0.4cm,yshift=-0.4cm] {};
      \node (v11) [vertex, above right of=v10,yshift=-2.5cm,xshift=1cm] {};
      \node (v12) [vertex, above of=v10,yshift=-0.6cm,xshift=1cm] {};
      \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
      \draw [local] (v2) -- (v1);
      \draw [local] (v1) -- (v3);
      \draw [local] (v3) -- (v2);
      \draw [local] (v2) -- (v5);
      \draw [local] (v1) -- (v4);
      \draw [local] (v5) -- (v4);
      \draw [local] (v5) -- (v6);
      \draw [local] (v4) -- (v7);
      \draw [local] (v8) -- (v4);
      \draw [local] (v8) -- (v9);
      \draw [local] (v9) -- (v11);
      \draw [local] (v10) -- (v7);
      \draw [local] (v10) -- (v12);
      \draw [local] (v12) -- (v13);
      \draw [local] (v10) -- (v6);
      \draw [local] (v10) -- (v11);
      \draw [local] (v13) -- (v10);
      \draw [local] (v11) -- (v12);
    \end{tikzpicture}
  }
  \caption{Connected graph}
\end{figure}

### 2020-04-16 slides/main.tex
copy epew slides as base

% ----------------------------------------------------------------------------------------
%	Configuration
%----------------------------------------------------------------------------------------

\documentclass[aspectratio=169]{beamer}

\usetheme[titleformat=regular,numbering=none,progressbar=frametitle]{metropolis}

% newcastle colour scheme
\definecolor{nclBlue}{HTML}{003A65}
\definecolor{nclRed}{HTML}{D91A35}

\setbeamercovered{transparent}
\setbeamercovered{invisible}
\setbeamercolor{progress bar}{fg=nclRed}
\setbeamercolor{frametitle}{bg=nclBlue}
\setbeamercolor{background canvas}{bg=white}
\setbeamercolor{block body}{bg=gray!30}
\setbeamercolor{block title}{bg=nclBlue,fg=white}
\setbeamercolor{palette primary}{fg=white, bg=nclBlue}

\linespread{1.5}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{mathspec}
\usepackage[font=footnotesize]{caption}

% Thin fonts
\setsansfont[BoldFont={Fira Sans}, Numbers={OldStyle}]{Fira Sans Light}
\setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Light}

% Thick fonts
% \setsansfont[BoldFont={Fira Sans SemiBold}, Numbers={OldStyle}]{Fira Sans}
% \setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Book}

% tikz configuration
\usetikzlibrary{shapes.geometric, arrows}
\usetikzlibrary{arrows, positioning, shapes}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{calc}

\tikzstyle{vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
\tikzstyle{mini_vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
\tikzstyle{server} = [rectangle, rounded corners,minimum width=3cm,minimum height=4cm,draw=black]
\tikzstyle{mini_server} = [rectangle, rounded corners,minimum width=1.5cm,minimum height=1.9cm,draw=black]
\tikzstyle{local} = [thick,->,>=stealth]
\tikzstyle{dist} = [thick,dashed,->,>=stealth]
\tikzstyle{bignode}=[draw, circle, minimum width=70pt,align=center]

%----------------------------------------------------------------------------------------
%	Title + Overview Slide
% ----------------------------------------------------------------------------------------

\title[]{Preserving Reciprocal Consistency in \\ Distributed Graph Databases}

\author[Waudby \& Ezhilchelvan \& Webber \& Mitrani]{\underline{\emph{Jack Waudby}}$^{1}$, Paul Ezhilchelvan$^{1}$, Jim Webber$^{2}$ \& Isi Mitrani$^{1}$}
\institute[NCL]
{
  $^{1}$Newcastle University\\
  $^{2}$Neo4j
}
\date{April 27, 2020}

\begin{document}

\begin{frame}
\maketitle
\begin{tikzpicture}[overlay, remember picture]
  \node[above left=2.8cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_newcastle_uni.png}};
  \node[above left=1.2cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_neo4j.png}};
\end{tikzpicture}
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
% Background
% ------------------------------------------------

\setcounter{framenumber}{0}
\metroset{numbering=counter}

\begin{frame}
  \frametitle{Graph Databases}
    \begin{figure}
      \includegraphics[scale=0.5]{images/property_graph.png}
      \caption{Labeled property graph\footnotemark}
    \end{figure}
  \footnotetext[1]{
    \url{https://neo4j.com}
  }
\end{frame}

\begin{frame}
  \frametitle{Graph Databases}
  \begin{itemize}
    \uncover<1->{
    \item \textbf{Efficient graph analysis:} reachability, pattern matching, shortest path search and clustering
    }
    \uncover<2->{
    \item \textbf{Use cases:} telecommunications, pharma, publishing, finance, social media
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Graph Database Popularity}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.3]{images/db_popularity.png}
    \caption{Database popularity by data model\footnotemark}
  \end{figure}
  \footnotetext[1]{
    \url{https://db-engines.com/en/ranking_categories}
  }
\end{frame}

\begin{frame}
  \frametitle{Distributed Graph Databases}
  \begin{itemize}
    \uncover<1->{
    \item Graph exceeds the storage capacity of a single machine
    }
    \uncover<2->{
    \item Partition the graph across multiple machines in a cluster, with partitions replicated for fault-tolerance and availability
    }
    \uncover<3->{
    \item Graph partitioning inevitably leads to \textbf{distributed edges}
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Distributed Graph Databases: Partitioning}
\input{./figures/graph}
\end{frame}

\begin{frame}
  \frametitle{Distributed Graph Databases: Partitioning}
  \begin{figure}[h!]
    \centering
    \resizebox{0.6\textwidth}{!}{
    \begin{tikzpicture}[node distance=2cm]
      \node (s1) [server,label=$S_1$] {};
      \node (s2) [server, right of=s1,xshift=2cm,yshift=2cm,label=$S_2$] {};
      \node (s3) [server, right of=s1,xshift=6cm,label=$S_3$] {};

      \node (v1) [vertex,xshift=-0.5cm,yshift=-1.5cm] {};
      \node (v2) [vertex, above of=v1,xshift=-0.3cm,yshift=1cm] {};
      \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
      \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.3cm] {};
      \node (v5) [vertex, above right of=v3,xshift=-0.6cm,yshift=-0.5cm] {};
      \node (v6) [vertex, right of=v5,xshift=0.3cm,yshift=0.8cm] {};
      \node (v7) [vertex, below of=v6, xshift=1cm,yshift=0.4cm] {};
      \node (v8) [vertex, right of=v4, yshift=-0.3cm,xshift=4.5cm] {};
      \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
      \node (v10) [vertex, above of=v9,xshift=-2.5cm,yshift=-0.3cm] {};
      \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
      \node (v12) [vertex, above of=v11,yshift=-1cm,xshift=-1cm] {};
      \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
      \draw [local] (v2) -- (v1);
      \draw [local] (v1) -- (v3);
      \draw [local] (v3) -- (v2);
      \draw [local] (v2) -- (v5);
      \draw [local] (v1) -- (v4);
      \draw [local] (v5) -- (v4);
      \draw [dist] (v5) -- (v6);
      \draw [dist] (v4) -- (v7);
      \draw [dist] (v8) -- (v4);
      \draw [local] (v8) -- (v9);
      \draw [local] (v9) -- (v11);
      \draw [local] (v10) -- (v7);
      \draw [dist] (v10) -- (v12);
      \draw [dist] (v12) -- (v13);
      \draw [local] (v10) -- (v6);
      \draw [dist] (v10) -- (v11);
      \draw [local] (v13) -- (v10);
      \draw [local] (v11) -- (v12);
    \end{tikzpicture}
  }
\caption{Graph partitioned across 3 servers}
  \end{figure}
\end{frame}

%------------------------------------------------
% Problem 1: Reciprocal Consistency}
% ------------------------------------------------

\begin{frame}
  \frametitle{Reciprocal Consistency}
  \begin{center}
  \textbf{A logical edge is represented by 2 physical records}
  \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}[node distance=7cm]
      \uncover<1-> {
      \node (n1) [bignode] {\texttt{Flight}};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:AVAILABLE}} (n2);
    }
      \uncover<2-> {
         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Seat Available In Flight}};
         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Flight Has Seat Available}};
         \draw [<->,dashed](q1) -- (q2);
      }
    \end{tikzpicture}
    \caption{A reciprocal consistent edge}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Reciprocal Consistency Violation}
  Two concurrent transactions:
  \begin{itemize}
  \item Mr Red requests to book the seat, $T_R$
  \item Mr Blue requests to book the seat, $T_B$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Reciprocal Consistency Violation}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}[node distance=2cm]
      \uncover<1-> {
        \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
        \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
        \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
      }
      \uncover<2-3> {
        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
      }
      \uncover<3-4> {
        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
      }
      \uncover<4-> {
        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
      }
      \uncover<5-> {
        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
      }
    \end{tikzpicture}
    \caption{Concurrent transactions interleave and violate reciprocal consistency}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Reciprocal Consistency Violation}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}[node distance=7cm]
      \node (n1) [bignode] {\texttt{Flight}};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);

      \uncover<2-> {
      \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
      \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
      \draw [<->,dashed](q1) -- (q2);
      \draw [<->,dashed](q1) -- (q2);
      \coordinate (a) at (4,-2);
      \coordinate (b) at (3,-3);
      \coordinate (c) at (3,-2);
      \coordinate (d) at (4,-3);
      \draw [very thick,red] ($ (a) $) -- ($ (b) $);
      \draw [very thick,red] ($ (c) $) -- ($ (d) $);
    }
    \end{tikzpicture}
    \caption{Reciprocal inconsistent distributed edge}
  \end{figure}
\end{frame}

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \uncover<1->{
%         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
%         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
%         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
%         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
%         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
%       }
%       \uncover<2->{
%         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
%         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
%         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
%         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
%         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
%       }
%     \end{tikzpicture}
%     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
%   \end{figure}
% \end{frame}

\begin{frame}
  \frametitle{Distributed Edge Reciprocal Consistency}
  \begin{itemize}
    \uncover<1->{
  \item Without concurrency control, distributed edges can become reciprocally inconsistent
    \begin{itemize}
    \item A distributed edge's reciprocal information is separated by the network
    \end{itemize}
  }
  \uncover<2->{
  \item Reciprocal inconsistency is a source of corruption\footnotemark
  }
  \uncover<3->{
  \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
    }
    \end{itemize}
    \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
\end{frame}

\begin{frame}
  \frametitle{Our Solution: Collision Detection}
  \begin{itemize}
    \uncover<1->{
    \item Carefully abort transaction(s)
    }
    \uncover<2->{
    \item No centralized control or synchronized clock
    }
    \uncover<3->{
    \item Permits interleavings that preserve reciprocal consistency
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Collision Detection General Rules}
  \begin{itemize}
    \uncover<1->{
    \item Updates are initially provisional
    }
    \uncover<2->{
    \item Transactions distinguish between their first and second updates using labels
    }
    \uncover<3->{
    \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Interference-free Update}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<1->{
        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
      }
      \uncover<2->{
        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
      }
      \uncover<3->{
        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
      }
      \uncover<3->{
        \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
      }
      \uncover<4->{
        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
      }
    \end{tikzpicture}
  \end{figure}

\end{frame}

\begin{frame}
  \frametitle{Collision Detection}
  \begin{center}
    \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
  \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<2->{
        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
      }
      \uncover<2->{
        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
      }
      \uncover<3->{
        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
      }
      \uncover<4-11>{
        \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
      }
      \uncover<5-10>{
        \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
      }
      \uncover<6-10>{
        \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
      }
      \uncover<7-7>{
        \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
      }
      \uncover<8-10>{
        \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
        \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
      }
      \uncover<9-9>{
        \node [] at (6,2) {$T_R$'s 2 is missing};
      }
      \uncover<10-10>{
        \node[] at (6,2) {$T_B$ ABORTS};
      }
      \uncover<12->{
        \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
      }
      \uncover<13->{
        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
      }
      \uncover<14->{
        \node[] at (6,2) {$T_R$ SUCCEEDS};
        }
    \end{tikzpicture}
  \end{figure}
\end{frame}

%------------------------------------------------
%Problem 2: Edge-Order Consistency
% ------------------------------------------------

\begin{frame}
  \frametitle{Edge-Order Consistency}
  A second consistency problem related to distributed edges:
  \begin{itemize}
    \uncover<2->{
    \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
    }
    \uncover<3->{
    \item Each update is reciprocally consistent
    }
    \uncover<4->{
    \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Edge-Order Consistency Violation}
  Consider two edges,
  \begin{figure}[h!]
    \centering
    \resizebox{1\textwidth}{!}{
      \begin{tikzpicture}[node distance=5.2cm]
        \node[] at (2.6,2.5) {\large{$E$}};
        \node[] at (13,2.5) {\large{$E'$}};
        \node (n1) [bignode] {\texttt{Flight}};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
        \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
         \draw [<->,dashed](q1) -- (q2);

        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
        \node (n4) [bignode, right of=n3] {\texttt{Room}};
        \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
        \draw [<->,dashed](q3) -- (q4);

      \end{tikzpicture}
    }
    \caption{Two distributed edges}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Edge-Order Consistency Violation}
  Two concurrent transactions by a travel agent:
  \begin{itemize}
  \item Requests to book room and seat for Mr Red, $T_R$
  \item Requests to book room and seat for Mr Blue, $T_B$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Edge-Order Consistency Violation}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<1->{

        \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
        \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
        \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
        \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
        \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
        \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
        \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
      }
      \uncover<2->{
        \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
      }
      \uncover<3->{
        \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
      }
      \uncover<4->{

        \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
      }
      \uncover<5->{

        \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
      }
      \uncover<6->{
        \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
        \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
        \draw (8.3,1) rectangle (11,2.5);
      }
    \end{tikzpicture}
    \caption{Edge-order consistency violation}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Edge-Order Consistency Violation}
  \begin{figure}[h!]
    \centering
    \resizebox{1\textwidth}{!}{
      \begin{tikzpicture}[node distance=5cm]
        % edge labels
        \node[] at (2.5,3) {\large{$E$}};
        \node[] at (12.5,3) {\large{$E'$}};
        % edge 1: flight -> seat
        \node (n1) [bignode] {\texttt{Flight}};
        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
        % edge 1: servers
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5cm,label=above:{$S_2$}] {};
        % edge 1
        \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
        % edge 1 information
        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
        \draw [<->,dashed](q1) -- (q2);

        % edge 2: hotel -> room
        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
        \node (n4) [bignode, right of=n3] {\texttt{Room}};
        \draw [local] (n3) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n4);
        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15cm,label=above:{$S_4$}] {};
        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10cm,label=above:{$S_3$}] {};
        \draw [<->,dashed](q3) -- (q4);

      \end{tikzpicture}
    }
    \caption{Edge-order consistency violation}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Our Solution: Order Arbitration}
  \begin{itemize}
  \uncover<1->{
  \item Transactions collect predecessors from each update
  }
  \uncover<2->{
  \item Centralized arbiter maintains some global state
  }
  \uncover<3->{
  \item Global state used to detect edge-order violations
  }
  \uncover<4->{
  \item Offending transactions are aborted
  }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Order Arbitration}
    \begin{center}
      \textbf{Rule: Collect predecessors after every successful update}
    \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<1->{
        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
        \draw [black,<-,>=stealth] (3,0) -- (3,4) ;
        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (3,4.5) node [label=above:{$S_2$}] {};
        \draw [black,<-,>=stealth] (7,0) -- (7,4) ;
        \draw [black,<-,>=stealth] (10,0) -- (10,4) ;
        \draw [black,dashed,->,>=stealth] (7,4.5) node [label=above:{$S_3$}] {} -- node [midway,above] {$E'$} (10,4.5) node [label=above:{$S_4$}] {};
      }
      \uncover<2->{
        \draw [blue,->,>=stealth] (0,3.2) node (TXS) [label=left:{\small{$T_B$}}] {} -- (3,3);
        \draw [red,->,>=stealth] (3,3.6) node (TXS) [label=right:{\small{$T_R$}}] {} -- (0,3.4);
        \draw (3.1, 2.8) node [label=right:{\small{\textcolor{black}{$T_B : \{{T_R}\}$}}}] {};
        \draw (3.2,2.3) rectangle (5,3.2);
      }
      \uncover<3->{
        \draw [blue,->,>=stealth] (7,2.3) node [label=left:{\small{$T_B$}}] {} -- (10,1.8);
        \draw [red,->,>=stealth] (7,1.8) node [label=left:{\small{$T_R$}}] {} -- (10,1.2) ;
        \draw (10.1, 0.8) node (TXS) [yshift=-1cm,label=right:{\small{\textcolor{black}{$T_R : \{{T_B}\}$}}}] {};
        \draw (10.2,-0.6) rectangle (12,0.3);
      }
      \uncover<4->{
        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
        (10.2,1.8) node[above,xshift=1cm, yshift=0.3cm] {ARBITER} -- (11,1.8);
        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
        (10.2,1.2)   -- (11,1.2);
      }
    \end{tikzpicture}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Order Arbitration}
      \begin{center}
      \textbf{Rule: If transaction exists in hit list, abort. Else, merge predecessors into hit list}
    \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<1->{ % 1 arbiter
        \node (s1) [server,label=$Arbiter$] {};
      }
      \uncover<2->{ % 2 hit list
        \node (a) [rectangle, label=above:{$Hit \ List$}, minimum width=2cm,minimum height=1cm,draw=black,xshift=0cm,yshift=0cm] {$T_z$ ,$T_a$ {\uncover<7->{,$T_R$}}};
      } % uncover T_R when T_B processed
      \uncover<3->{ % 3 queue
              \node (q) [rectangle, label=above:{$Queue$},left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=0cm] {{\uncover<6->{$...$}}};
              \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=0cm] {{\uncover<5-8>{$T_R$}}};
        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=0cm] {{\uncover<4-5>{$T_B$}}};
        \node (q1) [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=-1cm] {{\uncover<6->{$...$}}};
        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=-1cm] {{\uncover<5-8>{$T_B$}}};
        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=-1cm] {{\uncover<4-5>{$T_R$}}};
        \node [left of=q,xshift=0.2cm,label=left:{\small{\textcolor{black}{$tnx.id$}}}] {};
        \node [left of=q1,xshift=0.2cm,label=left:{\small{\textcolor{black}{$pred$}}}] {};
      }
      \uncover<4-4> {
        \node[right=0.5cm of s1] {$T_B$ \textbf{ENTERS QUEUE}};
      }
      \uncover<5-5> {
        \node[right=0.5cm of s1] {$T_R$ \textbf{ENTERS QUEUE}};
      }
      \uncover<6-8> {
        \node[right=0.5cm of s1,yshift=1cm] {$T_B$ \textbf{PROCESSED}};
      }
      \uncover<7-8> {
        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{ADDED TO HIT LIST}};
      }
      \uncover<8-8> {
        \node[right=0.5cm of s1,,yshift=-1cm] {$T_B$ \textbf{SUCCEEDS}};
      }
      \uncover<9-> {
        \node[right=0.5cm of s1,yshift=1cm] {$T_R$ \textbf{PROCESSED}};
      }
      \uncover<10-> {
        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{IN HIT LIST}};
      }
      \uncover<11-> {
        \node[right=0.5cm of s1,yshift=-1cm] {$T_R$ \textbf{ABORTS}};
      }
    \end{tikzpicture}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Order Arbitration Protocol}
  If a transaction updates $1$ or more distributed edges:
  \begin{itemize}
    \uncover<1->{
    \item Collect all predecessors
    }
    \uncover<2->{
    \item Go to arbiter
    }
    \uncover<3->{
    \item If not in hit list then proceed and enter predecessors into the hit list
    }
    \uncover<4->{
    \item Else in hit list then abort
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Edge Concurrency Control Protocol Summary}
    \begin{enumerate}
    \item Collision detection: enforces \textbf{reciprocal consistency} for distributed edges
    \item Order arbitration: enforces \textbf{edge-order consistency} between transactions
    \end{enumerate}
\end{frame}

%------------------------------------------------
% Performance Evaluation
% ------------------------------------------------

\begin{frame}
  \frametitle{Performance Evaluation}
  Performance measures of interest:
  \begin{itemize}
  \item Average number of transactions that are aborted
  \item Load at the arbiter
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Evaluation Approach}
  \begin{enumerate}
  \item Developed approximate model
  \item Measure accuracy of model through simulation
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Parameters}
  \begin{itemize}
  \item Database size
  \item Transaction arrival rate
  \item Average updates per transaction
  \item Average arbiter service time
  \item Average network delay
  \end{itemize}
\end{frame}

%------------------------------------------------
% Numerical and Simulation Results
% ------------------------------------------------

\begin{frame}
  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Database Size $(\boldsymbol{N})$}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.35]{images/varying_num_edges.png}
    \caption{TPS $= 1000$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
  \begin{center}
    Arbiter queue unstable at $\sim1500$ TPS
  \end{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.35]{images/varying_tnx_arrival_rate}
    \caption{\textcolor{red}{size $= 10K$}, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
  \begin{center}
    Arbiter queue unstable at $\sim 1100$ TPS
  \end{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.35]{images/larger_network_delay.png}
    \caption{size $= 10K$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, \textcolor{red}{av network delay $=10ms$}}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
  \begin{center}
    Arbiter queue unstable at $\sim 550$ TPS
  \end{center}
    \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.35]{images/varying_updates_per_tnx.png}
        \caption{size $= 10K$, \textcolor{red}{av updates/tnx $= 10$}, av arbiter service time $= 10ms$, av network delay $=5ms$}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Summary and Future Work}
  \begin{itemize}
    \uncover<1->{
    \item Model accuracy similar to simulations under a variety of parameter settings
    }
    \uncover<2->{
    \item Between 1-4\% transactions are aborted
    }
    \uncover<3->{
    \item Improve accuracy of simulation
    }
  \end{itemize}
\end{frame}

\begin{frame}[standout]
 Thanks for listening! \\
  \vspace{4mm}
  Any Questions? \\
  \vspace{4mm}
  {\normalsize Email: \href{mailto:j.waudby2@newcastle.ac.uk}{j.waudby2@newcastle.ac.uk}} \\
  {\normalsize Twitter: \href{https://twitter.com/waudberry_7}{@waudberry\_7}}
\end{frame}

\end{document}

### 2020-04-20 slides/config.tex
corrections

--- 2020-04-16 slides/main.tex
@@
-% ----------------------------------------------------------------------------------------
-%	Configuration
-%----------------------------------------------------------------------------------------
-
-\documentclass[aspectratio=169]{beamer}
-
 \usetheme[titleformat=regular,numbering=none,progressbar=frametitle]{metropolis}
 
-% newcastle colour scheme
+% colour scheme
 \definecolor{nclBlue}{HTML}{003A65}
 \definecolor{nclRed}{HTML}{D91A35}
+\definecolor{grey}{rgb}{0.52, 0.52, 0.51}
+\definecolor{red}{rgb}{0.7, 0.11, 0.11}
+\definecolor{blue}{rgb}{0.0, 0.0, 0.55}
+\definecolor{green}{rgb}{0.0, 0.42, 0.24}
 
+% beamer
 \setbeamercovered{transparent}
@@
 \setbeamercolor{palette primary}{fg=white, bg=nclBlue}
-
 \linespread{1.5}
 
+% packages
 \usepackage{graphicx}
@@
 \usepackage{url}
+\usepackage{xspace}
 \usepackage{amsmath}
@@
 \usepackage[font=footnotesize]{caption}
-
+\usepackage{subcaption}
 % Thin fonts
@@
 
+% commands
+\newcommand{\tDelta}{\textsf{Delta}\xspace}
+
 % tikz configuration
@@
 \usetikzlibrary{calc}
-
 \tikzstyle{vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
@@
 
-%----------------------------------------------------------------------------------------
-%	Title + Overview Slide
-% ----------------------------------------------------------------------------------------
+\tikzstyle{big_vertex} = [circle, minimum width=50pt,draw,inner sep=0pt]
 
-\title[]{Preserving Reciprocal Consistency in \\ Distributed Graph Databases}
+\def\record(#1,#2,#3,#4,#5,#6){
+  \node (v) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.2\linewidth,xshift=-7cm,align=center,yshift=#1cm] {#2};
 
-\author[Waudby \& Ezhilchelvan \& Webber \& Mitrani]{\underline{\emph{Jack Waudby}}$^{1}$, Paul Ezhilchelvan$^{1}$, Jim Webber$^{2}$ \& Isi Mitrani$^{1}$}
-\institute[NCL]
-{
-  $^{1}$Newcastle University\\
-  $^{2}$Neo4j
+  \node (b) [draw,black,minimum height=0.9cm,minimum width=15.9cm,text width=0.2\linewidth,xshift=2.85cm,yshift=#1cm] {};
+
+  \draw[thick,->] (v) -- (b);
+
+  \node (i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-3.1cm,text width=0.25\linewidth,yshift=#1cm,align=center] {#3};
+
+  \node (i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i1,yshift=0cm,text width=0.25\linewidth,align=center] {#4};
+
+  \node (i3)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i2,text width=0.25\linewidth,yshift=0cm,align=center] {#5};
+
+  \node (i4)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i3,text width=0.25\linewidth,yshift=0cm,align=center] {#6};
+
 }
-\date{April 27, 2020}
 
-\begin{document}
+        \def\greenRecord(#1,#2,#3,#4,#5,#6){
+  \node (v) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.2\linewidth,xshift=-7cm,align=center,yshift=#1cm] {#2};
 
-\begin{frame}
-\maketitle
-\begin{tikzpicture}[overlay, remember picture]
-  \node[above left=2.8cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_newcastle_uni.png}};
-  \node[above left=1.2cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_neo4j.png}};
-\end{tikzpicture}
-\end{frame}
+  \node (b) [draw,black,minimum height=0.9cm,minimum width=15.9cm,text width=0.2\linewidth,xshift=2.85cm,yshift=#1cm] {};
 
-%----------------------------------------------------------------------------------------
-%	PRESENTATION SLIDES
-%----------------------------------------------------------------------------------------
+  \draw[thick,->] (v) -- (b);
 
-%------------------------------------------------
-% Background
-% ------------------------------------------------
+  \node (i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-3.1cm,text width=0.25\linewidth,yshift=#1cm,align=center] {#3};
 
-\setcounter{framenumber}{0}
-\metroset{numbering=counter}
+  \node (i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i1,yshift=0cm,text width=0.25\linewidth,align=center,fill=green!30] {#4};
 
-\begin{frame}
-  \frametitle{Graph Databases}
-    \begin{figure}
-      \includegraphics[scale=0.5]{images/property_graph.png}
-      \caption{Labeled property graph\footnotemark}
-    \end{figure}
-  \footnotetext[1]{
-    \url{https://neo4j.com}
+  \node (i3)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i2,text width=0.25\linewidth,yshift=0cm,align=center] {#5};
+
+  \node (i4)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i3,text width=0.25\linewidth,yshift=0cm,align=center] {#6};
+
 }
-\end{frame}
 
-\begin{frame}
-  \frametitle{Graph Databases}
-  \begin{itemize}
-    \uncover<1->{
-    \item \textbf{Efficient graph analysis:} reachability, pattern matching, shortest path search and clustering
+        \def\greenRecordB(#1,#2,#3,#4,#5,#6){
+  \node (v) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.2\linewidth,xshift=-7cm,align=center,yshift=#1cm] {#2};
+
+  \node (b) [draw,black,minimum height=0.9cm,minimum width=15.9cm,text width=0.2\linewidth,xshift=2.85cm,yshift=#1cm] {};
+
+  \draw[thick,->] (v) -- (b);
+
+  \node (i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-3.1cm,text width=0.25\linewidth,yshift=#1cm,align=center] {#3};
+
+  \node (i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i1,yshift=0cm,text width=0.25\linewidth,align=center] {#4};
+
+  \node (i3)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i2,text width=0.25\linewidth,yshift=0cm,align=center,fill=green!30] {#5};
+
+  \node (i4)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i3,text width=0.25\linewidth,yshift=0cm,align=center] {#6};
+
 }
-    \uncover<2->{
-    \item \textbf{Use cases:} telecommunications, pharma, publishing, finance, social media
+
+\def\spaceRecord(#1){
+  \node (v) [black,minimum height=0.9cm,minimum width=1cm,text width=0.2\linewidth,xshift=-7cm,align=center,yshift=#1cm] {$\vdots$};
+  \node (b) [black,minimum height=0.9cm,minimum width=15.9cm,text width=0.2\linewidth,xshift=1.4cm,yshift=#1cm] {};
+  \node (i1)  [black,minimum height=0.6cm,minimum width=2cm,xshift=-3.1cm,text width=0.3\linewidth,yshift=#1cm,align=center] {$\vdots$};
+  \node (i2)  [black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i1,yshift=0cm,text width=0.3\linewidth,align=center] {$\vdots$};
+  \node (i3)  [minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i2,yshift=0cm,align=center] {$\vdots$};
+  \node (i4)  [minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,right of=i3,yshift=0cm,align=center] {$\vdots$};
 }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Graph Database Popularity}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.3]{images/db_popularity.png}
-    \caption{Database popularity by data model\footnotemark}
-  \end{figure}
-  \footnotetext[1]{
-    \url{https://db-engines.com/en/ranking_categories}
-  }
-\end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Graph Databases}
-  \begin{itemize}
-    \uncover<1->{
-    \item Graph exceeds the storage capacity of a single machine
-    }
-    \uncover<2->{
-    \item Partition the graph across multiple machines in a cluster, with partitions replicated for fault-tolerance and availability
-    }
-    \uncover<3->{
-    \item Graph partitioning inevitably leads to \textbf{distributed edges}
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Graph Databases: Partitioning}
-\input{./figures/graph}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Graph Databases: Partitioning}
-  \begin{figure}[h!]
-    \centering
-    \resizebox{0.6\textwidth}{!}{
-    \begin{tikzpicture}[node distance=2cm]
-      \node (s1) [server,label=$S_1$] {};
-      \node (s2) [server, right of=s1,xshift=2cm,yshift=2cm,label=$S_2$] {};
-      \node (s3) [server, right of=s1,xshift=6cm,label=$S_3$] {};
-
-      \node (v1) [vertex,xshift=-0.5cm,yshift=-1.5cm] {};
-      \node (v2) [vertex, above of=v1,xshift=-0.3cm,yshift=1cm] {};
-      \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-      \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.3cm] {};
-      \node (v5) [vertex, above right of=v3,xshift=-0.6cm,yshift=-0.5cm] {};
-      \node (v6) [vertex, right of=v5,xshift=0.3cm,yshift=0.8cm] {};
-      \node (v7) [vertex, below of=v6, xshift=1cm,yshift=0.4cm] {};
-      \node (v8) [vertex, right of=v4, yshift=-0.3cm,xshift=4.5cm] {};
-      \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-      \node (v10) [vertex, above of=v9,xshift=-2.5cm,yshift=-0.3cm] {};
-      \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-      \node (v12) [vertex, above of=v11,yshift=-1cm,xshift=-1cm] {};
-      \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-      \draw [local] (v2) -- (v1);
-      \draw [local] (v1) -- (v3);
-      \draw [local] (v3) -- (v2);
-      \draw [local] (v2) -- (v5);
-      \draw [local] (v1) -- (v4);
-      \draw [local] (v5) -- (v4);
-      \draw [dist] (v5) -- (v6);
-      \draw [dist] (v4) -- (v7);
-      \draw [dist] (v8) -- (v4);
-      \draw [local] (v8) -- (v9);
-      \draw [local] (v9) -- (v11);
-      \draw [local] (v10) -- (v7);
-      \draw [dist] (v10) -- (v12);
-      \draw [dist] (v12) -- (v13);
-      \draw [local] (v10) -- (v6);
-      \draw [dist] (v10) -- (v11);
-      \draw [local] (v13) -- (v10);
-      \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  }
-\caption{Graph partitioned across 3 servers}
-  \end{figure}
-\end{frame}
-
-%------------------------------------------------
-% Problem 1: Reciprocal Consistency}
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency}
-  \begin{center}
-  \textbf{A logical edge is represented by 2 physical records}
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=7cm]
-      \uncover<1-> {
-      \node (n1) [bignode] {\texttt{Flight}};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
-      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:AVAILABLE}} (n2);
-    }
-      \uncover<2-> {
-         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Seat Available In Flight}};
-         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Flight Has Seat Available}};
-         \draw [<->,dashed](q1) -- (q2);
-      }
-    \end{tikzpicture}
-    \caption{A reciprocal consistent edge}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  Two concurrent transactions:
-  \begin{itemize}
-  \item Mr Red requests to book the seat, $T_R$
-  \item Mr Blue requests to book the seat, $T_B$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \uncover<1-> {
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
-        \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
-      }
-      \uncover<2-3> {
-        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
-      }
-      \uncover<3-4> {
-        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
-      }
-      \uncover<4-> {
-        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
-      }
-      \uncover<5-> {
-        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
-      }
-    \end{tikzpicture}
-    \caption{Concurrent transactions interleave and violate reciprocal consistency}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=7cm]
-      \node (n1) [bignode] {\texttt{Flight}};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
-      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
-
-      \uncover<2-> {
-      \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
-      \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
-      \draw [<->,dashed](q1) -- (q2);
-      \draw [<->,dashed](q1) -- (q2);
-      \coordinate (a) at (4,-2);
-      \coordinate (b) at (3,-3);
-      \coordinate (c) at (3,-2);
-      \coordinate (d) at (4,-3);
-      \draw [very thick,red] ($ (a) $) -- ($ (b) $);
-      \draw [very thick,red] ($ (c) $) -- ($ (d) $);
-    }
-    \end{tikzpicture}
-    \caption{Reciprocal inconsistent distributed edge}
-  \end{figure}
-\end{frame}
-
-% \begin{frame}
-%   \frametitle{Reciprocal Consistency Violation}
-%   \begin{figure}[h!]
-%     \centering
-%     \begin{tikzpicture}[node distance=2cm]
-%       \uncover<1->{
-%         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
-%         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
-%         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
-%         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
-%         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
-%       }
-%       \uncover<2->{
-%         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
-%         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
-%         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
-%         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
-%         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
-%       }
-%     \end{tikzpicture}
-%     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
-%   \end{figure}
-% \end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Edge Reciprocal Consistency}
-  \begin{itemize}
-    \uncover<1->{
-  \item Without concurrency control, distributed edges can become reciprocally inconsistent
-    \begin{itemize}
-    \item A distributed edge's reciprocal information is separated by the network
-    \end{itemize}
-  }
-  \uncover<2->{
-  \item Reciprocal inconsistency is a source of corruption\footnotemark
-  }
-  \uncover<3->{
-  \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
-    }
-    \end{itemize}
-    \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Our Solution: Collision Detection}
-  \begin{itemize}
-    \uncover<1->{
-    \item Carefully abort transaction(s)
-    }
-    \uncover<2->{
-    \item No centralized control or synchronized clock
-    }
-    \uncover<3->{
-    \item Permits interleavings that preserve reciprocal consistency
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Collision Detection General Rules}
-  \begin{itemize}
-    \uncover<1->{
-    \item Updates are initially provisional
-    }
-    \uncover<2->{
-    \item Transactions distinguish between their first and second updates using labels
-    }
-    \uncover<3->{
-    \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Interference-free Update}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
-        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
-      }
-      \uncover<2->{
-        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
-      }
-      \uncover<3->{
-        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
-      }
-      \uncover<3->{
-        \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
-      }
-      \uncover<4->{
-        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
-      }
-    \end{tikzpicture}
-  \end{figure}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Collision Detection}
-  \begin{center}
-    \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<2->{
-        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
-        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
-      }
-      \uncover<2->{
-        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
-      }
-      \uncover<3->{
-        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
-      }
-      \uncover<4-11>{
-        \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
-      }
-      \uncover<5-10>{
-        \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
-      }
-      \uncover<6-10>{
-        \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
-      }
-      \uncover<7-7>{
-        \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
-      }
-      \uncover<8-10>{
-        \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
-        \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
-      }
-      \uncover<9-9>{
-        \node [] at (6,2) {$T_R$'s 2 is missing};
-      }
-      \uncover<10-10>{
-        \node[] at (6,2) {$T_B$ ABORTS};
-      }
-      \uncover<12->{
-        \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
-      }
-      \uncover<13->{
-        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
-      }
-      \uncover<14->{
-        \node[] at (6,2) {$T_R$ SUCCEEDS};
-        }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
-
-%------------------------------------------------
-%Problem 2: Edge-Order Consistency
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency}
-  A second consistency problem related to distributed edges:
-  \begin{itemize}
-    \uncover<2->{
-    \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
-    }
-    \uncover<3->{
-    \item Each update is reciprocally consistent
-    }
-    \uncover<4->{
-    \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  Consider two edges,
-  \begin{figure}[h!]
-    \centering
-    \resizebox{1\textwidth}{!}{
-      \begin{tikzpicture}[node distance=5.2cm]
-        \node[] at (2.6,2.5) {\large{$E$}};
-        \node[] at (13,2.5) {\large{$E'$}};
-        \node (n1) [bignode] {\texttt{Flight}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
-        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-        \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
-        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
-        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
-         \draw [<->,dashed](q1) -- (q2);
-
-        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
-        \node (n4) [bignode, right of=n3] {\texttt{Room}};
-        \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
-        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
-        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
-        \draw [<->,dashed](q3) -- (q4);
-
-      \end{tikzpicture}
-    }
-    \caption{Two distributed edges}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  Two concurrent transactions by a travel agent:
-  \begin{itemize}
-  \item Requests to book room and seat for Mr Red, $T_R$
-  \item Requests to book room and seat for Mr Blue, $T_B$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-
-        \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
-        \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
-        \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
-        \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
-        \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
-        \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
-        \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
-      }
-      \uncover<2->{
-        \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
-      }
-      \uncover<3->{
-        \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
-      }
-      \uncover<4->{
-
-        \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
-      }
-      \uncover<5->{
-
-        \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
-      }
-      \uncover<6->{
-        \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
-        \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
-        \draw (8.3,1) rectangle (11,2.5);
-      }
-    \end{tikzpicture}
-    \caption{Edge-order consistency violation}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \resizebox{1\textwidth}{!}{
-      \begin{tikzpicture}[node distance=5cm]
-        % edge labels
-        \node[] at (2.5,3) {\large{$E$}};
-        \node[] at (12.5,3) {\large{$E'$}};
-        % edge 1: flight -> seat
-        \node (n1) [bignode] {\texttt{Flight}};
-        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-        % edge 1: servers
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5cm,label=above:{$S_2$}] {};
-        % edge 1
-        \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
-        % edge 1 information
-        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
-        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
-        \draw [<->,dashed](q1) -- (q2);
-
-        % edge 2: hotel -> room
-        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
-        \node (n4) [bignode, right of=n3] {\texttt{Room}};
-        \draw [local] (n3) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n4);
-        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
-        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15cm,label=above:{$S_4$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10cm,label=above:{$S_3$}] {};
-        \draw [<->,dashed](q3) -- (q4);
-
-      \end{tikzpicture}
-    }
-    \caption{Edge-order consistency violation}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Our Solution: Order Arbitration}
-  \begin{itemize}
-  \uncover<1->{
-  \item Transactions collect predecessors from each update
-  }
-  \uncover<2->{
-  \item Centralized arbiter maintains some global state
-  }
-  \uncover<3->{
-  \item Global state used to detect edge-order violations
-  }
-  \uncover<4->{
-  \item Offending transactions are aborted
-  }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Order Arbitration}
-    \begin{center}
-      \textbf{Rule: Collect predecessors after every successful update}
-    \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (3,0) -- (3,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (3,4.5) node [label=above:{$S_2$}] {};
-        \draw [black,<-,>=stealth] (7,0) -- (7,4) ;
-        \draw [black,<-,>=stealth] (10,0) -- (10,4) ;
-        \draw [black,dashed,->,>=stealth] (7,4.5) node [label=above:{$S_3$}] {} -- node [midway,above] {$E'$} (10,4.5) node [label=above:{$S_4$}] {};
-      }
-      \uncover<2->{
-        \draw [blue,->,>=stealth] (0,3.2) node (TXS) [label=left:{\small{$T_B$}}] {} -- (3,3);
-        \draw [red,->,>=stealth] (3,3.6) node (TXS) [label=right:{\small{$T_R$}}] {} -- (0,3.4);
-        \draw (3.1, 2.8) node [label=right:{\small{\textcolor{black}{$T_B : \{{T_R}\}$}}}] {};
-        \draw (3.2,2.3) rectangle (5,3.2);
-      }
-      \uncover<3->{
-        \draw [blue,->,>=stealth] (7,2.3) node [label=left:{\small{$T_B$}}] {} -- (10,1.8);
-        \draw [red,->,>=stealth] (7,1.8) node [label=left:{\small{$T_R$}}] {} -- (10,1.2) ;
-        \draw (10.1, 0.8) node (TXS) [yshift=-1cm,label=right:{\small{\textcolor{black}{$T_R : \{{T_B}\}$}}}] {};
-        \draw (10.2,-0.6) rectangle (12,0.3);
-      }
-      \uncover<4->{
-        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
-        (10.2,1.8) node[above,xshift=1cm, yshift=0.3cm] {ARBITER} -- (11,1.8);
-        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
-        (10.2,1.2)   -- (11,1.2);
-      }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Order Arbitration}
-      \begin{center}
-      \textbf{Rule: If transaction exists in hit list, abort. Else, merge predecessors into hit list}
-    \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{ % 1 arbiter
-        \node (s1) [server,label=$Arbiter$] {};
-      }
-      \uncover<2->{ % 2 hit list
-        \node (a) [rectangle, label=above:{$Hit \ List$}, minimum width=2cm,minimum height=1cm,draw=black,xshift=0cm,yshift=0cm] {$T_z$ ,$T_a$ {\uncover<7->{,$T_R$}}};
-      } % uncover T_R when T_B processed
-      \uncover<3->{ % 3 queue
-              \node (q) [rectangle, label=above:{$Queue$},left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=0cm] {{\uncover<6->{$...$}}};
-              \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=0cm] {{\uncover<5-8>{$T_R$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=0cm] {{\uncover<4-5>{$T_B$}}};
-        \node (q1) [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=-1cm] {{\uncover<6->{$...$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=-1cm] {{\uncover<5-8>{$T_B$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=-1cm] {{\uncover<4-5>{$T_R$}}};
-        \node [left of=q,xshift=0.2cm,label=left:{\small{\textcolor{black}{$tnx.id$}}}] {};
-        \node [left of=q1,xshift=0.2cm,label=left:{\small{\textcolor{black}{$pred$}}}] {};
-      }
-      \uncover<4-4> {
-        \node[right=0.5cm of s1] {$T_B$ \textbf{ENTERS QUEUE}};
-      }
-      \uncover<5-5> {
-        \node[right=0.5cm of s1] {$T_R$ \textbf{ENTERS QUEUE}};
-      }
-      \uncover<6-8> {
-        \node[right=0.5cm of s1,yshift=1cm] {$T_B$ \textbf{PROCESSED}};
-      }
-      \uncover<7-8> {
-        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{ADDED TO HIT LIST}};
-      }
-      \uncover<8-8> {
-        \node[right=0.5cm of s1,,yshift=-1cm] {$T_B$ \textbf{SUCCEEDS}};
-      }
-      \uncover<9-> {
-        \node[right=0.5cm of s1,yshift=1cm] {$T_R$ \textbf{PROCESSED}};
-      }
-      \uncover<10-> {
-        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{IN HIT LIST}};
-      }
-      \uncover<11-> {
-        \node[right=0.5cm of s1,yshift=-1cm] {$T_R$ \textbf{ABORTS}};
-      }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Order Arbitration Protocol}
-  If a transaction updates $1$ or more distributed edges:
-  \begin{itemize}
-    \uncover<1->{
-    \item Collect all predecessors
-    }
-    \uncover<2->{
-    \item Go to arbiter
-    }
-    \uncover<3->{
-    \item If not in hit list then proceed and enter predecessors into the hit list
-    }
-    \uncover<4->{
-    \item Else in hit list then abort
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge Concurrency Control Protocol Summary}
-    \begin{enumerate}
-    \item Collision detection: enforces \textbf{reciprocal consistency} for distributed edges
-    \item Order arbitration: enforces \textbf{edge-order consistency} between transactions
-    \end{enumerate}
-\end{frame}
-
-%------------------------------------------------
-% Performance Evaluation
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Performance Evaluation}
-  Performance measures of interest:
-  \begin{itemize}
-  \item Average number of transactions that are aborted
-  \item Load at the arbiter
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Evaluation Approach}
-  \begin{enumerate}
-  \item Developed approximate model
-  \item Measure accuracy of model through simulation
-  \end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parameters}
-  \begin{itemize}
-  \item Database size
-  \item Transaction arrival rate
-  \item Average updates per transaction
-  \item Average arbiter service time
-  \item Average network delay
-  \end{itemize}
-\end{frame}
-
-%------------------------------------------------
-% Numerical and Simulation Results
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Database Size $(\boldsymbol{N})$}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_num_edges.png}
-    \caption{TPS $= 1000$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim1500$ TPS
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_tnx_arrival_rate}
-    \caption{\textcolor{red}{size $= 10K$}, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim 1100$ TPS
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/larger_network_delay.png}
-    \caption{size $= 10K$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, \textcolor{red}{av network delay $=10ms$}}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim 550$ TPS
-  \end{center}
-    \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_updates_per_tnx.png}
-        \caption{size $= 10K$, \textcolor{red}{av updates/tnx $= 10$}, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Summary and Future Work}
-  \begin{itemize}
-    \uncover<1->{
-    \item Model accuracy similar to simulations under a variety of parameter settings
-    }
-    \uncover<2->{
-    \item Between 1-4\% transactions are aborted
-    }
-    \uncover<3->{
-    \item Improve accuracy of simulation
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}[standout]
- Thanks for listening! \\
-  \vspace{4mm}
-  Any Questions? \\
-  \vspace{4mm}
-  {\normalsize Email: \href{mailto:j.waudby2@newcastle.ac.uk}{j.waudby2@newcastle.ac.uk}} \\
-  {\normalsize Twitter: \href{https://twitter.com/waudberry_7}{@waudberry\_7}}
-\end{frame}
-
-\end{document}

### 2020-04-20 slides/figures/dist-edge.tex
corrections

\begin{tikzpicture}[node distance=2.2cm]
  \node (rect)  [draw,rounded corners,minimum width=3.5cm,minimum height=4cm,label=above:{$S_i$}] {};
  \node (rect)  [draw,rounded corners,minimum width=3.5cm,minimum height=4cm,label=above:{$S_j$},xshift=6cm] {};
  \node (v1) [big_vertex,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
  \node (v2) [big_vertex,xshift=6cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
  \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
  \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
  \draw [thick,dashed,->,>=stealth] (0.9,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (5.1,0);
\end{tikzpicture}

### 2020-04-20 slides/figures/graph-part.tex
corrections

--- 2020-04-16 slides/main.tex
@@
-% ----------------------------------------------------------------------------------------
-%	Configuration
-%----------------------------------------------------------------------------------------
-
-\documentclass[aspectratio=169]{beamer}
-
-\usetheme[titleformat=regular,numbering=none,progressbar=frametitle]{metropolis}
-
-% newcastle colour scheme
-\definecolor{nclBlue}{HTML}{003A65}
-\definecolor{nclRed}{HTML}{D91A35}
-
-\setbeamercovered{transparent}
-\setbeamercovered{invisible}
-\setbeamercolor{progress bar}{fg=nclRed}
-\setbeamercolor{frametitle}{bg=nclBlue}
-\setbeamercolor{background canvas}{bg=white}
-\setbeamercolor{block body}{bg=gray!30}
-\setbeamercolor{block title}{bg=nclBlue,fg=white}
-\setbeamercolor{palette primary}{fg=white, bg=nclBlue}
-
-\linespread{1.5}
-
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{amssymb}
-\usepackage{pifont}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{mathspec}
-\usepackage[font=footnotesize]{caption}
-
-% Thin fonts
-\setsansfont[BoldFont={Fira Sans}, Numbers={OldStyle}]{Fira Sans Light}
-\setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Light}
-
-% Thick fonts
-% \setsansfont[BoldFont={Fira Sans SemiBold}, Numbers={OldStyle}]{Fira Sans}
-% \setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Book}
-
-% tikz configuration
-\usetikzlibrary{shapes.geometric, arrows}
-\usetikzlibrary{arrows, positioning, shapes}
-\usetikzlibrary{decorations.pathmorphing}
-\usetikzlibrary{calc}
-
-\tikzstyle{vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
-\tikzstyle{mini_vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=3cm,minimum height=4cm,draw=black]
-\tikzstyle{mini_server} = [rectangle, rounded corners,minimum width=1.5cm,minimum height=1.9cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{bignode}=[draw, circle, minimum width=70pt,align=center]
-
-%----------------------------------------------------------------------------------------
-%	Title + Overview Slide
-% ----------------------------------------------------------------------------------------
-
-\title[]{Preserving Reciprocal Consistency in \\ Distributed Graph Databases}
-
-\author[Waudby \& Ezhilchelvan \& Webber \& Mitrani]{\underline{\emph{Jack Waudby}}$^{1}$, Paul Ezhilchelvan$^{1}$, Jim Webber$^{2}$ \& Isi Mitrani$^{1}$}
-\institute[NCL]
-{
-  $^{1}$Newcastle University\\
-  $^{2}$Neo4j
-}
-\date{April 27, 2020}
-
-\begin{document}
-
-\begin{frame}
-\maketitle
-\begin{tikzpicture}[overlay, remember picture]
-  \node[above left=2.8cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_newcastle_uni.png}};
-  \node[above left=1.2cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_neo4j.png}};
-\end{tikzpicture}
-\end{frame}
-
-%----------------------------------------------------------------------------------------
-%	PRESENTATION SLIDES
-%----------------------------------------------------------------------------------------
-
-%------------------------------------------------
-% Background
-% ------------------------------------------------
-
-\setcounter{framenumber}{0}
-\metroset{numbering=counter}
-
-\begin{frame}
-  \frametitle{Graph Databases}
-    \begin{figure}
-      \includegraphics[scale=0.5]{images/property_graph.png}
-      \caption{Labeled property graph\footnotemark}
-    \end{figure}
-  \footnotetext[1]{
-    \url{https://neo4j.com}
-  }
-\end{frame}
-
-\begin{frame}
-  \frametitle{Graph Databases}
-  \begin{itemize}
-    \uncover<1->{
-    \item \textbf{Efficient graph analysis:} reachability, pattern matching, shortest path search and clustering
-    }
-    \uncover<2->{
-    \item \textbf{Use cases:} telecommunications, pharma, publishing, finance, social media
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Graph Database Popularity}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.3]{images/db_popularity.png}
-    \caption{Database popularity by data model\footnotemark}
-  \end{figure}
-  \footnotetext[1]{
-    \url{https://db-engines.com/en/ranking_categories}
-  }
-\end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Graph Databases}
-  \begin{itemize}
-    \uncover<1->{
-    \item Graph exceeds the storage capacity of a single machine
-    }
-    \uncover<2->{
-    \item Partition the graph across multiple machines in a cluster, with partitions replicated for fault-tolerance and availability
-    }
-    \uncover<3->{
-    \item Graph partitioning inevitably leads to \textbf{distributed edges}
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Graph Databases: Partitioning}
-\input{./figures/graph}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Graph Databases: Partitioning}
 \begin{figure}[h!]
@@
     \begin{tikzpicture}[node distance=2cm]
-      \node (s1) [server,label=$S_1$] {};
-      \node (s2) [server, right of=s1,xshift=2cm,yshift=2cm,label=$S_2$] {};
-      \node (s3) [server, right of=s1,xshift=6cm,label=$S_3$] {};
+      \node (s1) [server,label=$S_i$] {};
+      \node (s2) [server, right of=s1,xshift=2cm,yshift=2cm,label=$S_j$] {};
+      \node (s3) [server, right of=s1,xshift=6cm,label=$S_k$] {};
 
@@
   }
-\caption{Graph partitioned across 3 servers}
+  \caption{Partitioned graph}
 \end{figure}
-\end{frame}
-
-%------------------------------------------------
-% Problem 1: Reciprocal Consistency}
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency}
-  \begin{center}
-  \textbf{A logical edge is represented by 2 physical records}
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=7cm]
-      \uncover<1-> {
-      \node (n1) [bignode] {\texttt{Flight}};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
-      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:AVAILABLE}} (n2);
-    }
-      \uncover<2-> {
-         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Seat Available In Flight}};
-         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Flight Has Seat Available}};
-         \draw [<->,dashed](q1) -- (q2);
-      }
-    \end{tikzpicture}
-    \caption{A reciprocal consistent edge}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  Two concurrent transactions:
-  \begin{itemize}
-  \item Mr Red requests to book the seat, $T_R$
-  \item Mr Blue requests to book the seat, $T_B$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \uncover<1-> {
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
-        \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
-      }
-      \uncover<2-3> {
-        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
-      }
-      \uncover<3-4> {
-        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
-      }
-      \uncover<4-> {
-        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
-      }
-      \uncover<5-> {
-        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
-      }
-    \end{tikzpicture}
-    \caption{Concurrent transactions interleave and violate reciprocal consistency}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=7cm]
-      \node (n1) [bignode] {\texttt{Flight}};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
-      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
-
-      \uncover<2-> {
-      \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
-      \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
-      \draw [<->,dashed](q1) -- (q2);
-      \draw [<->,dashed](q1) -- (q2);
-      \coordinate (a) at (4,-2);
-      \coordinate (b) at (3,-3);
-      \coordinate (c) at (3,-2);
-      \coordinate (d) at (4,-3);
-      \draw [very thick,red] ($ (a) $) -- ($ (b) $);
-      \draw [very thick,red] ($ (c) $) -- ($ (d) $);
-    }
-    \end{tikzpicture}
-    \caption{Reciprocal inconsistent distributed edge}
-  \end{figure}
-\end{frame}
-
-% \begin{frame}
-%   \frametitle{Reciprocal Consistency Violation}
-%   \begin{figure}[h!]
-%     \centering
-%     \begin{tikzpicture}[node distance=2cm]
-%       \uncover<1->{
-%         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
-%         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
-%         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
-%         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
-%         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
-%       }
-%       \uncover<2->{
-%         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
-%         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
-%         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
-%         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
-%         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
-%       }
-%     \end{tikzpicture}
-%     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
-%   \end{figure}
-% \end{frame}
-
-\begin{frame}
-  \frametitle{Distributed Edge Reciprocal Consistency}
-  \begin{itemize}
-    \uncover<1->{
-  \item Without concurrency control, distributed edges can become reciprocally inconsistent
-    \begin{itemize}
-    \item A distributed edge's reciprocal information is separated by the network
-    \end{itemize}
-  }
-  \uncover<2->{
-  \item Reciprocal inconsistency is a source of corruption\footnotemark
-  }
-  \uncover<3->{
-  \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
-    }
-    \end{itemize}
-    \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Our Solution: Collision Detection}
-  \begin{itemize}
-    \uncover<1->{
-    \item Carefully abort transaction(s)
-    }
-    \uncover<2->{
-    \item No centralized control or synchronized clock
-    }
-    \uncover<3->{
-    \item Permits interleavings that preserve reciprocal consistency
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Collision Detection General Rules}
-  \begin{itemize}
-    \uncover<1->{
-    \item Updates are initially provisional
-    }
-    \uncover<2->{
-    \item Transactions distinguish between their first and second updates using labels
-    }
-    \uncover<3->{
-    \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Interference-free Update}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
-        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
-      }
-      \uncover<2->{
-        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
-      }
-      \uncover<3->{
-        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
-      }
-      \uncover<3->{
-        \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
-      }
-      \uncover<4->{
-        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
-      }
-    \end{tikzpicture}
-  \end{figure}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Collision Detection}
-  \begin{center}
-    \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<2->{
-        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
-        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
-      }
-      \uncover<2->{
-        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
-      }
-      \uncover<3->{
-        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
-      }
-      \uncover<4-11>{
-        \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
-      }
-      \uncover<5-10>{
-        \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
-      }
-      \uncover<6-10>{
-        \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
-      }
-      \uncover<7-7>{
-        \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
-      }
-      \uncover<8-10>{
-        \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
-        \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
-      }
-      \uncover<9-9>{
-        \node [] at (6,2) {$T_R$'s 2 is missing};
-      }
-      \uncover<10-10>{
-        \node[] at (6,2) {$T_B$ ABORTS};
-      }
-      \uncover<12->{
-        \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
-      }
-      \uncover<13->{
-        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
-      }
-      \uncover<14->{
-        \node[] at (6,2) {$T_R$ SUCCEEDS};
-        }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
-
-%------------------------------------------------
-%Problem 2: Edge-Order Consistency
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency}
-  A second consistency problem related to distributed edges:
-  \begin{itemize}
-    \uncover<2->{
-    \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
-    }
-    \uncover<3->{
-    \item Each update is reciprocally consistent
-    }
-    \uncover<4->{
-    \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  Consider two edges,
-  \begin{figure}[h!]
-    \centering
-    \resizebox{1\textwidth}{!}{
-      \begin{tikzpicture}[node distance=5.2cm]
-        \node[] at (2.6,2.5) {\large{$E$}};
-        \node[] at (13,2.5) {\large{$E'$}};
-        \node (n1) [bignode] {\texttt{Flight}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
-        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-        \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
-        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
-        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
-         \draw [<->,dashed](q1) -- (q2);
-
-        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
-        \node (n4) [bignode, right of=n3] {\texttt{Room}};
-        \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
-        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
-        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
-        \draw [<->,dashed](q3) -- (q4);
-
-      \end{tikzpicture}
-    }
-    \caption{Two distributed edges}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  Two concurrent transactions by a travel agent:
-  \begin{itemize}
-  \item Requests to book room and seat for Mr Red, $T_R$
-  \item Requests to book room and seat for Mr Blue, $T_B$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-
-        \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
-        \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
-        \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
-        \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
-        \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
-        \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
-        \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
-      }
-      \uncover<2->{
-        \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
-      }
-      \uncover<3->{
-        \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
-      }
-      \uncover<4->{
-
-        \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
-      }
-      \uncover<5->{
-
-        \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
-      }
-      \uncover<6->{
-        \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
-        \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
-        \draw (8.3,1) rectangle (11,2.5);
-      }
-    \end{tikzpicture}
-    \caption{Edge-order consistency violation}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \resizebox{1\textwidth}{!}{
-      \begin{tikzpicture}[node distance=5cm]
-        % edge labels
-        \node[] at (2.5,3) {\large{$E$}};
-        \node[] at (12.5,3) {\large{$E'$}};
-        % edge 1: flight -> seat
-        \node (n1) [bignode] {\texttt{Flight}};
-        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-        % edge 1: servers
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5cm,label=above:{$S_2$}] {};
-        % edge 1
-        \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
-        % edge 1 information
-        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
-        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
-        \draw [<->,dashed](q1) -- (q2);
-
-        % edge 2: hotel -> room
-        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
-        \node (n4) [bignode, right of=n3] {\texttt{Room}};
-        \draw [local] (n3) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n4);
-        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
-        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15cm,label=above:{$S_4$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10cm,label=above:{$S_3$}] {};
-        \draw [<->,dashed](q3) -- (q4);
-
-      \end{tikzpicture}
-    }
-    \caption{Edge-order consistency violation}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Our Solution: Order Arbitration}
-  \begin{itemize}
-  \uncover<1->{
-  \item Transactions collect predecessors from each update
-  }
-  \uncover<2->{
-  \item Centralized arbiter maintains some global state
-  }
-  \uncover<3->{
-  \item Global state used to detect edge-order violations
-  }
-  \uncover<4->{
-  \item Offending transactions are aborted
-  }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Order Arbitration}
-    \begin{center}
-      \textbf{Rule: Collect predecessors after every successful update}
-    \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (3,0) -- (3,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (3,4.5) node [label=above:{$S_2$}] {};
-        \draw [black,<-,>=stealth] (7,0) -- (7,4) ;
-        \draw [black,<-,>=stealth] (10,0) -- (10,4) ;
-        \draw [black,dashed,->,>=stealth] (7,4.5) node [label=above:{$S_3$}] {} -- node [midway,above] {$E'$} (10,4.5) node [label=above:{$S_4$}] {};
-      }
-      \uncover<2->{
-        \draw [blue,->,>=stealth] (0,3.2) node (TXS) [label=left:{\small{$T_B$}}] {} -- (3,3);
-        \draw [red,->,>=stealth] (3,3.6) node (TXS) [label=right:{\small{$T_R$}}] {} -- (0,3.4);
-        \draw (3.1, 2.8) node [label=right:{\small{\textcolor{black}{$T_B : \{{T_R}\}$}}}] {};
-        \draw (3.2,2.3) rectangle (5,3.2);
-      }
-      \uncover<3->{
-        \draw [blue,->,>=stealth] (7,2.3) node [label=left:{\small{$T_B$}}] {} -- (10,1.8);
-        \draw [red,->,>=stealth] (7,1.8) node [label=left:{\small{$T_R$}}] {} -- (10,1.2) ;
-        \draw (10.1, 0.8) node (TXS) [yshift=-1cm,label=right:{\small{\textcolor{black}{$T_R : \{{T_B}\}$}}}] {};
-        \draw (10.2,-0.6) rectangle (12,0.3);
-      }
-      \uncover<4->{
-        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
-        (10.2,1.8) node[above,xshift=1cm, yshift=0.3cm] {ARBITER} -- (11,1.8);
-        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
-        (10.2,1.2)   -- (11,1.2);
-      }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Order Arbitration}
-      \begin{center}
-      \textbf{Rule: If transaction exists in hit list, abort. Else, merge predecessors into hit list}
-    \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{ % 1 arbiter
-        \node (s1) [server,label=$Arbiter$] {};
-      }
-      \uncover<2->{ % 2 hit list
-        \node (a) [rectangle, label=above:{$Hit \ List$}, minimum width=2cm,minimum height=1cm,draw=black,xshift=0cm,yshift=0cm] {$T_z$ ,$T_a$ {\uncover<7->{,$T_R$}}};
-      } % uncover T_R when T_B processed
-      \uncover<3->{ % 3 queue
-              \node (q) [rectangle, label=above:{$Queue$},left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=0cm] {{\uncover<6->{$...$}}};
-              \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=0cm] {{\uncover<5-8>{$T_R$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=0cm] {{\uncover<4-5>{$T_B$}}};
-        \node (q1) [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=-1cm] {{\uncover<6->{$...$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=-1cm] {{\uncover<5-8>{$T_B$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=-1cm] {{\uncover<4-5>{$T_R$}}};
-        \node [left of=q,xshift=0.2cm,label=left:{\small{\textcolor{black}{$tnx.id$}}}] {};
-        \node [left of=q1,xshift=0.2cm,label=left:{\small{\textcolor{black}{$pred$}}}] {};
-      }
-      \uncover<4-4> {
-        \node[right=0.5cm of s1] {$T_B$ \textbf{ENTERS QUEUE}};
-      }
-      \uncover<5-5> {
-        \node[right=0.5cm of s1] {$T_R$ \textbf{ENTERS QUEUE}};
-      }
-      \uncover<6-8> {
-        \node[right=0.5cm of s1,yshift=1cm] {$T_B$ \textbf{PROCESSED}};
-      }
-      \uncover<7-8> {
-        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{ADDED TO HIT LIST}};
-      }
-      \uncover<8-8> {
-        \node[right=0.5cm of s1,,yshift=-1cm] {$T_B$ \textbf{SUCCEEDS}};
-      }
-      \uncover<9-> {
-        \node[right=0.5cm of s1,yshift=1cm] {$T_R$ \textbf{PROCESSED}};
-      }
-      \uncover<10-> {
-        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{IN HIT LIST}};
-      }
-      \uncover<11-> {
-        \node[right=0.5cm of s1,yshift=-1cm] {$T_R$ \textbf{ABORTS}};
-      }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Order Arbitration Protocol}
-  If a transaction updates $1$ or more distributed edges:
-  \begin{itemize}
-    \uncover<1->{
-    \item Collect all predecessors
-    }
-    \uncover<2->{
-    \item Go to arbiter
-    }
-    \uncover<3->{
-    \item If not in hit list then proceed and enter predecessors into the hit list
-    }
-    \uncover<4->{
-    \item Else in hit list then abort
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Edge Concurrency Control Protocol Summary}
-    \begin{enumerate}
-    \item Collision detection: enforces \textbf{reciprocal consistency} for distributed edges
-    \item Order arbitration: enforces \textbf{edge-order consistency} between transactions
-    \end{enumerate}
-\end{frame}
-
-%------------------------------------------------
-% Performance Evaluation
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Performance Evaluation}
-  Performance measures of interest:
-  \begin{itemize}
-  \item Average number of transactions that are aborted
-  \item Load at the arbiter
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Evaluation Approach}
-  \begin{enumerate}
-  \item Developed approximate model
-  \item Measure accuracy of model through simulation
-  \end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Parameters}
-  \begin{itemize}
-  \item Database size
-  \item Transaction arrival rate
-  \item Average updates per transaction
-  \item Average arbiter service time
-  \item Average network delay
-  \end{itemize}
-\end{frame}
-
-%------------------------------------------------
-% Numerical and Simulation Results
-% ------------------------------------------------
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Database Size $(\boldsymbol{N})$}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_num_edges.png}
-    \caption{TPS $= 1000$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim1500$ TPS
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_tnx_arrival_rate}
-    \caption{\textcolor{red}{size $= 10K$}, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim 1100$ TPS
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/larger_network_delay.png}
-    \caption{size $= 10K$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, \textcolor{red}{av network delay $=10ms$}}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim 550$ TPS
-  \end{center}
-    \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_updates_per_tnx.png}
-        \caption{size $= 10K$, \textcolor{red}{av updates/tnx $= 10$}, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Summary and Future Work}
-  \begin{itemize}
-    \uncover<1->{
-    \item Model accuracy similar to simulations under a variety of parameter settings
-    }
-    \uncover<2->{
-    \item Between 1-4\% transactions are aborted
-    }
-    \uncover<3->{
-    \item Improve accuracy of simulation
-    }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}[standout]
- Thanks for listening! \\
-  \vspace{4mm}
-  Any Questions? \\
-  \vspace{4mm}
-  {\normalsize Email: \href{mailto:j.waudby2@newcastle.ac.uk}{j.waudby2@newcastle.ac.uk}} \\
-  {\normalsize Twitter: \href{https://twitter.com/waudberry_7}{@waudberry\_7}}
-\end{frame}
-
-\end{document}

### 2020-04-20 slides/figures/hc-edge.tex
corrections

--- 2020-04-20 slides/figures/dist-edge.tex
@@
     \node (rect)  [draw,rounded corners,minimum width=3.5cm,minimum height=4cm,label=above:{$S_i$}] {};
-  \node (rect)  [draw,rounded corners,minimum width=3.5cm,minimum height=4cm,label=above:{$S_j$},xshift=6cm] {};
+  \node (rect)  [draw,rounded corners,minimum width=3.5cm,minimum height=4cm,label=above:{$S_j$},xshift=5.6cm] {};
   \node (v1) [big_vertex,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};
-  \node (v2) [big_vertex,xshift=6cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
+
+  \node (v2) [big_vertex,xshift=5.6cm,yshift=0cm] {\small{\texttt{b:\textcolor{green}{Book}}}};
+
   \node [below of=v1,yshift=1cm] {\small{\texttt{\textcolor{red}{name}:Tolkien}}};
   \node [below of=v2,yshift=1cm] {\small{\texttt{\textcolor{red}{title}:The Hobbit}}};
-  \draw [thick,dashed,->,>=stealth] (0.9,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (5.1,0);
+
+  \draw [thick,dashed,->,>=stealth] (0.9,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below,xshift=0.1cm] {\small{\texttt{\textcolor{red}{year}:1937}}} (2.7,0);
 \end{tikzpicture}

### 2020-04-20 slides/figures/hc-storage-2.tex
corrections

\begin{tikzpicture}[node distance=2cm,scale=0.6,every node/.style={transform shape}]

% 1

  \node (v1) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=0cm] {\small{\texttt{vertex id}}};

    \node (b1) [draw,black,minimum height=0.9cm,minimum width=8cm,text width=0.2\linewidth,xshift=1.4cm,yshift=0cm] {};
    \draw[thick,->] (v1) -- (b1);

      \node (1i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=0cm,align=center] {\small{\texttt{edge}}};

      \node (1i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 1i1,text width=0.6\linewidth,align=center] {\small{\texttt{edge}}};

  % 2
  \node (v) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=-1.2cm] {\small{\texttt{b:Book}}};

  \node (b) [draw,black,minimum height=0.9cm,minimum width=8cm,text width=0.2\linewidth,xshift=1.4cm,yshift=-1.2cm] {};
    \draw[thick,->] (v) -- (b);

  \node (2i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=-1.2cm,align=center] {\small{\texttt{title:The \hspace{-0.15cm} Hobbit}}};

  \node (2i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 2i1,text width=0.6\linewidth,align=center,fill=red!30] {};

  % 3
   \node (3v1) [black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=-2.1cm] {$\vdots$};

      \node (3i1)  [black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=-2.1cm,align=center] {$\vdots$};

  \node (3i2)  [black,minimum height=0.6cm,minimum width=2.1cm,xshift=1.9cm,yshift=0cm,right of = 3i1,text width=0.6\linewidth,align=center] {$\vdots$};

  % 4

    \node (4v1) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=-3.2cm] {\small{\texttt{vertex id}}};

    \node (4b1) [draw,black,minimum height=0.9cm,minimum width=8cm,text width=0.2\linewidth,xshift=1.4cm,yshift=-3.2cm] {};
    \draw[thick,->] (4v1) -- (4b1);

      \node (4i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=-3.2cm,align=center] {\small{\texttt{property}}};

  \node (4i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 4i1,text width=0.6\linewidth,align=center] {\small{\texttt{edge}}};

\end{tikzpicture}

### 2020-04-20 slides/figures/hc-storage.tex
corrections

--- 2020-04-20 slides/figures/hc-storage-2.tex
@@
 
-  \node (v1) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=0cm] {\small{\texttt{vertex id}}};
+  \node (v1) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=0cm] {\small{\texttt{a:Person}}};
 
@@
 
-      \node (1i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=0cm,align=center] {\small{\texttt{edge}}};
+      \node (1i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=0cm,align=center,fill=green!30] {\small{\texttt{$\rightarrow$ wrote b}}};
 
-      \node (1i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 1i1,text width=0.6\linewidth,align=center] {\small{\texttt{edge}}};
+  \node (1i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 1i1,text width=0.6\linewidth,align=center,fill=green!30] {\small{\texttt{year:1937}}};
 
   % 2
-  \node (v) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=-1.2cm] {\small{\texttt{b:Book}}};
+  \node (v) [draw,black,minimum height=0.9cm,minimum width=1cm,text width=0.4\linewidth,xshift=-4.3cm,align=center,yshift=-1.2cm] {\small{\texttt{vertex id}}};
 
@@
 
-  \node (2i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=-1.2cm,align=center] {\small{\texttt{title:The \hspace{-0.15cm} Hobbit}}};
+  \node (2i1)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=-0.5cm,text width=0.6\linewidth,yshift=-1.2cm,align=center] {\small{\texttt{edge}}};
 
-  \node (2i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 2i1,text width=0.6\linewidth,align=center,fill=red!30] {};
+  \node (2i2)  [draw,black,minimum height=0.6cm,minimum width=2cm,xshift=1.9cm,yshift=0cm,right of = 2i1,text width=0.6\linewidth,align=center]  {\small{\texttt{property}}};
 
@@
 
+    \node (t) [right of = b, xshift = 2cm] {};
+    \node (t1) [right of = b1, xshift = 2cm] {};
+  \node (t2) [right of = 4b1, xshift = 2cm] {};
+
 \end{tikzpicture}

### 2020-04-20 slides/main.tex
corrections

@@
-% ----------------------------------------------------------------------------------------
-%	Configuration
-%----------------------------------------------------------------------------------------
-
 \documentclass[aspectratio=169]{beamer}
 
-\usetheme[titleformat=regular,numbering=none,progressbar=frametitle]{metropolis}
+\input{config}
 
-% newcastle colour scheme
-\definecolor{nclBlue}{HTML}{003A65}
-\definecolor{nclRed}{HTML}{D91A35}
-
-\setbeamercovered{transparent}
-\setbeamercovered{invisible}
-\setbeamercolor{progress bar}{fg=nclRed}
-\setbeamercolor{frametitle}{bg=nclBlue}
-\setbeamercolor{background canvas}{bg=white}
-\setbeamercolor{block body}{bg=gray!30}
-\setbeamercolor{block title}{bg=nclBlue,fg=white}
-\setbeamercolor{palette primary}{fg=white, bg=nclBlue}
-
-\linespread{1.5}
-
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{amssymb}
-\usepackage{pifont}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{mathspec}
-\usepackage[font=footnotesize]{caption}
-
-% Thin fonts
-\setsansfont[BoldFont={Fira Sans}, Numbers={OldStyle}]{Fira Sans Light}
-\setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Light}
-
-% Thick fonts
-% \setsansfont[BoldFont={Fira Sans SemiBold}, Numbers={OldStyle}]{Fira Sans}
-% \setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Book}
-
-% tikz configuration
-\usetikzlibrary{shapes.geometric, arrows}
-\usetikzlibrary{arrows, positioning, shapes}
-\usetikzlibrary{decorations.pathmorphing}
-\usetikzlibrary{calc}
-
-\tikzstyle{vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
-\tikzstyle{mini_vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
-\tikzstyle{server} = [rectangle, rounded corners,minimum width=3cm,minimum height=4cm,draw=black]
-\tikzstyle{mini_server} = [rectangle, rounded corners,minimum width=1.5cm,minimum height=1.9cm,draw=black]
-\tikzstyle{local} = [thick,->,>=stealth]
-\tikzstyle{dist} = [thick,dashed,->,>=stealth]
-\tikzstyle{bignode}=[draw, circle, minimum width=70pt,align=center]
-
-%----------------------------------------------------------------------------------------
-%	Title + Overview Slide
-% ----------------------------------------------------------------------------------------
-
-\title[]{Preserving Reciprocal Consistency in \\ Distributed Graph Databases}
-
-\author[Waudby \& Ezhilchelvan \& Webber \& Mitrani]{\underline{\emph{Jack Waudby}}$^{1}$, Paul Ezhilchelvan$^{1}$, Jim Webber$^{2}$ \& Isi Mitrani$^{1}$}
-\institute[NCL]
-{
-  $^{1}$Newcastle University\\
-  $^{2}$Neo4j
-}
-\date{April 27, 2020}
+\input{title}
 
@@
 
-\begin{frame}
-\maketitle
-\begin{tikzpicture}[overlay, remember picture]
-  \node[above left=2.8cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_newcastle_uni.png}};
-  \node[above left=1.2cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_neo4j.png}};
-\end{tikzpicture}
-\end{frame}
-
-%----------------------------------------------------------------------------------------
-%	PRESENTATION SLIDES
-%----------------------------------------------------------------------------------------
-
-%------------------------------------------------
-% Background
-% ------------------------------------------------
+\input{title-images}
 
@@
 
-\begin{frame}
-  \frametitle{Graph Databases}
-    \begin{figure}
-      \includegraphics[scale=0.5]{images/property_graph.png}
-      \caption{Labeled property graph\footnotemark}
-    \end{figure}
-  \footnotetext[1]{
-    \url{https://neo4j.com}
-  }
-\end{frame}
+\input{s1-property-graph}
 
-\begin{frame}
-  \frametitle{Graph Databases}
-  \begin{itemize}
-    \uncover<1->{
-    \item \textbf{Efficient graph analysis:} reachability, pattern matching, shortest path search and clustering
-    }
-    \uncover<2->{
-    \item \textbf{Use cases:} telecommunications, pharma, publishing, finance, social media
-    }
-  \end{itemize}
-\end{frame}
+\input{s2-storage-layer}
 
-\begin{frame}
-  \frametitle{Graph Database Popularity}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.3]{images/db_popularity.png}
-    \caption{Database popularity by data model\footnotemark}
-  \end{figure}
-  \footnotetext[1]{
-    \url{https://db-engines.com/en/ranking_categories}
-  }
-\end{frame}
+\input{s3-reciprocal-consistency}
 
-\begin{frame}
-  \frametitle{Distributed Graph Databases}
-  \begin{itemize}
-    \uncover<1->{
-    \item Graph exceeds the storage capacity of a single machine
-    }
-    \uncover<2->{
-    \item Partition the graph across multiple machines in a cluster, with partitions replicated for fault-tolerance and availability
-    }
-    \uncover<3->{
-    \item Graph partitioning inevitably leads to \textbf{distributed edges}
-    }
-  \end{itemize}
-\end{frame}
+\input{s4-dist-graph}
 
-\begin{frame}
-  \frametitle{Distributed Graph Databases: Partitioning}
-\input{./figures/graph}
-\end{frame}
+\input{s6-half-corruption}
 
-\begin{frame}
-  \frametitle{Distributed Graph Databases: Partitioning}
-  \begin{figure}[h!]
-    \centering
-    \resizebox{0.6\textwidth}{!}{
-    \begin{tikzpicture}[node distance=2cm]
-      \node (s1) [server,label=$S_1$] {};
-      \node (s2) [server, right of=s1,xshift=2cm,yshift=2cm,label=$S_2$] {};
-      \node (s3) [server, right of=s1,xshift=6cm,label=$S_3$] {};
+\input{s7-delta}
 
-      \node (v1) [vertex,xshift=-0.5cm,yshift=-1.5cm] {};
-      \node (v2) [vertex, above of=v1,xshift=-0.3cm,yshift=1cm] {};
-      \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
-      \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.3cm] {};
-      \node (v5) [vertex, above right of=v3,xshift=-0.6cm,yshift=-0.5cm] {};
-      \node (v6) [vertex, right of=v5,xshift=0.3cm,yshift=0.8cm] {};
-      \node (v7) [vertex, below of=v6, xshift=1cm,yshift=0.4cm] {};
-      \node (v8) [vertex, right of=v4, yshift=-0.3cm,xshift=4.5cm] {};
-      \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
-      \node (v10) [vertex, above of=v9,xshift=-2.5cm,yshift=-0.3cm] {};
-      \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
-      \node (v12) [vertex, above of=v11,yshift=-1cm,xshift=-1cm] {};
-      \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
-      \draw [local] (v2) -- (v1);
-      \draw [local] (v1) -- (v3);
-      \draw [local] (v3) -- (v2);
-      \draw [local] (v2) -- (v5);
-      \draw [local] (v1) -- (v4);
-      \draw [local] (v5) -- (v4);
-      \draw [dist] (v5) -- (v6);
-      \draw [dist] (v4) -- (v7);
-      \draw [dist] (v8) -- (v4);
-      \draw [local] (v8) -- (v9);
-      \draw [local] (v9) -- (v11);
-      \draw [local] (v10) -- (v7);
-      \draw [dist] (v10) -- (v12);
-      \draw [dist] (v12) -- (v13);
-      \draw [local] (v10) -- (v6);
-      \draw [dist] (v10) -- (v11);
-      \draw [local] (v13) -- (v10);
-      \draw [local] (v11) -- (v12);
-    \end{tikzpicture}
-  }
-\caption{Graph partitioned across 3 servers}
-  \end{figure}
-\end{frame}
+\input{s8-evaluation}
 
-%------------------------------------------------
-% Problem 1: Reciprocal Consistency}
-% ------------------------------------------------
+\input{s9-summary}
 
-\begin{frame}
-  \frametitle{Reciprocal Consistency}
-  \begin{center}
-  \textbf{A logical edge is represented by 2 physical records}
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=7cm]
-      \uncover<1-> {
-      \node (n1) [bignode] {\texttt{Flight}};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
-      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:AVAILABLE}} (n2);
-    }
-      \uncover<2-> {
-         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Seat Available In Flight}};
-         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Flight Has Seat Available}};
-         \draw [<->,dashed](q1) -- (q2);
-      }
-    \end{tikzpicture}
-    \caption{A reciprocal consistent edge}
-  \end{figure}
-\end{frame}
+% %------------------------------------------------
+% % Problem 1: Reciprocal Consistency}
+% % ------------------------------------------------
 
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  Two concurrent transactions:
-  \begin{itemize}
-  \item Mr Red requests to book the seat, $T_R$
-  \item Mr Blue requests to book the seat, $T_B$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=2cm]
-      \uncover<1-> {
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
-        \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
-      }
-      \uncover<2-3> {
-        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
-      }
-      \uncover<3-4> {
-        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
-      }
-      \uncover<4-> {
-        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
-      }
-      \uncover<5-> {
-        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
-      }
-    \end{tikzpicture}
-    \caption{Concurrent transactions interleave and violate reciprocal consistency}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Reciprocal Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}[node distance=7cm]
-      \node (n1) [bignode] {\texttt{Flight}};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
-      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
-
-      \uncover<2-> {
-      \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
-      \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
-      \draw [<->,dashed](q1) -- (q2);
-      \draw [<->,dashed](q1) -- (q2);
-      \coordinate (a) at (4,-2);
-      \coordinate (b) at (3,-3);
-      \coordinate (c) at (3,-2);
-      \coordinate (d) at (4,-3);
-      \draw [very thick,red] ($ (a) $) -- ($ (b) $);
-      \draw [very thick,red] ($ (c) $) -- ($ (d) $);
-    }
-    \end{tikzpicture}
-    \caption{Reciprocal inconsistent distributed edge}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Reciprocal Consistency Violation}
+%   Two concurrent transactions:
+%   \begin{itemize}
+%   \item Mr Red requests to book the seat, $T_R$
+%   \item Mr Blue requests to book the seat, $T_B$
+%   \end{itemize}
+% \end{frame}
 
@@
 %     \begin{tikzpicture}[node distance=2cm]
-%       \uncover<1->{
-%         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
-%         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
-%         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
-%         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
-%         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
+%       \uncover<1-> {
+%         \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
+%         \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
+%         \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
 %       }
-%       \uncover<2->{
-%         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
-%         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
-%         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
-%         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
-%         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
+%       \uncover<2-3> {
+%         \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
+%       }
+%       \uncover<3-4> {
+%         \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
+%       }
+%       \uncover<4-> {
+%         \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
+%       }
+%       \uncover<5-> {
+%         \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
 %       }
 %     \end{tikzpicture}
-%     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
+%     \caption{Concurrent transactions interleave and violate reciprocal consistency}
 %   \end{figure}
@@
 
-\begin{frame}
-  \frametitle{Distributed Edge Reciprocal Consistency}
-  \begin{itemize}
-    \uncover<1->{
-  \item Without concurrency control, distributed edges can become reciprocally inconsistent
-    \begin{itemize}
-    \item A distributed edge's reciprocal information is separated by the network
-    \end{itemize}
-  }
-  \uncover<2->{
-  \item Reciprocal inconsistency is a source of corruption\footnotemark
-  }
-  \uncover<3->{
-  \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
-    }
-    \end{itemize}
-    \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Reciprocal Consistency Violation}
+%   \begin{figure}[h!]
+%     \centering
+%     \begin{tikzpicture}[node distance=7cm]
+%       \node (n1) [bignode] {\texttt{Flight}};
+%       \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
+%       \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
+%       \node (n2) [bignode, right of=n1] {\texttt{Seat}};
+%       \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
 
-\begin{frame}
-  \frametitle{Our Solution: Collision Detection}
-  \begin{itemize}
-    \uncover<1->{
-    \item Carefully abort transaction(s)
-    }
-    \uncover<2->{
-    \item No centralized control or synchronized clock
-    }
-    \uncover<3->{
-    \item Permits interleavings that preserve reciprocal consistency
-    }
-  \end{itemize}
-\end{frame}
+%       \uncover<2-> {
+%       \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
+%       \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
+%       \draw [<->,dashed](q1) -- (q2);
+%       \draw [<->,dashed](q1) -- (q2);
+%       \coordinate (a) at (4,-2);
+%       \coordinate (b) at (3,-3);
+%       \coordinate (c) at (3,-2);
+%       \coordinate (d) at (4,-3);
+%       \draw [very thick,red] ($ (a) $) -- ($ (b) $);
+%       \draw [very thick,red] ($ (c) $) -- ($ (d) $);
+%     }
+%     \end{tikzpicture}
+%     \caption{Reciprocal inconsistent distributed edge}
+%   \end{figure}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Collision Detection General Rules}
-  \begin{itemize}
-    \uncover<1->{
-    \item Updates are initially provisional
-    }
-    \uncover<2->{
-    \item Transactions distinguish between their first and second updates using labels
-    }
-    \uncover<3->{
-    \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
-    }
-  \end{itemize}
-\end{frame}
+% % \begin{frame}
+% %   \frametitle{Reciprocal Consistency Violation}
+% %   \begin{figure}[h!]
+% %     \centering
+% %     \begin{tikzpicture}[node distance=2cm]
+% %       \uncover<1->{
+% %         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
+% %         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
+% %         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
+% %         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
+% %         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
+% %       }
+% %       \uncover<2->{
+% %         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
+% %         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
+% %         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
+% %         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
+% %         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
+% %       }
+% %     \end{tikzpicture}
+% %     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
+% %   \end{figure}
+% % \end{frame}
 
-\begin{frame}
-  \frametitle{Interference-free Update}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
-        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
-      }
-      \uncover<2->{
-        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
-      }
-      \uncover<3->{
-        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
-      }
-      \uncover<3->{
-        \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
-      }
-      \uncover<4->{
-        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
-      }
-    \end{tikzpicture}
-  \end{figure}
+% \begin{frame}
+%   \frametitle{Distributed Edge Reciprocal Consistency}
+%   \begin{itemize}
+%     \uncover<1->{
+%   \item Without concurrency control, distributed edges can become reciprocally inconsistent
+%     \begin{itemize}
+%     \item A distributed edge's reciprocal information is separated by the network
+%     \end{itemize}
+%   }
+%   \uncover<2->{
+%   \item Reciprocal inconsistency is a source of corruption\footnotemark
+%   }
+%   \uncover<3->{
+%   \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
+%     }
+%     \end{itemize}
+%     \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
+% \end{frame}
 
-\end{frame}
+% \begin{frame}
+%   \frametitle{Our Solution: Collision Detection}
+%   \begin{itemize}
+%     \uncover<1->{
+%     \item Carefully abort transaction(s)
+%     }
+%     \uncover<2->{
+%     \item No centralized control or synchronized clock
+%     }
+%     \uncover<3->{
+%     \item Permits interleavings that preserve reciprocal consistency
+%     }
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Collision Detection}
-  \begin{center}
-    \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<2->{
-        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
-        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
-      }
-      \uncover<2->{
-        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
-      }
-      \uncover<3->{
-        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
-      }
-      \uncover<4-11>{
-        \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
-      }
-      \uncover<5-10>{
-        \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
-      }
-      \uncover<6-10>{
-        \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
-      }
-      \uncover<7-7>{
-        \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
-      }
-      \uncover<8-10>{
-        \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
-        \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
-      }
-      \uncover<9-9>{
-        \node [] at (6,2) {$T_R$'s 2 is missing};
-      }
-      \uncover<10-10>{
-        \node[] at (6,2) {$T_B$ ABORTS};
-      }
-      \uncover<12->{
-        \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
-      }
-      \uncover<13->{
-        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
-      }
-      \uncover<14->{
-        \node[] at (6,2) {$T_R$ SUCCEEDS};
-        }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Collision Detection General Rules}
+%   \begin{itemize}
+%     \uncover<1->{
+%     \item Updates are initially provisional
+%     }
+%     \uncover<2->{
+%     \item Transactions distinguish between their first and second updates using labels
+%     }
+%     \uncover<3->{
+%     \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
+%     }
+%   \end{itemize}
+% \end{frame}
 
-%------------------------------------------------
-%Problem 2: Edge-Order Consistency
-% ------------------------------------------------
+% \begin{frame}
+%   \frametitle{Interference-free Update}
+%   \begin{figure}[h!]
+%     \centering
+%     \begin{tikzpicture}
+%       \uncover<1->{
+%         \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
+%         \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
+%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
+%         \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
+%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
+%       }
+%       \uncover<2->{
+%         \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
+%       }
+%       \uncover<3->{
+%         \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
+%       }
+%       \uncover<3->{
+%         \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
+%       }
+%       \uncover<4->{
+%         \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
+%       }
+%     \end{tikzpicture}
+%   \end{figure}
 
-\begin{frame}
-  \frametitle{Edge-Order Consistency}
-  A second consistency problem related to distributed edges:
-  \begin{itemize}
-    \uncover<2->{
-    \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
-    }
-    \uncover<3->{
-    \item Each update is reciprocally consistent
-    }
-    \uncover<4->{
-    \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
-    }
-  \end{itemize}
-\end{frame}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  Consider two edges,
-  \begin{figure}[h!]
-    \centering
-    \resizebox{1\textwidth}{!}{
-      \begin{tikzpicture}[node distance=5.2cm]
-        \node[] at (2.6,2.5) {\large{$E$}};
-        \node[] at (13,2.5) {\large{$E'$}};
-        \node (n1) [bignode] {\texttt{Flight}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
-        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-        \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
-        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
-        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
-         \draw [<->,dashed](q1) -- (q2);
+% \begin{frame}
+%   \frametitle{Collision Detection}
+%   \begin{center}
+%     \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
+%   \end{center}
+%   \begin{figure}[h!]
+%     \centering
+%     \begin{tikzpicture}
+%       \uncover<2->{
+%         \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
+%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
+%         \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
+%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
+%         \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
+%       }
+%       \uncover<2->{
+%         \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
+%       }
+%       \uncover<3->{
+%         \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
+%       }
+%       \uncover<4-11>{
+%         \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
+%       }
+%       \uncover<5-10>{
+%         \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
+%       }
+%       \uncover<6-10>{
+%         \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
+%       }
+%       \uncover<7-7>{
+%         \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
+%       }
+%       \uncover<8-10>{
+%         \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
+%         \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
+%       }
+%       \uncover<9-9>{
+%         \node [] at (6,2) {$T_R$'s 2 is missing};
+%       }
+%       \uncover<10-10>{
+%         \node[] at (6,2) {$T_B$ ABORTS};
+%       }
+%       \uncover<12->{
+%         \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
+%       }
+%       \uncover<13->{
+%         \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
+%       }
+%       \uncover<14->{
+%         \node[] at (6,2) {$T_R$ SUCCEEDS};
+%         }
+%     \end{tikzpicture}
+%   \end{figure}
+% \end{frame}
 
-        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
-        \node (n4) [bignode, right of=n3] {\texttt{Room}};
-        \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
-        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
-        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
-        \draw [<->,dashed](q3) -- (q4);
+% %------------------------------------------------
+% %Problem 2: Edge-Order Consistency
+% % ------------------------------------------------
 
-      \end{tikzpicture}
-    }
-    \caption{Two distributed edges}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Edge-Order Consistency}
+%   A second consistency problem related to distributed edges:
+%   \begin{itemize}
+%     \uncover<2->{
+%     \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
+%     }
+%     \uncover<3->{
+%     \item Each update is reciprocally consistent
+%     }
+%     \uncover<4->{
+%     \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
+%     }
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  Two concurrent transactions by a travel agent:
-  \begin{itemize}
-  \item Requests to book room and seat for Mr Red, $T_R$
-  \item Requests to book room and seat for Mr Blue, $T_B$
-  \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Edge-Order Consistency Violation}
+%   Consider two edges,
+%   \begin{figure}[h!]
+%     \centering
+%     \resizebox{1\textwidth}{!}{
+%       \begin{tikzpicture}[node distance=5.2cm]
+%         \node[] at (2.6,2.5) {\large{$E$}};
+%         \node[] at (13,2.5) {\large{$E'$}};
+%         \node (n1) [bignode] {\texttt{Flight}};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
+%         \node (n2) [bignode, right of=n1] {\texttt{Seat}};
+%         \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
+%         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
+%         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
+%          \draw [<->,dashed](q1) -- (q2);
 
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
+%         \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
+%         \node (n4) [bignode, right of=n3] {\texttt{Room}};
+%         \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
+%         \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
+%         \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
+%         \draw [<->,dashed](q3) -- (q4);
 
-        \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
-        \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
-        \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
-        \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
-        \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
-        \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
-        \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
-      }
-      \uncover<2->{
-        \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
-      }
-      \uncover<3->{
-        \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
-      }
-      \uncover<4->{
+%       \end{tikzpicture}
+%     }
+%     \caption{Two distributed edges}
+%   \end{figure}
+% \end{frame}
 
-        \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
-      }
-      \uncover<5->{
+% \begin{frame}
+%   \frametitle{Edge-Order Consistency Violation}
+%   Two concurrent transactions by a travel agent:
+%   \begin{itemize}
+%   \item Requests to book room and seat for Mr Red, $T_R$
+%   \item Requests to book room and seat for Mr Blue, $T_B$
+%   \end{itemize}
+% \end{frame}
 
-        \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
-      }
-      \uncover<6->{
-        \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
-        \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
-        \draw (8.3,1) rectangle (11,2.5);
-      }
-    \end{tikzpicture}
-    \caption{Edge-order consistency violation}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Edge-Order Consistency Violation}
+%   \begin{figure}[h!]
+%     \centering
+%     \begin{tikzpicture}
+%       \uncover<1->{
 
-\begin{frame}
-  \frametitle{Edge-Order Consistency Violation}
-  \begin{figure}[h!]
-    \centering
-    \resizebox{1\textwidth}{!}{
-      \begin{tikzpicture}[node distance=5cm]
-        % edge labels
-        \node[] at (2.5,3) {\large{$E$}};
-        \node[] at (12.5,3) {\large{$E'$}};
-        % edge 1: flight -> seat
-        \node (n1) [bignode] {\texttt{Flight}};
-        \node (n2) [bignode, right of=n1] {\texttt{Seat}};
-        % edge 1: servers
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5cm,label=above:{$S_2$}] {};
-        % edge 1
-        \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
-        % edge 1 information
-        \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
-        \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
-        \draw [<->,dashed](q1) -- (q2);
+%         \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
+%         \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
+%         \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
+%         \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
+%         \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
+%         \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
+%         \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
+%       }
+%       \uncover<2->{
+%         \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
+%       }
+%       \uncover<3->{
+%         \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
+%       }
+%       \uncover<4->{
 
-        % edge 2: hotel -> room
-        \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
-        \node (n4) [bignode, right of=n3] {\texttt{Room}};
-        \draw [local] (n3) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n4);
-        \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
-        \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15cm,label=above:{$S_4$}] {};
-        \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10cm,label=above:{$S_3$}] {};
-        \draw [<->,dashed](q3) -- (q4);
+%         \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
+%       }
+%       \uncover<5->{
 
-      \end{tikzpicture}
-    }
-    \caption{Edge-order consistency violation}
-  \end{figure}
-\end{frame}
+%         \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
+%       }
+%       \uncover<6->{
+%         \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
+%         \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
+%         \draw (8.3,1) rectangle (11,2.5);
+%       }
+%     \end{tikzpicture}
+%     \caption{Edge-order consistency violation}
+%   \end{figure}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Our Solution: Order Arbitration}
-  \begin{itemize}
-  \uncover<1->{
-  \item Transactions collect predecessors from each update
-  }
-  \uncover<2->{
-  \item Centralized arbiter maintains some global state
-  }
-  \uncover<3->{
-  \item Global state used to detect edge-order violations
-  }
-  \uncover<4->{
-  \item Offending transactions are aborted
-  }
-  \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Edge-Order Consistency Violation}
+%   \begin{figure}[h!]
+%     \centering
+%     \resizebox{1\textwidth}{!}{
+%       \begin{tikzpicture}[node distance=5cm]
+%         % edge labels
+%         \node[] at (2.5,3) {\large{$E$}};
+%         \node[] at (12.5,3) {\large{$E'$}};
+%         % edge 1: flight -> seat
+%         \node (n1) [bignode] {\texttt{Flight}};
+%         \node (n2) [bignode, right of=n1] {\texttt{Seat}};
+%         % edge 1: servers
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5cm,label=above:{$S_2$}] {};
+%         % edge 1
+%         \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);
+%         % edge 1 information
+%         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
+%         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5cm,yshift=-2cm] {\textcolor{red}{\textit{Mr Red Booked Seat}}};
+%         \draw [<->,dashed](q1) -- (q2);
 
-\begin{frame}
-  \frametitle{Order Arbitration}
-    \begin{center}
-      \textbf{Rule: Collect predecessors after every successful update}
-    \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{
-        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
-        \draw [black,<-,>=stealth] (3,0) -- (3,4) ;
-        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (3,4.5) node [label=above:{$S_2$}] {};
-        \draw [black,<-,>=stealth] (7,0) -- (7,4) ;
-        \draw [black,<-,>=stealth] (10,0) -- (10,4) ;
-        \draw [black,dashed,->,>=stealth] (7,4.5) node [label=above:{$S_3$}] {} -- node [midway,above] {$E'$} (10,4.5) node [label=above:{$S_4$}] {};
-      }
-      \uncover<2->{
-        \draw [blue,->,>=stealth] (0,3.2) node (TXS) [label=left:{\small{$T_B$}}] {} -- (3,3);
-        \draw [red,->,>=stealth] (3,3.6) node (TXS) [label=right:{\small{$T_R$}}] {} -- (0,3.4);
-        \draw (3.1, 2.8) node [label=right:{\small{\textcolor{black}{$T_B : \{{T_R}\}$}}}] {};
-        \draw (3.2,2.3) rectangle (5,3.2);
-      }
-      \uncover<3->{
-        \draw [blue,->,>=stealth] (7,2.3) node [label=left:{\small{$T_B$}}] {} -- (10,1.8);
-        \draw [red,->,>=stealth] (7,1.8) node [label=left:{\small{$T_R$}}] {} -- (10,1.2) ;
-        \draw (10.1, 0.8) node (TXS) [yshift=-1cm,label=right:{\small{\textcolor{black}{$T_R : \{{T_B}\}$}}}] {};
-        \draw (10.2,-0.6) rectangle (12,0.3);
-      }
-      \uncover<4->{
-        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
-        (10.2,1.8) node[above,xshift=1cm, yshift=0.3cm] {ARBITER} -- (11,1.8);
-        \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
-        (10.2,1.2)   -- (11,1.2);
-      }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
+%         % edge 2: hotel -> room
+%         \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
+%         \node (n4) [bignode, right of=n3] {\texttt{Room}};
+%         \draw [local] (n3) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n4);
+%         \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
+%         \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15cm,yshift=-2cm] {\textcolor{blue}{\textit{Mr Blue Booked Room}}};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15cm,label=above:{$S_4$}] {};
+%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10cm,label=above:{$S_3$}] {};
+%         \draw [<->,dashed](q3) -- (q4);
 
-\begin{frame}
-  \frametitle{Order Arbitration}
-      \begin{center}
-      \textbf{Rule: If transaction exists in hit list, abort. Else, merge predecessors into hit list}
-    \end{center}
-  \begin{figure}[h!]
-    \centering
-    \begin{tikzpicture}
-      \uncover<1->{ % 1 arbiter
-        \node (s1) [server,label=$Arbiter$] {};
-      }
-      \uncover<2->{ % 2 hit list
-        \node (a) [rectangle, label=above:{$Hit \ List$}, minimum width=2cm,minimum height=1cm,draw=black,xshift=0cm,yshift=0cm] {$T_z$ ,$T_a$ {\uncover<7->{,$T_R$}}};
-      } % uncover T_R when T_B processed
-      \uncover<3->{ % 3 queue
-              \node (q) [rectangle, label=above:{$Queue$},left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=0cm] {{\uncover<6->{$...$}}};
-              \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=0cm] {{\uncover<5-8>{$T_R$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=0cm] {{\uncover<4-5>{$T_B$}}};
-        \node (q1) [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=-1cm] {{\uncover<6->{$...$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=-1cm] {{\uncover<5-8>{$T_B$}}};
-        \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=-1cm] {{\uncover<4-5>{$T_R$}}};
-        \node [left of=q,xshift=0.2cm,label=left:{\small{\textcolor{black}{$tnx.id$}}}] {};
-        \node [left of=q1,xshift=0.2cm,label=left:{\small{\textcolor{black}{$pred$}}}] {};
-      }
-      \uncover<4-4> {
-        \node[right=0.5cm of s1] {$T_B$ \textbf{ENTERS QUEUE}};
-      }
-      \uncover<5-5> {
-        \node[right=0.5cm of s1] {$T_R$ \textbf{ENTERS QUEUE}};
-      }
-      \uncover<6-8> {
-        \node[right=0.5cm of s1,yshift=1cm] {$T_B$ \textbf{PROCESSED}};
-      }
-      \uncover<7-8> {
-        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{ADDED TO HIT LIST}};
-      }
-      \uncover<8-8> {
-        \node[right=0.5cm of s1,,yshift=-1cm] {$T_B$ \textbf{SUCCEEDS}};
-      }
-      \uncover<9-> {
-        \node[right=0.5cm of s1,yshift=1cm] {$T_R$ \textbf{PROCESSED}};
-      }
-      \uncover<10-> {
-        \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{IN HIT LIST}};
-      }
-      \uncover<11-> {
-        \node[right=0.5cm of s1,yshift=-1cm] {$T_R$ \textbf{ABORTS}};
-      }
-    \end{tikzpicture}
-  \end{figure}
-\end{frame}
+%       \end{tikzpicture}
+%     }
+%     \caption{Edge-order consistency violation}
+%   \end{figure}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Order Arbitration Protocol}
-  If a transaction updates $1$ or more distributed edges:
-  \begin{itemize}
-    \uncover<1->{
-    \item Collect all predecessors
-    }
-    \uncover<2->{
-    \item Go to arbiter
-    }
-    \uncover<3->{
-    \item If not in hit list then proceed and enter predecessors into the hit list
-    }
-    \uncover<4->{
-    \item Else in hit list then abort
-    }
-  \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Our Solution: Order Arbitration}
+%   \begin{itemize}
+%   \uncover<1->{
+%   \item Transactions collect predecessors from each update
+%   }
+%   \uncover<2->{
+%   \item Centralized arbiter maintains some global state
+%   }
+%   \uncover<3->{
+%   \item Global state used to detect edge-order violations
+%   }
+%   \uncover<4->{
+%   \item Offending transactions are aborted
+%   }
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Edge Concurrency Control Protocol Summary}
-    \begin{enumerate}
-    \item Collision detection: enforces \textbf{reciprocal consistency} for distributed edges
-    \item Order arbitration: enforces \textbf{edge-order consistency} between transactions
-    \end{enumerate}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Order Arbitration}
+%     \begin{center}
+%       \textbf{Rule: Collect predecessors after every successful update}
+%     \end{center}
+%   \begin{figure}[h!]
+%     \centering
+%     \begin{tikzpicture}
+%       \uncover<1->{
+%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
+%         \draw [black,<-,>=stealth] (3,0) -- (3,4) ;
+%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (3,4.5) node [label=above:{$S_2$}] {};
+%         \draw [black,<-,>=stealth] (7,0) -- (7,4) ;
+%         \draw [black,<-,>=stealth] (10,0) -- (10,4) ;
+%         \draw [black,dashed,->,>=stealth] (7,4.5) node [label=above:{$S_3$}] {} -- node [midway,above] {$E'$} (10,4.5) node [label=above:{$S_4$}] {};
+%       }
+%       \uncover<2->{
+%         \draw [blue,->,>=stealth] (0,3.2) node (TXS) [label=left:{\small{$T_B$}}] {} -- (3,3);
+%         \draw [red,->,>=stealth] (3,3.6) node (TXS) [label=right:{\small{$T_R$}}] {} -- (0,3.4);
+%         \draw (3.1, 2.8) node [label=right:{\small{\textcolor{black}{$T_B : \{{T_R}\}$}}}] {};
+%         \draw (3.2,2.3) rectangle (5,3.2);
+%       }
+%       \uncover<3->{
+%         \draw [blue,->,>=stealth] (7,2.3) node [label=left:{\small{$T_B$}}] {} -- (10,1.8);
+%         \draw [red,->,>=stealth] (7,1.8) node [label=left:{\small{$T_R$}}] {} -- (10,1.2) ;
+%         \draw (10.1, 0.8) node (TXS) [yshift=-1cm,label=right:{\small{\textcolor{black}{$T_R : \{{T_B}\}$}}}] {};
+%         \draw (10.2,-0.6) rectangle (12,0.3);
+%       }
+%       \uncover<4->{
+%         \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
+%         (10.2,1.8) node[above,xshift=1cm, yshift=0.3cm] {ARBITER} -- (11,1.8);
+%         \draw [->,decorate,decoration={snake,amplitude=.4mm,segment length=2mm,post length=1mm}]
+%         (10.2,1.2)   -- (11,1.2);
+%       }
+%     \end{tikzpicture}
+%   \end{figure}
+% \end{frame}
 
-%------------------------------------------------
-% Performance Evaluation
-% ------------------------------------------------
+% \begin{frame}
+%   \frametitle{Order Arbitration}
+%       \begin{center}
+%       \textbf{Rule: If transaction exists in hit list, abort. Else, merge predecessors into hit list}
+%     \end{center}
+%   \begin{figure}[h!]
+%     \centering
+%     \begin{tikzpicture}
+%       \uncover<1->{ % 1 arbiter
+%         \node (s1) [server,label=$Arbiter$] {};
+%       }
+%       \uncover<2->{ % 2 hit list
+%         \node (a) [rectangle, label=above:{$Hit \ List$}, minimum width=2cm,minimum height=1cm,draw=black,xshift=0cm,yshift=0cm] {$T_z$ ,$T_a$ {\uncover<7->{,$T_R$}}};
+%       } % uncover T_R when T_B processed
+%       \uncover<3->{ % 3 queue
+%               \node (q) [rectangle, label=above:{$Queue$},left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=0cm] {{\uncover<6->{$...$}}};
+%               \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=0cm] {{\uncover<5-8>{$T_R$}}};
+%         \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=0cm] {{\uncover<4-5>{$T_B$}}};
+%         \node (q1) [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-4.5cm,yshift=-1cm] {{\uncover<6->{$...$}}};
+%         \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-3.5cm,yshift=-1cm] {{\uncover<5-8>{$T_B$}}};
+%         \node [rectangle, left of=s1, minimum width=1cm,minimum height=1cm,draw=black,xshift=-2.5cm,yshift=-1cm] {{\uncover<4-5>{$T_R$}}};
+%         \node [left of=q,xshift=0.2cm,label=left:{\small{\textcolor{black}{$tnx.id$}}}] {};
+%         \node [left of=q1,xshift=0.2cm,label=left:{\small{\textcolor{black}{$pred$}}}] {};
+%       }
+%       \uncover<4-4> {
+%         \node[right=0.5cm of s1] {$T_B$ \textbf{ENTERS QUEUE}};
+%       }
+%       \uncover<5-5> {
+%         \node[right=0.5cm of s1] {$T_R$ \textbf{ENTERS QUEUE}};
+%       }
+%       \uncover<6-8> {
+%         \node[right=0.5cm of s1,yshift=1cm] {$T_B$ \textbf{PROCESSED}};
+%       }
+%       \uncover<7-8> {
+%         \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{ADDED TO HIT LIST}};
+%       }
+%       \uncover<8-8> {
+%         \node[right=0.5cm of s1,,yshift=-1cm] {$T_B$ \textbf{SUCCEEDS}};
+%       }
+%       \uncover<9-> {
+%         \node[right=0.5cm of s1,yshift=1cm] {$T_R$ \textbf{PROCESSED}};
+%       }
+%       \uncover<10-> {
+%         \node[right=0.5cm of s1,yshift=0cm] {$T_R$ \textbf{IN HIT LIST}};
+%       }
+%       \uncover<11-> {
+%         \node[right=0.5cm of s1,yshift=-1cm] {$T_R$ \textbf{ABORTS}};
+%       }
+%     \end{tikzpicture}
+%   \end{figure}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Performance Evaluation}
-  Performance measures of interest:
-  \begin{itemize}
-  \item Average number of transactions that are aborted
-  \item Load at the arbiter
-  \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Order Arbitration Protocol}
+%   If a transaction updates $1$ or more distributed edges:
+%   \begin{itemize}
+%     \uncover<1->{
+%     \item Collect all predecessors
+%     }
+%     \uncover<2->{
+%     \item Go to arbiter
+%     }
+%     \uncover<3->{
+%     \item If not in hit list then proceed and enter predecessors into the hit list
+%     }
+%     \uncover<4->{
+%     \item Else in hit list then abort
+%     }
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Evaluation Approach}
-  \begin{enumerate}
-  \item Developed approximate model
-  \item Measure accuracy of model through simulation
-  \end{enumerate}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Edge Concurrency Control Protocol Summary}
+%     \begin{enumerate}
+%     \item Collision detection: enforces \textbf{reciprocal consistency} for distributed edges
+%     \item Order arbitration: enforces \textbf{edge-order consistency} between transactions
+%     \end{enumerate}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Parameters}
-  \begin{itemize}
-  \item Database size
-  \item Transaction arrival rate
-  \item Average updates per transaction
-  \item Average arbiter service time
-  \item Average network delay
-  \end{itemize}
-\end{frame}
+% %------------------------------------------------
+% % Performance Evaluation
+% % ------------------------------------------------
 
-%------------------------------------------------
-% Numerical and Simulation Results
-% ------------------------------------------------
+% \begin{frame}
+%   \frametitle{Performance Evaluation}
+%   Performance measures of interest:
+%   \begin{itemize}
+%   \item Average number of transactions that are aborted
+%   \item Load at the arbiter
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Database Size $(\boldsymbol{N})$}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_num_edges.png}
-    \caption{TPS $= 1000$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Evaluation Approach}
+%   \begin{enumerate}
+%   \item Developed approximate model
+%   \item Measure accuracy of model through simulation
+%   \end{enumerate}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim1500$ TPS
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_tnx_arrival_rate}
-    \caption{\textcolor{red}{size $= 10K$}, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Parameters}
+%   \begin{itemize}
+%   \item Database size
+%   \item Transaction arrival rate
+%   \item Average updates per transaction
+%   \item Average arbiter service time
+%   \item Average network delay
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim 1100$ TPS
-  \end{center}
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/larger_network_delay.png}
-    \caption{size $= 10K$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, \textcolor{red}{av network delay $=10ms$}}
-  \end{figure}
-\end{frame}
+% %------------------------------------------------
+% % Numerical and Simulation Results
+% % ------------------------------------------------
 
-\begin{frame}
-  \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
-  \begin{center}
-    Arbiter queue unstable at $\sim 550$ TPS
-  \end{center}
-    \begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.35]{images/varying_updates_per_tnx.png}
-        \caption{size $= 10K$, \textcolor{red}{av updates/tnx $= 10$}, av arbiter service time $= 10ms$, av network delay $=5ms$}
-  \end{figure}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Database Size $(\boldsymbol{N})$}
+%   \begin{figure}[h!]
+%     \centering
+%     \includegraphics[scale=0.35]{images/varying_num_edges.png}
+%     \caption{TPS $= 1000$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
+%   \end{figure}
+% \end{frame}
 
-\begin{frame}
-  \frametitle{Summary and Future Work}
-  \begin{itemize}
-    \uncover<1->{
-    \item Model accuracy similar to simulations under a variety of parameter settings
-    }
-    \uncover<2->{
-    \item Between 1-4\% transactions are aborted
-    }
-    \uncover<3->{
-    \item Improve accuracy of simulation
-    }
-  \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
+%   \begin{center}
+%     Arbiter queue unstable at $\sim1500$ TPS
+%   \end{center}
+%   \begin{figure}[h!]
+%     \centering
+%     \includegraphics[scale=0.35]{images/varying_tnx_arrival_rate}
+%     \caption{\textcolor{red}{size $= 10K$}, av updates/tnx $= 5$, av arbiter service time $= 10ms$, av network delay $=5ms$}
+%   \end{figure}
+% \end{frame}
 
-\begin{frame}[standout]
- Thanks for listening! \\
-  \vspace{4mm}
-  Any Questions? \\
-  \vspace{4mm}
-  {\normalsize Email: \href{mailto:j.waudby2@newcastle.ac.uk}{j.waudby2@newcastle.ac.uk}} \\
-  {\normalsize Twitter: \href{https://twitter.com/waudberry_7}{@waudberry\_7}}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
+%   \begin{center}
+%     Arbiter queue unstable at $\sim 1100$ TPS
+%   \end{center}
+%   \begin{figure}[h!]
+%     \centering
+%     \includegraphics[scale=0.35]{images/larger_network_delay.png}
+%     \caption{size $= 10K$, av updates/tnx $= 5$, av arbiter service time $= 10ms$, \textcolor{red}{av network delay $=10ms$}}
+%   \end{figure}
+% \end{frame}
+
+% \begin{frame}
+%   \frametitle{Aborts per second $(\boldsymbol{R})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
+%   \begin{center}
+%     Arbiter queue unstable at $\sim 550$ TPS
+%   \end{center}
+%     \begin{figure}[h!]
+%     \centering
+%     \includegraphics[scale=0.35]{images/varying_updates_per_tnx.png}
+%         \caption{size $= 10K$, \textcolor{red}{av updates/tnx $= 10$}, av arbiter service time $= 10ms$, av network delay $=5ms$}
+%   \end{figure}
+% \end{frame}
+
+% \begin{frame}
+%   \frametitle{Summary and Future Work}
+%   \begin{itemize}
+%     \uncover<1->{
+%     \item Model accuracy similar to simulations under a variety of parameter settings
+%     }
+%     \uncover<2->{
+%     \item Between 1-4\% transactions are aborted
+%     }
+%     \uncover<3->{
+%     \item Improve accuracy of simulation
+%     }
+%   \end{itemize}
+% \end{frame}
+
+% \begin{frame}[standout]
+%  Thanks for listening! \\
+%   \vspace{4mm}
+%   Any Questions? \\
+%   \vspace{4mm}
+%   {\normalsize Email: \href{mailto:j.waudby2@newcastle.ac.uk}{j.waudby2@newcastle.ac.uk}} \\
+%   {\normalsize Twitter: \href{https://twitter.com/waudberry_7}{@waudberry\_7}}
+% \end{frame}
 

### 2020-04-20 slides/s7-delta.tex
corrections

\begin{frame}
  \frametitle{Protocol Design Considerations}
  Design Considerations:
  \begin{enumerate}
  \item Graph workloads exhibit high contention.
  \item Graph transactions tend to be long-lived than those in other databases.
  \end{enumerate}
\textbf{Protocol must permit multiple updates on the same record provided they are they are \textit{sufficiently} apart in time in ensure reciprocal consistency}
\end{frame}

\begin{frame}
  \frametitle{\tDelta Protocol}
  \begin{itemize}
  \item Fact: a transaction updating a distributed edge must update one edge pointer then \textbf{immediately} update the other.
  \item Rule: an update is permitted if the immediately preceding update was done at least $\Delta$ time before. Else, abort.
  \item Assumption: the time interval that elapses between completing at update at one end and starting at the other can be estimated, $\delta$. Choosing $\Delta > \delta$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{\tDelta Protocol: Example}
  \begin{center}
    \textbf{\small{Rule: an update is permitted if the preceding update was done at least $\Delta$ time before. Else, abort}}
  \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<2->{
        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
        \draw [black,<-,>=stealth] (0,0) -- (0,4) node [label=above:{$S_i$}] {};
        \draw [black,<-,>=stealth] (4,0) -- (4,4)  node [label=above:{$S_j$}] {};
              \draw [white,<-,>=stealth] (7,1) -- (7,3) ;
      }
      \uncover<2->{
        \node [] at (-0.5,4) {$T_x$};
      }
      \uncover<3-6>{
        \draw [->,>=stealth] (0,3.5) -- (2,1.9);
      }
      \uncover<4->{
        \node [] at (4.5,3.3) {$T_y$};
      }
      \uncover<5-5>{
        \draw [->,>=stealth] (4,3.3) -- (2,2.9);
      }
      \uncover<6-8>{
        \draw [->,>=stealth] (4,3.5) -- (0,2.5);
      }
      \uncover<6-8>{
        \draw [<->,>=stealth] (-0.3,2.5) -- (-0.3,3.3);
        \node[] at (-0.8,3) {$< \Delta$};
        \node[] at (-1.5,2.5) {$T_y$ ABORTS};
      }
      \uncover<7->{
        \draw [->,>=stealth] (0,3.5)  -- (4,0.3);
      }
      \uncover<8-8>{
        \node[] at (6,1) {$T_x$ SUCCEEDS};
        \draw [<->,>=stealth] (4.3,0.3) -- (4.3,3.3);
        \node[] at (4.8,1.8) {$> \Delta$};
      }
    \end{tikzpicture}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{\tDelta Protocol}
  \begin{itemize}
  \item Reciprocal consistency is preserved if the time taken to complete am update at one end and start at other end remains less than  $\Delta$
  \item If $\Delta$ is exceeded then reciprocal inconsistency can occur
  \item Setting a large $\Delta$ tends to preserve consistency but leads to more aborted transactions
  \end{itemize}
\end{frame}

### 2020-04-20 slides/s8-evaluation.tex
corrections

\begin{frame}
  \frametitle{Performance Evaluation}
  Two metrics the following two metrics for various values of $\Delta$:
  \begin{itemize}
  \item Time taken for 10\% of a large database to be corrupt
    \item Number of transactions aborted per second
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Time until 10\% database corruption2 $\boldsymbol{log(U)}$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
  \begin{center}
    For $\Delta = 50ms$, time taken for 10\% database corruption is between to $1$-$75$ years
  \end{center}
    \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.5]{./figures/delta}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Fraction of Aborts vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
  \begin{center}
    For $\Delta = 50 ms$, the fraction of aborts is between 1 − 5\%
  \end{center}
    \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.5]{./figures/aborts}
  \end{figure}
\end{frame}

### 2020-04-21 slides/s7-delta.tex
corrections++

@@
   \frametitle{Protocol Design Considerations}
-  Design Considerations:
+     \uncover<1->{  Design considerations:}
   \begin{enumerate}
-  \item Graph workloads exhibit high contention.
-  \item Graph transactions tend to be long-lived than those in other databases.
+     \uncover<2->{  \item Graph workloads exhibit high contention}
+     \uncover<3->{  \item Graph transactions tend to be longer-lived than those in other databases}
   \end{enumerate}
-\textbf{Protocol must permit multiple updates on the same record provided they are they are \textit{sufficiently} apart in time in ensure reciprocal consistency}
+     \uncover<4->{\textbf{Protocol must permit multiple updates on the same distributed edge provided they are \textit{sufficiently} apart in time to ensure reciprocal consistency}}
 \end{frame}
@@
   \begin{itemize}
-  \item Fact: a transaction updating a distributed edge must update one edge pointer then \textbf{immediately} update the other.
-  \item Rule: an update is permitted if the immediately preceding update was done at least $\Delta$ time before. Else, abort.
-  \item Assumption: the time interval that elapses between completing at update at one end and starting at the other can be estimated, $\delta$. Choosing $\Delta > \delta$.
+  \uncover<1->{  \item \textbf{Fact:} a transaction updating a distributed edge must update one edge pointer then \textbf{immediately} update the other}
+  \uncover<2->{  \item \textbf{Rule:} an update is permitted if the immediately preceding update was done at least $\Delta$ time before. Else, abort}
+  \uncover<3->{  \item \textbf{Assumption:} the time interval that elapses between completing an update at one edge pointer and starting at the other can be estimated to be $\delta$.}
+    \uncover<4->{ \item \textbf{Parameter Selection:} Choose $\Delta > \delta$}
   \end{itemize}
@@
   \begin{itemize}
-  \item Reciprocal consistency is preserved if the time taken to complete am update at one end and start at other end remains less than  $\Delta$
-  \item If $\Delta$ is exceeded then reciprocal inconsistency can occur
-  \item Setting a large $\Delta$ tends to preserve consistency but leads to more aborted transactions
+    \uncover<1->{  \item Reciprocal consistency is preserved if the time taken to complete an update at one edge pointer and start at other remains less than $\Delta$}
+    \uncover<2->{  \item If $\Delta$ is exceeded then reciprocal inconsistency and hence semantic corruption can occur}
+      \uncover<3->{  \item Setting a large $\Delta$ tends to preserve consistency but leads to more aborted transactions}
   \end{itemize}

### 2020-04-21 slides/s8-evaluation.tex
corrections++

@@
   \frametitle{Performance Evaluation}
-  Two metrics the following two metrics for various values of $\Delta$:
+ Through simulations the following two metrics for various values of $\Delta$ were evaluated:
   \begin{itemize}
-  \item Time taken for 10\% of a large database to be corrupt
-    \item Number of transactions aborted per second
+  \item Time taken for 10\% of a large database to become semantically corrupt
+    \item Fraction of transactions aborted per second
   \end{itemize}
@@
 \begin{frame}
-  \frametitle{Time until 10\% database corruption2 $\boldsymbol{log(U)}$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
+  \frametitle{Time until 10\% semantic corruption $\boldsymbol{log(U)}$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
   \begin{center}
-    For $\Delta = 50ms$, time taken for 10\% database corruption is between to $1$-$75$ years
+    For $\Delta = 50ms$, time taken for 10\% of a large database to become semantically corrupt is between to $1$-$75$ years
   \end{center}
@@
 \begin{frame}
-  \frametitle{Fraction of Aborts vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
+  \frametitle{Fraction of Aborts $(\boldsymbol{\alpha})$ vs Transaction Arrival Rate $(\boldsymbol{\lambda})$}
   \begin{center}
